### 2017-10-18 paper/main_RA-L_v1.tex
commit message

%\documentclass[9pt,technote,fleqn]{IEEEtran}
\documentclass[letterpaper, 10 pt, conference,fleqn]{ieeeconf}

\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{url}
\usepackage[font={footnotesize}]{caption}
\usepackage[font={footnotesize}]{subcaption}
\usepackage[usenames, dvipsnames]{color}
\usepackage{amsmath}
\renewcommand{\labelenumii}{\Roman{enumii}} 
\usepackage{authblk}
\usepackage{hyperref}
% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
\usepackage{amssymb}

\let\labelindent\relax
\usepackage{enumitem}
\usepackage{dsfont}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}

\usepackage{stfloats}
\IEEEoverridecommandlockouts

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{Learning Joint-Space Stable Dynamical Systems for Task-Space Objectives}
\title{Learning Augmented Joint-Space Task-Oriented Dynamical Systems: \\ A
Linear Parameter Varying and Synergetic Control Approach}

%OR
%\title{Learning Augmented Joint-Space Task-Oriented \\ Dynamical Systems in Synergy Space}

%
%
\author{Yonadav~Shavit*,  Seyed~Sina~Mirrazavi~Salehian*, Nadia~Figueroa*, Aude~Billard
% <-this % stops a space
\thanks{*These authors contributed equally to this work.}
\thanks{Y. Shavit is with the Department
of Electrical Engineering and Computer Science, Massachusetts Institute of Technology, Cambridge,
MA, 02139 USA. E-mail: yonadav@mit.edu.}% <-this % stops a space
\thanks{S. S. M. Salehyan, N. Figueroa, and A. Billard are with the Swiss Federal Institute of Technology (EPFL), 1015 Lausanne, Switzerland. \\E-mail: \{sina.mirrazavi,nadia.figueroafernandez,aude.billard\}@epfl.ch.}
}% <-this %

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\begin{abstract}
\textcolor{red}{Too long, re-write}
In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space with respect to the robot's kinematic constraints. Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, we prove that it stably moves towards its target and is guaranteed to obey the robot's kinematic joint limits. To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system, enabling us to encode complex joint-space behaviors. We propose a method to train such complex behaviors from kinesthetic demonstrations, by learning the parameters of the LPV system. We begin by applying a latent-embedding and a probabilistic approach to discover and represent a set of local behaviors from the demonstrated motions. Then, the corresponding local joint-space behaviors are estimated by solving a convex semi-definite optimization problem that minimizes the joint velocity error from the demonstrations. Our proposed approach is validated on a variety of motions for a 7-DOF KUKA LWR4+ robot arm.\end{abstract}

%\begin{IEEEkeywords}
%Dynamical Systems, Imitation learning, Robot Kinematics, Gaussian Mixture Models
%\end{IEEEkeywords}

\IEEEpeerreviewmaketitle
\section{Introduction}
\label{sec:intro}
Robot motion planning in joint-space has long been a major field of study \cite{kelly2006control}. For manipulation problems with an objective defined in task space (i.e. target or desired trajectory), we can often find a myriad of joint-space trajectories to achieve the same task-space goal. In many cases, however, certain joint-space trajectories are favored over others; for example, when we expect the robot to follow a desired joint-space behavior or ``style", as illustrated in Fig. \ref{fig:robot_example}.

Learning robot motion generators that follow desired joint and/or task-space behaviors is, in fact, the central focus of Programming by Demonstration (PbD) \cite{billard2008robot} \cite{Argall:RAS:2009}. Historically, much of the original work in PbD focused on learning motions solely in joint-space by parameterizing motions using GMMs \cite{Calinon:HRI:2007}, HMMs \cite{Garrido:Neuro:2015}, GPs \cite{Shon:HUM:2005}, or other probabilistic models \cite{Schaal:IROS:2003,Schaal:AI:2002}. Concurrently, a significant body of work from the graphics community was targeted at learning joint-space motion ``styles" from human motion capture data, through a variety of latent-space and stylistic probabilistic motion models \cite{gielniak2010stylized,grochow2004style}. Yet, over time, research shifted to pure task-space learning \cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}, which has two main advantages over pure joint-space: \textit{generalization} and \textit{re-usability}. By learning motions in task-space, one can concisely encode the task-relevant features and generate new motions which meet these task-space criteria. Furthermore, a task-space trajectory is robot-independent, and thus can be reused across robots with differing morphologies.

A major drawback of these initial task-space methods was that despite learning motions similar to the demonstrations, there was no guarantee when applying the learned motion to a new problem (e.g. a different initial/target position) that the end-effector would stably converge to its ultimate task-space target. This led to the introduction and subsequent popularity of task-space learning techniques such as SEDS \cite{khansari2011learning}, $\tau$-SEDS \cite{Neumann:RAS:2015}  and LMDS \cite{Kronander:RAS:2015} which encode task-space behaviors as asymptotically stable dynamical systems (DS) guaranteed to converge towards a desired task-space target.

\begin{figure}[t]
\centering
\includegraphics[scale=.23,trim={0 0 0 14.0cm},clip]{Without_Obstcle.jpg}
\includegraphics[scale=.23,trim={0 0 0 9.cm},clip]{With_Obstcle.jpg}
\caption{Two robot motions in joint-space accomplishing the same behavior in task-space (pouring chips in a bowl). Though the \textcolor{cyan}{\textbf{task-space trajectories}} are \textit{similar}, the \textcolor{orange}{\textbf{joint-space trajectories}} are different; e.g. the bottom example moves in joint-space to avoid an obstacle that the top one does not.}
\label{fig:robot_example}
\vspace{-0.6cm}
\end{figure}

For all the previously mentioned task-space techniques, the robot is ultimately controlled by projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}. When the main focus is on executing a complex task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}. However, for other applications, such an approach yields significant problems \cite{buss2004introduction}. First of all, finding the pseudo-inverse is computationally taxing. Moreover, when the Jacobian matrix cannot be inverted (i.e. when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior. These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}.

Although DS-based task-space learning techniques ensure target convergence and stability, an advantageous feat within the LfD paradigm, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution. Further, for certain tasks, rather than blindly choosing the arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.  For example, throwing a ball is a motion less characterized by task-space behaviors (the ball moving backwards and then forwards) and more characterized by the joint behaviors: the elbow folds, the shoulder moves back and then swings forward, followed by the elbow straightening and the wrist snapping down. For a visual example of the need for such combined task/joint-space behavior learning, see Fig. \ref{fig:robot_example}. This is precisely the shortcoming associated with LfD in task-space: it fails to learn the joint-space behavior of the system. The current formulation of the DS-based learning techniques introduced thus far (e.g. SEDS, $\tau$-SEDS LMDS, etc.) do not allow for integrating combined task/joint-space behavior learning. \cite{calinon2008probabilistic} notably attempted to address this problem by learning separate motion policies in task space and the null space of the Jacobian (which would not affect task-space position), and driving the robot with a weighted sum of the two. However, this approach no longer guarantees convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian.

%\cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework, but does not propose a dynamical system for generating motions.

In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only complies with the task-space constraints, but also avoids the problems generated by pseudo-inverse approximations and takes into consideration the robot's kinematic limits. \cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities and joint kinematic limits. The resulting DS avoids singularities through generalization of the pseudo-inverse approximations and joint limits via online modulation. However, because the two DS have their own unique attractors, the non-linear interaction between them imposed by the kinematic constraints does not ensure that the combined DS has a unique attractor. This gives rise to spurious attractors or cycles, and thus must be carefully tuned in order to avoid them. In this work, we depart from any modulation or combination of DS and instead propose a novel dynamical system which:

\begin{enumerate}[leftmargin=*]
\item computes a \begin{bf}complex motion in joint space\end{bf} that provably and stably converges to a \begin{bf}fixed task-space target\end{bf}.
\item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic ``singularities".
\item is formulated such that complex \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.

\item is capable of complying with \textbf{kinematic joint limits}.
\end{enumerate}

The most similar approach to our proposed DS is the Jacobian transpose (JT) control \cite{wolovich1984computational} method. JT control is an inverse kinematics
method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably (in the sense of Lyapunov) over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations. It shares some of our approach's advantages: fast computation and provable task-space stability. However, despite some previous work integrating discontinuous joint limits \cite{sciavicco1988solution} and designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations while obeying the robots kinematic constraints.

This paper is organized as follows. Section \ref{Sec:Prob} formalizes the problem. The proposed dynamical system is introduced in Section \ref{Sec:DS}. In Section \ref{Sec:Learning}, a probabilistic model is introduced to approximate the parameters of the dynamical system. In addition, a convex optimization problem is  formalized to estimate these parameters. The effectiveness of the proposed method is shown through simulations and experiments on a 7-DOF robot-arm in Section \ref{Sec:Exp}. The paper is finalized with a discussion over our method and results in Section \ref{Sec:Dis}.

%\begin{figure}[t]
%\centering
%\includegraphics[width=0.95\linewidth,trim={0 20 450 0},clip]{Pic/Schematic.jpg}
%\includegraphics[width=0.9\linewidth,trim={580 20 0 0},clip]{Pic/Schematic.jpg}
%\caption{ Standard control architecture using Cartesian dynamical systems (DS)\textbf{(top)}  and our proposed approach \textbf{(bottom)}.}
%\label{fig:schematic}
%\end{figure}

\section{Problem Statement} \label{Sec:Prob}
Consider a robotic system with $d$ task-space dimensions and $m$ degrees of freedom. We direct the system using a joint position or joint velocity controller, which can have joint position limits and a maximum joint velocity. We are further provided with a set of $N$ demonstrated joint-space trajectories $D=\{\{q_{t,n},\dot{q}_{t,n}\}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$, where $T_n$ is the number of the sample points of the $n^{\text{th}}$ demonstration. We refer to the system's joint-space position as $q=\begin{bmatrix} q^1 & \dots & q^m
\end{bmatrix}^T \in \mathbb{R}^m$, and to its task-space position as $x\in \mathbb{R}^d$.\footnote{For sake of brevity and simplicity, the time index, $t$, is dropped throughout the paper. } The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{m\times d}$.\\
We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following criteria:
\renewcommand{\labelenumi}{(\Roman{enumi})}
\begin{enumerate}
\item The dynamical system must be Lyapunov stable\footnote{Unless otherwise specified, "stability" in this paper always refers to global Lyapunov stability. We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system. For example, any finite-size manipulator will always have regions beyond its reach that it cannot converge to.} with respect to a fixed task-space target $x^*$. This can be expressed by ensuring that the following Lyapunov function
\begin{equation}
V(q) = (H(q) - x^*)^TP(H(q) - x^*) 
\label{eq:Lyp}
\end{equation}
is stable; i.e. $\dot{V}(q)\leq0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$. Where  $Q=\{q|q^i_{min} < q^i < q^i_{max},~\forall i\in \{1,\dots,d\}\}$ and $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$. $P\in \mathbb{R}^{d\times d} $ is a symmetric and positive definite matrix. $V(.)$ can be thought of as a metric for the task-space distance-to-go.
\item The dynamical system must respect the kinematic joint constraints of the robot:
\begin{equation}
q(0)\in Q \rightarrow q(t)\in Q~~~~\forall t>0~
\end{equation}
\item The dynamical system should encapsulate the desired joint-space behaviors such that the following metric is minimized
\begin{equation}
\vspace{-5pt}
e_{total} = \frac{1}{NT_n}\sum_{n=1}^N\sum_{t=0}^{T_n} \left \| \dot{q}_{d; t,n} - f(q_{t,n}) \right \|
\label{eq:optim_error}
\end{equation}
  where $\dot{q}_d$ is the desired ``true" velocity, and $f(\cdot)$ is the motion generation policy.
\end{enumerate}

%Lastly, we wish to capture joint-space behaviors in our dynamical system.
We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}. First, the metric implicitly suggests that behaviors are defined by expert motions of the behavior, and that fulfilling a behavior means moving similarly to the experts' motion. Second, the type of feature of the trajectory that is being minimized is important. If instead the error feature were joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity. Alternatively, if the error feature were joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior). By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
\section{Augmented \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} \label{Sec:DS}
\label{sec:proposed_system}
We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the three criteria presented in (I), (II), (III).
\begin{equation}
\label{eq:ds}
\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
\end{equation}
where 
\begin{equation}
\label{eq:SAS}
\mathcal{A}(q) = S(q) A(q) S^T(q). 
\end{equation}
$S(q)=diag(s^1(q),\dots,s^d(q))$ is a diagonal matrix and $ s^i(\cdot)\in C^0~\forall i \in \{1,\dots d\} $. $A(q)\in \mathbb{R}^{d\times d}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a nonlinear combination of constant linear matrices, each of which encodes a local joint-space behavior. 
\begin{equation}
\label{eq:A_def}
A(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
\end{equation}
where $A_k\in \mathbb{R}^{d\times d} $ and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are constant matrices and the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. It is noteworthy that the presented stability and convergence proof can easily be extended for $\theta_k (t, q(t), d(t))$.} representing each local behavior, respectively. Based on \eqref{eq:ds}, one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.

\begin{figure}[t]
\centering
\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
\caption{Three example 3-DOF motions (A, B, C), each with a different constant behavior matrix $\mathcal{A}$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$. On the left, the task-space traces of each motion. On the right, the time-scaled joint positions of each joint. Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
\label{fig:A_example}
\vspace{-15pt}
\end{figure}

Before proving that the proposed dynamical system satisfies all of the criteria, let us establish an intuitive understanding of the components of the system. While it may seem daunting at first, each of the elements has a straightforward explanation. One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}). The positive semi-definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig. \ref{fig:A_example} illustrates the effects of  $A(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space. We refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities). Lastly, $S(q)$ intuitively enforces the joint constraints: as $q^i$ nears $q^i_{min}$, the velocity in that direction goes to $0$, and similarly when it nears $q^i_{max}$ the velocity goes to $0$.

%When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle). This ''augmentation" is simply a matrix multiplication ($A\dot{q}_{raw}$). The components of the outputted velocity are linear combinations of the ''raw" un-augmented joint velocities, where the entries in $A$ define the precise ratios. We can see the joint positions resulting from our experiment over time on the right side of Fig. \ref{fig:A_example}. Each motion tends to use its ``primary" joint most (the joint where the velocity is heavily scaled up) and uses the other available joints to compensate for what the primary joint cannot do. For example, Motion 1 starts out using mostly the hip, but once moving the hip no longer brings the system closer to the goal, the system uses the other joints to fold inwards and reach the goal. Ultimately, all three controllers converge on their target, but each has a different path through joint space depending on which joints we choose to emphasize in the behavior matrix $A$. The off-diagonal terms of $A$ have a more nuanced contribution; e.g. $A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.

\newtheorem{prop1}{Proposition}
\begin{prop1}
\label{prop:stability}
The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) and (II), if \eqref{eq:ds} meets the following constraints.

\begin{subequations}
\vspace{-5pt}
\begin{equation}
\label{eq:first_criteria_stability}
\begin{cases}
\begin{split}
&0 \prec P &\\
&0 \preceq A_k & &~\\
& 0 \leq \theta_k(\cdot) & &\\
%    & 0 \preceq S(q) &\forall q \in Q\\
\end{split}
\forall k \in \{1,\dots,K\} 
\end{cases}
\end{equation}    
\begin{equation}
\label{eq:first_criteria_joint}
\begin{cases}
\begin{split}
&s^{i}(q^i_{min})=0 & &\\
&s^{i}(q^i_{max})=0 & &
\end{split}
\forall i \in \{1,\dots,d\}
\end{cases}
\end{equation}
\end{subequations}\\
\end{prop1}
\vspace{-10pt}
\textbf{Proof}: See Appendix \ref{appendix:stability}. $\blacksquare$\\

It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  do not ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$, as $\exists q\in Q-Q^* $ such that both $\dot{V}(q)=0$ and $H(q) \neq x^*$. This happens if the target position is kinematically unreachable; i.e. for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or $S_{i,i}(q) = 0$ (the joint has reached its limit).  Note that in practice, because $s(q)$ is continuous and designed based on \eqref{eq:first_criteria_joint}, executing JT-DS \eqref{eq:ds} will likely never cause the manipulator to reach the joint limit. Instead, $s(q)$ significantly shrinks the magnitude of a joint's motion when a joint nears its motion boundary. For a simple method of constructing $S(q)$, see Appendix \ref{appendix:limits}.

Criterion (III) (i.e. encoding specific joint-space behaviors) is achieved by embedding the desired dynamics in the matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$. We describe in the next section an approach to automatically learn these matrices from demonstrated data.

%We achieve criterion (III), encoding joint-space behaviors in \eqref{eq:ds}, by approximating the joint augmentation matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$. It is important to understand how varying $A_k(q)$ will vary the system's joint space behavior. We know from \textit{Proposition  \ref{prop:stability}} that, regardless of choice of $A_k(q)$, the robot's motion is stable with respect to the target $x^*$. Yet this still leaves us with many different potential joint velocities to choose from, all of which would decrease $(H(q)-x^*)^TP(H(q)-x^*)$. The choice of $A_k(q)$ becomes our way to influence which of the many possible joint velocity motions is chosen. Thus we must clarify how the choice of $A$ affects the system's motion.

%By learning a position-dependent joint augmentation matrix $A(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints. \eqref{eq:A_def} constructs our overall $A(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion. In this paper the joint behavior matrix $A(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\

\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem in Synergy Space} 
\label{Sec:Learning}

\textcolor{red}{Talk about synergy space here, why we need it, etc.}\\

JT-DS \eqref{eq:ds} is parametrized by $\theta_k(q) $ and $A_k~\forall k\in\{1,\dots,K\}$. One is free to determine these parameters as long as that they satisfy constraints \eqref{eq:first_criteria_stability}. We propose a 3-step procedure as follows:
\begin{enumerate}
\item We first project the demonstrated data (i.e. collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$ through Principal Component Analysis (PCA) \cite{jolliffe1986pca}, so as to reduce the dimensionality and extract correlations across similar joint behavior regions.

\item We then jointly estimate the optimal number $K$ of local behavior regions and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$.

\item Finally, once the local behavior regions have been found (described by each of the Gaussian distributions $\theta_k(q)$), we compute the corresponding joint behavior matrices $A_k~\forall k\in\{1,\dots,K\}$ for each region by formulating a convex optimization problem that finds the optimal set of $A_k$'s that minimize the overall velocity error with respect to the demonstrations \eqref{eq:optim_error}.
\end{enumerate}

\subsection{Embedding Joint Configurations in Latent Space}
The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint motion styles" (defined in Section \ref{sec:intro}). It is important to understand that distances in joint space (defined as $\| \Delta q\|$) might not numerically reflect what a human intuitively perceives as the distance between joint configurations. Motor control studies have postulated that most human arm motions, be they reaching motions or following trajectories as straight/curved lines, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}. This suggests that human arm motion in general tends to move on a plane, and thus can be represented in such lower-dimensional space.

To this end, any type of topology-preserving latent-space embedding or dimensionality reduction approach can theoretically be used for $\phi(q)$. In this work, we choose to construct the projection $\phi(q)$ of $q$ through a linear map $W:q \in \mathbb{R}^{d} \rightarrow \phi(q) \in \mathbb{R}^{\delta\leq d}$ given by $\phi(q) = Wq$, where the projection matrix $W \in \mathbb{R}^{\delta \times d}$ is computed through Principal Component Analysis (PCA). Although the projected dimensions extracted from PCA are linear combinations of the original variables, this has been proven to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}.

\subsection{Discovering Local Behavior Regions}
Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local behaviors, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$. Moreover, we would like for our scheduling parameters $\theta'_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta'_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.

Scheduling parameters for LPV systems with such properties have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}. In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the scheduling parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and covariance matrices that parametrize the $k$-th multivariate Gaussian distribution. Each such distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ of our LPV system. We will define $\theta'_k(\phi(q))$ as $p(k|\phi(q))$:
\begin{equation}
\label{eq:theta}
\theta'_k(\phi(q))= \frac{\pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}
\end{equation}
which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th behavior. Therefore, each joint-space behavior region is associated with a Gaussian component of the GMM, cumulatively describing all the behavior regions of the dynamical system. We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM and choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.

\subsection{Estimating the Joint Behavior Matrices}
Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $A(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows: 
\begin{equation}
\label{eq:A}
A(q) = \frac{\sum_{k=1}^K A_k \pi_k\mathcal{N}(\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}.
\end{equation} 
Notice the resemblance of \eqref{eq:A} to the Nadaraya-Watson kernel estimator \cite{nadaraya1964regress,watson1964regress}\footnote{The Nadaraya-Watson kernel estimator is used to estimate an unknown regressive function $m(x) = \mathbb{E}\{Y|X\}$, which takes the general form of $\widehat{m}(x) = \frac{\sum_{i=1}^n y_i \mathcal{K}(x,x_i)}{\sum_{i=1}^n \mathcal{K}(x,x_i)}$ where $\mathcal{K}(x,x_i)$ is a kernel function denoting the distance or similarity of $x_i$ to the given location $x$. \cite{nadaraya1964regress,watson1964regress}} with a Gaussian pdf as its kernel function. Hence \eqref{eq:A} can be considered a type of kernel estimator, with the key distinction that the weighting functions $\theta_k'(\phi(q))$ are not determined by individual points (as in the original Nadaraya-Watson kernel estimator) but by the components of a GMM, similar to the weighting functions derived in Gaussian Mixture Regression (GMR) \cite{sung2004gmr}.

Intuitively, each ``region of behavior" in joint/latent space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's behavior ($A_k$) influences the robot's current joint-space motions. Finding the appropriate joint augmentation matrices $A_k$ for the set of local behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}. To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:
\begin{equation}
\vspace{-10pt}
\begin{aligned}
&\underset{A_1, \dots, A_K}{\text{min}} 
\sum_{n=1}^N\sum_{t=0}^{T_n} \left \| \dot{q}_{t,n} - f(q_{t,n}) \right \| \\
& \text{subject to}&\\
&  0 \preceq A_k, \; \forall k \in \{1, \dots, K\}. &
\end{aligned}
\label{eq:opt}
\end{equation}
where $f(q_{t,n})$ is calculated by combining our dynamical system formulation \eqref{eq:ds} with \eqref{eq:A}, and $x^*_{n}$ is defined as the endpoint of the $ n^{\text{th}} $ demonstrated trajectory.

\section{Experimental Validation} \label{Sec:Exp}

\subsection{JT-DS Learning  Evaluation}

The learning algorithm presented in the previous section, \eqref{eq:opt}, was implemented in MATLAB. In order to solve the semidefinite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used. In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis. The source code for learning (MATLAB) and execution (C++) of the Augmented Joint-Space Task-Oriented Dynamical System together with data generated in these experiments are available on-line: 
\centerline{\small \underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}}

\textcolor{red}{1. Introduce Datasets and metrics}\\

\textcolor{red}{2. Table with Evaluation of Learning Schemes for JTDS Models on Different Datasets  and discussion of results}

\newpage
\subsection{JT-DS Performance Evaluation}
The performance of the proposed framework was evaluated on a 7-DOF robot arm, the KUKA LWR 4+. The robot is controlled on the joint position level (linearly interpolating from joint velocities) at a rate of $ 500 $ Hz. The resultant joint angles are filtered by a critically damped filter to avoid high torques. Our empirical validation consists of three sections, each highlighting a different advantage of the proposed dynamical system: (i) moving in singular configurations, (ii) following desired behaviors in joint and task spaces simultaneously, and (iii) fast computation/convergence time.

\subsubsection{Systematic Assessment}
To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system. The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$. The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver. To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single Gaussian, i.e. uniform behaviors.  The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse. Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero. Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.

\begin{table}[t]
	\scriptsize
	\vspace{0.1cm}
	\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$. The convergence duration is the time required to move within $0.001$m of the target. The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
	\label{table:2}
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ \hline
		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline
	\end{tabular}
	\hspace{0.1cm}
	 \vspace{10pt}	
	\caption{Parameters and Performance of Comparative Experiments
	 \textcolor{red}{This table is wrong.. re-do} }
	 \vspace{-5pt}
	\centering
	\begin{tabular}{|l|c|c|c|c|}
		\hline
		& No. Dem. & No. K & $\phi(q)$ dim. & $e_{totol}$ [{rad}/{s}] \\ \hline
		Pour Obst. (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ \hline
		Foot (JT-DS) &       3               &     1             &          1 & 0.01344   \\ \hline
		Pour Obst. (SEDS) &      7                &     3             &        -  &   -   \\ \hline
		Foot (SEDS)&       3               &     1             &          - & -   \\ \hline
	\end{tabular}
	\label{table:1}
	\vspace{-5pt}
\end{table}

\subsubsection{Following Desired Joint Behaviors}

To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS \cite{khansari2011learning} algorithm using damped least square IK. In the first experiment, the robot was guided through a complex motion task, moving an object through an environment with an obstacle. 
Human experts guided the robot through a series of demonstrations, ensuring that neither the robot's joints nor the held object intersected the obstacle. The results can be seen in Fig. \ref{fig:snapshot:a} and Fig. \ref{fig:snapshot:b}. The JT-DS algorithm mimicked the joint-space behaviors of the demonstration (e.g. folding the elbow, lowering the shoulder), and therefore successfully avoided the obstacle while still converging to the desired Cartesian position. Meanwhile, SEDS only learned the demonstrated behavior in task-space (the position of the end-effector is signified by a black ball) meaning that the robot was not constrained in joint space. This ultimately led to one of its joints colliding with the obstacle. It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
In the second experiment the robot was taught a footstep-like motion, beginning with a straight leg, moving through a singularity, and finally bringing the knee up (Fig. \ref{fig:snapshot:c}). JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig. \ref{fig:foot}). This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.

Further inspection reveals another advantage of JT-DS over stable Cartesian motion generators: not needing to preprocess demonstrations. In Cartesian-space dynamical systems, in order to guarantee stability, every demonstration must have the same target position. As a result, all demonstrated trajectories are warped to end at a single position, introducing a non-negligible difference between the original demonstrations and those used to train the DS. JT-DS, on the other hand, does not require any demonstration warping to learn. Rather than shifting demonstrations, the algorithm learns behavior regions with respect to the base frame of the manipulator. This effect can be seen in the diagrams in Fig. \ref{fig:Sin}, which show a comparison between JT-DS and a Cartesian-space DS trained through SEDS \cite{khansari2011learning}. The Cartesian-space system warps the demonstrations and as a result fails to track them, while JT-DS does not modify the demonstrations and thus succeeds.

\subsubsection{Moving in singular configurations}
One of the main advantages of the proposed dynamical system is its ability to generate accurate paths in classical singular configurations. The first scenario, shown in Fig. \ref{fig:Sin}, was designed to assess this capability. The $10$ training demonstrations were constructed as follows: movement was constrained to the boundary of the workspace by fixing $ q^i=0~\forall i\in\{3,\dots,7\}$, the second joint was fixed to $q^2\in\{10^\circ,20^\circ,\dots,100^\circ\}$, and only the first joint was moved by a human demonstrator back-driving it. This restricted the motion to a series of arcs of different radii along the robot's motion boundary. In other words, the demonstrated motions were entirely within a classic kinematic singularity. Fig. \ref{fig:Sin} shows the \textit{demonstrated} motions and the motion  \textit{generated} by JT-DS \eqref{eq:ds}. The algorithm never requires the pseudo-inverse of the Jacobian matrix, so the generated motion perfectly follows the demonstrations throughout the workspace boundary.

\begin{figure}[h]
	\begin{subfigure}[t]{\linewidth}
		\includegraphics[width=\linewidth]{Pic/cropped_Sing_1.pdf}
		\caption{ }
		\label{fig:Sin}
	\end{subfigure}\\
	\begin{subfigure}[t]{\linewidth}
		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
		\caption{ }
		\label{fig:Det}
	\end{subfigure}\\
	\caption{The experiment generated $K=3$ Gaussian components. As the first joint is the only joint which was not fixed during the demonstration, the learned augmentation matrices had only one nonzero entry $A_k(1,1)\neq0~\forall k\in \{1, 2, 3\} $.  In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space. The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable). In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed. Due to sensory noise, the determinants are not exactly zero, but are very close.  }
	\vspace{-20pt}
\end{figure}

\section{Discussion and Future Work} 
\label{Sec:Dis}
In this paper, we have presented a dynamical system in joint space that is provably Lyapunov-stable in task space and which replicates demonstrated joint-space behaviors. The system provably obeys the manipulator's kinematic joint limits, is fast to compute, and smoothly handles singularities by avoiding the pseudo-inverse Jacobian. We showed the system's ability to learn different joint-space behaviors on a redundant robotic platform.

\begin{figure}
\vspace{-10pt}
	\centering
	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
	\caption{A plot of the end-effector trajectories for the footstep motion in Cartesian space. The JT-DS motion moves smoothly closely resembles the demonstrated trajectories. On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing due to becoming too large. \label{fig:foot}}
	\vspace{-24pt}
\end{figure}

One of the most important points when validating a learning from demonstration method is to evaluate the system's behavior away from demonstrations. When the current joint configuration is far from any of the local behavior regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target. Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably moving to the target. The reason for this is that rather than ``determining" the velocity of the system, our $A$ matrix only warps it. This means that adding multiple ``conflicting" behaviors together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.

One of the major advantages of the proposed dynamical system is that it avoids the undesirable effects of traditional Inverse Kinematic (IK) solvers. Fast dynamical motions require not only fast and accurate motion planning, but also precise IK mapping, which is not practically feasible. By utilizing the JT-DS method, even a fast system could reactively generate stable motions to its target.

The JTDS algorithm can also be expanded to accommodate joint velocity limits (in addition to joint position limits) by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits. Specifically, given some set of velocity constraints for each joint $i$ s.t. $\dot{q}_i < \dot{q}_{i, max}$, we can design a new joint augmentation matrix $\mathcal{A}' = r \mathcal{A}$ where $r = \min \left ( 1, \max_i \left ( \frac{\dot{q}_{i, max}}{\dot{q}_i} \right ) \right )$. This velocity scalar would slow down the joint trajectory when the robot dynamics could not execute them, but would not change the position profile of the joint motion, preserving a major part of the joint behavior.

A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists. As mentioned in Sec. \ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^TP(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits. If the only valid trajectory to reach the goal requires that we temporarily move away from the target (and increase the warped distance-to-goal), this controller will not find that trajectory. There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession. The smaller subtrajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits. Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.

One application that we have discussed for this algorithm is obstacle avoidance. In an environment with static obstacles, the system would, through demonstration, learn joint-space behaviors that avoid those obstacles. One can compare this form of joint-space obstacle avoidance with other methods from the literature \cite{sciavicco1988solution,petrivc2013smooth}. The main difference is that these previous methods require explicit knowledge of the obstacles' position, whereas our algorithm learns the obstacle positions only implicitly by learning motions that avoid those positions. This lack of explicit obstacle location can be an advantage: encoding obstacles' geometry can be expensive and cumbersome. However, it is also a drawback: without knowing the exact location of obstacles, the algorithm cannot be guaranteed to avoid collisions, especially away from demonstrations.

Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system. This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors,.
%\footnotesize
\appendices

\begin{figure}[t]
\vspace{-0.05cm}
	\begin{subfigure}[t]{\linewidth}
		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
		\caption{Desired Joint Behavior: obstacle avoidance motion. The luggage is the obstacle.}
		\label{fig:snapshot:a}
	\end{subfigure}\\
	\begin{subfigure}[t]{\linewidth}
		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}
		\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
		\includegraphics[width=0.32\linewidth]{Pic/Luggage3_SEDS.png}
		\caption{Desired Behavior: The Cartesian DS is used to generate the motion.}
		\label{fig:snapshot:b}
	\end{subfigure}\\
	\begin{subfigure}[t]{\linewidth}
	\centering	
		\includegraphics[width=0.45\linewidth]{Pic/1.png}~~
		\includegraphics[width=0.45\linewidth]{Pic/3.png}\\\vspace{0.1cm}
		\includegraphics[width=0.45\linewidth]{Pic/4.png}~~
		\includegraphics[width=0.45\linewidth]{Pic/2.png}
		\caption{Desired Joint Behavior: Foot step-like motion.}
		\label{fig:snapshot:c}
	\end{subfigure}\\	
	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{https://youtu.be/1dgKfmN1UgE}.}
	\label{fig:snapshot}
	\vspace{-0.8cm}
\end{figure}

\section{Proving Stability of the Dynamical System}
\label{appendix:stability}
We wish to prove Proposition \ref{prop:stability}, that is, that JT-DS \eqref{eq:ds} (reproduced below)
\begin{equation*}
\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
\end{equation*}
accomplishes criterion (I), where $\mathcal{A}$ is positive semi-definite, and $P$ is positive definite.
\newtheorem{theorem}{Theorem}[section]
\begin{theorem}[Proof of Lyapunov Stability]
JT-DS \eqref{eq:ds} is stable in the sense of Lyapunov with respect to the Lyapunov candidate
\[ V(q) = \frac{1}{2}(H(q) - x^*)^TP(H(q) - x^*) \]
That is, 
\begin{equation}
\begin{aligned}
& 0 \prec V(q)    &
V(q^*) = 0 & \quad \forall q \neq q^*\\
%    \[\dot{V}(q) = \frac{dV}{dq}\frac{dq}{dt}(q) \preceq 0 \quad \forall q \in Q\]
\end{aligned}
\end{equation}
where $q^*$ is any joint configuration such that $H(q^*) = x^*$.
\end{theorem}
The first two statements are trivially true because $P$ is positive definite and the rest of $V$ is a square that is only $0$ when $H(q) = x^*$. To prove the last statement, we find $\frac{dV}{dt}$:
\begin{equation}
\begin{aligned}
\frac{dV(q)}{dt} &= (H(q) - x^*)^TPJ(q)\dot{q}\\
&= -(H(q) - x^*)^TPJ(q)\mathcal{A}(q)J^T(q)P(H(q) - x^*)\\
&= -(H(q) - x^*)^TPJ(q)S(q) \sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot S^T(q)J^T(q)P(H(q) - x^*) \leq 0
\end{aligned}
\end{equation}
By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $A$, which is positive semi-definite. This means that the expression is guaranteed to be negative semi-definite. Therefore, JT-DS \eqref{eq:ds} is globally Lyapunov stable; i.e. $ \| H(q)-x^*\| $ and $ \dot{q} $ are bounded.

To prove the fulfillment of criterion (II), assume the contrapositive, that there exists some time $0<\tau$ such that $q^i_{max} < q^i(\tau) $. Because $\dot{q}^i$ given by \eqref{eq:ds} with respect to \eqref{eq:first_criteria_stability} and  \eqref{eq:first_criteria_joint}  is both bounded and continuous, we can invoke the intermediate value theorem to show that there must have been some $0<c<\tau$ s.t. $q^i(c) = q^i_{max}$. According to \eqref{eq:first_criteria_joint}, $s^i(q^i(c)) = 0$, which ensures that $\dot{q}^i(c) = f(q^i(c)) = 0$ and will be forever (by induction), which contradicts the assumption. The same proof holds for a $q^i(\tau) < q^i_{min} $.  $\blacksquare$
\section{Enforcing Joint Limits}
\label{appendix:limits}
We wish to find a joint-position-dependent diagonal matrix $S(q)$ which is bounded and satisfies:
\begin{equation}
\label{eq:last_criteria}
\begin{cases}
\begin{split}
& s^{i}(q^i_{max}) = 0&&&\\
& s^{i}(q^i_{min})= 0&&&
\end{split}
\forall i\in 1,\dots,m
\end{cases}
\end{equation}
as previously described in Section \ref{sec:proposed_system}. We propose the using the following functions:
\begin{equation}
s^{i}(q) =
1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{2d}  ~\forall i\in 1,\dots,m\\
\end{equation}
where $d$ is a positive integer. 
This function is continuous and satisfies all of the above criteria. We can understand it holistically as follows: the ratio $\frac{q(v) - q_{min}}{q_{max} - q_{min}}$ maps the set of valid $q$ values to the range between $0$ and $1$. The rest maps these values to a parabola-like polynomial, such that at $q^i = q^i_{max}$ or $q^i_{min}$ the value is $0$, and elsewhere the value is positive (with the maximum value, $s^i = 1$, at $q^i = \frac{q_{max} - q_{min}}{2}$). Note as $d$ gets bigger, $S(q)$ is flatter away from the joint limits and drops off more sharply near them, which allows us to preserve a wider range of motion. We found $d=2$ to be an effective choice.\\

\section*{Acknowledgment}
\footnotesize
This work was supported by EU project Cogimon H2020\textendash ICT \textendash 23\textendash2014. The authors would like to thank A. Karimi for his insightful comments about formulating the convex optimization problem.

\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{IEEEabrv,references}
%

% that's all folks
\end{document}

### 2017-10-18 devel/JTDS_py_lib/JSEDS.py

@@ -1,0 +1,138 @@
+import numpy as np
+from scipy.stats import multivariate_normal
+import sys
+sys.path.append('/home/yonadav/Research/joint-pbd/utils/robotics-toolbox-python/')
+from robot import *
+
+#JSEDS.MotionGenerator.CreateMotionGeneratorFromFile('/home/yonadav/Research/joint-pbd/kuka_demos/results/backhand_params.txt')
+def arrayFromNextLine(f):
+	return np.fromstring(f.readline(), sep=' ')
+
+class MotionGenerator:
+	def __init__(self, robotPlant, AGenerator, bound_type=None):
+		self.robotPlant = robotPlant
+		self.AGenerator = AGenerator
+		if bound_type != None and bound_type != 'discrete' and bound_type != 'continuous':
+			raise Exception('Motion generator bound type must be either "continuous" or "discrete"')
+		self.bound_type = bound_type
+
+		qmin = []
+		qmax = []
+		qdim = self.robotPlant.robot.n
+		for i in xrange(qdim):
+			qmin.append(self.robotPlant.robot.links[i].qlim[0,0])
+			qmax.append(self.robotPlant.robot.links[i].qlim[0,1])
+		self.qmin = np.asarray(qmin).reshape((qdim, 1))
+		self.qmax = np.asarray(qmax).reshape((qdim, 1))
+
+	@staticmethod
+	def CreateMotionGeneratorFromFile(filename, bound_type=None):
+		f = open(filename, 'r')
+		dimq, dimz, K = map(int, arrayFromNextLine(f)) # First line provides basic GMM parameters
+		Priors = arrayFromNextLine(f) # Next line is the priors
+		Mu = np.reshape(arrayFromNextLine(f), (dimz, K), order='F') # Next line is the mus
+		for i in xrange(K): # Next K lines are the sigmas, reshaped fortran style
+			sigma = np.reshape(arrayFromNextLine(f), (dimz, dimz), order='F')
+			if i == 0:
+				Sigma = sigma
+			else:
+				Sigma = np.dstack((Sigma, sigma))
+		for i in xrange(K): # Next K lines are the As, reshaped fortran style
+			A = np.reshape(arrayFromNextLine(f), (dimq, dimq), order='F')
+			if i == 0:
+				As = A
+			else:
+				As = np.dstack((As, A))
[…]

### 2017-10-18 devel/JTDS_py_lib/JSEDS_ROS.py

@@ -1,0 +1,42 @@
+#!/usr/bin/env python
+import rospy
+from std_msgs.msg import String
+from kuka_fri_bridge import JointStateImpedance
+
+q_initial = [-0.728371429443359,-1.3605418920517,2.69252680319093,0.620675325393677,0.955035626888275,0.141930669546127,-0.17979271709919]
+x_end = [-.5, -.5, 0.3]
+
+def listener_callback(data):
+
+    q_current = get_q_from_data(data)
+
+
+
+def talker():
+    pub = rospy.Publisher('/KUKA/joint_imp_cmd', JointStateImpedance, queue_size=50)
+    rospy.init_node('controller', anonymous=True)
+    rate = rospy.Rate(100) # 100hz
+
+
+    # Index of statuses:
+    # 0 - go to initial joint position
+    # 1 - do learned move to task target
+    status = 0 
+
+
+    while not rospy.is_shutdown():
+
+
+
+
+
+        hello_str = "hello world %s" % rospy.get_time()
+        rospy.loginfo(hello_str)
+        pub.publish(hello_str)
+        rate.sleep()
+
+if __name__ == '__main__':
+    try:
+        talker()
+    except rospy.ROSInterruptException:
+        pass

### 2017-10-18 src/JTDS_mat_lib/JTDS_Solver.m

@@ -1,0 +1,357 @@
+function [Priors, Mu, Sigma, As, latent_mapping] = JTDS_Solver(Data, robotplant, options)
+%JTDS_LEARNING 
+% This function takes as input a set of demonstrations and a kinematic
+% system, and outputs parameters for a Joint-space Task-oriented Dynamical
+% System model. 
+
+% Inputs -----------------------------------------------------------------
+%
+%   o Data:    A (2dimq + dimx) x N_Total matrix containing all demonstration data points.
+%              The rows are assembled as follows:
+%              1:dimq - joint angle values throughout the trajectories
+%              dimq+1:2dimq - first time derivative of joint angles
+%              2dimq+1:2dimq+dimx - target task position values
+%              Each column of Data stands
+%              for a datapoint. All demonstrations are put next to each other 
+%              along the second dimension. For example, if we have 3 demos
+%              D1, D2, and D3, then the matrix Data is:
+%                               Data = [D1 D2 D3]
+%              Not that this Data matrix is precisely the type outputted by
+%              preprocess_demos.m
+%
+%   o options: A structure to set the optional parameters of the
+%               joint-specific optimization. The options are specified
+%               below.
+%
+%
+%   ------------------- OUTPUTS ------------------
+%   o Priors: 1 x k, the priors of the GMM which encodes the spatial significance of each behavior
+%
+%   o Mu: qdim x k, the means of the GMM which encodes the spatial significance of each behavior
+%
+%   o Sigma: qdim x qdim x k, the covariance matrices of each component of 
+%               the GMM which encodes the spatial significance of each behavior
+%
+%   o As: qdim x qdim x k, the augmentation matrices defining each local behavior
+%
+%   o latent_mapping: a container, in the style of drtoolbox, for a 
+%               dimensionality reduction technique mapping a joint-space 
+%               position to alatent-space position
+%
+%
+%  ---------------------OPTIONS-------------------
+%
+%   o options.latent_mapping_type: the type of function used to reduce
+%           the dimensionality of the joint space, mapping from q to z.
+%           Must be one of:
+%               'PCA' - Principle Component Analysis
+%               'KPCA' - Kernel PCA with Gaussian kernel w/ variance 1 
[…]

### 2017-10-18 src/JTDS_mat_lib/JTDS_Solver_v2.m

@@ -1,0 +1,420 @@
+function [Priors, Mu, Sigma, As, latent_mapping] = JTDS_Solver_v2(Data, robotplant, options)
+%JTDS_LEARNING 
+% This function takes as input a set of demonstrations and a kinematic
+% system, and outputs parameters for a Joint-space Task-oriented Dynamical
+% System model. 
+
+% Inputs -----------------------------------------------------------------
+%
+%   o Data:    A (2dimq + dimx) x N_Total matrix containing all demonstration data points.
+%              The rows are assembled as follows:
+%              1:dimq - joint angle values throughout the trajectories
+%              dimq+1:2dimq - first time derivative of joint angles
+%              2dimq+1:2dimq+dimx - target task position values
+%              Each column of Data stands
+%              for a datapoint. All demonstrations are put next to each other 
+%              along the second dimension. For example, if we have 3 demos
+%              D1, D2, and D3, then the matrix Data is:
+%                               Data = [D1 D2 D3]
+%              Not that this Data matrix is precisely the type outputted by
+%              preprocess_demos.m
+%
+%   o options: A structure to set the optional parameters of the
+%               joint-specific optimization. The options are specified
+%               below.
+%
+%
+%   ------------------- OUTPUTS ------------------
+%   o Priors: 1 x k, the priors of the GMM which encodes the spatial significance of each behavior
+%
+%   o Mu: qdim x k, the means of the GMM which encodes the spatial significance of each behavior
+%
+%   o Sigma: qdim x qdim x k, the covariance matrices of each component of 
+%               the GMM which encodes the spatial significance of each behavior
+%
+%   o As: qdim x qdim x k, the augmentation matrices defining each local behavior
+%
+%   o latent_mapping: a container, in the style of drtoolbox, for a 
+%               dimensionality reduction technique mapping a joint-space 
+%               position to alatent-space position
+%
+%
+%  ---------------------OPTIONS-------------------
+%
+%   o options.latent_mapping_type: the type of function used to reduce
+%           the dimensionality of the joint space, mapping from q to z.
+%           Must be one of:
+%               'PCA' - Principle Component Analysis
+%               'KPCA' - Kernel PCA with Gaussian kernel w/ variance 1 
[…]

### 2017-10-18 src/JTDS_mat_lib/crossvalidation_dr_kuka.m

@@ -1,0 +1,189 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Evaluation of Learning Schemes for JTDS Models on Different Datasets     %%
+%  Train and compare a series of JTDS models with different dimensionality  %%   
+%  reduction schemes and GMM model fitting                                  %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'pour_obst'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour'
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
+        demos_location = strcat(data_path,'singularity/data.mat');
+        demo_ids = [1:10];  
+        fprintf('Loading demonstrations from %s \n', demos_location);
+        load(demos_location)
+end
+
+if ~strcmp(choosen_dataset,'singularity')
+    fprintf('Loading demonstrations from %s \n', demos_location);
+    [Qs_, Ts_] = ImportDemonstrations(demos_location);
[…]

### 2017-10-18 src/JTDS_mat_lib/demo_footstep.m

@@ -1,0 +1,174 @@
+close all
+clear
+clc
+[this_folder,~,~] = fileparts(mfilename);
+addpath(genpath(strcat(this_folder,'../../')));
+
+% This demo creates a 3-DOF robot leg, and a number of hand-designed
+% trajectories with added random perturbations. It then trains the
+% Joint-Space Task-Directed Dynamical System model on these demonstrations,
+% and visualizes the resulting motions. Near the end, the file can generate
+% a number of insightful figures on the behavior in latent-space and
+% velocity-space of the learned controller.
+dimq = 3;
+% The following are the DH parameters for our 3 DOF robot
+A = [.4, .4, .1];
+D = [0, 0,0]; % All components are in the 2D plane
+Alpha = [0, 0,0]; 
+Qmin = -pi*[1, 1, -.25]; % No joint limits except for on the ankle
+Qmax = pi*[1, 1, .75];
+% Create a model of the robot
+robot = initialize_robot(A,D,Alpha,Qmin,Qmax);
+fig = initialize_robot_figure(robot);             
+% Create a model plant for the robot's motor controller
+robotplant = RobotPlant(robot, 'end_trans');
+
+% The time between demonstration steps
+dt = .005;
+
+% We create two demonstrated motions, corresponding to the forward swing 
+% and backward swing of a footstep.
+
+% In the forward swing trajectory, the knee upward first, followed
+% afterward by the foot (as occurs in human footsteps).
+demo_basic{1} = [
+    linspace(-pi/2 -.4, - pi/2 + .6, 50) linspace( - pi/2 + .6, -pi/2 + .5, 50);
+    linspace(-.2, -pi/2, 50) linspace( -pi/2, -.5, 50);
+    linspace(pi/2 - .3, pi/2, 50) linspace(pi/2, pi/2 + .05, 50)
+    ];
+
+% In teh backward swing trajectory, the whole foot moves backward in one
+% large motion.
+demo_basic{2} = [
+    linspace(-pi/2 + .5, -pi/2 -.4, 100);
+    linspace(-.5, -.2, 100);
+    linspace(pi/2 + .05,pi/2 - .3, 100);
+    ];
+
+% Since we need sufficient training data, we'll generate random
+% perturbations of these two base trajectories.
+num_demos_generated = 10;
[…]

### 2017-10-18 src/JTDS_mat_lib/demo_kuka.m

@@ -1,0 +1,103 @@
+%% KUKA training demo
+% This file implements a JT-DS training script for the KUKA LWR 4+.
+% To get some pre-recorded KUKA demonstration data, check out
+% github.com/epfl-lasa/rtk_JT_DS . In particular, you can use a
+% demonstration like:
+%
+% https://github.com/epfl-lasa/rtk_JT_DS/tree/master/data/mat/fore_hand/raw_data.mat
+%
+% Whatever you use, replace the line below importing the demonstrations
+% with your local demonstration data file.
+
+close all
+clear
+clc
+% 
+% DH parameters for the KUKA LWR 4+ robot
+dimq = 7;
+A = [0 0 0 0 0 0 0.05];
+Alpha = pi/2*[1 -1 -1 1 1 -1 0];
+D = [.34 0 .4 0 .4 0 .279];
+Qmin = 2*pi/180*[-85, -90, -100, -110, -140, -90, -120];
+Qmax = 2*pi/180*[85, 90, 100, 110, 140, 90, 120];
+% Create a model of the robot
+robot = initialize_robot(A,D,Alpha,Qmin,Qmax);
+fig = initialize_robot_figure(robot);             
+% Create a model plant for the robot's motor controller
+robotplant = RobotPlant(robot, 'end_trans');
+%% Upload the demonstration data
+% --------------- REPLACE THIS WITH KUKA DEMONSTRATIONS LOCATION ---------
+%demos_location = 'path/to/kuka/demo/raw_data.mat';
+demos_location = '~/Downloads/pour_no_obst/data.mat';
+% ------------------------------------------------------------------------
+[Qs, Ts] = ImportDemonstrations(demos_location);
+
+% If the data is very dense, initializing the semidefinite program may take
+% a long time. In this case, it may help to thin down the number of
+% demonstrated points (by varying "thinning_ratio", so long as there are still sufficient points to
+% satisfactorily reconstruct the shape of the trajectory.
+thinning_ratio = 50; % In the KUKA case, we get 500 datapoints per second, so we shrink the data density considerably
+for i = 1:length(Qs)
+    Qs{i} = Qs{i}(:, 1:thinning_ratio:end);
+    Ts{i} = Ts{i}(:, 1:thinning_ratio:end);
+end
+
+%%  Execute the learning phase
+
+time = Ts;
+demos = Qs;
+tol_cutting = .1;
[…]

### 2017-10-18 src/JTDS_mat_lib/demo_varying_A.m

@@ -1,0 +1,75 @@
+% This file allows you to observe differing behaviors of the robot motion
+% controller based on different values of the joint augmentation matrix A.
+
+% First, we define the parameters of the system:
+dimq = 3;
+% The following are the DH parameters for our 3 DOF robot
+A = [.4, .4, .1];
+D = [0, 0,0]; % All components are in the 2D plane
+Alpha = [0, 0,0]; 
+Qmin = -pi*[1, 1, -.25]; % No joint limits except for on the ankle
+Qmax = pi*[1, 1, .75];
+% Create a model of the robot
+robot = initialize_robot(A,D,Alpha,Qmin,Qmax);
+fig = initialize_robot_figure(robot);             
+% Create a model plant for the robot's motor controller
+robotplant = RobotPlant(robot, 'end_trans');
+
+
+% We write out a number of different A matrices (each emphasizing a
+% different joint) and build controllers to match them.
+q_init = [-pi/2;-pi/6;0]; x_targets = [0.5; -.3; 0];
+identity_mapping = compute_mapping(eye(dimq), 'None');
+mg_vanilla = MotionGeneratorBounded(robotplant, zeros(dimq, 1), eye(dimq), 1, eye(dimq), identity_mapping);
+A_hip = [5,0,0; 0, 1, 0; 0,0, 1];
+mg_hip = MotionGeneratorBounded(robotplant, zeros(dimq, 1), eye(dimq), 1, A_hip, identity_mapping);
+A_knee = [1,0,0; 0, 5, 0; 0,0, 1];
+mg_knee = MotionGeneratorBounded(robotplant, zeros(dimq, 1), eye(dimq), 1, A_knee, identity_mapping);
+A_wrist = [1,0,0; 0, 1, 0; 0,0, 5];
+mg_wrist = MotionGeneratorBounded(robotplant, zeros(dimq, 1), eye(dimq), 1, A_wrist, identity_mapping);
+
+% We interpolate the resulting trajectories forward
+[Q_vanilla, T_vanilla]...
+ = computeFullTrajectory(q_init, x_targets, mg_vanilla);
+[Q_hip, T_hip] = computeFullTrajectory(q_init, x_targets, mg_hip);
+[Q_knee, T_knee] = computeFullTrajectory(q_init, x_targets, mg_knee);
[…]

### 2017-10-18 src/JTDS_mat_lib/hyperparam_kpca_kuka.m

@@ -1,0 +1,248 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Evaluation of Learning Schemes for JTDS Models on Different Datasets     %%
+%  Train and compare a series of JTDS models with different dimensionality  %%   
+%  reduction schemes and GMM model fitting                                  %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour'
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
+        demos_location = strcat(data_path,'singularity/data.mat');
+        demo_ids = [1:10];  
+        fprintf('Loading demonstrations from %s \n', demos_location);
+        load(demos_location)
+end
+
+if ~strcmp(choosen_dataset,'singularity')
+    fprintf('Loading demonstrations from %s \n', demos_location);
[…]

### 2017-10-18 src/JTDS_mat_lib/kuka_experiment.m

@@ -1,0 +1,80 @@
+function [motion_generator, rmse] = kuka_experiment(demos, times, options)
+% This function is a wrapper that sets up and trains a JTDS model with a 
+% given set of hyperparameters.
+%
+% Inputs -----------------------------------------------------------------
+%
+%
+%   o demos:   A variable containing all demonstrations (only
+%              joint trajectories). The variable 'demos' should follow the
+%              following format:
+%              - demos{n}: qdim x T^n matrix representing the qdim dimensional
+%                          joint trajectories. T^n is the number of datapoints in
+%                          each demonstration (1 < n < N)
+%
+%   o time:    This variable can be provided in two ways. If time steps
+%              between all demonstration points are the same, 'time' could be
+%              given as a positive scalar (i.e. time = dt). If not, 'time'
+%              should follow the following format:
+%              - time{n}: 1 x T^n vector representing the time array of length
+%                         T^n cor
+%
+%   o options: This variable contains the options struct passed into
+%              JTDS_Solver. For more information, call 'doc JTDS_Solver'.
+%              One can specify additional options, including:
+%              o options.tol_cutting - used to trim the ends of the
+%                  demonstrations. For more info, call 
+%                  'doc preprocess_demos_jts'. Default 0.1.
+%              o options.robot - used to provide a custom robot. If not,
+%                  given the default DH params for the KUKA LWR 4+.
+%
+% Outputs -----------------------------------------------------------------
+%
+%
+%   o motion_generator: 
+%               a MotionGenerator object (or MotionGeneratorBounded
+%               object if options.learn_with_bounds = true) with parameters
+%               learned through the JTDS optimization.
+%   
+%   o rmse:     The root mean-squared error between the true joint velocity
+%               and the estimated joint velocity (using motion_generator)
+%               evaluated pointwise at each location along the original
+%               demos.
+
+
+    if ~(isfield(options, 'tol_cutting'))
+        options.tol_cutting = 0.1;
+    end
+    if ~(isfield(options, 'learn_with_bounds'))
+        options.learn_with_bounds = true;
+    end
+    if ~(isfield(options, 'robot'))
[…]

### 2017-10-18 src/JTDS_mat_lib/model/MotionGenerator.m

@@ -1,0 +1,88 @@
+classdef MotionGenerator
+    %MOTIONGENERATOR The class that contains the JT-DS dynamical system,
+    %though without built-in joint limit functionality. For a controller
+    %that also integrates joint limits, see the child class
+    %"MotionGeneratorBounded".
+    
+    properties
+        plant % a RobotPlant used to compute kinematics properties
+        Mu % the means of the learned behavior GMM
+        Sigma % the sigmas of the GMM
+        Priors % the scaling of relative Gaussians of the GMM
+        As % the augmentation matrices of the components of the GMM
+        latent_mapping % the container for the details of the function mapping from the joint space q to its latent embedding space z
+    end
+    
+    methods
+        function obj = MotionGenerator(robotplant, mu, sigma, priors, A, latent_mapping)
+            obj.plant = robotplant;
+            if nargin == 1 % if only given the robot, create a default controller with no learned values
+                dimq = robotplant.robot.n;
+                obj.Mu = zeros(dimq, 1);
+                obj.Sigma = eye(dimq);
+                obj.Priors = 1;
+                obj.As = eye(4);
+                [~, obj.latent_mapping] = compute_mapping(eye(q), 'None');
+            else
+                obj.Mu = mu;
+                obj.Sigma = sigma;
+                obj.Priors = priors;
+                obj.As = A;
+                obj.latent_mapping = latent_mapping;
+            end
+        end
+        
+        function A = compute_A(obj, q)
+        % This function computes the weighted A matrix based on the current
+        % robot configuration. This combines the contributions of each of
+        % the local augmentation matrices (As).
+            k = length(obj.Priors);
+            for i = 1:k
+                h_raw(i) = obj.Priors(i)*mvnpdf(out_of_sample(q', obj.latent_mapping)', obj.Mu(:, i), obj.Sigma(:, :, i));
+            end
+            htotal = sum(h_raw);
+            if htotal == 0 % if our point is too far from any of the Gaussians, weight them equally
+                h = 1./k*ones(k,1);
+            else
+                h = h_raw/htotal;
+            end
+            A_total = zeros(size(obj.As(:,:,1)));
+            for i = 1:k
+                A_total = A_total + h(i)*obj.As(:, :, i);
+            end
+            A = A_total;
+        end
+        
+        function qd = get_next_motion(obj, q, xt, ~) % Generates the next joint velocity given the joint position,
+            % desired position, a GMM model, and an augmentation matrix for
[…]

### 2017-10-18 src/JTDS_mat_lib/model/MotionGeneratorBounded.m

@@ -1,0 +1,50 @@
+classdef MotionGeneratorBounded < MotionGenerator
+    %MOTIONGENERATOR The class which implements the dynamical system
+    %specified using our model, and also ensures the system's joint limits are
+    %respected.
+    %When given the parameters of the system, can be used to generate the
+    %next instantaneous velocity from a current position.
+    
+    properties
+        Qmin % the minimum and maximum angles
+        Qmax
+    end
+    
+    methods
+        
+        function obj = MotionGeneratorBounded(robotplant, mu, sigma, priors, A, latent_mapping)
+            obj@MotionGenerator(robotplant, mu, sigma, priors, A, latent_mapping);
+            qlim = robotplant.robot.qlim;
+            obj.Qmin = zeros(size(qlim, 1),1);
+            obj.Qmax = zeros(size(qlim, 1),1);
+            for i = 1:size(qlim, 1)
+                obj.Qmin(i) = qlim(i,1);
+                obj.Qmax(i) = qlim(i,2);
+            end
+        end
+        
+        function qd = get_next_motion(obj, q_old, xt, dt) % Generates the next joint velocity given the joint position,
+            % desired position, a GMM model, and an augmentation matrix for
+            % each Gaussian. dt is the timestep for our next movement
+            % We compute the velocity by integrating our joint limitation
+            % matrix "S" into our classic MotionGenerator controller, all
+            % the while ensuring positive-semidefiniteness.
+            S = obj.plant.compute_S(q_old);
+            qd = S*obj.compute_A(q_old)*S'*obj.plant.qd_basis(q_old, xt);
+        end
+        
+        function qd_fun = ODE_fun(obj, xt, dt)  % Yields a function handle
+            % which computes the first order differential equation
+            % associated with this system, i.e. f where qd = f(t, q)
+            % Used as input to an ODE solver.
+            function qd = qd_fun_tmp(t, q)
+                S = obj.plant.compute_S(q);
+                qd = S*obj.compute_A(q)*S*obj.plant.qd_basis(q,xt);
+            end
+            qd_fun = @qd_fun_tmp;
+        end
+    end
+    
+end
+
+

### 2017-10-18 src/JTDS_mat_lib/model/RobotPlant.m

@@ -1,0 +1,75 @@
+classdef RobotPlant
+    %ROBOTPLANT A class that encodes the robot's kinematic model, and also
+    % for generating various convenient properties relating to the
+    % kinematic model.
+    
+    properties
+        robot % a Corke robotics-toolbox type robot to generate kinematics
+        forward_kinematics % function that maps a joint configuration q to a task position x
+        jacobian % takes as input q and gives corresponding dx/dq as output
+        dimx % the dimension of x
+    end
+    
+    methods
+        function obj = RobotPlant(robot, task_type, custom_forward_kinematics, custom_jacobian, dimx)
+            % the task-space description, task_type can be one of the following:
+            % 'end_trans' = end-effector xyz coordinates, or 'custom'
+            % if 'custom' must specify the forward kinematics and custom
+            % jacobian using the 3rd and 4th arguments
+            % dimx is the dimensionality of the task space
+            % custom_qd_basis takes as input the object, q, and dt,
+            % and outputs the basis for our velocity policy
+            obj.robot = robot;
+                
+            if nargin<5 % If the number of arguments suggests standard non-custom kinematics
+                if strcmp(task_type, 'end_trans')
+                    obj.forward_kinematics = @obj.end_trans_forward_kinematics;
+                    obj.jacobian = @obj.end_trans_jacobian;
+                    obj.dimx = 3;
+                else
+                    error('Custom kinematics/jacobian functions were not specified');
+                end
+            else
+                obj.forward_kinematics = custom_forward_kinematics;
+                obj.jacobian = custom_jacobian;
+                obj.dimx = dimx;
+            end
+        end
+        
+        function basis = qd_basis(obj, q, xt)
+            % Computes the base joint velocity to move at a joint position q
+            % toward a task position xt using the jacobian transpose
+            % method (i.e. qd = J^T(H(q) - xt) ).
+            % In general, this will be multiplied by an A matrix to get the
+            % augmented joint velocity.
+            basis = -1*transpose(obj.jacobian(q))*(obj.forward_kinematics(q) - xt);
+        end
+        
+        function x = end_trans_forward_kinematics(obj,q)
+            T = obj.robot.fkine(q);
+            x = T(1:3, 4);
+        end
+        
+        function j = end_trans_jacobian(obj, q)
+            j = obj.robot.jacob0(q, 'trans');
+        end
+        
+        function output = compute_S(obj, q)
[…]

### 2017-10-18 src/JTDS_mat_lib/model/initialize_robot.m

@@ -1,0 +1,27 @@
+function robot = initialize_robot(A, D, Alpha, Qmin, Qmax, plotopt)
+%INITIALIZE_ROBOT Creates a robot object using Peter Corke's Robotics Toolbox 
+%   This object can be used to generate jacobians and forward kinematics,
+%   and to pass around robot parameters. The first 3 parameters are column
+%   vectors of Denavit-Hartenberg parameters (with the "offset q" assumed to be 0).
+%   The last two are minimum and maximum joint angles with respect to the
+%   default "0" angle.
+%   Plotopt is a struct containing plotting options as specified in Corke's
+%   robot-toolkit
+    qdim = length(A);
+    if nargin < 5% if joint limits unspecified, set them to the max possible
+        Qmin = 2*pi*ones(qdim, 1);
+        Qmax = -2*pi*ones(qdim, 1);
+        warning('No joint limits specified, may cause unwanted behavior if using limit-enforcement methods.');
+    end
+    
+    for i = 1:qdim
+        L(i) = Link('d', D(i), 'a', A(i), 'alpha', Alpha(i), 'standard', 'revolute', 'qlim', [Qmin(i) Qmax(i)]);
+    end
+    robot = SerialLink(L, 'name','my_robot');
+    if nargin == 6
+        robot.plotopt = plotopt;
+    else
+        robot.plotopt = {'noshadow','nojaxes', 'nowrist', 'noname','jointdiam', 2, 'basewidth', 1, 'linkcolor',0.7*[1,1,1], 'ortho','noshading','notiles','jointcolor',0.6*[1,1,1]};
+        %,'nojaxes', 'nowrist', 'noname'
+    end
+end

### 2017-10-18 src/JTDS_mat_lib/run_kuka_experiments.m

@@ -1,0 +1,36 @@
+%% Train and compare a series of JTDS models
+demos_location = '~/Downloads/fore_hand/data.mat';
+fprintf('Loading demonstrations from %s \n', demos_location);
+[Qs, Ts] = ImportDemonstrations(demos_location);
+% If the data is very dense, initializing the semidefinite program may take
+% a long time. In this case, it may help to thin down the number of
+% demonstrated points (by varying "thinning_ratio", so long as there are still sufficient points to
+% satisfactorily reconstruct the shape of the trajectory.
+% In the KUKA case, we get 500 datapoints per second, so we recommend shrinking the data density considerably
+thinning_ratio = 50;
+for i = 1:length(Qs)
+    Qs{i} = Qs{i}(:, 1:thinning_ratio:end);
+    Ts{i} = Ts{i}(:, 1:thinning_ratio:end);
+end
+demos = Qs;
+times = Ts;
+
+mappings_to_compare = {'None', 'PCA', 'KPCA'};
+options.kpca_sigma = 0.1;
+options.thinning_ratio = 80;
+options.GMM_sigma_type = 'full'; % Can be either 'full' or 'diagonal'
+options.explained_variance_threshold = .90;
+options.GMM_maximize_BIC = false;
+options.fixed_num_gaussians = 5;
+%options.max_gaussians = 8;
+options.BIC_regularization = 2; % this quantity should probably be between 1 and 3
+options.learn_with_bounds = false;
+options.verbose = false;
+
+for i = 1:size(mappings_to_compare, 2)
+    mapping_name = mappings_to_compare{i};
+    fprintf('Training JTDS generator using %s mapping...\n', mapping_name);
+    options.latent_mapping_type = mapping_name;
+    [motion_generator, resulting_rmse] = kuka_experiment(demos, times, options);
+    fprintf('Using %s mapping, got prediction RMSE: %d \n', mapping_name, resulting_rmse);
+end

### 2017-10-18 src/JTDS_mat_lib/run_modified_kuka_experiments.m

@@ -1,0 +1,50 @@
+%% Train and compare a series of JTDS models, on a simulated modification of a kuka robot
+demos_location = '~/Downloads/fore_hand/data.mat';
+fprintf('Loading demonstrations from %s \n', demos_location);
+[Qs, Ts] = ImportDemonstrations(demos_location);
+% If the data is very dense, initializing the semidefinite program may take
+% a long time. In this case, it may help to thin down the number of
+% demonstrated points (by varying "thinning_ratio", so long as there are still sufficient points to
+% satisfactorily reconstruct the shape of the trajectory.
+% In the KUKA case, we get 500 datapoints per second, so we recommend shrinking the data density considerably
+% Now create a modified, "nuka" robot
+dimq_nuka = 11;
+A = [0 0 0 0 0 0 0 0 0 0 0.05];
+Alpha = pi/2*[1 -1 -1 1 1 -1 -1 1 1 -1 0];
+D = [.34 0 .4 0 .15 0 .15 0 .15 0 .279];
+Qmin = -2*pi/180*ones(1, dimq_nuka);
+Qmax = 2*pi/180*ones(1, dimq_nuka);
+nuka_robot = initialize_robot(A,D,Alpha,Qmin,Qmax);
+kuka_to_nuka_demonstration = @(x) [x(1:4, :); x(5:6, :) + 0.1*rand(size(x(5:6, :))); x(5:6, :) + 0.1*rand(size(x(5:6, :))); x(5:6, :) + 0.1*rand(size(x(5:6, :))); x(7, :)];
+
+
+
+thinning_ratio = 50;
+for i = 1:length(Qs)
+    Qs{i} = Qs{i}(:, 1:thinning_ratio:end);
+    Qs{i} = Qs{i} + 0.02*rand(size(Qs{i}));
+    %Qs{i} = kuka_to_nuka_demonstration(Qs{i});
+    Ts{i} = Ts{i}(:, 1:thinning_ratio:end);
+end
+demos = Qs;
+times = Ts;
+
+mappings_to_compare = {'None', 'PCA', 'KPCA'};
+options.kpca_sigma = 1;
+options.GMM_sigma_type = 'full'; % Can be either 'full' or 'diagonal'
+options.explained_variance_threshold = .90;
+options.GMM_maximize_BIC = true;
+%options.fixed_num_gaussians = 5;
+options.max_gaussians = 8;
+options.BIC_regularization = 2; % this quantity should probably be between 1 and 3
+options.learn_with_bounds = false;
[…]

### 2017-10-18 src/JTDS_mat_lib/simulate/GraphRobot.m

@@ -1,0 +1,35 @@
+function robot_copy_names = GraphRobot(robot, Q, fig, robot_plot_options )
+% Overlays the robot in the given joint configuration on the figure
+% Arguments:
+% robot - a Corke robotics-toolkit style robot
+% Q - a dimq x n matrix of robot joint positions to plot, each with a
+% separate robot
+% fig - if specified, plot the robot in this figure
+% robot_plot
+% robot_plot_options - a Corke robotics-toolbox plotopt structure to
+% specify the plot type
+%
+% Output:
+% robot_copy_names - the names of each of the generated robots plotted
+    robot_copy_names = [];
+    if nargin < 3
+        fig = [];
+    end
+    for i = 1:size(Q, 2)
+        q = Q(:, i);
+        new_robot = SerialLink(robot);
+        new_robot.name = sprintf('robot copy %d', rand*10000);
+        robot_copy_names = [robot_copy_names new_robot.name];
+        if nargin < 4 || isempty(robot_plot_options)
+            new_robot.plotopt = {'jointdiam', 3,'basewidth', .5, 'noshadow', 'notiles', 'noshading','noname', 'nowrist', 'linkcolor', (0.5 + 0.3*i/size(Q,2))*[1,1,1], 'jointcolor', (0.4 + 0.3*i/size(Q,2))*[1,1,1]};
+        else
+            new_robot.plotopt = robot_plot_options;
+        end
+        if isempty(fig)
+            fig = initialize_robot_figure(new_robot);
+        end
+        figure(fig);
+        new_robot.plot(q');
+    end
+end
+

### 2017-10-18 src/JTDS_mat_lib/simulate/Simulation.m

@@ -1,0 +1,53 @@
+function [q, qd] = Simulation(q0, xt, motion_generator, dt,  fig, markertype)
+%SIMULATION Simulates forward in real-time a DS policy in a given system
+
+% Arguments
+% q0 - dimq x 1, the initial joint-space configuration of the robot
+% xt - dimx x 1, the ultimate task-space target for the robot
+% motion_generator - a MotionGenerator/(child class) object for computing
+% the simulatable policy
+% dt - optional, provides the timestep to step forward between each frame
+% fig - optional, the figure to plot the simulation in, if not specified a new one is
+% generated
+% markertype - input to plot like 'ro', 'bx', for plotting the task-space
+% trace of the robot
+
+% Output
+% q - the joint-space positions achieved by the robot throughout the
+% simulation
+% qd - the joint-space velocities achieved by the robot throughout the
+% simulation
+
+    if nargin < 4 || isempty(dt) % If a timestep isn't specified, default is .01
+        dt = .01;
+    end
+    goal_threshold = .05; % Determines how far from the goal the simulation should accept
+    
+    robot = motion_generator.plant.robot;
+    if nargin < 5
+        fig = initialize_robot_figure(robot);
+    else
+        figure(fig);
+    end
+    q = q0;
+    robot.animate(q');
+    if nargin < 6
+        markertype = 'ro';
+    end
+    plot(xt(1), xt(2), markertype, 'markersize', 20);
+    x = motion_generator.plant.forward_kinematics(q0);
+    while(norm(x - xt) > goal_threshold) % at each time step, system is integrated using the Newton metho
+        qd = motion_generator.get_next_motion(q, xt, dt);
+        q_candidate = q + qd*dt;
+        q = q_candidate;
+        x = motion_generator.plant.forward_kinematics(q);
+        robot.plot(q');
+        if ~ishandle(fig)
+            error('Figure closed prematurely');
+        end
+        plot(x(1), x(2), 'b.', 'markersize', 10);
+        pause(.2);
+    end
+    disp('Reached goal.');
+end
+

### 2017-10-18 src/JTDS_mat_lib/simulate/computeFullTrajectory.m

@@ -1,0 +1,50 @@
+function [Q, T] = computeFullTrajectory(q_initial, x_targets, motion_generator, goal_tolerance, max_duration)
+%COMPUTEFULLTRAJECTORY Interpolates a JT-DS trajectory
+%   This function integrates a JT-DS problem and computes the evolution of
+%   the system over a given timespan, using an ODE solver. Each target is
+%   integrated for up to a fixed timespan, or until the target reaches the
+%   goal.
+
+%   Arguments:
+%   q_initial - a dimq x 1 vector denoting the initial joint configuration
+%   x_targets - a dimx x #targets matrix denoting the desired task-space
+%   positions for the system to move to
+%   motion_generator - a MotionGenerator/(child class) obj obj to compute the DS motion
+%   goal_tolerance - the distance from the target position necessary for
+%   the controller to "accept"
+%   max_duration - the maximum period of time a trajectory will be
+%   interpolated. After this period elapses, the controller begins
+%   integrating the motion to the next target on the list.
+%   Outputs:
+%   Q is an n x dimq matrix denoting the interpolated robot positions
+%   T is an n x 1 matrix denoting the times corresponding to each position
+    if nargin < 4
+        goal_tolerance = 0.03;
+    end
+    if nargin < 5
+        max_duration = 60; %seconds
+    end
+    
+    T = []; Q = [];
+    q_start = q_initial;
+    for i = 1:size(x_targets, 2)
+        x_target = x_targets(:, i);
+        [T_subtraj, Q_subtraj] = ode15s(motion_generator.ODE_fun(x_target, 1), ...
+            [0, max_duration], q_start, ...
+            motion_generator.ODE_options(x_target, goal_tolerance));
+        if isempty(T)
+            T = T_subtraj;
+        else
+            T = [T; T_subtraj];
+        end
+        if sum(sum(isnan(Q_subtraj)))
+            error('Generated a NaN... something is wrong with ODE solution. Consider switching between ode15s and ode45 in computeFullTrajectory.')
+        end
+        Q = [Q; Q_subtraj];
+        q_start = Q_subtraj(end, :)';
+    end
+    T = T';
+    Q = Q';
+
+end
+

### 2017-10-18 src/JTDS_mat_lib/simulate/initialize_robot_figure.m

@@ -1,0 +1,16 @@
+function fig = initialize_robot_figure(robot,varargin)
+% Creates a simple visualization of the specified robot
+% If the second argument is a figure, then plots the robot in that figure
+
+if nargin>1
+    % figure exists and we should activate axes to plot in
+    axes(varargin{1})
+else
+    % create plot in new figure window
+    fig = figure();
+end
+robot.plot(zeros(1, robot.n));
+view([0 90]) % This positions the camera above the xy plane looking downward
+hold on
+
+end

### 2017-10-18 src/JTDS_mat_lib/test_modified_kuka.m

@@ -1,0 +1,51 @@
+% A script for generating a modified kuka robot (with 4 extra links).
+% Real demonstrations are converted to demonstrations on the fake "nuka"
+% robot by copying the 5th and 6th links to the 7th and 8th and 9th and
+% 10th respectively, and the 7th to the 11th. These "new" link
+% demonstrations are also given a small amount of noise.
+
+
+close all
+clear
+clc
+% 
+% DH parameters for the KUKA LWR 4+ robot
+dimq_kuka = 7;
+A = [0 0 0 0 0 0 0.05];
+Alpha = pi/2*[1 -1 -1 1 1 -1 0];
+D = [.34 0 .4 0 .4 0 .279];
+Qmin = 2*pi/180*[-85, -90, -100, -110, -140, -90, -120];
+Qmax = 2*pi/180*[85, 90, 100, 110, 140, 90, 120];
+% Create a model of the robot
+kuka_robot = initialize_robot(A,D,Alpha,Qmin,Qmax);
+% Now create a modified, "nuka" robot
+dimq_nuka = 11;
+A = [0 0 0 0 0 0 0 0 0 0 0.05];
+Alpha = pi/2*[1 -1 -1 1 1 -1 -1 1 1 -1 0];
+D = [.34 0 .4 0 .15 0 .15 0 .15 0 .279];
+Qmin = -2*pi/180*ones(1, dimq_nuka);
+Qmax = 2*pi/180*ones(1, dimq_nuka);
+nuka_robot = initialize_robot(A,D,Alpha,Qmin,Qmax);
+kuka_to_nuka_demonstration = @(x) [x(1:4, :); x(5:6, :) + 0.1*rand(size(x(5:6, :))); x(5:6, :) + 0.1*rand(size(x(5:6, :))); x(5:6, :) + 0.1*rand(size(x(5:6, :))); x(7, :)];
+
+%fig = initialize_robot_figure(kuka_robot);
+% Create a model plant for the robot's motor controller
+kuka_robotplant = RobotPlant(kuka_robot, 'end_trans');
+nuka_robotplant = RobotPlant(nuka_robot, 'end_trans');
+
+demos_location = '~/Downloads/fore_hand/data.mat';
+[Qs, Ts] = ImportDemonstrations(demos_location);
+i = 1;
+disp(size(Qs{i}));
+Qs_nuka = kuka_to_nuka_demonstration(Qs{i});
[…]

### 2017-10-18 src/JTDS_mat_lib/utils/ImportDemonstrations.m

@@ -1,0 +1,13 @@
+function [Qs, Ts] = ImportDemonstrations(filename)
+%IMPORTDEMONSTRATION Converts from a demonstration text file
+%   to a set of joint positions Q and times T
+    file = load(filename);
+    N = length(file.data);
+    Qs = cell(N, 1);
+    Ts = cell(N, 1);
+    for i = 1:N
+        Qs{i} = file.data{i}.kuka.joint_states(1:end-1, :);
+        Ts{i} = file.data{i}.kuka.joint_states(end, :);
+    end
+end
+

### 2017-10-18 src/JTDS_mat_lib/utils/LogLikelihood_GMM.m

@@ -1,0 +1,45 @@
+function [ loglik ] = LogLikelihood_GMM(X,Priors,Mu,Sigma,w)
+%LOGLIKELIHOOD_GMM computes the likelihood of data given a GMM
+% Used as a utility function for GMM_BIC
+
+%   input------------------------------------------------------------------
+%
+%       o X : (D x N), set of N data points of dimension  D
+%
+%       o w : (1 x N), set of weights if the data points are weighted
+%
+%   output ----------------------------------------------------------------
+%
+%       o loglik : (N x 1) , loglikelihood
+%
+
+
+N = size(X,2);
+K = size(Priors,2);
+
+if isempty(w)
+    w = ones(1,length(X));
+end
+
+Pxi = zeros(N,K);
+for i=1:K
+    Pxi(:,i) = gaussPDF(X, Mu(:,i), Sigma(:,:,i));
+end
+
+if exist('w','var') && ~isempty(w)
+    %               (N x K)
+    %             (N x 1)
+    F            = (repmat(w(:),1,K) .* Pxi) * Priors';
+else
+    
+    %Compute the log likelihood
+    F            = Pxi*Priors';
+   
+end
+
+F(F<realmin) = realmin;
+F            = Pxi*Priors';
+loglik       = sum(log(F));
+
+
+end

### 2017-10-18 src/JTDS_mat_lib/utils/PlaybackTrajectory.m

@@ -1,0 +1,75 @@
+function [fig, h] = PlaybackTrajectory(robotplant, Q, T, old_fig, color)
+% Visualizes a robot's trajectory, displaying a series of robot joint positions and their corresponding
+% times. If clicked before completion, will restart the playback.
+
+% Arguments:
+% robotplant - a RobotPlant object specifying the kinematics of the robot being simulated
+% Q - a dimq x n matrix of robot joint positions
+% T - either a single value corresponding to the timestep, OR
+%     a 1 x n matrix of times corresponding to each joint position
+% old_fig - optional, a figure in which to plot the trajectory (if not
+%     specified another is generated)
+% color - optional, the color of the trace being plotted (r, g, b, y,
+%     etc.)
+% 
+% Outputs:
+% fig - the figure in which the simulation was displated
+% h - the axes on which the trace was plotted.
+
+    restart_flag = 0;
+    complete_flag = 0;
+    if nargin < 4 % if axes unspecified
+        fig = initialize_robot_figure(robotplant.robot);
+    else
+        fig = figure(old_fig);
+    end
+    if nargin < 5 % if color unspecified
+        color = 'r';
+    end
+    
+    n = size(Q, 2);
+    if size(T) == [1 1]
+        dT = [repmat(T, 1, n-1) 0];
+    else
+        dT = [diff(T) 0];
+    end
+    h = [];
+    %disable unwanted figure modes 
+    zoom off
+    rotate3d off
+    pan off
+    brush off
+    datacursormode off
+
+    set(fig,'WindowButtonDownFcn',@(h,e)restart_animation(h,e));
+    set(fig,'WindowButtonUpFcn',[]);
+    
+    frame_ind = 1;
+    while 1 
+        if ~complete_flag
+            q = Q(:, frame_ind);
+            robotplant.robot.plot(transpose(q));
+            x = robotplant.forward_kinematics(q);
+            h_tmp = plot3(x(1), x(2), x(3), sprintf('.%s', color));
+            h = [h h_tmp];
+            robotplant.robot.delay = dT(frame_ind);
+            frame_ind = frame_ind + 1;
+            if frame_ind > n
+                complete_flag = 1;
+                break
+            end
+        end
+        if restart_flag
+            restart_flag = 0;
+            frame_ind = 1;
+            complete_flag = 0;
+        end
+    end
+        
+        
+        function restart_animation(h,e)
+            restart_flag = 1;
+        end
+
+end
+

### 2017-10-18 src/JTDS_mat_lib/utils/computePairwiseDistances.m

@@ -1,0 +1,32 @@
+function [D, mean_D, max_D] = computePairwiseDistances(X, display_hist)
+
+maxSamples = 10000;
+if length(X) < maxSamples
+    X_train = X;
+    hist_distances = 1;
+else
+    X_train = X(1:maxSamples, :);
+    hist_distances = 10;
+end
+
+%%%%% Compute Element-wise pairwise euclidean distances %%%%%%
+%%% Throughout ALL training points %%%
+tic;
+D = pdist(X_train, 'euclidean');
+toc;
+mean_D = mean(D(:));
+max_D  = max(D(:));
+
+% Visualize pairwise distances as Histogram
+if (display_hist == 1)
+    figure('Color',[1 1 1])
+        
+    D_v = D(:);           
+    H = histfit(D_v(1:hist_distances:end,:));
+    title('Demonstrations Pairwise Distances', 'Interpreter','LaTex')
+    xlabel('$L_2$ Norm', 'Interpreter','LaTex')
+    grid on
+    axis tight
+end
+
+end

### 2017-10-18 src/JTDS_mat_lib/utils/export2JSEDS_Cpp_lib.m

@@ -1,0 +1,55 @@
+function out = export2JSEDS_Cpp_lib(filename,Priors,Mu,Sigma,robot, As,M)
+% Export the parameters of a JT-DS model to a given file.
+% 
+% Arguments:
+% filename - the filename to write the parameters to
+% Priors - the GMM priors
+% Mu - the GMM means of the JT-DS model
+% Sigma - the GMM variances of the JT-DS model
+% robot - the Corke robot-toolkit robot which the JT-DS model was built for
+% As - the joint augmentation matrices in the JT-DS model
+% M - the PCA matrix for embedding the joint-space positions into a
+%   lower-dimensional latent space in the JT-DS model
+
+% Outputs:
+% out - whether the file was successfully fully written
+
+
+
+% The parameters are written to the file in the following order:
+% dimq, dimz, K
+% priors
+% Mus (reshaped horizontally)
+% Sigmas (one matrix per each line, reshaped horizontally)
+% As (one matrix per each line, reshaped horizontally)
+% M (the PCA matrix) reshaped horizontally
+% Robot DH parameters: all d's, then all a's, then all alpha's, then all
+% qmin's, then all qmax's
+
+dlmwrite(filename, [size(M, 2), size(Mu)],'Delimiter',' ','precision','%i');
+dlmwrite(filename, Priors,'newline','pc','-append','Delimiter',' ','precision','%.6f');
+%dlmwrite(name, reshape(Mu/1000,1,[]),'newline','pc','-append','Delimiter',' ','precision','%.6f');
+dlmwrite(filename, reshape(Mu,1,[]),'newline','pc','-append','Delimiter',' ','precision','%.6f');
+for i=1:size(Mu,2)
+    %dlmwrite(name, reshape(Sigma(:,:,i)/1000^2,1,[]),'newline','pc','-append','Delimiter',' ','precision','%.6f');
+    dlmwrite(filename, reshape(Sigma(:,:,i)',1,[]),'newline','pc','-append','Delimiter',' ','precision','%.6f');
+end
+for i=1:size(Mu,2)
+    %dlmwrite(name, reshape(Sigma(:,:,i)/1000^2,1,[]),'newline','pc','-append','Delimiter',' ','precision','%.6f');
+    dlmwrite(filename, reshape(As(:,:,i)',1,[]),'newline','pc','-append','Delimiter',' ','precision','%.6f');
+end
[…]

### 2017-10-18 src/JTDS_mat_lib/utils/preprocess_demos_jtds.m

@@ -1,0 +1,108 @@
+function [Data, index] = preprocess_demos_jtds(robotplant, demos, time, tol_cutting)
+% This function preprocesses raw joint demonstration data and molds it into
+% a suitable format.
+% The function computes the first time derivative of demonstrations and
+% trims the data. The function can be
+% called using: 
+%
+%          [Data, index] = preprocess_demos(demos,time,tol_cutting)
+%
+% Inputs -----------------------------------------------------------------
+%
+%
+%   o demos:   A variable containing all demonstrations (only
+%              joint trajectories). The variable 'demos' should follow the
+%              following format:
+%              - demos{n}: qdim x T^n matrix representing the qdim dimensional
+%                          joint trajectories. T^n is the number of datapoints in
+%                          each demonstration (1 < n < N)
+%
+%   o time:    This variable can be provided in two ways. If time steps
+%              between all demonstration points are the same, 'time' could be
+%              given as a positive scalar (i.e. time = dt). If not, 'time'
+%              should follow the following format:
+%              - time{n}: 1 x T^n vector representing the time array of length
+%                         T^n corresponding to each demo  (1 < n < N)
+%
+%   o tol_cutting:  A small positive scalar that is used to trim data. It
+%                   removes the redundant datapoint from the begining and
+%                   the end of each demonstration that their first time
+%                   derivative is less than 'tol_cutting'. This is not
+%                   strictly necessary for JT-DS; however from practical point of
+%                   view, it is very useful. There are always lots of noisy
+%                   data at the begining (before the user starts the
+%                   demonstration) and the end (after the user finished the
+%                   demonstration) of each demosntration that are not
+%                   useful.
+%
+% Outputs ----------------------------------------------------------------
+%
+%   o xT:      qdim x N array representing the each of the demonstration's
+%              final points (target points).
+%
+%   o Data:    A (2*qdim + xdim) x N_Total matrix containing all demonstration data points.
+%              The rows are layed out as follows:
+%              1:qdim = joint angles
+%              qdim + 1: 2*qdim = joint velocities
[…]

### 2017-10-18 src/JTDS_mat_lib/utils/trajectory_error.m

@@ -1,0 +1,8 @@
+function trajectory_RMSE = trajectory_error(motion_generator, Q_true, Qd_true, Xt)
+    Qd = zeros(size(Qd_true));
+    for i = 1:size(Q_true, 2)
+        Qd(:, i) = motion_generator.get_next_motion(Q_true(:, i), Xt(:, i));
+    end
+    trajectory_RMSE = sqrt(mean((Qd - Qd_true).^2, 2));
+end
+

### 2017-10-18 paper/main_RA-L_v1.tex
Latest changes in paper
Latest changes

@@
 \begin{abstract}
-\textcolor{red}{Too long, re-write}
+\textcolor{red}{Too long, re-write --}
 In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space with respect to the robot's kinematic constraints.
@@
 \centering
-\includegraphics[scale=.23,trim={0 0 0 14.0cm},clip]{Without_Obstcle.jpg}
-\includegraphics[scale=.23,trim={0 0 0 9.cm},clip]{With_Obstcle.jpg}
-\caption{Two robot motions in joint-space accomplishing the same behavior in task-space (pouring chips in a bowl).
-Though the \textcolor{cyan}{\textbf{task-space trajectories}} are \textit{similar}, the \textcolor{orange}{\textbf{joint-space trajectories}} are different; e.g.
-the bottom example moves in joint-space to avoid an obstacle that the top one does not.}
+\includegraphics[scale=.23,trim={0 0 0 0cm},clip]{Without_Obstcle_white_new.png}
+\includegraphics[scale=.23,trim={0 0 0 0cm},clip]{With_Obstcle.png}
+\caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl).
+The bottom example avoids a known obstacle in its workspace, while the top one does not.}
 \label{fig:robot_example}
-\vspace{-0.6cm}
+\vspace{-10pt}
 \end{figure}

### 2017-10-24 paper/main_RA-L_v1.tex
Updated abstract

@@
 \begin{abstract}
-\textcolor{red}{Too long, re-write --}
 In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space with respect to the robot's kinematic constraints.
 Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, we prove that it stably moves towards its target and is guaranteed to obey the robot's kinematic joint limits.
-To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system, enabling us to encode complex joint-space behaviors.
-We propose a method to train such complex behaviors from kinesthetic demonstrations, by learning the parameters of the LPV system.
-We begin by applying a latent-embedding and a probabilistic approach to discover and represent a set of local behaviors from the demonstrated motions.
-Then, the corresponding local joint-space behaviors are estimated by solving a convex semi-definite optimization problem that minimizes the joint velocity error from the demonstrations.
+To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system, and we provide a method for learning behavior matrices from demonstrations.
+Specifically, we use dimensionality reduction to find low-dimensional behavior regions, and then estimate the parameters of the corresponding joint-space behaviors by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
 Our proposed approach is validated on a variety of motions for a 7-DOF KUKA LWR4+ robot arm.\end{abstract}

### 2017-10-24 paper/main_RA-L_v1.tex
Removed all joint-limits mentions (except for discussion) - need to reincorporate use of discontinuous joint limits in experiment section

@@
 of Electrical Engineering and Computer Science, Massachusetts Institute of Technology, Cambridge,
-MA, 02139 USA. E-mail: yonadav@mit.edu.}% <-this % stops a space
+MA, 02139 USA. E-mail: yonadav.shavit@gmail.com.}% <-this % stops a space
 \thanks{S.
@@
 \begin{abstract}
-In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space with respect to the robot's kinematic constraints.
-Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, we prove that it stably moves towards its target and is guaranteed to obey the robot's kinematic joint limits.
+In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space.
+Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it Lyapunov-stably converges towards its task-space attractor.
 To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system, and we provide a method for learning behavior matrices from demonstrations.
@@
 
-In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only complies with the task-space constraints, but also avoids the problems generated by pseudo-inverse approximations and takes into consideration the robot's kinematic limits.
-\cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities and joint kinematic limits.
-The resulting DS avoids singularities through generalization of the pseudo-inverse approximations and joint limits via online modulation.
+In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
+\cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities.
+The resulting DS avoids singularities through generalization of the pseudo-inverse approximations.
 However, because the two DS have their own unique attractors, the non-linear interaction between them imposed by the kinematic constraints does not ensure that the combined DS has a unique attractor.
@@
 \item is formulated such that complex \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
-
-\item is capable of complying with \textbf{kinematic joint limits}.
 \end{enumerate}
@@
 It shares some of our approach's advantages: fast computation and provable task-space stability.
-However, despite some previous work integrating discontinuous joint limits \cite{sciavicco1988solution} and designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations while obeying the robots kinematic constraints.
+However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations while obeying the robots kinematic constraints.
 
@@
 } The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{m\times d}$.\\
-We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following criteria:
+We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following two criteria:
 \renewcommand{\labelenumi}{(\Roman{enumi})}
@@
 $V(.)$ can be thought of as a metric for the task-space distance-to-go.
-\item The dynamical system must respect the kinematic joint constraints of the robot:
-\begin{equation}
-q(0)\in Q \rightarrow q(t)\in Q~~~~\forall t>0~
-\end{equation}
 \item The dynamical system should encapsulate the desired joint-space behaviors such that the following metric is minimized
@@
 \label{sec:proposed_system}
-We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the three criteria presented in (I), (II), (III).
+We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \ref{eq:Lyp} and \ref{eq:optim_error}.
 \begin{equation}
 \label{eq:ds}
-\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
 \end{equation}
-where 
-\begin{equation}
-\label{eq:SAS}
-\mathcal{A}(q) = S(q) A(q) S^T(q). 
-\end{equation}
-$S(q)=diag(s^1(q),\dots,s^d(q))$ is a diagonal matrix and $ s^i(\cdot)\in C^0~\forall i \in \{1,\dots d\} $.
-$A(q)\in \mathbb{R}^{d\times d}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a nonlinear combination of constant linear matrices, each of which encodes a local joint-space behavior.
+where $P\in \mathbb{R}^{d\times d}$, and $A(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a nonlinear combination of constant linear matrices, each of which encodes a local joint-space behavior.
+$P$ is
 \begin{equation}
@@
 \end{equation}
-where $A_k\in \mathbb{R}^{d\times d} $ and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are constant matrices and the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
+where $A_k\in \mathbb{R}^{m\times m} $ and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are constant matrices and the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
 system $q $ or external signals $d(t)$, i.e.
@@
 the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
-The positive semi-definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
+The positive semi-definite matrix $A(q)$ warps the resulting joint-space velocity; Fig.
 \ref{fig:A_example} illustrates the effects of  $A(q)$ on the generated motion.
 Thus the controller can be thought of as a proportional controller in joint space.
-We refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
-Lastly, $S(q)$ intuitively enforces the joint constraints: as $q^i$ nears $q^i_{min}$, the velocity in that direction goes to $0$, and similarly when it nears $q^i_{max}$ the velocity goes to $0$.
-
+Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
 %When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle).
@@
 \label{prop:stability}
-The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) and (II), if \eqref{eq:ds} meets the following constraints.
+The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if \eqref{eq:ds} meets the following constraints.
 
@@
 & 0 \leq \theta_k(\cdot) & &\\
-%    & 0 \preceq S(q) &\forall q \in Q\\
+%    & 0 \preceq and &\forall q \in Q\\
 \end{split}
 \forall k \in \{1,\dots,K\} 
-\end{cases}
-\end{equation}    
-\begin{equation}
-\label{eq:first_criteria_joint}
-\begin{cases}
-\begin{split}
-&s^{i}(q^i_{min})=0 & &\\
-&s^{i}(q^i_{max})=0 & &
-\end{split}
-\forall i \in \{1,\dots,d\}
 \end{cases}
@@
 This happens if the target position is kinematically unreachable; i.e.
-for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or $S_{i,i}(q) = 0$ (the joint has reached its limit).
-Note that in practice, because $s(q)$ is continuous and designed based on \eqref{eq:first_criteria_joint}, executing JT-DS \eqref{eq:ds} will likely never cause the manipulator to reach the joint limit.
-Instead, $s(q)$ significantly shrinks the magnitude of a joint's motion when a joint nears its motion boundary.
-For a simple method of constructing $S(q)$, see Appendix \ref{appendix:limits}.
+for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or the joint has reached a kinematic limit.
+Instead, the proof in Appendix \ref{appendix:stability} shows that the system converges in the sense of Lyapunov, meaning that its distance to the attractor is always monotonically decreasing.
 
-Criterion (III) (i.e.
+Criterion (II) (i.e.
 encoding specific joint-space behaviors) is achieved by embedding the desired dynamics in the matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$.
@@
 In this paper, we have presented a dynamical system in joint space that is provably Lyapunov-stable in task space and which replicates demonstrated joint-space behaviors.
-The system provably obeys the manipulator's kinematic joint limits, is fast to compute, and smoothly handles singularities by avoiding the pseudo-inverse Jacobian.
+The desired motions are fast to compute, and smoothly handle singularities by avoiding the pseudo-inverse Jacobian.
 We showed the system's ability to learn different joint-space behaviors on a redundant robotic platform.
@@
 
-The JTDS algorithm can also be expanded to accommodate joint velocity limits (in addition to joint position limits) by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits.
+JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$.
+Previous work proposed adding discontinuous limits to JT control \cite{sciavicco1988solution}, and we additionally propose a continuous modification as follows: \textcolor{red}{Remove this part?
+Should we not talk about continuous joint limiting?} Replace the joint behavior matrix $A(q)$ with a joint-limited joint behavior matrix $\mathcal{A} = S(q)A(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that
+\begin{equation}
+\label{eq:last_criteria}
+\begin{cases}
+\begin{split}
+& s^{i}(q^i_{max}) = 0&&&\\
+& s^{i}(q^i_{min})= 0&&&
+\end{split}
+\forall i\in 1,\dots,m
+\end{cases}
+\end{equation}
+$\mathbb{A}$ is also positive semidefinite, so the system is still provably stable w.r.t.
+the attractor.
+The structure of $S$ enforces kinematic joint limits because as a joint nears one of its limits, its magnitude decreases such that it will never actually collide with the limit.
+One way of constructing $S$ is to use the following function:
+\begin{equation}
+s^{i}(q) =
+1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{4}  ~\forall i\in 1,\dots,m\\
+\end{equation}
+The algorithm can similarly be expanded to accommodate joint velocity limits by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits.
 Specifically, given some set of velocity constraints for each joint $i$ s.t.
 $\dot{q}_i < \dot{q}_{i, max}$, we can design a new joint augmentation matrix $\mathcal{A}' = r \mathcal{A}$ where $r = \min \left ( 1, \max_i \left ( \frac{\dot{q}_{i, max}}{\dot{q}_i} \right ) \right )$.
-This velocity scalar would slow down the joint trajectory when the robot dynamics could not execute them, but would not change the position profile of the joint motion, preserving a major part of the joint behavior.
+This velocity scalar would slow down the joint trajectory when the robot dynamics could not execute them, but would not change the position profile of the joint motion, preserving a major aspect of the joint behavior.
 
@@
 Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system.
-This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors,.
+This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors.
+\textcolor{red}{Are we actually still doing this?}
 %\footnotesize
@@
 \begin{equation*}
-\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
 \end{equation*}
-accomplishes criterion (I), where $\mathcal{A}$ is positive semi-definite, and $P$ is positive definite.
+accomplishes criterion (I), where $A$ is positive semi-definite, and $P$ is positive definite.
 \newtheorem{theorem}{Theorem}[section]
@@
 \frac{dV(q)}{dt} &= (H(q) - x^*)^TPJ(q)\dot{q}\\
-&= -(H(q) - x^*)^TPJ(q)\mathcal{A}(q)J^T(q)P(H(q) - x^*)\\
-&= -(H(q) - x^*)^TPJ(q)S(q) \sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot S^T(q)J^T(q)P(H(q) - x^*) \leq 0
+&= -(H(q) - x^*)^TPJ(q)A(q)J^T(q)P(H(q) - x^*)\\
+&= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) \leq 0
 \end{aligned}
@@
 $ \| H(q)-x^*\| $ and $ \dot{q} $ are bounded.
-
-To prove the fulfillment of criterion (II), assume the contrapositive, that there exists some time $0<\tau$ such that $q^i_{max} < q^i(\tau) $.
-Because $\dot{q}^i$ given by \eqref{eq:ds} with respect to \eqref{eq:first_criteria_stability} and  \eqref{eq:first_criteria_joint}  is both bounded and continuous, we can invoke the intermediate value theorem to show that there must have been some $0<c<\tau$ s.t.
-$q^i(c) = q^i_{max}$.
-According to \eqref{eq:first_criteria_joint}, $s^i(q^i(c)) = 0$, which ensures that $\dot{q}^i(c) = f(q^i(c)) = 0$ and will be forever (by induction), which contradicts the assumption.
-The same proof holds for a $q^i(\tau) < q^i_{min} $.
-$\blacksquare$
-\section{Enforcing Joint Limits}
-\label{appendix:limits}
-We wish to find a joint-position-dependent diagonal matrix $S(q)$ which is bounded and satisfies:
-\begin{equation}
-\label{eq:last_criteria}
-\begin{cases}
-\begin{split}
-& s^{i}(q^i_{max}) = 0&&&\\
-& s^{i}(q^i_{min})= 0&&&
-\end{split}
-\forall i\in 1,\dots,m
-\end{cases}
-\end{equation}
-as previously described in Section \ref{sec:proposed_system}. We propose the using the following functions:
-\begin{equation}
-s^{i}(q) =
-1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{2d}  ~\forall i\in 1,\dots,m\\
-\end{equation}
-where $d$ is a positive integer. 
-This function is continuous and satisfies all of the above criteria.
-We can understand it holistically as follows: the ratio $\frac{q(v) - q_{min}}{q_{max} - q_{min}}$ maps the set of valid $q$ values to the range between $0$ and $1$.
-The rest maps these values to a parabola-like polynomial, such that at $q^i = q^i_{max}$ or $q^i_{min}$ the value is $0$, and elsewhere the value is positive (with the maximum value, $s^i = 1$, at $q^i = \frac{q_{max} - q_{min}}{2}$).
-Note as $d$ gets bigger, $S(q)$ is flatter away from the joint limits and drops off more sharply near them, which allows us to preserve a wider range of motion.
-We found $d=2$ to be an effective choice.\\
 

### 2017-10-26 paper/main_RA-L_v1.tex
Added synergy references throughout paper, especially in the beginning of Section IV. The subheadings of IV still need work.

@@
 Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it Lyapunov-stably converges towards its task-space attractor.
-To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system, and we provide a method for learning behavior matrices from demonstrations.
-Specifically, we use dimensionality reduction to find low-dimensional behavior regions, and then estimate the parameters of the corresponding joint-space behaviors by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
+To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
+Specifically, we use dimensionality reduction to find low-dimensional synergy regions, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
 Our proposed approach is validated on a variety of motions for a 7-DOF KUKA LWR4+ robot arm.\end{abstract}
@@
 \end{equation}
-where $P\in \mathbb{R}^{d\times d}$, and $A(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a nonlinear combination of constant linear matrices, each of which encodes a local joint-space behavior.
-$P$ is
+where $P\in \mathbb{R}^{d\times d}$, and $A(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
 \begin{equation}
@@
 \end{equation}
-where $A_k\in \mathbb{R}^{m\times m} $ and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are constant matrices and the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
+where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
 system $q $ or external signals $d(t)$, i.e.
@@
 In this paper, we only consider it as a function of the states of the system.
-It is noteworthy that the presented stability and convergence proof can easily be extended for $\theta_k (t, q(t), d(t))$.} representing each local behavior, respectively.
-Based on \eqref{eq:ds}, one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.
+It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$.} modulating each local synergy through time as well as space.
+Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more complex multi-behavior motion.
 
@@
 \includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
-\caption{Three example 3-DOF motions (A, B, C), each with a different constant behavior matrix $\mathcal{A}$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
+\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $A$ (emphasizing the hip, knee, and ankle respectively), A: $A(q)=diag(5,1,1) $, B: $A(q)=diag(1,5,1)$, and C: $A(q)=diag(1,1,5)$.
 On the left, the task-space traces of each motion.
@@
 
-\textcolor{red}{Talk about synergy space here, why we need it, etc.}\\
+The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{petrivc2013muscle}.
+In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors.
+In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and the combination of synergies used to control the system is $A(q)$ (through \eqref{eq:ds}).
+We are thus faced with two problems: how to determine the parameters for the synergies themselves ($A_k~\forall k\in\{1,\dots,K\}$) and how to combine the different synergies appropriately throughout different parts of the motion to yield our desired behavior (by intelligent choice of $\theta_k(q)$).
+Of course, we must also choose these parameters subject to the constraints laid out in \eqref{eq:first_criteria_stability}.
 
-JT-DS \eqref{eq:ds} is parametrized by $\theta_k(q) $ and $A_k~\forall k\in\{1,\dots,K\}$.
-One is free to determine these parameters as long as that they satisfy constraints \eqref{eq:first_criteria_stability}.
-We propose a 3-step procedure as follows:
+We propose a 3-step procedure to find these parameters.
+Counter-intuitively, we first find an appropriate way to combine the synergies (without knowing the synergies themselves) by learning a "synergy-space" for commanding the robot, and then identifying appropriate regions to place synergies in synergy-space.
+After, we will learn parameters for each synergy that would work best for each of the regions previously identified.
+More precisely:
+
 \begin{enumerate}
-\item We first project the demonstrated data (i.e.
-collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$ through Principal Component Analysis (PCA) \cite{jolliffe1986pca}, so as to reduce the dimensionality and extract correlations across similar joint behavior regions.
+\item We first construct our "synergy-space", which provides us a lower-dimensional manifold through which to control the robot, by projecting the demonstration data (i.e.
+collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$.
+In each experiment, we used either Principal Component Analysis (PCA) \cite{jolliffe1986pca} or Kernel PCA (KPCA) \cite{scholkopf1997kernel}.
 
-\item We then jointly estimate the optimal number $K$ of local behavior regions and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$.
+\item We then jointly estimate the optimal number $K$ of local synergies and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$.
 
-\item Finally, once the local behavior regions have been found (described by each of the Gaussian distributions $\theta_k(q)$), we compute the corresponding joint behavior matrices $A_k~\forall k\in\{1,\dots,K\}$ for each region by formulating a convex optimization problem that finds the optimal set of $A_k$'s that minimize the overall velocity error with respect to the demonstrations \eqref{eq:optim_error}.
+\item Finally, once the local synergy regions have been found (described by each of the Gaussian distributions $\theta_k(q)$), we compute the corresponding joint synergy matrices $A_k~\forall k\in\{1,\dots,K\}$ for each region by formulating a convex optimization problem that finds the optimal set of $A_k$'s that minimize the overall velocity error with respect to the demonstrations \eqref{eq:optim_error}.
 \end{enumerate}
 
-\subsection{Embedding Joint Configurations in Latent Space}
-The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint motion styles" (defined in Section \ref{sec:intro}).
+\subsection{Embedding Joint Configurations in Synergy Space}
+The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint behaviors" (defined in Section \ref{sec:intro}), in which it would be appropriate to define synergies.
 It is important to understand that distances in joint space (defined as $\| \Delta q\|$) might not numerically reflect what a human intuitively perceives as the distance between joint configurations.
@@
 
-\subsection{Discovering Local Behavior Regions}
-Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local behaviors, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
+\nocite{*}
+
+\subsection{Discovering Local Synergy Regions}
+Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
 Moreover, we would like for our scheduling parameters $\theta'_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta'_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
@@
 \end{equation}
-which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th behavior.
-Therefore, each joint-space behavior region is associated with a Gaussian component of the GMM, cumulatively describing all the behavior regions of the dynamical system.
+which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th synergy.
+Therefore, each synergy region is associated with a Gaussian component of the GMM, cumulatively describing all the synergy regions of the dynamical system.
 We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM and choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.
 
-\subsection{Estimating the Joint Behavior Matrices}
+\subsection{Estimating the Synergy Matrices}
 Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $A(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
@@
 
-Intuitively, each ``region of behavior" in joint/latent space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's behavior ($A_k$) influences the robot's current joint-space motions.
-Finding the appropriate joint augmentation matrices $A_k$ for the set of local behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
+Intuitively, each ``synergy region" in synergy space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
+Finding the appropriate synergy matrices $A_k$ for the set of local behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
 To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:
@@
 The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
-To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single Gaussian, i.e.
+To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
 uniform behaviors.
@@
 One of the most important points when validating a learning from demonstration method is to evaluate the system's behavior away from demonstrations.
-When the current joint configuration is far from any of the local behavior regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target.
+When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target.
 Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably moving to the target.
 The reason for this is that rather than ``determining" the velocity of the system, our $A$ matrix only warps it.
-This means that adding multiple ``conflicting" behaviors together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
+This means that adding multiple ``conflicting" synergies together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
 
@@
 Previous work proposed adding discontinuous limits to JT control \cite{sciavicco1988solution}, and we additionally propose a continuous modification as follows: \textcolor{red}{Remove this part?
-Should we not talk about continuous joint limiting?} Replace the joint behavior matrix $A(q)$ with a joint-limited joint behavior matrix $\mathcal{A} = S(q)A(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that
+Should we not talk about continuous joint limiting?} Replace the combined synergy matrix $A(q)$ with a joint-limited joint-behavior matrix $\mathcal{A} = S(q)A(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that
 \begin{equation}

### 2017-10-27 paper/main_RA-L_v1.tex
Reformulated the idea of synergy space

@@
 In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors.
-In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and the combination of synergies used to control the system is $A(q)$ (through \eqref{eq:ds}).
-We are thus faced with two problems: how to determine the parameters for the synergies themselves ($A_k~\forall k\in\{1,\dots,K\}$) and how to combine the different synergies appropriately throughout different parts of the motion to yield our desired behavior (by intelligent choice of $\theta_k(q)$).
-Of course, we must also choose these parameters subject to the constraints laid out in \eqref{eq:first_criteria_stability}.
+In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $A(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
 
-We propose a 3-step procedure to find these parameters.
-Counter-intuitively, we first find an appropriate way to combine the synergies (without knowing the synergies themselves) by learning a "synergy-space" for commanding the robot, and then identifying appropriate regions to place synergies in synergy-space.
-After, we will learn parameters for each synergy that would work best for each of the regions previously identified.
-More precisely:
+A central question becomes how to modulate the synergies in different regions to yield our desired behavior.
+First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that a latent mapping $\phi(\cdot)$ and latent space $B: \{b = \phi(a) | a \in A$\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
+We would like to define the robot's policy in synergy space such that in different regions of the space, we will utilize different synergies.
+We are thus left with three problems: finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(q)~\forall k\in\{1,\dots,K\}$), and finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
+Our choice of parameters must also obey the constraints laid out in \eqref{eq:first_criteria_stability}.
+We propose the following 3-step procedure.
 
@@
 
-\item We then jointly estimate the optimal number $K$ of local synergies and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$.
+\item We then jointly estimate the optimal number $K$ of synergy "regions" and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$ seen in the demonstrations.
 
@@
 \subsection{Embedding Joint Configurations in Synergy Space}
+
+%WE ASSUME THAT, BECAUSE OUR DEMONSTRATIONS ARE SIMILAR, POINTS IN SIMILAR REGIONS WILL HAVE SIMILAR MOTION POLICIES, SO FINDING As IS EQUIVALENT TO FINDING PRIMARY SYNERGY IN REGION, AND THEN INTERPOLATING THE SYNERGIES IN THE REGIONS IN-BETWEEN
+
 The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint behaviors" (defined in Section \ref{sec:intro}), in which it would be appropriate to define synergies.

### 2017-10-27 paper/main_RA-L_v1.tex
My comments and modifications.

@@
 \usepackage{authblk}
-\usepackage{hyperref}
+\usepackage[hidelinks]{hyperref}
 % *** MATH PACKAGES ***
@@
 \label{sec:intro}
+\textcolor{red}{We'll remove the subsection headings later..\\ Some comments:\\- Change PbD to LfD everywhere}
+\subsection{\textcolor{red}{Motivation for learning joint+task-space behaviors}}
+\textcolor{red}{*See comments in tex*}
+% % TO YO: OUTLINE FOR RE-WRITING THIS PART
+%1. Start with motivation of LfD  
+%2. Mention that we can learn behaviors in either joint or task space 
+%3. briefly summarize the approaches in each without going into detail; i.e. probabilistic models, DS, etc 
+%4.
+Motivate why we want to learn behaviors in joint+task space: 1 is because of the implications of IK as we already discuss, 2 to mimic joint trajectories..
+here give your example of why we would want that and mention applications such as table tennis[1], grasping[2] and bi-manual coordination [3].
+%
+%[1] Y. Huang, D. B ̈uchler, O. Koc, B. Scholkopf and J. Peters, “Jointly
+%learning trajectory generation and hitting point prediction in robot table
+%tennis,” in Proc. IEEE International Conference on Humanoid Robots,
+%2016, pp. 650-655.
+%
+%[2] S. Calinon and A. Billard, “Statistical learning by imitation of competing constraints in joint space and task space,” Advanced Robotics,vol. 23, pp. 2059-2076, 2009.
+%
+%[3] J. Silverio, S. Calinon, L. Rozo and D. G. Caldwell, “Learning competing constraints and task priorities from demonstrations of bimanual skills,” arXiv:1707.06791, 2017.
+
 Robot motion planning in joint-space has long been a major field of study \cite{kelly2006control}.
@@
 \label{fig:robot_example}
-\vspace{-10pt}
 \end{figure}
@@
 
-Although DS-based task-space learning techniques ensure target convergence and stability, an advantageous feat within the LfD paradigm, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution.
-Further, for certain tasks, rather than blindly choosing the arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.
+Although DS-based task-space learning techniques ensure target convergence and stability, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution.
+Further, for certain tasks, rather than generating arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.
 For example, throwing a ball is a motion less characterized by task-space behaviors (the ball moving backwards and then forwards) and more characterized by the joint behaviors: the elbow folds, the shoulder moves back and then swings forward, followed by the elbow straightening and the wrist snapping down.
@@
 SEDS, $\tau$-SEDS LMDS, etc.) do not allow for integrating combined task/joint-space behavior learning.
+
+\subsection{\textcolor{red}{Approaches that tackle the joint+task problem}}
+\textcolor{red}{*See comments in tex*}
+% TO YO:
+% ADD THESE THREE REFERENCES HERE AND SUMMARIZE THEM
+%[1] Y. Huang, D. B ̈uchler, O. Koc, B. Scholkopf and J. Peters, “Jointly
+%learning trajectory generation and hitting point prediction in robot table
+%tennis,” in Proc. IEEE International Conference on Humanoid Robots,
+%2016, pp. 650-655.
+%
+%[2] S. Calinon and A. Billard, “Statistical learning by imitation of competing constraints in joint space and task space,” Advanced Robotics,vol. 23, pp. 2059-2076, 2009.
+%
+%[3] J. Silverio, S. Calinon, L. Rozo and D. G. Caldwell, “Learning competing constraints and task priorities from demonstrations of bimanual skills,” arXiv:1707.06791, 2017.
+
 \cite{calinon2008probabilistic} notably attempted to address this problem by learning separate motion policies in task space and the null space of the Jacobian (which would not affect task-space position), and driving the robot with a weighted sum of the two.
@@
 
-%\cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework, but does not propose a dynamical system for generating motions.
+\cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework, but does not propose a dynamical system for generating motions.
 
+\subsection{Our proposed Approach}
+\textcolor{red}{Dynamical System, Why?\\Linear Parameter Varying, Why?\\Lower-Dimensional Manifold, Why?\\}
 In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
@@
 It shares some of our approach's advantages: fast computation and provable task-space stability.
-However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations while obeying the robots kinematic constraints.
+However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations.
+%while obeying the robots kinematic constraints.
 
@@
 %\end{figure}
-
 \section{Problem Statement} \label{Sec:Prob}
@@
 \label{sec:proposed_system}
+\textcolor{red}{Change semi-definite to definite everywhere\\Change overall $A(q)$ to $\mathcal{A}(q)$\\ Introduce 'behavior synergies' here\\}
 We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \ref{eq:Lyp} and \ref{eq:optim_error}.
@@
 A central question becomes how to modulate the synergies in different regions to yield our desired behavior.
-First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that a latent mapping $\phi(\cdot)$ and latent space $B: \{b = \phi(a) | a \in A$\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
+First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that a latent mapping $\phi(\cdot)$ and latent space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
 We would like to define the robot's policy in synergy space such that in different regions of the space, we will utilize different synergies.
@@
 
-\subsection{Embedding Joint Configurations in Synergy Space}
+\subsection{Embedding Joint Configurations in Lower-Dimensional Space}
 
@@
 
-\subsection{Discovering Local Synergy Regions}
+\subsection{Discovering Local Behavior Synergies}
 Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
@@
 
+\newpage
 \section{Experimental Validation} \label{Sec:Exp}
@@
 In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis.
-The source code for learning (MATLAB) and execution (C++) of the Augmented Joint-Space Task-Oriented Dynamical System together with data generated in these experiments are available on-line:
-\centerline{\small \underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}}
+The source code for learning (MATLAB) and execution (C++) of the Augmented Joint-Space Task-Oriented Dynamical System together with data generated in these experiments are available on-line: \\
+\footnotesize Learning:  \underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}\\
+\footnotesize Execution:  \underline{\url{https://github.com/epfl-lasa/JT-DS-lib}}\\
+\normalsize
 
-\textcolor{red}{1. Introduce Datasets and metrics}\\
+\begin{table}[t]
+\centering
+\scriptsize
+\caption{Datasets used to Evaluate JT-DS learning approach }
+\label{tab:datasets}
+\begin{tabular}{l|l|l}
+\hline\hline
+Task Behavior Dataset            & \# Demos & \# Samples (25Hz) \\ \hline
+Backhand Stroke     & 11       & 1223                    \\
+Forehand Stroke     & 10       & 1424                    \\
+Pouring - Free      & 9        & 1032                    \\
+Pouring - Obstacle  & 11       & 1232                    \\
+Foot Swing          & 8        & 1058                    \\
+Singularity Motions & 10       & 1467                 \\ \hline\hline
+\end{tabular}
+\end{table}
+
+To evaluate the proposed JT-DS learning algorithm we collected demonstrations from 6 different tasks (see Table \ref{tab:datasets}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space.
+These $M$ demonstrations consist of ($14$-D) time-series $\boldsymbol{\xi}^{(i)} = \{\xi_1^{(i)},\dots,\xi_T^{(i)}\}$ for $i= \{1,\dots,M\}$.
+The $14$ dimensions in $\xi_t = [q_t,\dot{q_t}]^T$ correspond to joints-space positions and velocities.
 
 \textcolor{red}{2. Table with Evaluation of Learning Schemes for JTDS Models on Different Datasets and discussion of results}
-
 \newpage

### 2017-10-27 paper/main_RA-L_v1.tex
Changed A to script A

@@
 \label{sec:proposed_system}
-\textcolor{red}{Change semi-definite to definite everywhere\\Change overall $A(q)$ to $\mathcal{A}(q)$\\ Introduce 'behavior synergies' here\\}
+\textcolor{red}{Change semi-definite to definite everywhere\\ Introduce 'behavior synergies' here\\}
 We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \ref{eq:Lyp} and \ref{eq:optim_error}.
@@
 \label{eq:ds}
-\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
 \end{equation}
-where $P\in \mathbb{R}^{d\times d}$, and $A(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
+where $P\in \mathbb{R}^{d\times d}$, and $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
 \begin{equation}
 \label{eq:A_def}
-A(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
+\mathcal{A}(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
 \end{equation}
@@
 \includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
-\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $A$ (emphasizing the hip, knee, and ankle respectively), A: $A(q)=diag(5,1,1) $, B: $A(q)=diag(1,5,1)$, and C: $A(q)=diag(1,1,5)$.
+\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
 On the left, the task-space traces of each motion.
@@
 the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
-The positive semi-definite matrix $A(q)$ warps the resulting joint-space velocity; Fig.
-\ref{fig:A_example} illustrates the effects of  $A(q)$ on the generated motion.
+The positive semi-definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
+\ref{fig:A_example} illustrates the effects of  $\mathcal{A}(q)$ on the generated motion.
 Thus the controller can be thought of as a proportional controller in joint space.
-Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
+Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
 %When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle).
@@
 
-%By learning a position-dependent joint augmentation matrix $A(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints.
-\eqref{eq:A_def} constructs our overall $A(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.
-In this paper the joint behavior matrix $A(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\
+%By learning a position-dependent joint augmentation matrix $\mathcal{A}(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints.
+\eqref{eq:A_def} constructs our overall $\mathcal{A}(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.
+In this paper the joint behavior matrix $\mathcal{A}(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\
 
@@
 In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors.
-In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $A(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
+In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $\mathcal{A}(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
 
@@
 \subsection{Estimating the Synergy Matrices}
-Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $A(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
+Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $\mathcal{A}(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
 \begin{equation}
 \label{eq:A}
-A(q) = \frac{\sum_{k=1}^K A_k \pi_k\mathcal{N}(\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}.
+\mathcal{A}(q) = \frac{\sum_{k=1}^K A_k \pi_k\mathcal{N}(\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}.
 \end{equation} 
@@
 Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably moving to the target.
-The reason for this is that rather than ``determining" the velocity of the system, our $A$ matrix only warps it.
+The reason for this is that rather than ``determining" the velocity of the system, our $\mathcal{A}$ matrix only warps it.
 This means that adding multiple ``conflicting" synergies together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
@@
 
-JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$.
+%JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$.
 Previous work proposed adding discontinuous limits to JT control \cite{sciavicco1988solution}, and we additionally propose a continuous modification as follows: \textcolor{red}{Remove this part?
-Should we not talk about continuous joint limiting?} Replace the combined synergy matrix $A(q)$ with a joint-limited joint-behavior matrix $\mathcal{A} = S(q)A(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that
-\begin{equation}
-\label{eq:last_criteria}
-\begin{cases}
-\begin{split}
-& s^{i}(q^i_{max}) = 0&&&\\
-& s^{i}(q^i_{min})= 0&&&
-\end{split}
-\forall i\in 1,\dots,m
-\end{cases}
-\end{equation}
-$\mathbb{A}$ is also positive semidefinite, so the system is still provably stable w.r.t.
+Should we not talk about continuous joint limiting?} Replace the combined synergy matrix $\mathcal{A}(q)$ with a joint-limited joint-behavior matrix $\mathcal{A}_l = S(q)\mathcal{A}(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that
+%\begin{equation}
+%\label{eq:last_criteria}
+%\begin{cases}
+%\begin{split}
+%& s^{i}(q^i_{max}) = 0&&&\\
+%& s^{i}(q^i_{min})= 0&&&
+%\end{split}
+%\forall i\in 1,\dots,m
+%\end{cases}
+%\end{equation}
+%$\mathbb{A}$ is also positive semidefinite, so the system is still provably stable w.r.t.
 the attractor.
@@
 One way of constructing $S$ is to use the following function:
-\begin{equation}
-s^{i}(q) =
-1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{4}  ~\forall i\in 1,\dots,m\\
-\end{equation}
-The algorithm can similarly be expanded to accommodate joint velocity limits by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits.
+%\begin{equation}
+%s^{i}(q) =
+%1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{4}  ~\forall i\in 1,\dots,m\\
+%\end{equation}
+%The algorithm can similarly be expanded to accommodate joint velocity limits by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits.
 Specifically, given some set of velocity constraints for each joint $i$ s.t.
@@
 \begin{equation*}
-\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
 \end{equation*}
-accomplishes criterion (I), where $A$ is positive semi-definite, and $P$ is positive definite.
+accomplishes criterion (I), where $\mathcal{A}(q)$ is positive semi-definite, and $P$ is positive definite.
 \newtheorem{theorem}{Theorem}[section]
@@
 \frac{dV(q)}{dt} &= (H(q) - x^*)^TPJ(q)\dot{q}\\
-&= -(H(q) - x^*)^TPJ(q)A(q)J^T(q)P(H(q) - x^*)\\
+&= -(H(q) - x^*)^TPJ(q)\mathcal{A}(q)J^T(q)P(H(q) - x^*)\\
 &= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) \leq 0
@@
 \end{equation}
-By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $A$, which is positive semi-definite.
+By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $\mathcal{A}(q)$, which is positive semi-definite.
 This means that the expression is guaranteed to be negative semi-definite.

### 2017-10-27 paper/main_RA-L_v1_nadia.tex
Created another _nadia.tex file with my changes.. Still have to finish Section 5.. discuss results.

--- 2017-10-27 paper/main_RA-L_v1.tex
@@
 \usepackage{authblk}
+\usepackage{tabularx}
+\usepackage{multirow}
 \usepackage[hidelinks]{hyperref}
@@
 \usepackage{amssymb}
+
+\usepackage{xcolor,colortbl}
+
+\newcommand{\mc}[2]{\multicolumn{#1}{c}{#2}}
+\definecolor{Gray}{gray}{0.85}
+\definecolor{LightCyan}{rgb}{0.88,1,1}
 
@@
 
-\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem in Synergy Space} 
+\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} 
 \label{Sec:Learning}
@@
 
-\newpage
+\clearpage
 \section{Experimental Validation} \label{Sec:Exp}
+\textcolor{red}{We should try to have page 5 start with this section, i.e. we must reduce 1 full column from the previous sections.}
+\subsection{JT-DS Learning Performance  Evaluation}
+To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 6 different tasks (see Table \ref{tab:learning}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space.
+These behaviors include: a (1) backhand and (2) forehand stroke, a pouring motion (3) with and (4) without an obstacle in its workspace, (5) a foot swing and (6) singularity motions.
+For each behavior, we collected $M$ demonstrations consisting of ($14$-D) time-series $\boldsymbol{\xi}^{(i)} = \{\xi_1^{(i)},\dots,\xi_T^{(i)}\}$ for $i= \{1,\dots,M\}$.
+The $14$ dimensions in $\xi_t = [q_t,\dot{q_t}]^T$ correspond to joints angles and velocities.
+We also record the final task-space configurations ${x^*}^{(i)}$, which we assume is the attractor in task-space.
+The recordings are initially collected at a rate of 500 Hz.
+However, our learning approach does not require such a dense representation of the trajectories, hence we down-sample them to 25 Hz.
+The total number of demonstrations $M$ and samples per \textit{behavior dataset} $M^*T$ are provided in Table \ref{tab:learning}.
+Given these datasets, we seek to:
+\begin{enumerate}[leftmargin=*]
+\item Verify that the learned \textit{behavior synergies} are capable of generating the desired joint-space behaviors.  
+\item Find the dimensionality approach which yields the best trade-off between accuracy and model complexity; i.e. the least dimensions $\delta$ and synergies $K$.
+\end{enumerate}
+We thus evaluate three learning approaches with different dimensionality reduction algorithms: (1) None, (2) PCA and (3) K-PCA with RBF kernel.
+We perform 10-fold cross-validation on all datasets, for each learning approach, with a training/testing ratio of $60 \%$.
+To evaluate the performance of each learning approach, we take the square root of the MSE defined in \eqref{eq:optim_error}, which we refer to as the joint-velocity RMSE.
+Performing such cross-validation in our case, is however not straight-forward, as our $\mathcal{A}(q)$ formulation has several hyper-parameters, namely the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
+Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$ in the RBF kernel.
+Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
+\textcolor{red}{\begin{enumerate}[leftmargin=*]
+\item Explain how optimal $\delta$ is computed for PCA. 
+\item Explain how optimal $\delta$ and $\sigma$ are found for K-PCA.
+\item Explain how optimal $K$ is computed for GMM.
+\end{enumerate}}
 
-\subsection{JT-DS Learning  Evaluation}
+\begin{table*}[!t]
+\scriptsize
+\caption{Performance Comparison of Learning Approach with different Dimensionality Reduction Schemes.
+We present the \texttt{mean(std)} for the optimal lower-dimensional $d$ and behavior synergies $K$ and RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
+\begin{center}
+\begin{tabular}{cc|cc|cc}
+    \hline
+    \hline
+    \multicolumn{1}{c}{\multirow{4}{*}{Behavior Dataset}} & \multicolumn{1}{c}{\multirow{4}{*}{Dim.
+Red.
+Approach}} & \multicolumn{4}{c}{\multirow{2}{*}{Optimal Parameters | Joint Velocity RMSE [$rad/s$]}}
+	\\  
+    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
 
-The learning algorithm presented in the previous section, \eqref{eq:opt}, was implemented in MATLAB.
-In order to solve the semidefinite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used.
+    \multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
+	& K-PCA ($\sigma = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
+	\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
+		& K-PCA ($\sigma = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
+		\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($M=9,M^*T=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
+	& K-PCA ($\sigma = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
+				\hline
+	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
+				& K-PCA ($\sigma = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
+				\hline	
+	\multicolumn{1}{c}{\multirow{3}{*}{Foot Swing}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{($M=8,M^*T=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
+	& K-PCA ($\sigma = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
+				\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
+	& K-PCA ($\sigma = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
+\end{tabular}
+\end{center}
+\vspace{-10pt}
+
+\end{table*}
+
+Once these optimal hyper-parameters are estimated for each fold, we solve for the definite convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
+We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from the training/testing datasets, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing set trajectories.
+
+\textcolor{red}{Discuss results: As can be seen in Table \ref{tab:learning}, ...}
+
+This learning pipeline was implemented in MATLAB.
+In order to solve the definite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used.
 In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis.
@@
 \footnotesize Learning:  \underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}\\
-\footnotesize Execution:  \underline{\url{https://github.com/epfl-lasa/JT-DS-lib}}\\
 \normalsize
+\textcolor{red}{In bla, we provide the necessary code to run this learning approach on our test datasets as well as on your own. We also provide code for the cross-validation evaluation.}
 
-\begin{table}[t]
-\centering
-\scriptsize
-\caption{Datasets used to Evaluate JT-DS learning approach }
-\label{tab:datasets}
-\begin{tabular}{l|l|l}
-\hline\hline
-Task Behavior Dataset            & \# Demos & \# Samples (25Hz) \\ \hline
-Backhand Stroke     & 11       & 1223                    \\
-Forehand Stroke     & 10       & 1424                    \\
-Pouring - Free      & 9        & 1032                    \\
-Pouring - Obstacle  & 11       & 1232                    \\
-Foot Swing          & 8        & 1058                    \\
-Singularity Motions & 10       & 1467                 \\ \hline\hline
-\end{tabular}
-\end{table}
-
-To evaluate the proposed JT-DS learning algorithm we collected demonstrations from 6 different tasks (see Table \ref{tab:datasets}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space.
-These $M$ demonstrations consist of ($14$-D) time-series $\boldsymbol{\xi}^{(i)} = \{\xi_1^{(i)},\dots,\xi_T^{(i)}\}$ for $i= \{1,\dots,M\}$.
-The $14$ dimensions in $\xi_t = [q_t,\dot{q_t}]^T$ correspond to joints-space positions and velocities.
-
-\textcolor{red}{2. Table with Evaluation of Learning Schemes for JTDS Models on Different Datasets and discussion of results}
-\newpage
-\subsection{JT-DS Performance Evaluation}
-The performance of the proposed framework was evaluated on a 7-DOF robot arm, the KUKA LWR 4+.
+\subsection{Comparison with Cartesian DS + IK Solver}
+We now evaluate the performance of the properties of JT-DS\footnote{With the optimal learning approach found in the previous section.} against a standard Cartesian motion generator + IK solver approach.
+In this evaluation we seek to compare the capabilities (or lack thereof) of both approaches to:
+\begin{enumerate}
+\item Follow desired behaviors in both joint and task spaces.
+\item Achieve fast computation/convergence time.
+\item Transit through singular configurations. 
+\end{enumerate}
+We continue our evaluation on a 7-DOF robot arm, the KUKA LWR 4+.
 The robot is controlled on the joint position level (linearly interpolating from joint velocities) at a rate of $ 500 $ Hz.
 The resultant joint angles are filtered by a critically damped filter to avoid high torques.
-Our empirical validation consists of three sections, each highlighting a different advantage of the proposed dynamical system: (i) moving in singular configurations, (ii) following desired behaviors in joint and task spaces simultaneously, and (iii) fast computation/convergence time.
 
-\subsubsection{Systematic Assessment}
-To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system.
-The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
-The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
-To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
-uniform behaviors.
-The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
-Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
-Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.
-
-\begin{table}[t]
-	\scriptsize
-	\vspace{0.1cm}
-	\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
-The convergence duration is the time required to move within $0.001$m of the target.
-The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
-	\label{table:2}
-	\centering
-	\begin{tabular}{|l|l|l|l|}
-		\hline
-		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
-		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ \hline
-		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline
-	\end{tabular}
-	\hspace{0.1cm}
-	 \vspace{10pt}	
-	\caption{Parameters and Performance of Comparative Experiments
-	 \textcolor{red}{This table is wrong.. re-do} }
+\begin{figure}[t]
 \vspace{-5pt}
 	\centering
-	\begin{tabular}{|l|c|c|c|c|}
-		\hline
-		& No. Dem. & No. K & $\phi(q)$ dim. & $e_{totol}$ [{rad}/{s}] \\ \hline
-		Pour Obst. (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ \hline
-		Foot (JT-DS) &       3               &     1             &          1 & 0.01344   \\ \hline
-		Pour Obst. (SEDS) &      7                &     3             &        -  &   -   \\ \hline
-		Foot (SEDS)&       3               &     1             &          - & -   \\ \hline
-	\end{tabular}
-	\label{table:1}
-	\vspace{-5pt}
-\end{table}
+	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
+	\caption{End-effector trajectories for the footstep motion in Cartesian space.
+The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
+On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
+\label{fig:foot}}
+\end{figure}
 
 \subsubsection{Following Desired Joint Behaviors}
-
 To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS \cite{khansari2011learning} algorithm using damped least square IK.
@@
 
-\subsubsection{Moving in singular configurations}
+\begin{table}[t]
+	\tiny
+	\vspace{0.1cm}
+	\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
+The convergence duration is the time required to move within $0.001$m of the target.
+The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
+	\label{table:2}
+	\centering
+	\begin{tabular}{l|l|l|l}
+		\hline\hline
+		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
+		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
+		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
+	\end{tabular}
+	\hspace{0.1cm}
+\end{table}
+
+\subsubsection{Systematic Assessment}
+To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system.
+The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
+The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
+To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
+uniform behaviors.
+The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
+Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
+Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.
+
+\subsubsection{Transiting in singular configurations}
 One of the main advantages of the proposed dynamical system is its ability to generate accurate paths in classical singular configurations.
@@
 }
-	\vspace{-20pt}
+	\vspace{-10pt}
 \end{figure}
@@
 We showed the system's ability to learn different joint-space behaviors on a redundant robotic platform.
-
-\begin{figure}
-\vspace{-10pt}
-	\centering
-	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
-	\caption{A plot of the end-effector trajectories for the footstep motion in Cartesian space.
-The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
-On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing due to becoming too large.
-\label{fig:foot}}
-	\vspace{-24pt}
-\end{figure}
 
@@
 
-JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$.
+%JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$.
 Previous work proposed adding discontinuous limits to JT control \cite{sciavicco1988solution}, and we additionally propose a continuous modification as follows: \textcolor{red}{Remove this part?
 Should we not talk about continuous joint limiting?} Replace the combined synergy matrix $A(q)$ with a joint-limited joint-behavior matrix $\mathcal{A} = S(q)A(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that
-\begin{equation}
-\label{eq:last_criteria}
-\begin{cases}
-\begin{split}
-& s^{i}(q^i_{max}) = 0&&&\\
-& s^{i}(q^i_{min})= 0&&&
-\end{split}
-\forall i\in 1,\dots,m
-\end{cases}
-\end{equation}
-$\mathbb{A}$ is also positive semidefinite, so the system is still provably stable w.r.t.
+%\begin{equation}
+%\label{eq:last_criteria}
+%\begin{cases}
+%\begin{split}
+%& s^{i}(q^i_{max}) = 0&&&\\
+%& s^{i}(q^i_{min})= 0&&&
+%\end{split}
+%\forall i\in 1,\dots,m
+%\end{cases}
+%\end{equation}
+%$\mathbb{A}$ is also positive semidefinite, so the system is still provably stable w.r.t.
 the attractor.
@@
 
-Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system.
+%Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system.
 This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors.
@@
 \label{appendix:stability}
+\small
 We wish to prove Proposition \ref{prop:stability}, that is, that JT-DS \eqref{eq:ds} (reproduced below)

### 2017-10-28 paper/main_RA-L_v1_nadia.tex
Latest changes in Section 5 on _nadia.tex.. also restructured directory and have new figures

@@
 \section{Experimental Validation} \label{Sec:Exp}
-\textcolor{red}{We should try to have page 5 start with this section, i.e. we must reduce 1 full column from the previous sections.}
+%\textcolor{red}{We should try to have page 5 start with this section, i.e. we must reduce 1 full column from the previous sections.}
 \subsection{JT-DS Learning Performance  Evaluation}
 To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 6 different tasks (see Table \ref{tab:learning}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space.
-These behaviors include: a (1) backhand and (2) forehand stroke, a pouring motion (3) with and (4) without an obstacle in its workspace, (5) a foot swing and (6) singularity motions.
+These behaviors include (as shown in Fig.
+\ref{fig:behaviors} ):
+\begin{itemize}[leftmargin=*]
+\item \textit{(1-2) Tennis strokes:} For these particular behaviors, we guide the robot to perform a forehand and backhand stroke.  
+\item \textit{(3-4) Pouring motions:} 
+\item \textit{(5) Footstep-like motion:} This behavior begins with a straight leg, moving through a singularity, and finally bringing the knee up (Fig. \ref{fig:snapshot:c})
+\item \textit{(6) Singularity motions:} 
+\end{itemize}
+
 For each behavior, we collected $M$ demonstrations consisting of ($14$-D) time-series $\boldsymbol{\xi}^{(i)} = \{\xi_1^{(i)},\dots,\xi_T^{(i)}\}$ for $i= \{1,\dots,M\}$.
@@
 \end{enumerate}}
+
+\begin{figure}[t]
+	\centering	
+		\begin{subfigure}[t]{\linewidth}
+		\centering
+			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
+					\caption{Tennis Strokes (left) Forehand (right) Backhand.}
+		\end{subfigure}
+		\begin{subfigure}[t]{\linewidth}
+\centering
+	\includegraphics[width=0.45\linewidth]{new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{new-pics/pour-obst_crop.png}
+						\caption{Pouring Motions (left) without obstacle (right) with obstacle.}
+			\end{subfigure}
+			\begin{subfigure}[t]{\linewidth}		
+\centering
+	\includegraphics[width=0.475\linewidth]{new-pics/footstep.png}\hspace{5pt}\includegraphics[width=0.4475\linewidth]{new-pics/singularity_crop.png}
+	\caption{(left) Foot-step motion (right) Singularity motions.}
+				\end{subfigure}
+	\caption{Demonstrated Joint-Space Behaviors with Task-Space Targets. \label{fig:behaviors}}
+	\vspace{-10pt}
+\end{figure}
 
@@
 In the first experiment, the robot was guided through a complex motion task, moving an object through an environment with an obstacle.
-Human experts guided the robot through a series of demonstrations, ensuring that neither the robot's joints nor the held object intersected the obstacle.
+%Human experts guided the robot through a series of demonstrations, ensuring that neither the robot's joints nor the held object intersected the obstacle. 
 The results can be seen in Fig.
@@
 It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
-In the second experiment the robot was taught a footstep-like motion, beginning with a straight leg, moving through a singularity, and finally bringing the knee up (Fig.
-\ref{fig:snapshot:c}).
+.
 JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig.
@@
 \begin{table}[t]
-	\tiny
 	\vspace{0.1cm}
@@
 	\centering
-	\begin{tabular}{l|l|l|l}
+	\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
 		\hline\hline
@@
 		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
-	\end{tabular}
-	\hspace{0.1cm}
+	\end{tabular}}
+	\vspace{-10pt}
 \end{table}
+
+\begin{figure}[t]
+\vspace{-0.05cm}
+	\begin{subfigure}[t]{\linewidth}
+		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
+		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
+		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
+		\caption{Desired Joint Behavior: obstacle avoidance motion. The luggage is the obstacle.}
+		\label{fig:snapshot:a}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{\linewidth}
+		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}
+		\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
+		\includegraphics[width=0.32\linewidth]{Pic/Luggage3_SEDS.png}
+		\caption{Desired Behavior: The Cartesian DS is used to generate the motion.}
+		\label{fig:snapshot:b}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{\linewidth}
+	\centering	
+		\includegraphics[width=0.45\linewidth]{Pic/1.png}~~
+		\includegraphics[width=0.45\linewidth]{Pic/3.png}\\\vspace{0.1cm}
+		\includegraphics[width=0.45\linewidth]{Pic/4.png}~~
+		\includegraphics[width=0.45\linewidth]{Pic/2.png}
+		\caption{Desired Joint Behavior: Foot step-like motion.}
+		\label{fig:snapshot:c}
+	\end{subfigure}\\	
+	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{https://youtu.be/1dgKfmN1UgE}.}
+	\label{fig:snapshot}
+	\vspace{-0.8cm}
+\end{figure}
 
@@
 \textcolor{red}{Are we actually still doing this?}
-%\footnotesize
+
+\footnotesize
 \appendices
-
-\begin{figure}[t]
-\vspace{-0.05cm}
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
-		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
-		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
-		\caption{Desired Joint Behavior: obstacle avoidance motion. The luggage is the obstacle.}
-		\label{fig:snapshot:a}
-	\end{subfigure}\\
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}
-		\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
-		\includegraphics[width=0.32\linewidth]{Pic/Luggage3_SEDS.png}
-		\caption{Desired Behavior: The Cartesian DS is used to generate the motion.}
-		\label{fig:snapshot:b}
-	\end{subfigure}\\
-	\begin{subfigure}[t]{\linewidth}
-	\centering	
-		\includegraphics[width=0.45\linewidth]{Pic/1.png}~~
-		\includegraphics[width=0.45\linewidth]{Pic/3.png}\\\vspace{0.1cm}
-		\includegraphics[width=0.45\linewidth]{Pic/4.png}~~
-		\includegraphics[width=0.45\linewidth]{Pic/2.png}
-		\caption{Desired Joint Behavior: Foot step-like motion.}
-		\label{fig:snapshot:c}
-	\end{subfigure}\\	
-	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{https://youtu.be/1dgKfmN1UgE}.}
-	\label{fig:snapshot}
-	\vspace{-0.8cm}
-\end{figure}
-
 \section{Proving Stability of the Dynamical System}

### 2017-10-28 paper/main_RA-L_v1_nadia.tex
Final changes of the day.. will continue tomorrow

@@
 \item \textit{(1-2) Tennis strokes:} For these particular behaviors, we guide the robot to perform a forehand and backhand stroke.  
-\item \textit{(3-4) Pouring motions:} 
-\item \textit{(5) Footstep-like motion:} This behavior begins with a straight leg, moving through a singularity, and finally bringing the knee up (Fig. \ref{fig:snapshot:c})
-\item \textit{(6) Singularity motions:} 
+\item \textit{(3-4) Pouring motions:} We guide the robot to pour chips on a bowl.
+The first instance is without an obstacle in the workspace, which is demonstrated as a linear joint interpolation.
+The second instance is with an obstacle, where the human is seen folding the elbow and lowering the shoulder of the robot arm to avoid the obstacle.
+\item \textit{(5) Footstep-like motion:} We demonstrate a footstep motion which begins with a straight leg, moving through a singularity, and finally bringing the knee up.
+\item \textit{(6) Singularity motions:} For this behavior we simulate joint trajectories in which the base rotates and all of the other joints are locked in a singular configuration.
 \end{itemize}
@@
 We perform 10-fold cross-validation on all datasets, for each learning approach, with a training/testing ratio of $60 \%$.
-To evaluate the performance of each learning approach, we take the square root of the MSE defined in \eqref{eq:optim_error}, which we refer to as the joint-velocity RMSE.
+To evaluate performance, we take the square root of the MSE defined in \eqref{eq:optim_error}, which we refer to as the joint-velocity RMSE.
+
 Performing such cross-validation in our case, is however not straight-forward, as our $\mathcal{A}(q)$ formulation has several hyper-parameters, namely the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
@@
 \scriptsize
-\caption{Performance Comparison of Learning Approach with different Dimensionality Reduction Schemes.
+\caption{\scriptsize Performance Comparison of Learning Approach with different Dimensionality Reduction Schemes.
 We present the \texttt{mean(std)} for the optimal lower-dimensional $d$ and behavior synergies $K$ and RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
@@
 \subsection{Comparison with Cartesian DS + IK Solver}
-We now evaluate the performance of the properties of JT-DS\footnote{With the optimal learning approach found in the previous section.} against a standard Cartesian motion generator + IK solver approach.
+We now evaluate the performance of JT-DS\footnote{With the optimal learning approach found in the previous section.} against a standard Cartesian motion generator + IK solver approach.
 In this evaluation we seek to compare the capabilities (or lack thereof) of both approaches to:

### 2017-10-28 paper/main_RA-L_v1.tex
Rewrote introduction, added discussion section on why we don't control directly in embedding space, still need to rewrite Sec. IV

@@
 \ref{fig:robot_example}.
+In this paper, we will explore the problem of learning a preferred joint-space behavior from previously demonstrated trajectories while still accomplishing a task-space goal, through the Learning from Demonstrations (LfD) paradigm \cite{billard2008robot,Argall:RAS:2009}.
 
-Learning robot motion generators that follow desired joint and/or task-space behaviors is, in fact, the central focus of Programming by Demonstration (PbD) \cite{billard2008robot} \cite{Argall:RAS:2009}.
-Historically, much of the original work in PbD focused on learning motions solely in joint-space by parameterizing motions using GMMs \cite{Calinon:HRI:2007}, HMMs \cite{Garrido:Neuro:2015}, GPs \cite{Shon:HUM:2005}, or other probabilistic models \cite{Schaal:IROS:2003,Schaal:AI:2002}.
+Many LfD approaches learn motions in either joint space \cite{Calinon:HRI:2007, Garrido:Neuro:2015, Schaal:IROS:2003,Schaal:AI:2002} or task space \cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}.
+These approaches generally learn probabilistic model of the demonstrated motions, and construct a dynamical system that chooses behaviors according to the learned model.
+Both joint and task space are independently useful: joint-space behaviors let us directly influence the behavior of the physical robot, while task-space behaviors let us control the most task-relevant component of the robot's behavior.
+In particular, a number of approaches have pioneered stable convergence to an attractor in task space \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015} as a desirable capability for a dynamical system.
+
+In many cases, one may want to specify a joint-space behavior while simultaneously fulfilling a task-space objective.
+For example, throwing a ball is a motion characterized both by a task-space behavior (the ball moves backwards and then forwards) and also more obviously by its joint-space behavior (the elbow folding, the shoulder moving back and then swinging forward, and then the elbow straightening and the wrist snapping down).
+Similarly, learned joint-space behaviors are helpful in playing ping pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bimanual manipulators\cite{silverio2017learning}.
+For a visual example of a task-space problem in which joint behavior is essential, see Fig.
+\ref{fig:robot_example}.
+
+A second reason for learning a motion in joint space is that this allows us to avoid inverse kinematic (IK) approximations.
+The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}.
+When the main focus is on executing a complex task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
+However, for other applications, such an approach yields significant problems \cite{buss2004introduction}.
+First of all, finding the pseudo-inverse is computationally taxing.
+Moreover, when the Jacobian matrix cannot be inverted (i.e.
+when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior.
+These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}.
+By defining a controller in joint space directly, one can avoid IK and all its associated drawbacks.
+%Further, for certain tasks, rather than generating arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.
+For example, throwing a ball is a motion less characterized by task-space behaviors (the ball moving backwards and then forwards) and more characterized by the joint behaviors: the elbow folds, the shoulder moves back and then swings forward, followed by the elbow straightening and the wrist snapping down.
+For a visual example of the need for such combined task/joint-space behavior learning, see Fig.
+\ref{fig:robot_example}.
+%In this work, we propose a method for learning a joint-space dynamical system that learns behaviors from demonstrations in joint space, while provably converging to an attractor in task space.
+%For all the previously mentioned task-space techniques, the robot is ultimately controlled by projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}.
+When the main focus is on executing a complex task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
+However, for other applications, such an approach yields significant problems \cite{buss2004introduction}.
+First of all, finding the pseudo-inverse is computationally taxing.
+Moreover, when the Jacobian matrix cannot be inverted (i.e.
+when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior.
+These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}.
+%Historically, much of the original work in LfD focused on learning motions solely in joint-space by parameterizing motions using GMMs \cite{Calinon:HRI:2007}, HMMs \cite{Garrido:Neuro:2015}, GPs \cite{Shon:HUM:2005}, or other probabilistic models \cite{Schaal:IROS:2003,Schaal:AI:2002}.
 Concurrently, a significant body of work from the graphics community was targeted at learning joint-space motion ``styles" from human motion capture data, through a variety of latent-space and stylistic probabilistic motion models \cite{gielniak2010stylized,grochow2004style}.
@@
 Furthermore, a task-space trajectory is robot-independent, and thus can be reused across robots with differing morphologies.
-
-A major drawback of these initial task-space methods was that despite learning motions similar to the demonstrations, there was no guarantee when applying the learned motion to a new problem (e.g.
+%A major drawback of these initial task-space methods was that despite learning motions similar to the demonstrations, there was no guarantee when applying the learned motion to a new problem (e.g.
 a different initial/target position) that the end-effector would stably converge to its ultimate task-space target.
@@
 \end{figure}
-
-For all the previously mentioned task-space techniques, the robot is ultimately controlled by projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}.
+%For all the previously mentioned task-space techniques, the robot is ultimately controlled by projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}.
 When the main focus is on executing a complex task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
@@
 These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}.
-
-Although DS-based task-space learning techniques ensure target convergence and stability, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution.
+%Although DS-based task-space learning techniques ensure target convergence and stability, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution.
 Further, for certain tasks, rather than generating arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.
@@
 
-\subsection{\textcolor{red}{Approaches that tackle the joint+task problem}}
+\subsection{Prior Work on Combined Joint and Task Space Control}
 \textcolor{red}{*See comments in tex*}
@@
 \cite{calinon2008probabilistic} notably attempted to address this problem by learning separate motion policies in task space and the null space of the Jacobian (which would not affect task-space position), and driving the robot with a weighted sum of the two.
-However, this approach no longer guarantees convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian.
+However, this approach does not seek convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian.
+\cite{calinon2010learning} and \cite{silverio2017learning} expand on this approach by projecting task-space constraints into joint space using IK, and then learning a joint-space policy that incorporates both task and joint space constraints, but this similarly relies on IK approximations and does not ensure convergence to an attractor.
+% FACT-CHECK: is this what Calinon 2009 and Silverio et al. 2017 are doing? They don't ensure convergence, right? TP-GMM never seems to address it
 
 \cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework, but does not propose a dynamical system for generating motions.
+\cite{huang2016jointly} learns a controller for playing table tennis by combining joint-space Dynamic Movement Primitives \cite{schaal2006dynamic} to mimic a desired task-space racket swing, but does not learn joint-space behaviors.
+% Do DMPs count as learning joint space behaviors? If yes, I'm having a hard time differentiating huang2016 from us.
 
-\subsection{Our proposed Approach}
-\textcolor{red}{Dynamical System, Why?\\Linear Parameter Varying, Why?\\Lower-Dimensional Manifold, Why?\\}
-In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
 \cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities.
@@
 This gives rise to spurious attractors or cycles, and thus must be carefully tuned in order to avoid them.
-In this work, we depart from any modulation or combination of DS and instead propose a novel dynamical system which:
 
+\subsection{Our Proposed Approach}
+In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
+We specifically chose to focus on linear parameter-varying dynamical systems because they are fast to compute and thus allow a controller to quickly react in a changing environment, a capability that is further improved by our not needing to compute pseudo-inverses.
+Our dynamical system:
 \begin{enumerate}[leftmargin=*]
@@
 \end{enumerate}
+In brief, our algorithm learns a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behaviors in joint space, and modulates the use of these synergies throughout joint-space using a learned LPV system.
+We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
 
-The most similar approach to our proposed DS is the Jacobian transpose (JT) control \cite{wolovich1984computational} method. JT control is an inverse kinematics
-method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably (in the sense of Lyapunov) over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations.
+The most similar approach to our proposed DS is the Jacobian transpose (JT) control \cite{wolovich1984computational} method.
+JT control is an inverse kinematics method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably (in the sense of Lyapunov) over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations.
 It shares some of our approach's advantages: fast computation and provable task-space stability.
-However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations.
+However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is to the best of our knowledge the first work to employ a JT system to learn behaviors from demonstrations.
 %while obeying the robots kinematic constraints.
@@
 A central question becomes how to modulate the synergies in different regions to yield our desired behavior.
-First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that a latent mapping $\phi(\cdot)$ and latent space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
+First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that an embedding $\phi(\cdot)$ and embedding space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
 We would like to define the robot's policy in synergy space such that in different regions of the space, we will utilize different synergies.
@@
 
+Since we learn a lower-dimensional embedding space $Z := \{\phi(q) ~|~q \in Q\}$, which we suggest provides a better representation for the learned joint behaviors, a careful reader might wonder: why not define a motion policy $\pi_Z$ directly in $Z$-space and then map the learned policy back out into joint space using the inverse embedding\footnote{Such an inverse is trivial to extract in the case where $\phi$ is obtained through PCA.} $\phi^{-1}(z)$?
+The answer is that we would lose any guarantee of stability with respect to a target attractor $x_t$.
+By construction, $\phi$ maps from a joint-position vector of size $d$ to a low-dimensional vector of size $u<d$.
+Thus the inverse $\phi^{-1}$ must be of less than full rank, so the resulting $\mathcal{A}$ matrix (derived as in \eqref{eq:ds}) will also be of less than full rank, and thus no longer positive definite.
+This means that our controller no longer provably converges to the attractor (since our convergence proof no longer holds).
+Intuitively, this is because any policy defined in a lower-dimensional space than the actuation space will forfeit certain degrees of freedom, and thus may not be able to span the configuration space.
+Instead, we use the embedding space to modulate synergies defined in joint space, and which we can thus guarantee will be positive definite and lead us to convergence to the target.
 %JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$.

### 2017-10-29 paper/main_RA-L_v1_nadia.tex
Added demo videos link for further reference
Latest updates on Section 5 in _nadia.tex

@@
 
-\clearpage
+\begin{figure*}[!ht]
+  \begin{minipage}[b]{0.3\textwidth}
+	\centering	
+		\begin{subfigure}[t]{\linewidth}
+		\centering
+			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
+					\caption{Tennis Strokes (1-2).}
+		\end{subfigure}
+		\begin{subfigure}[t]{\linewidth}
+\centering
+	\includegraphics[width=0.45\linewidth]{new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{new-pics/pour-obst_crop.png}
+						\caption{ Pouring Motions (3-4)}
+			\end{subfigure}
+			\begin{subfigure}[t]{\linewidth}		
+\centering
+	\includegraphics[width=0.45\linewidth]{new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{new-pics/footstep.png}
+	\caption{ Pouring Motion (5) and Foot-step (6).}
+				\end{subfigure}
+    \captionof{figure}{Demonstrated Joint-Space Behaviors with Task-Space Targets. \href{https://www.dropbox.com/sh/3vx512oldgbnkq5/AABP5iEqO0oYn7dlE-gvK1kPa?dl=0}{[videos]} \label{fig:behaviors}}
+  \end{minipage}
+  \hfill
+  \begin{minipage}[b]{0.7\textwidth}
+    \scriptsize
+    \centering
+\resizebox{\linewidth}{!}{\begin{tabular}{cc|cc|cc}
+    \hline
+    \hline
+    \multicolumn{1}{c}{\multirow{4}{*}{Behavior Dataset}} & \multicolumn{1}{c}{\multirow{4}{*}{Dim.
+Red.
+Approach}} & \multicolumn{4}{c}{\multirow{2}{*}{Optimal Parameters | Joint Velocity RMSE [$rad/s$]}}
+	\\  
+    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
+		& K-PCA ($\sigma = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
+		\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
+		    \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
+			& K-PCA ($\sigma = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
+			\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($M=9,M^*T=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
+	& K-PCA ($\sigma = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
+				\hline
+	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle 1}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
+				& K-PCA ($\sigma = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
+				\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{\textcolor{red}{Pouring - Obstacle 2}}} & None & 7 & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($M=X,M^*T=Y$)}} & PCA & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) \\	
+				& K-PCA ($\sigma = X$) & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y)  \\
+				\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Foot Step}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{($M=8,M^*T=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
+	& K-PCA ($\sigma = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
+				\hline
+
+	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
+	& K-PCA ($\sigma = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
+\end{tabular}}
+      \captionof{table}{\scriptsize Performance Comparison of Learning Approach with different Dim.
+Red.
+Schemes.
+We present the \texttt{mean(std)} for the optimal $\delta$, $K$ and joint-velocity RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
+    \end{minipage}
+\end{figure*}
+
+\newpage
 \section{Experimental Validation} \label{Sec:Exp}
-%\textcolor{red}{We should try to have page 5 start with this section, i.e. we must reduce 1 full column from the previous sections.}
 \subsection{JT-DS Learning Performance  Evaluation}
-To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 6 different tasks (see Table \ref{tab:learning}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space.
+To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 7 different tasks (see Table \ref{tab:learning}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space.
 These behaviors include (as shown in Fig.
@@
 \item \textit{(1-2) Tennis strokes:} For these particular behaviors, we guide the robot to perform a forehand and backhand stroke.  
-\item \textit{(3-4) Pouring motions:} We guide the robot to pour chips on a bowl.
-The first instance is without an obstacle in the workspace, which is demonstrated as a linear joint interpolation.
-The second instance is with an obstacle, where the human is seen folding the elbow and lowering the shoulder of the robot arm to avoid the obstacle.
-\item \textit{(5) Footstep-like motion:} We demonstrate a footstep motion which begins with a straight leg, moving through a singularity, and finally bringing the knee up.
-\item \textit{(6) Singularity motions:} For this behavior we simulate joint trajectories in which the base rotates and all of the other joints are locked in a singular configuration.
+\item \textit{(3-5) Pouring motions:} We guide the robot to pour chips in a bowl with three different joint-space behaviors.
+The first is without an obstacle in the workspace.
+The second is with an obstacle in the upper hemisphere of the workspace, where the human is seen folding the elbow and lowering the shoulder of the robot arm to avoid the obstacle.
+The third instance is with an obstacle in the lower hemisphere, and now the human raises the shoulder to avoid it.
+\item \textit{(6) Footstep-like motion:} We demonstrate a footstep motion which begins with a straight leg, moving through a singularity, and finally bringing the knee up.
+\item \textit{(7) Singularity motions:} For this behavior we simulate joint trajectories in which the base rotates and all of the other joints are locked in singular configurations.
 \end{itemize}
 
-For each behavior, we collected $M$ demonstrations consisting of ($14$-D) time-series $\boldsymbol{\xi}^{(i)} = \{\xi_1^{(i)},\dots,\xi_T^{(i)}\}$ for $i= \{1,\dots,M\}$.
-The $14$ dimensions in $\xi_t = [q_t,\dot{q_t}]^T$ correspond to joints angles and velocities.
-We also record the final task-space configurations ${x^*}^{(i)}$, which we assume is the attractor in task-space.
-The recordings are initially collected at a rate of 500 Hz.
+\noindent For each behavior, we collected $N$ demonstrations consisting of ($14$-D) time-series $D=\{\{ q_{t,n},\dot{q}_{t,n} \}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where each demonstration has duration $T_n$.
+We also record the final task-space configurations ${x^*}^{(n)}$, which we assume is the attractor in task-space.
+The recordings are collected at a rate of 500 Hz.
 However, our learning approach does not require such a dense representation of the trajectories, hence we down-sample them to 25 Hz.
-The total number of demonstrations $M$ and samples per \textit{behavior dataset} $M^*T$ are provided in Table \ref{tab:learning}.
+The total number of demonstrations $N$ and samples per \textit{dataset} $M = \sum_{n=1}^{N}T_n$ are provided in Table \ref{tab:learning}.
 Given these datasets, we seek to:
 \begin{enumerate}[leftmargin=*]
-\item Verify that the learned \textit{behavior synergies} are capable of generating the desired joint-space behaviors.  
+\item Verify that the learned \textit{behavior synergies} are capable of generating the demonstrated joint-space behaviors.  
 \item Find the dimensionality approach which yields the best trade-off between accuracy and model complexity; i.e. the least dimensions $\delta$ and synergies $K$.
@@
 
-Performing such cross-validation in our case, is however not straight-forward, as our $\mathcal{A}(q)$ formulation has several hyper-parameters, namely the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
+Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters, namely the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
 Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$ in the RBF kernel.
@@
 
-\begin{figure}[t]
-	\centering	
-		\begin{subfigure}[t]{\linewidth}
-		\centering
-			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
-					\caption{Tennis Strokes (left) Forehand (right) Backhand.}
-		\end{subfigure}
-		\begin{subfigure}[t]{\linewidth}
-\centering
-	\includegraphics[width=0.45\linewidth]{new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{new-pics/pour-obst_crop.png}
-						\caption{Pouring Motions (left) without obstacle (right) with obstacle.}
-			\end{subfigure}
-			\begin{subfigure}[t]{\linewidth}		
-\centering
-	\includegraphics[width=0.475\linewidth]{new-pics/footstep.png}\hspace{5pt}\includegraphics[width=0.4475\linewidth]{new-pics/singularity_crop.png}
-	\caption{(left) Foot-step motion (right) Singularity motions.}
-				\end{subfigure}
-	\caption{Demonstrated Joint-Space Behaviors with Task-Space Targets. \label{fig:behaviors}}
-	\vspace{-10pt}
-\end{figure}
-
-\begin{table*}[!t]
-\scriptsize
-\caption{\scriptsize Performance Comparison of Learning Approach with different Dimensionality Reduction Schemes.
-We present the \texttt{mean(std)} for the optimal lower-dimensional $d$ and behavior synergies $K$ and RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
-\begin{center}
-\begin{tabular}{cc|cc|cc}
-    \hline
-    \hline
-    \multicolumn{1}{c}{\multirow{4}{*}{Behavior Dataset}} & \multicolumn{1}{c}{\multirow{4}{*}{Dim.
+%\begin{figure}[t]
+%	\centering	
+%		\begin{subfigure}[t]{\linewidth}
+%		\centering
+%			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
+%					\caption{Tennis Strokes (left) Forehand (right) Backhand.}
+%		\end{subfigure}
+%		\begin{subfigure}[t]{\linewidth}
+%\centering
+%	\includegraphics[width=0.45\linewidth]{new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{new-pics/pour-obst_crop.png}
+%						\caption{Pouring Motions (left) without obstacle (right) with obstacle 1.}
+%			\end{subfigure}
+%			\begin{subfigure}[t]{\linewidth}		
+%\centering
+%	\includegraphics[width=0.45\linewidth]{new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{new-pics/footstep.png}
+%	\caption{(left) Pouring Motions with obstacle (right) with obstacle 2 (right) Foot-step motion.}
+%				\end{subfigure}
+%	\caption{Demonstrated Joint-Space Behaviors with Task-Space Targets. \href{https://www.dropbox.com/sh/3vx512oldgbnkq5/AABP5iEqO0oYn7dlE-gvK1kPa?dl=0}{[videos]} \label{fig:behaviors}}
+%	\vspace{-10pt}
+%\end{figure}
+%
+%\begin{table*}[!t]
+%\scriptsize
+%\begin{center}
+%\begin{tabular}{cc|cc|cc}
+%    \hline
+%    \hline
+%    \multicolumn{1}{c}{\multirow{4}{*}{Behavior Dataset}} & \multicolumn{1}{c}{\multirow{4}{*}{Dim.
 Red.
 Approach}} & \multicolumn{4}{c}{\multirow{2}{*}{Optimal Parameters | Joint Velocity RMSE [$rad/s$]}}
-	\\  
-    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
-
-    \multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
-	& K-PCA ($\sigma = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
-	\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
-		& K-PCA ($\sigma = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
-		\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($M=9,M^*T=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
-	& K-PCA ($\sigma = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
-				\hline
-	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
-				& K-PCA ($\sigma = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
-				\hline	
-	\multicolumn{1}{c}{\multirow{3}{*}{Foot Swing}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
-	\multicolumn{1}{c}{\multirow{3}{*}{($M=8,M^*T=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
-	& K-PCA ($\sigma = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
-				\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
-	& K-PCA ($\sigma = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
-\end{tabular}
-\end{center}
-\vspace{-10pt}
-
-\end{table*}
+%	\\  
+%    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
+%            
+%    \multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
+%    \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
+%	& K-PCA ($\sigma = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
+%	\hline
+%	
+%	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
+%    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
+%		& K-PCA ($\sigma = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
+%		\hline
+%		
+%	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
+%   \multicolumn{1}{c}{\multirow{3}{*}{($M=9,M^*T=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
+%	& K-PCA ($\sigma = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
+%				\hline
+%	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
+%   \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
+%				& K-PCA ($\sigma = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
+%				\hline	
+%	\multicolumn{1}{c}{\multirow{3}{*}{Foot Swing}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
+%	\multicolumn{1}{c}{\multirow{3}{*}{($M=8,M^*T=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
+%	& K-PCA ($\sigma = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
+%				\hline		
+%	
+%	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
+%    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
+%	& K-PCA ($\sigma = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
+%\end{tabular}
+%\caption{\scriptsize Performance Comparison of Learning Approach with different Dimensionality Reduction Schemes.
+We present the \texttt{mean(std)} for the optimal lower-dimensional $d$ and behavior synergies $K$ and RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
+%\end{center}
+%\vspace{-10pt}
+%\end{table*}
 
@@
 
-\begin{figure}[t]
-\vspace{-5pt}
-	\centering
-	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
-	\caption{End-effector trajectories for the footstep motion in Cartesian space.
-The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
-On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
-\label{fig:foot}}
-\end{figure}
-
 \subsubsection{Following Desired Joint Behaviors}
@@
 This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.
+
+\begin{figure}[t]
+\vspace{-5pt}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
+	\caption{End-effector trajectories for the footstep motion in Cartesian space.
+The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
+On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
+\label{fig:foot}}
+\end{figure}
 
@@
 
+\subsubsection{Systematic Assessment}
+To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system.
+The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
+The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
+To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
+uniform behaviors.
+The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
+Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
+Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.
+
+\begin{table}[t]
+	\centering
+	\resizebox{\linewidth}{!}{\begin{tabular}{l|c|c|c|c}
+		\hline \hline
+		& No. Dem. & No. K & $\phi(q)$ dim. & $e_{totol}$ [{rad}/{s}] \\ \hline
+		Obstacle avoidance (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ 
+		Obstacle avoidance (SEDS) &      7                &     3             &        -  &   -   \\		
+		\hline
+		Foot step  (JT-DS) &       3               &     1             &          1 & 0.01344   \\ 
+		Foot step (SEDS)&       3               &     1             &          - & -   \\ \hline \hline
+	\end{tabular}}
+	\caption{Parameters and Performance of Comparative Experiments \textcolor{red}{(REDO)} }
+	\label{table:1}
+\end{table}
+
 \begin{figure}[t]
@@
 		\centering	
-		\includegraphics[width=0.45\linewidth]{Pic/1.png}~~
-		\includegraphics[width=0.45\linewidth]{Pic/3.png}\\\vspace{0.1cm}
-		\includegraphics[width=0.45\linewidth]{Pic/4.png}~~
-		\includegraphics[width=0.45\linewidth]{Pic/2.png}
+		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/1.png}~~
+		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/3.png}\\\vspace{0.1cm}
+		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/4.png}~~
+		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/2.png}
 		\caption{Desired Joint Behavior: Foot step-like motion.}
@@
 	\end{subfigure}\\	
-	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{https://youtu.be/1dgKfmN1UgE}.}
+	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{[video link]}. \textcolor{red}{REDO} }
 	\label{fig:snapshot}
@@
 \end{figure}
-
-\subsubsection{Systematic Assessment}
-To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system.
-The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
-The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
-To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
-uniform behaviors.
-The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
-Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
-Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\begin{subfigure}[t]{\linewidth}

### 2017-10-29 paper/main_RA-L_v1_nadia.tex
Latest changes on Section 5 _nadia.tex, added a new dataset to learning eval (pour obstacle 2). Need to finish discussion on results and re-do comparison to SEDS+IK section.

@@
 Therefore, each synergy region is associated with a Gaussian component of the GMM, cumulatively describing all the synergy regions of the dynamical system.
-We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM and choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.
+We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM.
 
@@
 \begin{figure*}[!ht]
-  \begin{minipage}[b]{0.3\textwidth}
+  \begin{minipage}[b]{0.30\textwidth}
 	\centering	
@@
 	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
-		& K-PCA ($\sigma = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
+		& K-PCA ($\sigma_{opt} = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
 		\hline
@@
 	\multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
-		    \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
-			& K-PCA ($\sigma = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
+		    \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
+			& K-PCA ($\sigma_{opt} = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
 			\hline
@@
 	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($M=9,M^*T=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
-	& K-PCA ($\sigma = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($N=9,M=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
+	& K-PCA ($\sigma_{opt} = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
 				\hline
 	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle 1}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
-				& K-PCA ($\sigma = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
+				& K-PCA ($\sigma_{opt} = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
 				\hline
 
-	\multicolumn{1}{c}{\multirow{3}{*}{\textcolor{red}{Pouring - Obstacle 2}}} & None & 7 & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($M=X,M^*T=Y$)}} & PCA & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) \\	
-				& K-PCA ($\sigma = X$) & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y) & X $\pm$ (Y)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{\textcolor{red}{Pouring - Obstacle 2}}} & None & 7 & 4.3 $\pm$ (1.418) & \textbf{ 0.122 $\pm$ (0.014) } & 3.095 $\pm$ (3.406) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($N=7,M=1406$)}} & PCA & \textbf{ 2.9 $\pm$ 0.316 } & 3.7 $\pm$ (2.213) & 0.126 $\pm$ (0.02) & \cellcolor{blue!15} 0.825 $\pm$ (0.764) \\	
+	& K-PCA ($\sigma_{opt} = 7.25$) & 3 $\pm$ (0.421) & \textbf{3} & 0.091 $\pm$ (0.007) & \cellcolor{blue!10} 1.089 $\pm$ (1.218)  \\
 				\hline
@@
 	\multicolumn{1}{c}{\multirow{3}{*}{Foot Step}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
-	\multicolumn{1}{c}{\multirow{3}{*}{($M=8,M^*T=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
-	& K-PCA ($\sigma = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
+	\multicolumn{1}{c}{\multirow{3}{*}{($N=8,M=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
+	& K-PCA ($\sigma_{opt} = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
 				\hline
@@
 	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
-	& K-PCA ($\sigma = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
+    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
+	& K-PCA ($\sigma_{opt} = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
 \end{tabular}}
@@
     \end{minipage}
+    \vspace{-20pt}
 \end{figure*}
@@
 These behaviors include (as shown in Fig.
-\ref{fig:behaviors} ):
+\ref{fig:behaviors}):
 \begin{itemize}[leftmargin=*]
@@
 
-\noindent For each behavior, we collected $N$ demonstrations consisting of ($14$-D) time-series $D=\{\{ q_{t,n},\dot{q}_{t,n} \}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where each demonstration has duration $T_n$.
+\noindent For each behavior, we collected $N$ demonstrations consisting of ($14$-D) time-series $D=\{\{ q_{t,n},\dot{q}_{t,n} \}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where each demonstration has different durations $T_n$.
 We also record the final task-space configurations ${x^*}^{(n)}$, which we assume is the attractor in task-space.
@@
 The total number of demonstrations $N$ and samples per \textit{dataset} $M = \sum_{n=1}^{N}T_n$ are provided in Table \ref{tab:learning}.
-Given these datasets, we seek to:
+By performing our evaluation on all these datasets, we seek to:
 \begin{enumerate}[leftmargin=*]
-\item Verify that the learned \textit{behavior synergies} are capable of generating the demonstrated joint-space behaviors.  
-\item Find the dimensionality approach which yields the best trade-off between accuracy and model complexity; i.e. the least dimensions $\delta$ and synergies $K$.
+\item Verify that the learned \textit{behavior synergies} are mimicking the demonstrated joint-space behaviors and generalize them to new initial joint configurations.
+\item Analyze the role of dimensionality reduction in our proposed approach and further find the method which yields the best trade-off between accuracy and model complexity; i.e.
+the least number of dimensions $\delta$ to represent our activation function $\theta(q)$ and the least number of local behavior synergies $K$.
 \end{enumerate}
@@
 Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
-\textcolor{red}{\begin{enumerate}[leftmargin=*]
-\item Explain how optimal $\delta$ is computed for PCA. 
-\item Explain how optimal $\delta$ and $\sigma$ are found for K-PCA.
-\item Explain how optimal $K$ is computed for GMM.
-\end{enumerate}}
-
-%\begin{figure}[t]
-%	\centering	
-%		\begin{subfigure}[t]{\linewidth}
-%		\centering
-%			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
-%					\caption{Tennis Strokes (left) Forehand (right) Backhand.}
-%		\end{subfigure}
-%		\begin{subfigure}[t]{\linewidth}
-%\centering
-%	\includegraphics[width=0.45\linewidth]{new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{new-pics/pour-obst_crop.png}
-%						\caption{Pouring Motions (left) without obstacle (right) with obstacle 1.}
-%			\end{subfigure}
-%			\begin{subfigure}[t]{\linewidth}		
-%\centering
-%	\includegraphics[width=0.45\linewidth]{new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{new-pics/footstep.png}
-%	\caption{(left) Pouring Motions with obstacle (right) with obstacle 2 (right) Foot-step motion.}
-%				\end{subfigure}
-%	\caption{Demonstrated Joint-Space Behaviors with Task-Space Targets. \href{https://www.dropbox.com/sh/3vx512oldgbnkq5/AABP5iEqO0oYn7dlE-gvK1kPa?dl=0}{[videos]} \label{fig:behaviors}}
-%	\vspace{-10pt}
-%\end{figure}
-%
-%\begin{table*}[!t]
-%\scriptsize
-%\begin{center}
-%\begin{tabular}{cc|cc|cc}
-%    \hline
-%    \hline
-%    \multicolumn{1}{c}{\multirow{4}{*}{Behavior Dataset}} & \multicolumn{1}{c}{\multirow{4}{*}{Dim.
-Red.
-Approach}} & \multicolumn{4}{c}{\multirow{2}{*}{Optimal Parameters | Joint Velocity RMSE [$rad/s$]}}
-%	\\  
-%    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
-%            
-%    \multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
-%    \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
-%	& K-PCA ($\sigma = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
-%	\hline
-%	
-%	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
-%    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
-%		& K-PCA ($\sigma = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
-%		\hline
-%		
-%	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
-%   \multicolumn{1}{c}{\multirow{3}{*}{($M=9,M^*T=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
-%	& K-PCA ($\sigma = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
-%				\hline
-%	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
-%   \multicolumn{1}{c}{\multirow{3}{*}{($M=11,M^*T=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
-%				& K-PCA ($\sigma = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
-%				\hline	
-%	\multicolumn{1}{c}{\multirow{3}{*}{Foot Swing}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
-%	\multicolumn{1}{c}{\multirow{3}{*}{($M=8,M^*T=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
-%	& K-PCA ($\sigma = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
-%				\hline		
-%	
-%	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
-%    \multicolumn{1}{c}{\multirow{3}{*}{($M=10,M^*T=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
-%	& K-PCA ($\sigma = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
-%\end{tabular}
-%\caption{\scriptsize Performance Comparison of Learning Approach with different Dimensionality Reduction Schemes.
-We present the \texttt{mean(std)} for the optimal lower-dimensional $d$ and behavior synergies $K$ and RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
-%\end{center}
-%\vspace{-10pt}
-%\end{table*}
+\begin{enumerate}[leftmargin=*]
+\item \textit{$\delta$ for PCA:} We choose $\delta$ such that the projection is capable of explaining $95\%$ of the variance in the data.
+\item \textit{$\delta$ and $\sigma$ for K-PCA:} The interaction of these two parameters plays a major roll in the resulting projection obtained from K-PCA.
+Hence, we do a grid search on a log-spaced range of $\sigma_{range} = [\sigma_{min}:\sigma_{max}]$ values, where
+$\sigma_{min} = \frac{1}{\kappa\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ and $\sigma_{max} = \frac{2\kappa}{\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ for $\kappa = cte.$ This yields a feasible range for $\sigma$ that is guided by the pairwise Euclidean distances between all points in the dataset.
+Moreover, K-PCA is not limited to providing a $\delta \ll d$, in fact it can generate $\delta \le M$, this is a nuisance as in our datasets $M\approx1000$.
+To alleviate this we truncate $\sigma_{range}$ by compute the explained variance of the eigenvectors in feature space for different values of  $\sigma \in \sigma_{range}$.
+We then remove $\sigma$ values from $\sigma_{range}$ whose number of \textit{optimal} eigenvectors $\delta > d$ and resample it.
+By doing this procedure, we can ensure that for all values in truncated $\bar{\sigma}_{range}$ we will obtain $\delta \le d$.
+We then choose $\sigma_{opt}$ by running 10-fold cross-validation for all $\sigma \in \bar{\sigma}_{range}$.
+\item \textit{$K$ for GMM:} We choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.
+Typically, one chooses the optimal $K$ manually, by visually identifying the point at which the BIC curve produced from $K_{range}$ stops changing or plateaus.
+To automate this process, we devised an approach which selects the optimal $K$ as the one which yields the highest inflection point on the second order derivative of the BIC curve.
+\end{enumerate}
 
 Once these optimal hyper-parameters are estimated for each fold, we solve for the definite convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
-We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from the training/testing datasets, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing set trajectories.
+We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}
 
-\textcolor{red}{Discuss results: As can be seen in Table \ref{tab:learning}, ...}
+\textcolor{red}{Discuss results: As can be seen,  ... generalization capabilities, etc.}
 
-This learning pipeline was implemented in MATLAB.
-In order to solve the definite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used.
+\noindent\texttt{Implementation Details:} The learning pipeline is implemented in MATLAB.
+\textcolor{red}{Mention Dimensionalrit Toolbox here.} In order to solve the definite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used.
 In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis.
-The source code for learning (MATLAB) and execution (C++) of the Augmented Joint-Space Task-Oriented Dynamical System together with data generated in these experiments are available on-line: \\
-\footnotesize Learning:  \underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}\\
+Source code for the learning pipeline and cross-validation with simulation of multi-DOF robot arms can be found in:
+\footnotesize  \centerline{\underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}}
 \normalsize
-\textcolor{red}{In bla, we provide the necessary code to run this learning approach on our test datasets as well as on your own. We also provide code for the cross-validation evaluation.}
 
-\subsection{Comparison with Cartesian DS + IK Solver}
-We now evaluate the performance of JT-DS\footnote{With the optimal learning approach found in the previous section.} against a standard Cartesian motion generator + IK solver approach.
-In this evaluation we seek to compare the capabilities (or lack thereof) of both approaches to:
-\begin{enumerate}
-\item Follow desired behaviors in both joint and task spaces.
-\item Achieve fast computation/convergence time.
-\item Transit through singular configurations. 
-\end{enumerate}
-We continue our evaluation on a 7-DOF robot arm, the KUKA LWR 4+.
-The robot is controlled on the joint position level (linearly interpolating from joint velocities) at a rate of $ 500 $ Hz.
+For execution of the learnt JT-DS models on a real 7-DOF the KUKA LWR 4+ robot arm, we provide the following code in C++:
+\footnotesize \underline{\url{https://github.com/epfl-lasa/JT-DS-lib}}\\
+\normalsize
+The robot is controlled on the joint position level (linearly interpolating from joint velocities computed from JT-DS) at a rate of $ 500 $ Hz.
 The resultant joint angles are filtered by a critically damped filter to avoid high torques.
 
-\subsubsection{Following Desired Joint Behaviors}
-To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS \cite{khansari2011learning} algorithm using damped least square IK.
-In the first experiment, the robot was guided through a complex motion task, moving an object through an environment with an obstacle.
-%Human experts guided the robot through a series of demonstrations, ensuring that neither the robot's joints nor the held object intersected the obstacle. 
+\newpage
+\subsection{Comparison with Standard Approaches}
+In the previous subsection we have shown that our JT-DS model and learning algorithm are capable of following and generalizing the demonstrated joint-space behaviors.
+We now seek to elucidate the distinctive properties of our JT-DS model by comparing it against a standard DS-based Cartesian motion generator + IK solver approach for specific behaviors.
+
+\subsubsection{Following Desired Joint and Task-Based Behaviors}
+To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS \cite{khansari2011learning} algorithm using damped least square IK for behaviors (5) and (6).
+
 The results can be seen in Fig.
@@
 
-\begin{figure}[t]
-\vspace{-5pt}
-	\centering
-	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
-	\caption{End-effector trajectories for the footstep motion in Cartesian space.
-The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
-On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
-\label{fig:foot}}
-\end{figure}
-
-Further inspection reveals another advantage of JT-DS over stable Cartesian motion generators: not needing to preprocess demonstrations.
+%Further inspection reveals another advantage of JT-DS over stable Cartesian motion generators: not needing to preprocess demonstrations.
 In Cartesian-space dynamical systems, in order to guarantee stability, every demonstration must have the same target position.
@@
 The Cartesian-space system warps the demonstrations and as a result fails to track them, while JT-DS does not modify the demonstrations and thus succeeds.
-
-\begin{table}[t]
-	\vspace{0.1cm}
-	\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
-The convergence duration is the time required to move within $0.001$m of the target.
-The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
-	\label{table:2}
-	\centering
-	\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
-		\hline\hline
-		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
-		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
-		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
-	\end{tabular}}
-	\vspace{-10pt}
-\end{table}
 
@@
 
+\begin{figure}[t]
+\vspace{-5pt}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
+	\caption{End-effector trajectories for the footstep motion in Cartesian space.
+The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
+On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
+\label{fig:foot}}
+\end{figure}
+
 \begin{table}[t]
@@
 		& No. Dem. & No. K & $\phi(q)$ dim. & $e_{totol}$ [{rad}/{s}] \\ \hline
-		Obstacle avoidance (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ 
-		Obstacle avoidance (SEDS) &      7                &     3             &        -  &   -   \\		
+		Pour Obstacle 2 (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ 
+		Pour Obstacle 2 (SEDS) &      7                &     3             &        -  &   -   \\		
 		\hline
@@
 	\label{table:1}
+	\vspace{-10pt}
 \end{table}
@@
 	\label{fig:snapshot}
-	\vspace{-0.8cm}
+	\vspace{-10pt}
 \end{figure}
+
+\begin{table}[h]
+	\vspace{0.1cm}
+	\label{table:2}
+	\centering
+	\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
+		\hline\hline
+		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
+		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
+		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
+	\end{tabular}}
+		\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
+The convergence duration is the time required to move within $0.001$m of the target.
+The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
+\end{table}
 

### 2017-10-29 paper/main_RA-L_v1_nadia.tex

@@
 
-	\multicolumn{1}{c}{\multirow{3}{*}{Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(1) Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
     \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
@@
 
-	\multicolumn{1}{c}{\multirow{3}{*}{Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(2) Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
 		    \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
@@
 
-	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(3) Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
    \multicolumn{1}{c}{\multirow{3}{*}{($N=9,M=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
@@
 				\hline
-	\multicolumn{1}{c}{\multirow{3}{*}{Pouring - Obstacle 1}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(4) Pouring - Obstacle 1}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
    \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
-				& K-PCA ($\sigma_{opt} = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & 3.1 $\pm$ (0.316) & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
+				& K-PCA ($\sigma_{opt} = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & \textbf{ 3.1 $\pm$ (0.316) } & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
 				\hline
 
-	\multicolumn{1}{c}{\multirow{3}{*}{\textcolor{red}{Pouring - Obstacle 2}}} & None & 7 & 4.3 $\pm$ (1.418) & \textbf{ 0.122 $\pm$ (0.014) } & 3.095 $\pm$ (3.406) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($N=7,M=1406$)}} & PCA & \textbf{ 2.9 $\pm$ 0.316 } & 3.7 $\pm$ (2.213) & 0.126 $\pm$ (0.02) & \cellcolor{blue!15} 0.825 $\pm$ (0.764) \\	
-	& K-PCA ($\sigma_{opt} = 7.25$) & 3 $\pm$ (0.421) & \textbf{3} & 0.091 $\pm$ (0.007) & \cellcolor{blue!10} 1.089 $\pm$ (1.218)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(5) Pouring - Obstacle 2}} & None & 7 & 4.3 $\pm$ (1.418) & \textbf{ 0.122 $\pm$ (0.014) } & 3.095 $\pm$ (3.406) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($N=7,M=1406$)}} & PCA & \textbf{ 2.9 $\pm$ 0.316 } & 3.7 $\pm$ (2.213) & 0.126 $\pm$ (0.02) & \cellcolor{blue!10} 0.825 $\pm$ (0.764) \\	
+	& K-PCA ($\sigma_{opt} = 2.86$) & 3.9 $\pm$ (0.316) & \textbf{3} & 0.129 $\pm$ (0.013) & \cellcolor{blue!15} 0.738 $\pm$ (0.507)  \\
 				\hline
 
-	\multicolumn{1}{c}{\multirow{3}{*}{Foot Step}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(6) Foot Step}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
 	\multicolumn{1}{c}{\multirow{3}{*}{($N=8,M=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
@@
 
-	\multicolumn{1}{c}{\multirow{3}{*}{Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(7) Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
     \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
@@
 \subsection{JT-DS Learning Performance  Evaluation}
-To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 7 different tasks (see Table \ref{tab:learning}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space.
-These behaviors include (as shown in Fig.
-\ref{fig:behaviors}):
+To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 7 different tasks (see Table \ref{tab:learning} and Fig.
+\ref{fig:behaviors}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space:
 \begin{itemize}[leftmargin=*]
@@
 
-Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters, namely the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
-Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$ in the RBF kernel.
+Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
+Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$.
 Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
@@
 $\sigma_{min} = \frac{1}{\kappa\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ and $\sigma_{max} = \frac{2\kappa}{\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ for $\kappa = cte.$ This yields a feasible range for $\sigma$ that is guided by the pairwise Euclidean distances between all points in the dataset.
-Moreover, K-PCA is not limited to providing a $\delta \ll d$, in fact it can generate $\delta \le M$, this is a nuisance as in our datasets $M\approx1000$.
+Moreover, K-PCA is not limited to providing a $\delta \ll d$, in fact it can generate $\delta \le M$.
+This is a nuisance as in our datasets $M\approx1000$.
 To alleviate this we truncate $\sigma_{range}$ by compute the explained variance of the eigenvectors in feature space for different values of  $\sigma \in \sigma_{range}$.
@@
 
-\textcolor{red}{Discuss results: As can be seen,  ... generalization capabilities, etc.}
+As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
+This suggests that using DR to represent our activation functions $\theta(q)$ yields better generalization capabilities than the contrary.
+This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learnt without DR is an order of magnitude higher than with DR.
+Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the number of $K$ needed to represent the motion.
+For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
+By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
+Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
 
-\noindent\texttt{Implementation Details:} The learning pipeline is implemented in MATLAB.
-\textcolor{red}{Mention Dimensionalrit Toolbox here.} In order to solve the definite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used.
-In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis.
-Source code for the learning pipeline and cross-validation with simulation of multi-DOF robot arms can be found in:
+Both PCA and K-PCA yield comparable results, with K-PCA providing a slight improvement on some datasets.
+This suggests that perhaps a linear DR method might be sufficient for such tasks.
+However, if we seek to maximize accuracy a non-linear DR method should be employed.
+One of the drawbacks of K-PCA is its computational complexity for out-of-sample evaluations, which involves computing the kernel function between the new data-point and all the samples $M$.
+This is, however, not so taxing for our method as our datasets range in $M\approx1000$ and in previous work \cite{Mirrazavi:IJRR:2017} we have experimentally found that evaluating $<3000$ RBF kernel computations (on a 3.4-GHz i7 PC with 8GB RAM)) in a control-loop with 2ms rate is feasible.
+If one requires extremely fast computation; i.e.
+a control-loop rate of $<2$ms then PCA should be used as opposed to K-PCA.
+
+\begin{figure*}[!t]
+	\vspace{-0.05cm}
+	\begin{subfigure}[t]{0.42\linewidth}
+		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
+		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
+		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
+		\caption{Execution of Pour Obstacle 2 learned through JT-DS.}
+		\label{fig:snapshot:a}
+	\end{subfigure}
+	\begin{subfigure}[t]{0.567\linewidth}
+		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
+		\caption{Execution of Pour Obstacle 2 learned through SEDS.}
+		\label{fig:snapshot:b}
+	\end{subfigure}\\	
+	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{[video link]}.}
+	\label{fig:snapshot}
+	\vspace{-10pt}
+\end{figure*}
+
+\textit{Implementation Details:} The learning pipeline is implemented in MATLAB.
+We used the Matlab Toolbox for Dimensionality Reduction \cite{Maaten08dimensionalityreduction} for implementations of PCA, K-PCA and its out-of-sample extension.
+The YALMIP framework \cite{lofberg2004yalmip} was used to solve the definite convex optimization problem.
+Source code for the entire learning pipeline and cross-validation with simulation of multi-DOF robot arms can be found in:
 \footnotesize  \centerline{\underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}}
@@
 
-\newpage
-\subsection{Comparison with Standard Approaches}
-In the previous subsection we have shown that our JT-DS model and learning algorithm are capable of following and generalizing the demonstrated joint-space behaviors.
-We now seek to elucidate the distinctive properties of our JT-DS model by comparing it against a standard DS-based Cartesian motion generator + IK solver approach for specific behaviors.
+\subsection{JT-DS Execution Performance Evaluation}
+We now seek to elucidate the distinctive properties of our JT-DS model by comparing its performance to a DS-based Cartesian motion generator + IK solver approach for behaviors \textit{(5-7)}.
+For the Cartesian motion generator we use the SEDS approach \cite{khansari2011learning} which learns an asymptotically stable DS in Cartesian space from demonstrations.
+We then generate joint trajectories through a damped least square IK solver.
+From herein we refer to this approach as SEDS+IK.
 
 \subsubsection{Following Desired Joint and Task-Based Behaviors}
-To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS \cite{khansari2011learning} algorithm using damped least square IK for behaviors (5) and (6).
+To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS  algorithm using  for behaviors (5) and (6).
 
@@
 folding the elbow, lowering the shoulder), and therefore successfully avoided the obstacle while still converging to the desired Cartesian position.
-Meanwhile, SEDS only learned the demonstrated behavior in task-space (the position of the end-effector is signified by a black ball) meaning that the robot was not constrained in joint space.
+Meanwhile, SEDS only learned the demonstrated behavior in task-space meaning that the robot was not constrained in joint space.
 This ultimately led to one of its joints colliding with the obstacle.
@@
 
-\begin{figure}[t]
-\vspace{-5pt}
-	\centering
-	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
-	\caption{End-effector trajectories for the footstep motion in Cartesian space.
-The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
-On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
-\label{fig:foot}}
-\end{figure}
-
 \begin{table}[t]
@@
 \begin{figure}[t]
-	\vspace{-0.05cm}
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
-		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
-		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
-		\caption{Desired Joint Behavior: obstacle avoidance motion. The luggage is the obstacle.}
-		\label{fig:snapshot:a}
-	\end{subfigure}\\
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}
-		\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
-		\includegraphics[width=0.32\linewidth]{Pic/Luggage3_SEDS.png}
-		\caption{Desired Behavior: The Cartesian DS is used to generate the motion.}
-		\label{fig:snapshot:b}
-	\end{subfigure}\\
-	\begin{subfigure}[t]{\linewidth}
+\begin{minipage}[b]{\linewidth}
 \centering
-		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/1.png}~~
-		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/3.png}\\\vspace{0.1cm}
-		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/4.png}~~
-		\includegraphics[angle=180,origin=c,width=0.45\linewidth]{Pic/2.png}
-		\caption{Desired Joint Behavior: Foot step-like motion.}
+		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{new-pics/singularity_crop.png}
+		\caption{Execution of \textit{problematic} joint-space behaviors learned through JT-DS: (left) Foot step-like motion and (right) Singularity motions.}
 		\label{fig:snapshot:c}
-	\end{subfigure}\\	
-	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{[video link]}. \textcolor{red}{REDO} }
-	\label{fig:snapshot}
+\end{minipage}
+\begin{minipage}[b]{\linewidth}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
+	\caption{End-effector trajectories for the footstep motion in Cartesian space.
+The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
+On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
+\label{fig:foot}}
 	\vspace{-10pt}
+	\end{minipage}
+	\vspace{-20pt}
 \end{figure}
@@
 The algorithm never requires the pseudo-inverse of the Jacobian matrix, so the generated motion perfectly follows the demonstrations throughout the workspace boundary.
-
-\begin{figure}[t]
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=\linewidth]{Pic/cropped_Sing_1.pdf}
-		\caption{ }
-		\label{fig:Sin}
-	\end{subfigure}\\
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
-		\caption{ }
-		\label{fig:Det}
-	\end{subfigure}\\
-	\caption{The experiment generated $K=3$ Gaussian components.
-As the first joint is the only joint which was not fixed during the demonstration, the learned augmentation matrices had only one nonzero entry $A_k(1,1)\neq0~\forall k\in \{1, 2, 3\} $.
-In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space.
-The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).
-In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
-Due to sensory noise, the determinants are not exactly zero, but are very close.
-}
-	\vspace{-10pt}
-\end{figure}
 
@@
 
+\begin{figure}[t]
+	\begin{subfigure}[t]{\linewidth}
+		\includegraphics[width=\linewidth]{Pic/cropped_Sing_1.pdf}
+		\caption{ }
+		\label{fig:Sin}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{\linewidth}
+		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
+		\caption{ }
+		\label{fig:Det}
+	\end{subfigure}\\
+	\caption{The experiment generated $K=3$ Gaussian components.
+As the first joint is the only joint which was not fixed during the demonstration, the learned augmentation matrices had only one nonzero entry $A_k(1,1)\neq0~\forall k\in \{1, 2, 3\} $.
+In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space.
+The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).
+In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
+Due to sensory noise, the determinants are not exactly zero, but are very close.
+}
+	\vspace{-10pt}
+\end{figure}
+
 \footnotesize

### 2017-10-29 paper/main_RA-L_v1_nadia.tex
Latest change of the day, finished with Section 5.

@@
 \centering
-\includegraphics[scale=.23,trim={0 0 0 0cm},clip]{Without_Obstcle_white_new.png}
-\includegraphics[scale=.23,trim={0 0 0 0cm},clip]{With_Obstcle.png}
+\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{With_Obstcle.png}
 \caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl).
-The bottom example avoids a known obstacle in its workspace, while the top one does not.}
+The left example avoids a known obstacle in its workspace, while the right one does not.}
 \label{fig:robot_example}
+\vspace{-15pt}
 \end{figure}
@@
 \item \textit{(6) Footstep-like motion:} We demonstrate a footstep motion which begins with a straight leg, moving through a singularity, and finally bringing the knee up.
-\item \textit{(7) Singularity motions:} For this behavior we simulate joint trajectories in which the base rotates and all of the other joints are locked in singular configurations.
+\item \textit{(7) Singularity motions:} Movement was constrained to the boundary of the workspace by fixing $ q^i=0~\forall i\in\{3,\dots,7\}$, the second joint was fixed to $q^2\in\{10^\circ,20^\circ,\dots,100^\circ\}$, and only the first joint was moved by a human.
+This restricted the motion to a series of arcs of different radii along the robot's motion boundary.
 \end{itemize}
@@
 
-Both PCA and K-PCA yield comparable results, with K-PCA providing a slight improvement on some datasets.
-This suggests that perhaps a linear DR method might be sufficient for such tasks.
-However, if we seek to maximize accuracy a non-linear DR method should be employed.
-One of the drawbacks of K-PCA is its computational complexity for out-of-sample evaluations, which involves computing the kernel function between the new data-point and all the samples $M$.
-This is, however, not so taxing for our method as our datasets range in $M\approx1000$ and in previous work \cite{Mirrazavi:IJRR:2017} we have experimentally found that evaluating $<3000$ RBF kernel computations (on a 3.4-GHz i7 PC with 8GB RAM)) in a control-loop with 2ms rate is feasible.
-If one requires extremely fast computation; i.e.
-a control-loop rate of $<2$ms then PCA should be used as opposed to K-PCA.
-
-\begin{figure*}[!t]
+\begin{figure*}[ht]
 	\vspace{-0.05cm}
@@
 \end{figure*}
+
+Both PCA and K-PCA yield comparable results, with K-PCA providing a slight improvement on some datasets.
+This suggests that perhaps a linear DR method might be sufficient for such tasks.
+However, if we seek to maximize accuracy a non-linear DR method should be employed.
+One of the drawbacks of K-PCA is its computational complexity for out-of-sample evaluations, which involves computing the kernel function between the new data-point and all the samples $M$.
+This is, however, not so taxing for our method as our datasets range in $M\approx1000$ and in previous work \cite{Mirrazavi:IJRR:2017} we have experimentally found that evaluating $<3000$ RBF kernel computations (on a 3.4-GHz i7 PC with 8GB RAM)) in a control-loop with 2ms rate is feasible.
+If one requires extremely fast computation; i.e.
+a control-loop rate of $<2$ms then PCA should be used as opposed to K-PCA.
 
@@
 \subsubsection{Following Desired Joint and Task-Based Behaviors}
-To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS  algorithm using  for behaviors (5) and (6).
-
-The results can be seen in Fig.
-\ref{fig:snapshot:a} and Fig.
-\ref{fig:snapshot:b}.
-The JT-DS algorithm mimicked the joint-space behaviors of the demonstration (e.g.
-folding the elbow, lowering the shoulder), and therefore successfully avoided the obstacle while still converging to the desired Cartesian position.
-Meanwhile, SEDS only learned the demonstrated behavior in task-space meaning that the robot was not constrained in joint space.
+We compare tracking capabilities of our JT-DS method with those of SEDS+IK for behaviors \textit{(5)} and \textit{(6)}.
+In Fig.
+\ref{fig:snapshot:a} and \ref{fig:snapshot:b} we can qualitatively see the difference between these two approaches for behavior \textit{(5)}.
+The JT-DS algorithm mimicked the joint-space behavior of the demonstration (e.g.
+folding the elbow, raising the shoulder), successfully avoiding the obstacle while still converging to the desired Cartesian position.
+Meanwhile, SEDS+IK only learned the demonstrated behavior in task-space, it does not have the capabilities of constraining motion in joint-space.
 This ultimately led to one of its joints colliding with the obstacle.
 It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
-.
-JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig.
-\ref{fig:foot}).
+For the foot-step behavior \textit{(6)} JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig.
+\ref{fig:snapshot:c} and \ref{fig:foot}).
 This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.
-
-%Further inspection reveals another advantage of JT-DS over stable Cartesian motion generators: not needing to preprocess demonstrations.
-In Cartesian-space dynamical systems, in order to guarantee stability, every demonstration must have the same target position.
-As a result, all demonstrated trajectories are warped to end at a single position, introducing a non-negligible difference between the original demonstrations and those used to train the DS.
-JT-DS, on the other hand, does not require any demonstration warping to learn.
-Rather than shifting demonstrations, the algorithm learns behavior regions with respect to the base frame of the manipulator.
-This effect can be seen in the diagrams in Fig.
-\ref{fig:Sin}, which show a comparison between JT-DS and a Cartesian-space DS trained through SEDS \cite{khansari2011learning}.
-The Cartesian-space system warps the demonstrations and as a result fails to track them, while JT-DS does not modify the demonstrations and thus succeeds.
-
-\subsubsection{Systematic Assessment}
-To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system.
-The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
-The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
-To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
-uniform behaviors.
-The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
-Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
-Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.
-
-\begin{table}[t]
-	\centering
-	\resizebox{\linewidth}{!}{\begin{tabular}{l|c|c|c|c}
-		\hline \hline
-		& No. Dem. & No. K & $\phi(q)$ dim. & $e_{totol}$ [{rad}/{s}] \\ \hline
-		Pour Obstacle 2 (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ 
-		Pour Obstacle 2 (SEDS) &      7                &     3             &        -  &   -   \\		
-		\hline
-		Foot step  (JT-DS) &       3               &     1             &          1 & 0.01344   \\ 
-		Foot step (SEDS)&       3               &     1             &          - & -   \\ \hline \hline
-	\end{tabular}}
-	\caption{Parameters and Performance of Comparative Experiments \textcolor{red}{(REDO)} }
-	\label{table:1}
-	\vspace{-10pt}
-\end{table}
 
@@
 
-\begin{table}[h]
-	\vspace{0.1cm}
+\subsubsection{Systematic Assessment}
+To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a SEDS+IK, we simulated $400$ simple motions on each system.
+The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
+The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
+To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
+uniform behaviors.
+The results, summarized in Table \ref{tab:sys}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
+As the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
+Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than SEDS+IK.
+
+\begin{table}[t]
 	\label{table:2}
@@
 	\end{tabular}}
-		\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
+		\caption{Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
 The convergence duration is the time required to move within $0.001$m of the target.
-The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
+The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.\label{tab:sys}}
+		\vspace{-10pt}
 \end{table}
 
-\subsubsection{Transiting in singular configurations}
-One of the main advantages of the proposed dynamical system is its ability to generate accurate paths in classical singular configurations.
-The first scenario, shown in Fig.
-\ref{fig:Sin}, was designed to assess this capability.
-The $10$ training demonstrations were constructed as follows: movement was constrained to the boundary of the workspace by fixing $ q^i=0~\forall i\in\{3,\dots,7\}$, the second joint was fixed to $q^2\in\{10^\circ,20^\circ,\dots,100^\circ\}$, and only the first joint was moved by a human demonstrator back-driving it.
-This restricted the motion to a series of arcs of different radii along the robot's motion boundary.
+\subsubsection{Transiting through singular configurations}
+One of the main advantages of the proposed DS is its ability to generate accurate paths in classical singular configurations.
+To evaluate this we generated behavior \textit{(7)}.
+The $10$ training demonstrations were constructed as follows:
+
 In other words, the demonstrated motions were entirely within a classic kinematic singularity.
@@
 
-\section{Discussion and Future Work} 
+\section{Discussion} 
 \label{Sec:Dis}
@@
 
-%JTDS can be expanded to incorporate joint limits of the form $q^i_{max} > q^i > q^i_{min}$.
-Previous work proposed adding discontinuous limits to JT control \cite{sciavicco1988solution}, and we additionally propose a continuous modification as follows: \textcolor{red}{Remove this part?
-Should we not talk about continuous joint limiting?} Replace the combined synergy matrix $A(q)$ with a joint-limited joint-behavior matrix $\mathcal{A} = S(q)A(q)S^T(q)$, where $S(q)$ is a bounded positive semidefinite diagonal matrix such that
-%\begin{equation}
-%\label{eq:last_criteria}
-%\begin{cases}
-%\begin{split}
-%& s^{i}(q^i_{max}) = 0&&&\\
-%& s^{i}(q^i_{min})= 0&&&
-%\end{split}
-%\forall i\in 1,\dots,m
-%\end{cases}
-%\end{equation}
-%$\mathbb{A}$ is also positive semidefinite, so the system is still provably stable w.r.t.
-the attractor.
-The structure of $S$ enforces kinematic joint limits because as a joint nears one of its limits, its magnitude decreases such that it will never actually collide with the limit.
-One way of constructing $S$ is to use the following function:
-%\begin{equation}
-%s^{i}(q) =
-%1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{4}  ~\forall i\in 1,\dots,m\\
-%\end{equation}
-%The algorithm can similarly be expanded to accommodate joint velocity limits by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits.
-Specifically, given some set of velocity constraints for each joint $i$ s.t.
-$\dot{q}_i < \dot{q}_{i, max}$, we can design a new joint augmentation matrix $\mathcal{A}' = r \mathcal{A}$ where $r = \min \left ( 1, \max_i \left ( \frac{\dot{q}_{i, max}}{\dot{q}_i} \right ) \right )$.
-This velocity scalar would slow down the joint trajectory when the robot dynamics could not execute them, but would not change the position profile of the joint motion, preserving a major aspect of the joint behavior.
-
 A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists.
@@
 There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession.
-The smaller subtrajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits.
+The smaller sub-trajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits.
 Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.
@@
 However, it is also a drawback: without knowing the exact location of obstacles, the algorithm cannot be guaranteed to avoid collisions, especially away from demonstrations.
-
-%Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system.
-This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors.
-\textcolor{red}{Are we actually still doing this?}
 

### 2017-10-29 paper/main_RA-L_v1_nadia.tex
Some corrections

@@
 One of the main advantages of the proposed DS is its ability to generate accurate paths in classical singular configurations.
-To evaluate this we generated behavior \textit{(7)}.
-The $10$ training demonstrations were constructed as follows:
-
-In other words, the demonstrated motions were entirely within a classic kinematic singularity.
+To evaluate this we generated behavior \textit{(7)}, which were demonstrated motions are entirely within a classic kinematic singularity.
 Fig.
@@
 The algorithm never requires the pseudo-inverse of the Jacobian matrix, so the generated motion perfectly follows the demonstrations throughout the workspace boundary.
+In Fig.
+\ref{fig:snapshot:c} we show the learnt singular motion successfully  executed on the real robot.
 

### 2017-10-30 paper/main_RA-L_v1.tex
Rewrote part 4

@@
 \label{sec:intro}
-\textcolor{red}{We'll remove the subsection headings later..\\ Some comments:\\- Change PbD to LfD everywhere}
 \subsection{\textcolor{red}{Motivation for learning joint+task-space behaviors}}
-\textcolor{red}{*See comments in tex*}
 % % TO YO: OUTLINE FOR RE-WRITING THIS PART
@@
 \subsection{Prior Work on Combined Joint and Task Space Control}
-\textcolor{red}{*See comments in tex*}
 % TO YO:
@@
 The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{petrivc2013muscle}.
-In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors.
+In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the joint-space that are sufficient to accurately recreate the desired robotic behaviors.
 In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $\mathcal{A}(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
@@
 A central question becomes how to modulate the synergies in different regions to yield our desired behavior.
-First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that an embedding $\phi(\cdot)$ and embedding space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
-We would like to define the robot's policy in synergy space such that in different regions of the space, we will utilize different synergies.
+First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that an embedding $\phi(\cdot)$ and embedding space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{embedding space}.
+We would like to define the robot's policy in embedding space such that in different regions of the space, we will prioritize different synergies.
 We are thus left with three problems: finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(q)~\forall k\in\{1,\dots,K\}$), and finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
@@
 \begin{enumerate}
-\item We first construct our "synergy-space", which provides us a lower-dimensional manifold through which to control the robot, by projecting the demonstration data (i.e.
+\item We first construct our embedding, which provides us a lower-dimensional manifold through which to control the robot, by projecting the demonstration data (i.e.
 collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$.
@@
 
-%WE ASSUME THAT, BECAUSE OUR DEMONSTRATIONS ARE SIMILAR, POINTS IN SIMILAR REGIONS WILL HAVE SIMILAR MOTION POLICIES, SO FINDING As IS EQUIVALENT TO FINDING PRIMARY SYNERGY IN REGION, AND THEN INTERPOLATING THE SYNERGIES IN THE REGIONS IN-BETWEEN
+The search for a lower-dimensional embedding of the joint space stems from the desire to identify a simplified coordinate system in which each principal component corresponded to an important source of variation in the demonstrated trajectories, and which is thus suitable for learning the demonstrated behavior.
+For example, motor control studies have postulated that human arm motions like reaching or following straight/curved line trajectories, rather than utilizing the full joint space $|q|$, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
+This suggests that human arm motion in general tends to move on a plane, and thus can be represented in such a lower-dimensional space.
 
-The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint behaviors" (defined in Section \ref{sec:intro}), in which it would be appropriate to define synergies.
-It is important to understand that distances in joint space (defined as $\| \Delta q\|$) might not numerically reflect what a human intuitively perceives as the distance between joint configurations.
-Motor control studies have postulated that most human arm motions, be they reaching motions or following trajectories as straight/curved lines, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
-This suggests that human arm motion in general tends to move on a plane, and thus can be represented in such lower-dimensional space.
-
-To this end, any type of topology-preserving latent-space embedding or dimensionality reduction approach can theoretically be used for $\phi(q)$.
-In this work, we choose to construct the projection $\phi(q)$ of $q$ through a linear map $W:q \in \mathbb{R}^{d} \rightarrow \phi(q) \in \mathbb{R}^{\delta\leq d}$ given by $\phi(q) = Wq$, where the projection matrix $W \in \mathbb{R}^{\delta \times d}$ is computed through Principal Component Analysis (PCA).
-Although the projected dimensions extracted from PCA are linear combinations of the original variables, this has been proven to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}.
+In this work, we assume that configurations that are nearby in joint-space should exhibit similar behaviors, and thus it is natural that we choose a low-dimensional embedding that preserves the variance in our demonstrations.
+To this end, we construct an embedding $\phi(q)$ by training dimensionality reduction techniques on the demonstrated trajectories.
+The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^d$ into a lower-dimensional configuration $z \in \mathbb{R}^u$, where $u < d$.
+Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}.
+In previous work, PCA has shown to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}, and this work further explores the usefulness of joint-space dimensionality reduction.
 
@@
 \subsection{Discovering Local Behavior Synergies}
-Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
+The next step is identifying the regions of space in which to activate different synergies (though the synergies' parameters have not yet been specified).
+Specifically, given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
 Moreover, we would like for our scheduling parameters $\theta'_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta'_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
@@
 Scheduling parameters for LPV systems with such properties have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}.
+Intuitively, we search for a probabilistic model that "explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a local behavior, which the synergy will then approximate.
 In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the scheduling parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and covariance matrices that parametrize the $k$-th multivariate Gaussian distribution.
-Each such distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ of our LPV system.
+Each such distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ of our LPV system relating to the $k$th synergy.
 We will define $\theta'_k(\phi(q))$ as $p(k|\phi(q))$:
@@
 \subsection{Estimating the Synergy Matrices}
-Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $\mathcal{A}(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
+Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $\mathcal{A}(q)$ as a linear combination of local $A_k$ synergy matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
 \begin{equation}
@@
 Intuitively, each ``synergy region" in synergy space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
-Finding the appropriate synergy matrices $A_k$ for the set of local behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
+Finding the appropriate synergy matrices $A_k$ to accurately reproduce the observed behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
 To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:

### 2017-10-30 paper/main_RA-L_v1.tex
Changed a lot of mentions of A as PSD to positive definite, and rewrote discussions of stability to asymptotic stability

@@
 \begin{abstract}
-In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space.
-Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it Lyapunov-stably converges towards its task-space attractor.
+In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space.
+Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.
 To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
-Specifically, we use dimensionality reduction to find low-dimensional synergy regions, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
+Specifically, we use dimensionality reduction to find a low-dimensional embedding space for modulating joint synergies, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
 Our proposed approach is validated on a variety of motions for a 7-DOF KUKA LWR4+ robot arm.\end{abstract}
@@
 \begin{enumerate}[leftmargin=*]
-\item computes a \begin{bf}complex motion in joint space\end{bf} that provably and stably converges to a \begin{bf}fixed task-space target\end{bf}.
+\item computes a \begin{bf}complex motion in joint space\end{bf} that provably and asymptotically stably converges to a \begin{bf}fixed task-space target\end{bf}.
 \item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic ``singularities".
@@
 The most similar approach to our proposed DS is the Jacobian transpose (JT) control \cite{wolovich1984computational} method.
-JT control is an inverse kinematics method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably (in the sense of Lyapunov) over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations.
+JT control is an inverse kinematics method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations.
 It shares some of our approach's advantages: fast computation and provable task-space stability.
@@
 \begin{enumerate}
-\item The dynamical system must be Lyapunov stable\footnote{Unless otherwise specified, "stability" in this paper always refers to global Lyapunov stability.
-We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system.
-For example, any finite-size manipulator will always have regions beyond its reach that it cannot converge to.} with respect to a fixed task-space target $x^*$.
+\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified, "stability" in this paper always refers to asymptotic stability within the workspace of the robot, and assumes no joint limits.
+We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system.} with respect to a fixed task-space target $x^*$.
 This can be expressed by ensuring that the following Lyapunov function
@@
 is stable; i.e.
-$\dot{V}(q)\leq0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$.
+$\dot{V}(q)<0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$.
 Where  $Q=\{q|q^i_{min} < q^i < q^i_{max},~\forall i\in \{1,\dots,d\}\}$ and $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$.
@@
 the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
-The positive semi-definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
+The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
 \ref{fig:A_example} illustrates the effects of  $\mathcal{A}(q)$ on the generated motion.
@@
 &0 \prec P &\\
-&0 \preceq A_k & &~\\
+&0 \prec A_k & &~\\
 & 0 \leq \theta_k(\cdot) & &\\
@@
 
-It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  do not ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$, as $\exists q\in Q-Q^* $ such that both $\dot{V}(q)=0$ and $H(q) \neq x^*$.
+It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  only ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$ if the robot does not have kinematic joint limits.
+If we introduce kinematic joint limits, then it's possible we may find $\exists q\in Q-Q^* $ such that both $\dot{V}(q)=0$ and $H(q) \neq x^*$.
 This happens if the target position is kinematically unreachable; i.e.
 for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or the joint has reached a kinematic limit.
-Instead, the proof in Appendix \ref{appendix:stability} shows that the system converges in the sense of Lyapunov, meaning that its distance to the attractor is always monotonically decreasing.
+In this case, the proof in Appendix \ref{appendix:stability} shows instead that the system converges in the sense of Lyapunov, meaning that its distance to the attractor is always monotonically decreasing.
 
@@
 To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:
+% SINA/NADIA: can you rewrite this to use the "positive definite" optimization formulation that you're thinking of?
 \begin{equation}
@@
 \end{equation*}
-accomplishes criterion (I), where $\mathcal{A}(q)$ is positive semi-definite, and $P$ is positive definite.
+accomplishes criterion (I), where $\mathcal{A}(q)$ is positive definite, and $P$ is positive definite.
 \newtheorem{theorem}{Theorem}[section]
@@
 &= -(H(q) - x^*)^TPJ(q)\mathcal{A}(q)J^T(q)P(H(q) - x^*)\\
-&= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) \leq 0
+&= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\prec}\underbrace{A_k}_{0\prec}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) < 0
 \end{aligned}
 \end{equation}
-By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $\mathcal{A}(q)$, which is positive semi-definite.
-This means that the expression is guaranteed to be negative semi-definite.
+By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $\mathcal{A}(q)$, which is positive definite.
+This means that the expression is guaranteed to be negative definite.
 Therefore, JT-DS \eqref{eq:ds} is globally Lyapunov stable; i.e.

### 2017-10-30 paper/main_RA-L_v1.tex
A few more adjustments about stability language

@@
 When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target.
-Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably moving to the target.
+Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably converging to the target.
 The reason for this is that rather than ``determining" the velocity of the system, our $\mathcal{A}$ matrix only warps it.
@@
 \begin{theorem}[Proof of Lyapunov Stability]
-JT-DS \eqref{eq:ds} is stable in the sense of Lyapunov with respect to the Lyapunov candidate
+JT-DS \eqref{eq:ds} is asymptotically stable with respect to the Lyapunov candidate
 \[ V(q) = \frac{1}{2}(H(q) - x^*)^TP(H(q) - x^*) \]
@@
 This means that the expression is guaranteed to be negative definite.
-Therefore, JT-DS \eqref{eq:ds} is globally Lyapunov stable; i.e.
+Therefore, JT-DS \eqref{eq:ds} is globally asymptotically stable with respect to a task-space attractor; i.e.
 $ \| H(q)-x^*\| $ and $ \dot{q} $ are bounded.

### 2017-10-30 paper/main_RA-L_v1_merge.tex
Merged Yo and my re-writes.. need to get rid of 3 columns!

--- 2017-10-29 paper/main_RA-L_v1_nadia.tex
@@
 \label{sec:intro}
-\textcolor{red}{We'll remove the subsection headings later..\\ Some comments:\\- Change PbD to LfD everywhere}
-\subsection{\textcolor{red}{Motivation for learning joint+task-space behaviors}}
-\textcolor{red}{*See comments in tex*}
-% % TO YO: OUTLINE FOR RE-WRITING THIS PART
-%1. Start with motivation of LfD  
-%2. Mention that we can learn behaviors in either joint or task space 
-%3. briefly summarize the approaches in each without going into detail; i.e. probabilistic models, DS, etc 
-%4.
-Motivate why we want to learn behaviors in joint+task space: 1 is because of the implications of IK as we already discuss, 2 to mimic joint trajectories..
-here give your example of why we would want that and mention applications such as table tennis[1], grasping[2] and bi-manual coordination [3].
-%
-%[1] Y. Huang, D. B ̈uchler, O. Koc, B. Scholkopf and J. Peters, “Jointly
-%learning trajectory generation and hitting point prediction in robot table
-%tennis,” in Proc. IEEE International Conference on Humanoid Robots,
-%2016, pp. 650-655.
-%
-%[2] S. Calinon and A. Billard, “Statistical learning by imitation of competing constraints in joint space and task space,” Advanced Robotics,vol. 23, pp. 2059-2076, 2009.
-%
-%[3] J. Silverio, S. Calinon, L. Rozo and D. G. Caldwell, “Learning competing constraints and task priorities from demonstrations of bimanual skills,” arXiv:1707.06791, 2017.
-
 Robot motion planning in joint-space has long been a major field of study \cite{kelly2006control}.
@@
 \ref{fig:robot_example}.
+In this paper, we will explore the problem of learning a preferred joint-space behavior from previously demonstrated trajectories while still accomplishing a task-space goal, through the Learning from Demonstrations (LfD) paradigm \cite{billard2008robot,Argall:RAS:2009}.
 
-Learning robot motion generators that follow desired joint and/or task-space behaviors is, in fact, the central focus of Programming by Demonstration (PbD) \cite{billard2008robot} \cite{Argall:RAS:2009}.
-Historically, much of the original work in PbD focused on learning motions solely in joint-space by parameterizing motions using GMMs \cite{Calinon:HRI:2007}, HMMs \cite{Garrido:Neuro:2015}, GPs \cite{Shon:HUM:2005}, or other probabilistic models \cite{Schaal:IROS:2003,Schaal:AI:2002}.
-Concurrently, a significant body of work from the graphics community was targeted at learning joint-space motion ``styles" from human motion capture data, through a variety of latent-space and stylistic probabilistic motion models \cite{gielniak2010stylized,grochow2004style}.
-Yet, over time, research shifted to pure task-space learning \cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}, which has two main advantages over pure joint-space: \textit{generalization} and \textit{re-usability}.
-By learning motions in task-space, one can concisely encode the task-relevant features and generate new motions which meet these task-space criteria.
-Furthermore, a task-space trajectory is robot-independent, and thus can be reused across robots with differing morphologies.
+Many LfD approaches learn motions in either joint space \cite{Calinon:HRI:2007, Garrido:Neuro:2015, Schaal:IROS:2003,Schaal:AI:2002} or task space \cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}.
+These approaches generally learn probabilistic model of the demonstrated motions, and construct a dynamical system that chooses behaviors according to the learned model.
+Both joint and task space are independently useful: joint-space behaviors let us directly influence the behavior of the physical robot, while task-space behaviors let us control the most task-relevant component of the robot's behavior.
+In particular, a number of approaches have pioneered stable convergence to an attractor in task space \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015} as a desirable capability for a dynamical system.
 
-A major drawback of these initial task-space methods was that despite learning motions similar to the demonstrations, there was no guarantee when applying the learned motion to a new problem (e.g.
-a different initial/target position) that the end-effector would stably converge to its ultimate task-space target.
-This led to the introduction and subsequent popularity of task-space learning techniques such as SEDS \cite{khansari2011learning}, $\tau$-SEDS \cite{Neumann:RAS:2015}  and LMDS \cite{Kronander:RAS:2015} which encode task-space behaviors as asymptotically stable dynamical systems (DS) guaranteed to converge towards a desired task-space target.
+In many cases, one may want to specify a joint-space behavior while simultaneously fulfilling a task-space objective.
+For example, throwing a ball is a motion characterized both by a task-space behavior (the ball moves backwards and then forwards) and also more obviously by its joint-space behavior (the elbow folding, the shoulder moving back and then swinging forward, and then the elbow straightening and the wrist snapping down).
+Similarly, learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}.
+For a visual example of a task-space problem in which joint behavior is essential, see Fig.
+\ref{fig:robot_example}.
 
@@
 
-For all the previously mentioned task-space techniques, the robot is ultimately controlled by projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}.
+Furthermore, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations.
+The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}.
 When the main focus is on executing a complex task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
@@
 These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}.
+By defining a controller in joint space directly, one can avoid IK and all its associated drawbacks.
 
-Although DS-based task-space learning techniques ensure target convergence and stability, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution.
-Further, for certain tasks, rather than generating arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.
-For example, throwing a ball is a motion less characterized by task-space behaviors (the ball moving backwards and then forwards) and more characterized by the joint behaviors: the elbow folds, the shoulder moves back and then swings forward, followed by the elbow straightening and the wrist snapping down.
-For a visual example of the need for such combined task/joint-space behavior learning, see Fig.
-\ref{fig:robot_example}.
-This is precisely the shortcoming associated with LfD in task-space: it fails to learn the joint-space behavior of the system.
-The current formulation of the DS-based learning techniques introduced thus far (e.g.
-SEDS, $\tau$-SEDS LMDS, etc.) do not allow for integrating combined task/joint-space behavior learning.
-
-\subsection{\textcolor{red}{Approaches that tackle the joint+task problem}}
-\textcolor{red}{*See comments in tex*}
-% TO YO:
-% ADD THESE THREE REFERENCES HERE AND SUMMARIZE THEM
-%[1] Y. Huang, D. B ̈uchler, O. Koc, B. Scholkopf and J. Peters, “Jointly
-%learning trajectory generation and hitting point prediction in robot table
-%tennis,” in Proc. IEEE International Conference on Humanoid Robots,
-%2016, pp. 650-655.
-%
-%[2] S. Calinon and A. Billard, “Statistical learning by imitation of competing constraints in joint space and task space,” Advanced Robotics,vol. 23, pp. 2059-2076, 2009.
-%
-%[3] J. Silverio, S. Calinon, L. Rozo and D. G. Caldwell, “Learning competing constraints and task priorities from demonstrations of bimanual skills,” arXiv:1707.06791, 2017.
-
+Several approaches have tackled the problem of learning joint-space behaviors with task-space objectives.
 \cite{calinon2008probabilistic} notably attempted to address this problem by learning separate motion policies in task space and the null space of the Jacobian (which would not affect task-space position), and driving the robot with a weighted sum of the two.
-However, this approach no longer guarantees convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian.
-
+However, this approach does not seek convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian.
+\cite{calinon2010learning} and \cite{silverio2017learning} expand on this approach by projecting task-space constraints into joint space using IK, and then learning a joint-space policy that incorporates both task and joint space constraints, but this similarly relies on IK approximations and does not ensure convergence to an attractor.
+% FACT-CHECK: is this what Calinon 2009 and Silverio et al. 2017 are doing? They don't ensure convergence, right? TP-GMM never seems to address it ->CORRECT
 \cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework, but does not propose a dynamical system for generating motions.
-
-\subsection{Our proposed Approach}
-\textcolor{red}{Dynamical System, Why?\\Linear Parameter Varying, Why?\\Lower-Dimensional Manifold, Why?\\}
-In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
 \cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities.
@@
 This gives rise to spurious attractors or cycles, and thus must be carefully tuned in order to avoid them.
-In this work, we depart from any modulation or combination of DS and instead propose a novel dynamical system which:
+%Most recently, \cite{huang2016jointly} proposed an approach that learns a controller for playing table tennis by combining joint-space Dynamic Movement Primitives (DMP) \cite{schaal2006dynamic} to mimic a desired task-space racket swing..
+\textcolor{red}{ must re-read this...
+I think they are doing something very similar...
+can remove if problematic}
+% Do DMPs count as learning joint space behaviors? If yes, I'm having a hard time differentiating huang2016 from us -> Yes they do
 
+%\subsection{Our Proposed Approach}
+In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
+We specifically chose to focus on linear parameter-varying dynamical systems because they are fast to compute and thus allow a controller to quickly react in a changing environment, a capability that is further improved by our not needing to compute pseudo-inverses.
+Our dynamical system:
 \begin{enumerate}[leftmargin=*]
-\item computes a \begin{bf}complex motion in joint space\end{bf} that provably and stably converges to a \begin{bf}fixed task-space target\end{bf}.
+\item computes a \begin{bf}complex motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf} target.
 \item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic ``singularities".
@@
 \end{enumerate}
+In brief, our algorithm learns a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behaviors in joint space, and modulates the use of these synergies throughout joint-space using a learned LPV system.
+We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
 
-The most similar approach to our proposed DS is the Jacobian transpose (JT) control \cite{wolovich1984computational} method. JT control is an inverse kinematics
-method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably (in the sense of Lyapunov) over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations.
+The most similar approach to our proposed DS is the Jacobian transpose (JT) control \cite{wolovich1984computational} method.
+JT control is an inverse kinematics method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations.
 It shares some of our approach's advantages: fast computation and provable task-space stability.
-However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations.
+However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is, to the best of our knowledge the first work to employ a JT system to learn behaviors from demonstrations.
 %while obeying the robots kinematic constraints.
@@
 In addition, a convex optimization problem is  formalized to estimate these parameters.
-The effectiveness of the proposed method is shown through simulations and experiments on a 7-DOF robot-arm in Section \ref{Sec:Exp}.
-The paper is finalized with a discussion over our method and results in Section \ref{Sec:Dis}.
+In Section \ref{Sec:Exp} we provide a thorough validation of our proposed DS and learning approach.
+We finalize with a discussion in Section \ref{Sec:Dis}.
 
@@
 \label{sec:proposed_system}
-\textcolor{red}{Change semi-definite to definite everywhere\\Change overall $A(q)$ to $\mathcal{A}(q)$\\ Introduce 'behavior synergies' here\\}
-We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \ref{eq:Lyp} and \ref{eq:optim_error}.
+We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \eqref{eq:Lyp} and \eqref{eq:optim_error}.
 \begin{equation}
 \label{eq:ds}
-\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
 \end{equation}
-where $P\in \mathbb{R}^{d\times d}$, and $A(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
+where $P\in \mathbb{R}^{d\times d}$, and $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
 \begin{equation}
 \label{eq:A_def}
-A(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
+\mathcal{A}(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
 \end{equation}
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
-\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $A$ (emphasizing the hip, knee, and ankle respectively), A: $A(q)=diag(5,1,1) $, B: $A(q)=diag(1,5,1)$, and C: $A(q)=diag(1,1,5)$.
-On the left, the task-space traces of each motion.
-On the right, the time-scaled joint positions of each joint.
-Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
-\label{fig:A_example}
-\vspace{-15pt}
-\end{figure}
-
 Before proving that the proposed dynamical system satisfies all of the criteria, let us establish an intuitive understanding of the components of the system.
@@
 the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
-The positive semi-definite matrix $A(q)$ warps the resulting joint-space velocity; Fig.
-\ref{fig:A_example} illustrates the effects of  $A(q)$ on the generated motion.
+The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
+\ref{fig:A_example} illustrates the effects of  $\mathcal{A}(q)$ on the generated motion.
 Thus the controller can be thought of as a proportional controller in joint space.
-Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
+Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(as it augments the task error, and thus the direction of motion) and $\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (as it augments the outputted joint velocities).
 %When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle).
@@
 $A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
+\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
+On the left, the task-space traces of each motion.
+On the right, the time-scaled joint positions of each joint.
+Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
+\label{fig:A_example}
+\vspace{-15pt}
+\end{figure}s
 
@@
 
-%By learning a position-dependent joint augmentation matrix $A(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints.
-\eqref{eq:A_def} constructs our overall $A(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.
-In this paper the joint behavior matrix $A(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\
+%By learning a position-dependent joint augmentation matrix $\mathcal{A}(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints.
+\eqref{eq:A_def} constructs our overall $\mathcal{A}(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.
+In this paper the joint behavior matrix $\mathcal{A}(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\
 
-\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} 
+\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem in Synergy Space} 
 \label{Sec:Learning}
@@
 The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{petrivc2013muscle}.
-In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors.
-In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $A(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
+In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the joint-space that are sufficient to accurately recreate the desired robotic behaviors.
+In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $\mathcal{A}(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
 
-A central question becomes how to modulate the synergies in different regions to yield our desired behavior.
-First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that a latent mapping $\phi(\cdot)$ and latent space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
-We would like to define the robot's policy in synergy space such that in different regions of the space, we will utilize different synergies.
+A central question becomes, how to modulate the synergies in different regions to yield our desired behavior?
+First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that an embedding $\phi(\cdot)$ and embedding space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using a sub-manifold of the joint-space, called the \emph{embedding space}.
+We would like to define the robot's policy in embedding space such that in different regions of the space, we will prioritize different synergies.
 We are thus left with three problems: finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(q)~\forall k\in\{1,\dots,K\}$), and finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
 Our choice of parameters must also obey the constraints laid out in \eqref{eq:first_criteria_stability}.
-We propose the following 3-step procedure.
+We propose the following procedure:
 
 \begin{enumerate}
-\item We first construct our "synergy-space", which provides us a lower-dimensional manifold through which to control the robot, by projecting the demonstration data (i.e.
+\item We first construct our embedding, which provides us a lower-dimensional manifold through which to control the robot, by projecting the demonstration data (i.e.
 collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$.
-In each experiment, we used either Principal Component Analysis (PCA) \cite{jolliffe1986pca} or Kernel PCA (KPCA) \cite{scholkopf1997kernel}.
+In each experiment, we used either Principal Component Analysis (PCA) \cite{jolliffe1986pca} or Kernel PCA (KPCA) with RBF kernel \cite{scholkopf1997kernel}.
 
@@
 
-\subsection{Embedding Joint Configurations in Lower-Dimensional Space}
+\subsection{Embedding Joint Configurations in Low-Dim. Space}
+The search for a lower-dimensional embedding of the joint space stems from the desire to identify a simplified coordinate system in which each principal component corresponded to an important source of variation in the demonstrated trajectories, and which is thus suitable for learning the demonstrated behavior.
+For example, motor control studies have postulated that human arm motions like reaching or following straight/curved line trajectories, rather than utilizing the full joint space $|q|$, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
+This suggests that human arm motion in general tends to move on a plane, and thus can be represented in such a lower-dimensional space.
 
-%WE ASSUME THAT, BECAUSE OUR DEMONSTRATIONS ARE SIMILAR, POINTS IN SIMILAR REGIONS WILL HAVE SIMILAR MOTION POLICIES, SO FINDING As IS EQUIVALENT TO FINDING PRIMARY SYNERGY IN REGION, AND THEN INTERPOLATING THE SYNERGIES IN THE REGIONS IN-BETWEEN
-
-The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint behaviors" (defined in Section \ref{sec:intro}), in which it would be appropriate to define synergies.
-It is important to understand that distances in joint space (defined as $\| \Delta q\|$) might not numerically reflect what a human intuitively perceives as the distance between joint configurations.
-Motor control studies have postulated that most human arm motions, be they reaching motions or following trajectories as straight/curved lines, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
-This suggests that human arm motion in general tends to move on a plane, and thus can be represented in such lower-dimensional space.
-
-To this end, any type of topology-preserving latent-space embedding or dimensionality reduction approach can theoretically be used for $\phi(q)$.
-In this work, we choose to construct the projection $\phi(q)$ of $q$ through a linear map $W:q \in \mathbb{R}^{d} \rightarrow \phi(q) \in \mathbb{R}^{\delta\leq d}$ given by $\phi(q) = Wq$, where the projection matrix $W \in \mathbb{R}^{\delta \times d}$ is computed through Principal Component Analysis (PCA).
-Although the projected dimensions extracted from PCA are linear combinations of the original variables, this has been proven to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}.
+In this work, we assume that configurations that are nearby in joint-space should exhibit similar behaviors, and thus it is natural that we choose a low-dimensional embedding that preserves the variance in our demonstrations.
+To this end, we construct an embedding $\phi(q)$ by training dimensionality reduction techniques on the demonstrated trajectories.
+The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^d$ into a lower-dimensional configuration $z \in \mathbb{R}^\delta$, where $\delta < d$.
+Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}.
+%In previous work, PCA has shown to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}, and this work further explores the usefulness of joint-space dimensionality reduction.
 
@@
 \subsection{Estimating the Synergy Matrices}
-Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $A(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
+Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $\mathcal{A}(q)$ as a linear combination of local $A_k$ synergy matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
 \begin{equation}
 \label{eq:A}
-A(q) = \frac{\sum_{k=1}^K A_k \pi_k\mathcal{N}(\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}.
+\mathcal{A}(q) = \frac{\sum_{k=1}^K A_k \pi_k\mathcal{N}(\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}.
 \end{equation} 
@@
 Intuitively, each ``synergy region" in synergy space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
-Finding the appropriate synergy matrices $A_k$ for the set of local behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
+Finding the appropriate synergy matrices $A_k$ to accurately reproduce the observed behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
 To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:
+% SINA/NADIA: can you rewrite this to use the "positive definite" optimization formulation that you're thinking of?
 \begin{equation}
@@
 
-\newpage
 \section{Experimental Validation} \label{Sec:Exp}
@@
 	\end{minipage}
-	\vspace{-20pt}
+	\vspace{-10pt}
 \end{figure}
@@
 The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.\label{tab:sys}}
-		\vspace{-10pt}
+		\vspace{-15pt}
 \end{table}
@@
 When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target.
-Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably moving to the target.
-The reason for this is that rather than ``determining" the velocity of the system, our $A$ matrix only warps it.
+Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably converging to the target.
+The reason for this is that rather than ``determining" the velocity of the system, our $\mathcal{A}$ matrix only warps it.
 This means that adding multiple ``conflicting" synergies together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
@@
 
-A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists.
-As mentioned in Sec.
-\ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^TP(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits.
-If the only valid trajectory to reach the goal requires that we temporarily move away from the target (and increase the warped distance-to-goal), this controller will not find that trajectory.
-There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession.
-The smaller sub-trajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits.
-Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.
-
-One application that we have discussed for this algorithm is obstacle avoidance.
-In an environment with static obstacles, the system would, through demonstration, learn joint-space behaviors that avoid those obstacles.
-One can compare this form of joint-space obstacle avoidance with other methods from the literature \cite{sciavicco1988solution,petrivc2013smooth}.
-The main difference is that these previous methods require explicit knowledge of the obstacles' position, whereas our algorithm learns the obstacle positions only implicitly by learning motions that avoid those positions.
-This lack of explicit obstacle location can be an advantage: encoding obstacles' geometry can be expensive and cumbersome.
-However, it is also a drawback: without knowing the exact location of obstacles, the algorithm cannot be guaranteed to avoid collisions, especially away from demonstrations.
+Since we learn a lower-dimensional embedding space $Z := \{\phi(q) ~|~q \in Q\}$, which we suggest provides a better representation for the learned joint behaviors, a careful reader might wonder: why not define a motion policy $\pi_Z$ directly in $Z$-space and then map the learned policy back out into joint space using the inverse embedding\footnote{Such an inverse is trivial to extract in the case where $\phi$ is obtained through PCA.} $\phi^{-1}(z)$?
+The answer is that we would lose any guarantee of stability with respect to a target attractor $x_t^*$.
+By construction, $\phi$ maps from a joint-position vector of size $d$ to a low-dimensional vector of size $u<d$.
+Thus the inverse $\phi^{-1}$ must be of less than full rank, so the resulting $\mathcal{A}$ matrix (derived as in \eqref{eq:ds}) will also be of less than full rank, and thus no longer positive definite.
+This means that our controller no longer provably converges to the attractor (since our convergence proof no longer holds).
+Intuitively, this is because any policy defined in a lower-dimensional space than the actuation space will forfeit certain degrees of freedom, and thus may not be able to span the configuration space.
+Instead, we use the embedding space to modulate synergies defined in joint space, and which we can thus guarantee will be positive definite and lead us to convergence to the target.
 
@@
 
+A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists.
+As mentioned in Sec.
+\ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^TP(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits.
+If the only valid trajectory to reach the goal requires that we temporarily move away from the target (and increase the warped distance-to-goal), this controller will not find that trajectory.
+There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession.
+The smaller subtrajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits.
+Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.
+
+One application that we have discussed for this algorithm is obstacle avoidance.
+In an environment with static obstacles, the system would, through demonstration, learn joint-space behaviors that avoid those obstacles.
+One can compare this form of joint-space obstacle avoidance with other methods from the literature \cite{sciavicco1988solution,petrivc2013smooth}.
+The main difference is that these previous methods require explicit knowledge of the obstacles' position, whereas our algorithm learns the obstacle positions only implicitly by learning motions that avoid those positions.
+This lack of explicit obstacle location can be an advantage: encoding obstacles' geometry can be expensive and cumbersome.
+However, it is also a drawback: without knowing the exact location of obstacles, the algorithm cannot be guaranteed to avoid collisions, especially away from demonstrations.
+
 \footnotesize
@@
 \label{appendix:stability}
-\small
 We wish to prove Proposition \ref{prop:stability}, that is, that JT-DS \eqref{eq:ds} (reproduced below)
 \begin{equation*}
-\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
 \end{equation*}
-accomplishes criterion (I), where $A$ is positive semi-definite, and $P$ is positive definite.
+accomplishes criterion (I), where $\mathcal{A}(q)$ is positive definite, and $P$ is positive definite.
 \newtheorem{theorem}{Theorem}[section]
 \begin{theorem}[Proof of Lyapunov Stability]
-JT-DS \eqref{eq:ds} is stable in the sense of Lyapunov with respect to the Lyapunov candidate
+JT-DS \eqref{eq:ds} is asymptotically stable with respect to the Lyapunov candidate
 \[ V(q) = \frac{1}{2}(H(q) - x^*)^TP(H(q) - x^*) \]
@@
 \frac{dV(q)}{dt} &= (H(q) - x^*)^TPJ(q)\dot{q}\\
-&= -(H(q) - x^*)^TPJ(q)A(q)J^T(q)P(H(q) - x^*)\\
-&= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) \leq 0
+&= -(H(q) - x^*)^TPJ(q)\mathcal{A}(q)J^T(q)P(H(q) - x^*)\\
+&= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\prec}\underbrace{A_k}_{0\prec}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) < 0
 \end{aligned}
 \end{equation}
-By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $A$, which is positive semi-definite.
-This means that the expression is guaranteed to be negative semi-definite.
-Therefore, JT-DS \eqref{eq:ds} is globally Lyapunov stable; i.e.
+By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $\mathcal{A}(q)$, which is positive definite.
+This means that the expression is guaranteed to be negative definite.
+Therefore, JT-DS \eqref{eq:ds} is globally asymptotically stable with respect to a task-space attractor; i.e.
 $ \| H(q)-x^*\| $ and $ \dot{q} $ are bounded.

### 2017-10-30 paper/main_RA-L_v1_merge.tex
.

@@
 M.
-Salehyan, N.
+Salehian, N.
 Figueroa, and A.
@@
 \begin{abstract}
-In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space.
-Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.
-To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
+In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while stably converging towards a its task-space attractor.
+% Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.
+ To encode complex joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
 Specifically, we use dimensionality reduction to find a low-dimensional embedding space for modulating joint synergies, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
-Our proposed approach is validated on a variety of motions for a 7-DOF KUKA LWR4+ robot arm.\end{abstract}
+Our proposed approach is empirically validated on a variety of motions.
+%for a 7-DOF KUKA LWR4+ robot arm.
+ \end{abstract}
 
@@
 Robot motion planning in joint-space has long been a major field of study \cite{kelly2006control}.
-For manipulation problems with an objective defined in task space (i.e.
-target or desired trajectory), we can often find a myriad of joint-space trajectories to achieve the same task-space goal.
+For manipulation problems with an objective defined in the task space, %(i.e.
+target or desired trajectory)
+ we can often find a myriad of joint-space trajectories to achieve the same task-space goal.
 In many cases, however, certain joint-space trajectories are favored over others; for example, when we expect the robot to follow a desired joint-space behavior or ``style", as illustrated in Fig.
@@
 
-Many LfD approaches learn motions in either joint space \cite{Calinon:HRI:2007, Garrido:Neuro:2015, Schaal:IROS:2003,Schaal:AI:2002} or task space \cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}.
-These approaches generally learn probabilistic model of the demonstrated motions, and construct a dynamical system that chooses behaviors according to the learned model.
-Both joint and task space are independently useful: joint-space behaviors let us directly influence the behavior of the physical robot, while task-space behaviors let us control the most task-relevant component of the robot's behavior.
-In particular, a number of approaches have pioneered stable convergence to an attractor in task space \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015} as a desirable capability for a dynamical system.
+Many LfD approaches learn motions in either joint space %\cite{Calinon:HRI:2007, Garrido:Neuro:2015, Schaal:IROS:2003,Schaal:AI:2002}
+\cite{Garrido:Neuro:2015,Schaal:AI:2002} or task space %\cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}
+\cite{Pastor:ICRA:2009,Calinon:ISR:2015}.
+These approaches generally learn probabilistic model of the demonstrated motions.
+%, and construct a dynamical system that chooses behaviors according to the learned model.
+Both %joint and task 
+spaces are independently useful:  task-space behaviors let us control the most task-relevant component of the robot's behavior.
+In particular, a number of approaches have pioneered stable convergence to an attractor in task space % \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015}
+ \cite{khansari2011learning, Neumann:RAS:2015}as a desirable capability for a dynamical system.
 
-In many cases, one may want to specify a joint-space behavior while simultaneously fulfilling a task-space objective.
-For example, throwing a ball is a motion characterized both by a task-space behavior (the ball moves backwards and then forwards) and also more obviously by its joint-space behavior (the elbow folding, the shoulder moving back and then swinging forward, and then the elbow straightening and the wrist snapping down).
-Similarly, learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}.
-For a visual example of a task-space problem in which joint behavior is essential, see Fig.
-\ref{fig:robot_example}.
+	In many cases, one may want to specify a joint-space behavior to directly define the behavior of the physical robot while simultaneously fulfilling a task-space objective.
+For example, %throwing a ball is a motion characterized both by a task-space behavior (the ball moves backwards and then forwards) and also more obviously by its joint-space behavior (the elbow folding, the shoulder moving back and then swinging forward, and then the elbow straightening and the wrist snapping down).
+Similarly,
+learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}.
+For a visual example of a task-space problem in which joint behavior is essential, see Fig.\ref{fig:robot_example}.
 
@@
 \label{fig:robot_example}
-\vspace{-15pt}
+\vspace{-20pt}
 \end{figure}
@@
 Furthermore, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations.
-The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse based Inverse-Kinematic (IK) approximations and variants thereof \cite{kelly2006control}.
+The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}.
 When the main focus is on executing a complex task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
@@
 % FACT-CHECK: is this what Calinon 2009 and Silverio et al. 2017 are doing? They don't ensure convergence, right? TP-GMM never seems to address it ->CORRECT
-\cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework, but does not propose a dynamical system for generating motions.
+\cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework.
+% but does not propose a dynamical system for generating motions.
 \cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities.
@@
 In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
-We specifically chose to focus on linear parameter-varying dynamical systems because they are fast to compute and thus allow a controller to quickly react in a changing environment, a capability that is further improved by our not needing to compute pseudo-inverses.
-Our dynamical system:
+To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behaviors in joint space, and modulates the use of these synergies throughout joint-space using a learned LPV system.
+We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
+%We specifically chose to focus on linear parameter-varying dynamical systems because they are fast to compute and thus allow a controller to quickly react in a changing environment, a capability that is further improved by our not needing to compute pseudo-inverses.
+Hence, our dynamical system:
 \begin{enumerate}[leftmargin=*]
 \item computes a \begin{bf}complex motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf} target.
-\item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic ``singularities".
+%\item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic ``singularities".
 \item is formulated such that complex \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
 \end{enumerate}
-In brief, our algorithm learns a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behaviors in joint space, and modulates the use of these synergies throughout joint-space using a learned LPV system.
-We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
 
-The most similar approach to our proposed DS is the Jacobian transpose (JT) control \cite{wolovich1984computational} method.
-JT control is an inverse kinematics method that yields a dynamical system in joint space $\dot{q} = f(q)$ which converges stably over time to a desired end-effector position $x^*$, without the need for pseudo-inverse computations.
+The most similar approach to our proposed DS is the Jacobian transpose (JT) control method \cite{wolovich1984computational} .
+The JT control is an IK method that yields a dynamical system in joint space %$\dot{q} = f(q)$
+which converges stably over time to a desired end-effector position, %$x^*$
+ without the need for pseudo-inverse computations.
 It shares some of our approach's advantages: fast computation and provable task-space stability.
@@
 \begin{equation}
-V(q) = (H(q) - x^*)^TP(H(q) - x^*) 
+V(q) = (H(q) - x^*)^T(H(q) - x^*) 
 \label{eq:Lyp}
@@
 is stable; i.e.
-$\dot{V}(q)<0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$.
+$\dot{V}(q)\leq 0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$.
 Where  $Q=\{q|q^i_{min} < q^i < q^i_{max},~\forall i\in \{1,\dots,d\}\}$ and $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$.
@@
 %Lastly, we wish to capture joint-space behaviors in our dynamical system.
-We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}.
+
+%If instead of \eqref{eq:optim_error}, the error feature was the joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity.
+Alternatively, if the error feature was the joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior).
+By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
+
+The error metric \eqref{eq:optim_error} is advantageous in a way that it mimics the magnitude and direction of the demonstrated motions.
+Hence, as suggested by \cite{gielniak2010stylized}, it is visually most similar to the human definition of "joint motion style".
+
+% ways. First, it implicitly suggests that behaviors are defined by expert motions of the behavior
+
+%We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}.
 First, the metric implicitly suggests that behaviors are defined by expert motions of the behavior, and that fulfilling a behavior means moving similarly to the experts' motion.
@@
 \label{eq:ds}
-\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)(H(q) - x^*)
 \end{equation}
-where $P\in \mathbb{R}^{d\times d}$, and $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
+where  $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the LPV system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
 \begin{equation}
@@
 \end{equation}
-where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
-system $q $ or external signals $d(t)$, i.e.
-$\theta_k (t, q(t), d(t))$.
-In this paper, we only consider it as a function of the states of the system.
-It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$.} modulating each local synergy through time as well as space.
+where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{%The scheduling parameters can be a function of time $t$, the states of the
+%system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. 
+It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$, where $d(t)$ is an external signal.} modulating each local synergy through time as well as space.
 Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more complex multi-behavior motion.
@@
 One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
-the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
+the task target and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
 The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
@@
 Thus the controller can be thought of as a proportional controller in joint space.
-Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(as it augments the task error, and thus the direction of motion) and $\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (as it augments the outputted joint velocities).
+Lastly, we refer to %the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(as it augments the task error, and thus the direction of motion) and
+$\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} as it augments the outputted joint velocities.
 %When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle).
@@
 \vspace{-15pt}
-\end{figure}s
+\end{figure}
 
@@
 \label{prop:stability}
-The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if \eqref{eq:ds} meets the following constraints.
+The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ \eqref{eq:ds} meets the following constraints.
 
-\begin{subequations}
-\vspace{-5pt}
 \begin{equation}
@@
 \begin{split}
-&0 \prec P &\\
-&0 \prec A_k & &~\\
-& 0 \leq \theta_k(\cdot) & &\\
-%    & 0 \preceq and &\forall q \in Q\\
+ & 0 \prec A_k, & ~~~~ 0 \leq \theta_k(\cdot) \\
 \end{split}
-\forall k \in \{1,\dots,K\} 
 \end{cases}
 \end{equation}    
-\end{subequations}\\
 \end{prop1}
-\vspace{-10pt}
-\textbf{Proof}: See Appendix \ref{appendix:stability}. $\blacksquare$\\
 
-It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  only ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$ if the robot does not have kinematic joint limits.
+\textbf{Proof}: See Appendix \ref{appendix:stability}. $\blacksquare$
+
+%It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  only ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$ if the robot does not have kinematic joint limits.
 If we introduce kinematic joint limits, then it's possible we may find $\exists q\in Q-Q^* $ such that both $\dot{V}(q)=0$ and $H(q) \neq x^*$.
@@
 
-The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{petrivc2013muscle}.
+The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{7954741}.
 In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the joint-space that are sufficient to accurately recreate the desired robotic behaviors.
@@
 & \text{subject to}&\\
-&  0 \preceq A_k, \; \forall k \in \{1, \dots, K\}. &
+&  0 \prec A_k, \; \forall k \in \{1, \dots, K\}. &
 \end{aligned}
@@
 \end{equation}
-where $f(q_{t,n})$ is calculated by combining our dynamical system formulation \eqref{eq:ds} with \eqref{eq:A}, and $x^*_{n}$ is defined as the endpoint of the $ n^{\text{th}} $ demonstrated trajectory.
+where $f(q_{t,n})$ is calculated by %combining our dynamical system formulation
+ \eqref{eq:ds} with \eqref{eq:A}, and $x^*_{n}$ is defined as the endpoint of the $ n^{\text{th}} $ demonstrated trajectory.
 
@@
 Schemes.
-We present the \texttt{mean(std)} for the optimal $\delta$, $K$ and joint-velocity RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
+We present the \texttt{mean(std)} for the optimal $\delta$, $K$ and joint-velocity RMSE on training and testing set, found for every learning scheme over 10 runs.
+$M = \sum_{n=1}^{N}T_n$.\label{tab:learning}}
     \end{minipage}
@@
 
-\noindent For each behavior, we collected $N$ demonstrations consisting of ($14$-D) time-series $D=\{\{ q_{t,n},\dot{q}_{t,n} \}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where each demonstration has different durations $T_n$.
+%\noindent For each behavior, we collected $N$ demonstrations consisting of ($14$-D) time-series $D=\{\{ q_{t,n},\dot{q}_{t,n} \}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where each demonstration has different durations $T_n$.
 We also record the final task-space configurations ${x^*}^{(n)}$, which we assume is the attractor in task-space.
@@
 By performing our evaluation on all these datasets, we seek to:
+
+\noindent   The recordings are collected at a rate of 500 Hz.
+However, our learning approach does not require such a dense representation of the trajectories, hence we down-sample them to 25 Hz.
+By performing our evaluation on all these datasets, we seek to:
+
 \begin{enumerate}[leftmargin=*]
@@
 
-Once these optimal hyper-parameters are estimated for each fold, we solve for the definite convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
+Once these optimal hyper-parameters are estimated for each fold, we solve for the convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
 We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}
@@
 This ultimately led to one of its joints colliding with the obstacle.
-It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
+It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as expected), but did ultimately converge to its target position.
 For the foot-step behavior \textit{(6)} JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig.
 \ref{fig:snapshot:c} and \ref{fig:foot}).
-This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.
+This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities.
+%, overcoming the Achilles heel of equivalent Cartesian-based motion generators.
 
@@
 	\end{minipage}
-	\vspace{-10pt}
+	\vspace{-20pt}
 \end{figure}
@@
 One of the most important points when validating a learning from demonstration method is to evaluate the system's behavior away from demonstrations.
-When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target.
+When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\frac{1}{K}\sum\limits_{k=1}^{K}A_k$ is used to move the robot, which is still guaranteed to move towards the target.
 Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably converging to the target.
@@
 
-One of the major advantages of the proposed dynamical system is that it avoids the undesirable effects of traditional Inverse Kinematic (IK) solvers.
+%One of the major advantages of the proposed dynamical system is that it avoids the undesirable effects of traditional Inverse Kinematic (IK) solvers.
 Fast dynamical motions require not only fast and accurate motion planning, but also precise IK mapping, which is not practically feasible.
@@
 \begin{figure}[t]
-	\begin{subfigure}[t]{\linewidth}
+%	\begin{subfigure}[t]{\linewidth}
 		\includegraphics[width=\linewidth]{Pic/cropped_Sing_1.pdf}
-		\caption{ }
-		\label{fig:Sin}
-	\end{subfigure}\\
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
-		\caption{ }
-		\label{fig:Det}
-	\end{subfigure}\\
+%		\caption{ }
+%	\end{subfigure}\\
+%	\begin{subfigure}[t]{\linewidth} %
+%		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
+%		\caption{ }
+%		\label{fig:Det}
+%	\end{subfigure}\\
 	\caption{The experiment generated $K=3$ Gaussian components.
@@
 In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space.
-The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).
-In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
+The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).} %In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
 Due to sensory noise, the determinants are not exactly zero, but are very close.
 }
-	\vspace{-10pt}
+			\label{fig:Sin}
+	\vspace{-20pt}
 \end{figure}
 
-A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists.
+%A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists.
 As mentioned in Sec.
-\ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^TP(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits.
+\ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^T(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits.
 If the only valid trajectory to reach the goal requires that we temporarily move away from the target (and increase the warped distance-to-goal), this controller will not find that trajectory.
@@
 
-One application that we have discussed for this algorithm is obstacle avoidance.
+Finally, as the proposed DS might get stuck at its kinematic joint limits, we are currently working on improving the performance of the proposed DS, by learning its parameters wrt.
+the kinematic constraints.
+In this way, we can ensure that the performance of the DS is kinematically feasible for the robot to  follow.
+
+%One application that we have discussed for this algorithm is obstacle avoidance.
 In an environment with static obstacles, the system would, through demonstration, learn joint-space behaviors that avoid those obstacles.
@@
 
-\footnotesize
+{\footnotesize
 \appendices
@@
 \label{appendix:stability}
-We wish to prove Proposition \ref{prop:stability}, that is, that JT-DS \eqref{eq:ds} (reproduced below)
-\begin{equation*}
-\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
-\end{equation*}
-accomplishes criterion (I), where $\mathcal{A}(q)$ is positive definite, and $P$ is positive definite.
-\newtheorem{theorem}{Theorem}[section]
-\begin{theorem}[Proof of Lyapunov Stability]
+%We wish to prove Proposition \ref{prop:stability}, that is, that JT-DS \eqref{eq:ds} (reproduced below)
+%\begin{equation*}
+%\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
+%\end{equation*}
+%accomplishes criterion (I), where $\mathcal{A}(q)$ is positive definite, and $P$ is positive definite.
+%\newtheorem{theorem}{Theorem}[section]
+%\begin{theorem}[Proof of Lyapunov Stability]
 JT-DS \eqref{eq:ds} is asymptotically stable with respect to the Lyapunov candidate
-\[ V(q) = \frac{1}{2}(H(q) - x^*)^TP(H(q) - x^*) \]
-That is, 
+\[ V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*) \]
+That is, $ 0 \prec V(q),~ \forall q \neq q^*$ and $ V(q^*) = 0 $. Where $q^*$ is any joint configuration such that $H(q^*) = x^*$.  The derivative of $ V $ wrt. time is as follows:
+
 \begin{equation}
 \begin{aligned}
-& 0 \prec V(q)    &
-V(q^*) = 0 & \quad \forall q \neq q^*\\
-%    \[\dot{V}(q) = \frac{dV}{dq}\frac{dq}{dt}(q) \preceq 0 \quad \forall q \in Q\]
+\frac{dV(q)}{dt} &= (H(q) - x^*)^TJ(q)\dot{q}\\
+&= -(H(q) - x^*)^TJ(q)\mathcal{A}(q)J^T(q)(H(q) - x^*)\\
+&= -(H(q) - x^*)^TJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\prec}\underbrace{A_k}_{0\prec}J^T(q)(H(q) - x^*)\leq 0
 \end{aligned}
 \end{equation}
-where $q^*$ is any joint configuration such that $H(q^*) = x^*$.
-\end{theorem}
-The first two statements are trivially true because $P$ is positive definite and the rest of $V$ is a square that is only $0$ when $H(q) = x^*$. To prove the last statement, we find $\frac{dV}{dt}$:
-\begin{equation}
-\begin{aligned}
-\frac{dV(q)}{dt} &= (H(q) - x^*)^TPJ(q)\dot{q}\\
-&= -(H(q) - x^*)^TPJ(q)\mathcal{A}(q)J^T(q)P(H(q) - x^*)\\
-&= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\prec}\underbrace{A_k}_{0\prec}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) < 0
-\end{aligned}
-\end{equation}
-By observation, each and every term in the final expression is multiplied by its transpose (creating a square) except for $\mathcal{A}(q)$, which is positive definite.
-This means that the expression is guaranteed to be negative definite.
-Therefore, JT-DS \eqref{eq:ds} is globally asymptotically stable with respect to a task-space attractor; i.e.
-$ \| H(q)-x^*\| $ and $ \dot{q} $ are bounded.
+
+ Therefore, JT-DS \eqref{eq:ds} is stable and as $ q=q^* $ is the largest invariance in $ \mathbb{R}^m $, it is  globally asymptotically stable with respect to a task-space attractor; i.e.
+$\lim\limits_{t\rightarrow \infty} \| H(q)-x^*\|=0 $.}
 

### 2017-10-30 paper/main_RA-L_v1_merge.tex
final merge!

@@
 %
-\author{Yonadav~Shavit*,  Seyed~Sina~Mirrazavi~Salehian*, Nadia~Figueroa*, Aude~Billard
+\author{Yonadav~Shavit*, Nadia~Figueroa*, Seyed~Sina~Mirrazavi~Salehian*, Aude~Billard
 % <-this % stops a space
@@
 MA, 02139 USA. E-mail: yonadav.shavit@gmail.com.}% <-this % stops a space
-\thanks{S.
+\thanks{N.
+Figueroa, S.
 S.
 M.
-Salehian, N.
-Figueroa, and A.
+Salehian, and A.
 Billard are with the Swiss Federal Institute of Technology (EPFL), 1015 Lausanne, Switzerland.
-\\E-mail: \{sina.mirrazavi,nadia.figueroafernandez,aude.billard\}@epfl.ch.}
+\\E-mail: \{nadia.figueroafernandez,sina.mirrazavi,aude.billard\}@epfl.ch.}
 }% <-this %
@@
 These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}.
-By defining a controller in joint space directly, one can avoid IK and all its associated drawbacks.
+By defining a controller in joint space, one can avoid IK and all its associated drawbacks.
 
@@
 % FACT-CHECK: is this what Calinon 2009 and Silverio et al. 2017 are doing? They don't ensure convergence, right? TP-GMM never seems to address it ->CORRECT
-\cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework.
+\cite{lee2014unifying} learns from demonstration in both joint and task space within the ``thin-plate spline" trajectory warping framework.
 % but does not propose a dynamical system for generating motions.
@@
 \begin{enumerate}[leftmargin=*]
-\item computes a \begin{bf}complex motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf} target.
-%\item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic ``singularities".
+\item computes a \begin{bf}complex motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
 \item is formulated such that complex \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
+\item can cleanly moves through kinematic \textbf{singularities}.
 \end{enumerate}
@@
 \begin{enumerate}
-\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified, "stability" in this paper always refers to asymptotic stability within the workspace of the robot, and assumes no joint limits.
+\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified,``stability" in this paper always refers to asymptotic stability within the workspace of the robot, and assumes no joint limits.
 We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system.} with respect to a fixed task-space target $x^*$.
@@
 The error metric \eqref{eq:optim_error} is advantageous in a way that it mimics the magnitude and direction of the demonstrated motions.
-Hence, as suggested by \cite{gielniak2010stylized}, it is visually most similar to the human definition of "joint motion style".
+Hence, as suggested by \cite{gielniak2010stylized}, it is visually most similar to the human definition of ``joint motion style".
 
@@
 The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ \eqref{eq:ds} meets the following constraints. 
-
 \begin{equation}
@@
 \end{prop1}
-
+\vspace{5pt}
 \textbf{Proof}: See Appendix \ref{appendix:stability}. $\blacksquare$
-
 %It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  only ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$ if the robot does not have kinematic joint limits.
@@
 We would like to define the robot's policy in embedding space such that in different regions of the space, we will prioritize different synergies.
-We are thus left with three problems: finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(q)~\forall k\in\{1,\dots,K\}$), and finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
+We are thus left with three problems: (I) finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), (II) finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(q)~\forall k\in\{1,\dots,K\}$), and (III) finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
 Our choice of parameters must also obey the constraints laid out in \eqref{eq:first_criteria_stability}.
-We propose the following procedure:
+We thus propose the following 3-step learning procedure:
 
@@
 
-\item We then jointly estimate the optimal number $K$ of synergy "regions" and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$ seen in the demonstrations.
+\item We then jointly estimate the optimal number $K$ of synergy ``regions" and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$ seen in the demonstrations.
 
@@
 The search for a lower-dimensional embedding of the joint space stems from the desire to identify a simplified coordinate system in which each principal component corresponded to an important source of variation in the demonstrated trajectories, and which is thus suitable for learning the demonstrated behavior.
-For example, motor control studies have postulated that human arm motions like reaching or following straight/curved line trajectories, rather than utilizing the full joint space $|q|$, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
+Motor control studies have postulated that human arm motions like reaching or following straight/curved line trajectories, rather than utilizing the full joint space $|q|$, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
 This suggests that human arm motion in general tends to move on a plane, and thus can be represented in such a lower-dimensional space.
@@
 \subsection{Discovering Local Behavior Synergies}
-The next step is identifying the regions of space in which to activate different synergies (though the synergies' parameters have not yet been specified).
-Specifically, given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
+The next step is identifying the regions of space in which to activate different synergies.
+Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
 Moreover, we would like for our scheduling parameters $\theta'_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta'_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
@@
 Scheduling parameters for LPV systems with such properties have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}.
-Intuitively, we search for a probabilistic model that "explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a local behavior, which the synergy will then approximate.
+Intuitively, we search for a probabilistic model that ``explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a local behavior, which the synergy will then approximate.
 In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the scheduling parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and covariance matrices that parametrize the $k$-th multivariate Gaussian distribution.
-Each such distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ of our LPV system relating to the $k$th synergy.
-We will define $\theta'_k(\phi(q))$ as $p(k|\phi(q))$:
+Each distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ to the $k$th synergy of \eqref{eq:A_def}.
+We define $\theta'_k(\phi(q))$ as $p(k|\phi(q))$:
 \begin{equation}
@@
 \begin{figure*}[!ht]
-  \begin{minipage}[b]{0.30\textwidth}
+  \begin{minipage}[b]{0.31\textwidth}
 	\centering	
@@
   \hfill
-  \begin{minipage}[b]{0.7\textwidth}
+  \begin{minipage}[b]{0.69\textwidth}
     \scriptsize
@@
 
-As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
-This suggests that using DR to represent our activation functions $\theta(q)$ yields better generalization capabilities than the contrary.
-This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learnt without DR is an order of magnitude higher than with DR.
-Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the number of $K$ needed to represent the motion.
-For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
-By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
-Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
-
-\begin{figure*}[ht]
+\begin{figure*}[th]
 	\vspace{-0.05cm}
@@
 \end{figure*}
+
+As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
+This suggests that using DR to represent our activation functions $\theta(q)$ yields better generalization capabilities than the contrary.
+This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learnt without DR is an order of magnitude higher than with DR.
+Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the number of $K$ needed to represent the motion.
+For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
+By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
+Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
 

### 2017-11-15 paper/Pic/Sources/Schematic.m
the schematic is added

@@ -1,0 +1,75 @@
+clc
+close all
+clear
+startup_rvc
+
+
+l1=1.5;
+l2=1.5;
+
+q1_limit=[-3.14:0.01:3.14];
+q2_limit=[-3.14:0.01:3.14];
+x_limit=[0:0.01:4];
+y_limit=[0:0.01:4];
+X_d=[2;2];
+
+
+
+
+for i=1:size(y_limit,2)
+    H=[x_limit;repmat(y_limit(i),1,size(y_limit,2))];
+    V(i,:)=sum((H-repmat(X_d,1,size(H,2))).*(H-repmat(X_d,1,size(H,2))),1);
+end
+
+Xlim=repmat(x_limit,size(y_limit,2),1);
+Ylim=repmat(y_limit',1,size(x_limit,2));
+
+subplot1 = subplot(1,2,1);
+hold(subplot1,'on');
+
+L(1) = Link([0 0 l1 0],'standard')
+L(2)= Link([0 0 l2 0],'standard')
+L(3)= Link([0 0 0 0],'standard')
+r = SerialLink(L,'name','two link')
+r.plotopt = {'noshadow','nojaxes', 'nowrist','noname','linkcolor',0.7*[1,1,1], 'ortho','noshading','notiles','jointcolor',0.4*[1,1,1]};
+r.plot([pi/3,-pi/4,0])
+hold on
+contourf(Xlim,Ylim,V)
+colormap(cool)
+axis equal
+xlabel('$X~ [m]$','Interpreter','latex');
+
+% Create ylabel
+ylabel('$Y~ [m]$','Interpreter','latex');
+box(subplot1,'on');
+axis(subplot1,'tight');
+set(subplot1,'BoxStyle','full','FontSize',24,'Layer','top',...
+    'TickLabelInterpreter','latex');
+xlim(subplot1,[x_limit(1) x_limit(end)]);
+ylim(subplot1,[y_limit(1) y_limit(end)]);
+
+q1lim=repmat(q1_limit,size(q1_limit,2),1);
+q2lim=repmat(q2_limit',1,size(q1_limit,2));
+
+
+for i=1:size(q1_limit,2)
+    for j=1:size(q2_limit,2)
+    q1lim(i,j)= q1_limit(i);
+    q2lim(i,j)= q2_limit(j);
[…]

### 2017-11-16 paper/Pic/Sources/Schematic.m
Final schematic

@@ -7,13 +7,17 @@ startup_rvc
 l1=1.5;
 l2=1.5;
 
-q1_limit=[-3.14:0.01:3.14];
-q2_limit=[-3.14:0.01:3.14];
+q1_limit=[-3.14:0.005:3.14];
+q2_limit=[-3.14:0.005:3.14];
 x_limit=[0:0.01:4];
 y_limit=[0:0.01:4];
 X_d=[2;2];
 
-
+% Plot the SVM Contours
+level = 100; n = ceil(level/2);
+cmap1 = [linspace(1, 1, n); linspace(0, 1, n); linspace(0, 1, n)]';
+cmap2 = [linspace(1, 0, n); linspace(1, 0, n); linspace(1, 1, n)]';
+cmap = [cmap1; cmap2(2:end, :)];
 
 
 for i=1:size(y_limit,2)
@@ -34,8 +38,12 @@ r = SerialLink(L,'name','two link')
 r.plotopt = {'noshadow','nojaxes', 'nowrist','noname','linkcolor',0.7*[1,1,1], 'ortho','noshading','notiles','jointcolor',0.4*[1,1,1]};
 r.plot([pi/3,-pi/4,0])
 hold on
-contourf(Xlim,Ylim,V)
-colormap(cool)
+contourf(Xlim,Ylim,V,15,'LineWidth', 0.001)
+% contourf(Xlim,Ylim,V,10,'LineStyle', 'none'); 
+colormap(vivid(cmap, [.5, .5]));
+% colormap(vivid('n',256,[.5, 1]));
+% colormap(cool);
+colorbar
 axis equal
 xlabel('$X~ [m]$','Interpreter','latex');
 
@@ -43,7 +51,7 @@ xlabel('$X~ [m]$','Interpreter','latex');
 ylabel('$Y~ [m]$','Interpreter','latex');
 box(subplot1,'on');
 axis(subplot1,'tight');
-set(subplot1,'BoxStyle','full','FontSize',24,'Layer','top',...
+set(subplot1,'BoxStyle','full','FontSize',18,'Layer','top',...
     'TickLabelInterpreter','latex');
 xlim(subplot1,[x_limit(1) x_limit(end)]);
 ylim(subplot1,[y_limit(1) y_limit(end)]);
@@ -62,14 +70,17 @@ for i=1:size(q1_limit,2)
 end
 subplot1 = subplot(1,2,2);
 hold(subplot1,'on');
-contourf(q1lim,q2lim,V_q)
[…]

### 2017-11-17 paper/main_RA-L_v1_final.tex
Final draft, cleaned up folder, need to get rid of some text on the 9th page.

--- 2017-10-30 paper/main_RA-L_v1_merge.tex
@@
 % Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.
- To encode complex joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
+ To encode joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
 Specifically, we use dimensionality reduction to find a low-dimensional embedding space for modulating joint synergies, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
@@
 The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}.
-When the main focus is on executing a complex task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
+When the main focus is on executing a non-linear task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
 However, for other applications, such an approach yields significant problems \cite{buss2004introduction}.
@@
 In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
-To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behaviors in joint space, and modulates the use of these synergies throughout joint-space using a learned LPV system.
+To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system.
 We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
-%We specifically chose to focus on linear parameter-varying dynamical systems because they are fast to compute and thus allow a controller to quickly react in a changing environment, a capability that is further improved by our not needing to compute pseudo-inverses.
 Hence, our dynamical system:
 \begin{enumerate}[leftmargin=*]
-\item computes a \begin{bf}complex motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
-\item is formulated such that complex \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
-\item can cleanly moves through kinematic \textbf{singularities}.
+\item computes a \begin{bf}motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
+\item is formulated such that \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
+\item can transit through kinematic \textbf{singularities}.
 \end{enumerate}
@@
 The most similar approach to our proposed DS is the Jacobian transpose (JT) control method \cite{wolovich1984computational} .
-The JT control is an IK method that yields a dynamical system in joint space %$\dot{q} = f(q)$
-which converges stably over time to a desired end-effector position, %$x^*$
- without the need for pseudo-inverse computations.
+The JT control is an IK method that yields a dynamical system in joint space
+which converges stably over time to a desired end-effector position, without the need for pseudo-inverse computations.
 It shares some of our approach's advantages: fast computation and provable task-space stability.
 However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is, to the best of our knowledge the first work to employ a JT system to learn behaviors from demonstrations.
-%while obeying the robots kinematic constraints.
+Furthermore, by formulating our LPV system on a latent embedding (via dimensionality reduction schemes), we are capable of discovering meaningful local behavior synergy regions, while being robust to outliers, noise and redundancies that might arise from raw demonstrations.
+This leads to a compelling improvement in generalization of the demonstrated behavior, as opposed to learning the LPV system solely in joint space.
 
@@
 It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$, where $d(t)$ is an external signal.} modulating each local synergy through time as well as space.
-Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more complex multi-behavior motion.
+Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more non-linear multi-behavior motion.
 
@@
 One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
-the task target and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
+the task target and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig.
+\ref{fig:error_schematic}.
 The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
@@
 $A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
-\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
-On the left, the task-space traces of each motion.
-On the right, the time-scaled joint positions of each joint.
-Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
-\label{fig:A_example}
-\vspace{-15pt}
-\end{figure}
 
@@
 
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{./Pic/Sources/Schematic_final.pdf}
+	\caption{Illustration of our proposed Lyapunov function and error projection.
+On the left we show the task-space potential function that we are trying to descend on for a 2-DOF robot with target $x^*=[2,2]^T$.
+On the right we show the corresponding Lyapunuv function in joint space \eqref{eq:Lyp}.
+The colors on each plot correspond to each other, one can see how the error, and consequently the \textit{attractive} regions, in task-space are warped in joint-space.}
+	\label{fig:error_schematic}
+	\vspace{-15pt}
+\end{figure}
+
 \section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem in Synergy Space} 
@@
 We thus propose the following 3-step learning procedure:
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
+	\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
+On the left, the task-space traces of each motion.
+On the right, the time-scaled joint positions of each joint.
+Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
+	\label{fig:A_example}
+	\vspace{-15pt}
+\end{figure}
 
@@
 The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^d$ into a lower-dimensional configuration $z \in \mathbb{R}^\delta$, where $\delta < d$.
-Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}.
+%Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}. 
 %In previous work, PCA has shown to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}, and this work further explores the usefulness of joint-space dimensionality reduction.
@@
 \cite{nadaraya1964regress,watson1964regress}} with a Gaussian pdf as its kernel function.
-Hence \eqref{eq:A} can be considered a type of kernel estimator, with the key distinction that the weighting functions $\theta_k'(\phi(q))$ are not determined by individual points (as in the original Nadaraya-Watson kernel estimator) but by the components of a GMM, similar to the weighting functions derived in Gaussian Mixture Regression (GMR) \cite{sung2004gmr}.
+Hence \eqref{eq:A} can be considered a type of kernel estimator, with the key distinction that the weighting functions $\theta_k'(\phi(q))$ are not determined by individual points (as in the original Nadaraya-Watson kernel estimator) but by the components of a GMM, similar to the weighting functions derived in Gaussian Mixture Regression (GMR).
 
@@
 To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:
-% SINA/NADIA: can you rewrite this to use the "positive definite" optimization formulation that you're thinking of?
 \begin{equation}
@@
 			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
-					\caption{Tennis Strokes (1-2).}
+					\caption{Forward/Backward Reaching (1-2).\label{fig:forback}}
 		\end{subfigure}
@@
 
-	\multicolumn{1}{c}{\multirow{3}{*}{(1) Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(1) Forward Reaching}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
     \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
@@
 
-	\multicolumn{1}{c}{\multirow{3}{*}{(2) Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(2) Backward Reaching}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
 		    \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
@@
 \begin{itemize}[leftmargin=*]
-\item \textit{(1-2) Tennis strokes:} For these particular behaviors, we guide the robot to perform a forehand and backhand stroke.  
-\item \textit{(3-5) Pouring motions:} We guide the robot to pour chips in a bowl with three different joint-space behaviors.
+\item \textit{(1-2) Forward/Backward reaching:} We guide the robot to  forward and backward reaching motions towards a specific target in task-space with joint-space trajectories mimicking forehand/backhand strokes as shown in Fig.
+\ref{fig:forback}.
+\item \textit{(3-5) Pouring motions:} We guide the robot to emulate pouring motions with three different environmental constraints, leading to different joint-space trajectories.
 The first is without an obstacle in the workspace.
@@
 \end{itemize}
-
-%\noindent For each behavior, we collected $N$ demonstrations consisting of ($14$-D) time-series $D=\{\{ q_{t,n},\dot{q}_{t,n} \}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where each demonstration has different durations $T_n$.
-We also record the final task-space configurations ${x^*}^{(n)}$, which we assume is the attractor in task-space.
-The recordings are collected at a rate of 500 Hz.
-However, our learning approach does not require such a dense representation of the trajectories, hence we down-sample them to 25 Hz.
-The total number of demonstrations $N$ and samples per \textit{dataset} $M = \sum_{n=1}^{N}T_n$ are provided in Table \ref{tab:learning}.
-By performing our evaluation on all these datasets, we seek to:
 
@@
 
-Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
-Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$.
-Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
-\begin{enumerate}[leftmargin=*]
-\item \textit{$\delta$ for PCA:} We choose $\delta$ such that the projection is capable of explaining $95\%$ of the variance in the data.
-\item \textit{$\delta$ and $\sigma$ for K-PCA:} The interaction of these two parameters plays a major roll in the resulting projection obtained from K-PCA.
-Hence, we do a grid search on a log-spaced range of $\sigma_{range} = [\sigma_{min}:\sigma_{max}]$ values, where
-$\sigma_{min} = \frac{1}{\kappa\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ and $\sigma_{max} = \frac{2\kappa}{\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ for $\kappa = cte.$ This yields a feasible range for $\sigma$ that is guided by the pairwise Euclidean distances between all points in the dataset.
-Moreover, K-PCA is not limited to providing a $\delta \ll d$, in fact it can generate $\delta \le M$.
-This is a nuisance as in our datasets $M\approx1000$.
-To alleviate this we truncate $\sigma_{range}$ by compute the explained variance of the eigenvectors in feature space for different values of  $\sigma \in \sigma_{range}$.
-We then remove $\sigma$ values from $\sigma_{range}$ whose number of \textit{optimal} eigenvectors $\delta > d$ and resample it.
-By doing this procedure, we can ensure that for all values in truncated $\bar{\sigma}_{range}$ we will obtain $\delta \le d$.
-We then choose $\sigma_{opt}$ by running 10-fold cross-validation for all $\sigma \in \bar{\sigma}_{range}$.
-\item \textit{$K$ for GMM:} We choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.
-Typically, one chooses the optimal $K$ manually, by visually identifying the point at which the BIC curve produced from $K_{range}$ stops changing or plateaus.
-To automate this process, we devised an approach which selects the optimal $K$ as the one which yields the highest inflection point on the second order derivative of the BIC curve.
-\end{enumerate}
-
-Once these optimal hyper-parameters are estimated for each fold, we solve for the convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
-We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}
-
-\begin{figure*}[th]
+\begin{figure*}[t]
 	\vspace{-0.05cm}
@@
 
+Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
+Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$.
+Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
+\begin{enumerate}[leftmargin=*]
+\item \textit{$\delta$ for PCA:} We choose $\delta$ such that the projection is capable of explaining $95\%$ of the variance in the data.
+\item \textit{$\delta$ and $\sigma$ for K-PCA:} The interaction of these two parameters plays a major roll in the resulting projection obtained from K-PCA.
+Hence, we do a grid search on a log-spaced range of $\sigma_{range} = [\sigma_{min}:\sigma_{max}]$ values, where
+$\sigma_{min} = \frac{1}{\kappa\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ and $\sigma_{max} = \frac{2\kappa}{\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ for $\kappa = cte.$ This yields a feasible range for $\sigma$ that is guided by the pairwise Euclidean distances between all points in the dataset.
+Moreover, K-PCA is not limited to providing a $\delta \ll d$, in fact it can generate $\delta \le M$.
+This is a nuisance as in our datasets $M\approx1000$.
+To alleviate this we truncate $\sigma_{range}$ by computing the explained variance of the eigenvectors in feature space for different values of  $\sigma \in \sigma_{range}$.
+We then remove $\sigma$ values from $\sigma_{range}$ whose number of \textit{optimal} eigenvectors $\delta > d$ and resample it.
+By doing this procedure, we can ensure that for all values in truncated $\bar{\sigma}_{range}$ we will obtain $\delta \le d$.
+We then choose $\sigma_{opt}$ by running 10-fold cross-validation for all $\sigma \in \bar{\sigma}_{range}$.
+\item \textit{$K$ for GMM:} We choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.
+Typically, one chooses the optimal $K$ manually, by visually identifying the point at which the BIC curve produced from $K_{range}$ stops changing or plateaus.
+To automate this process, we devised an approach which selects the optimal $K$ as the one which yields the highest inflection point on the second order derivative of the BIC curve.
+\end{enumerate}
+
+Once these optimal hyper-parameters are estimated for each fold, we solve for the convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
+We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}
+
 As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
@@
 Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
+\begin{figure}[t]
+	\begin{minipage}[b]{\linewidth}
+		\centering
+		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{new-pics/singularity_crop.png}
+		\caption{Execution of \textit{problematic} joint-space behaviors learned through JT-DS: (left) Foot step-like motion and (right) Singularity motions.}
+		\label{fig:snapshot:c}
+	\end{minipage}
+	\begin{minipage}[b]{\linewidth}
+		\centering
+		\includegraphics[width=0.7\linewidth]{Pic/Foot.pdf}
+		\caption{End-effector trajectories for the footstep motion in Cartesian space.
+The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
+On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
+\label{fig:foot}}
+		\vspace{-10pt}
+	\end{minipage}
+	\vspace{-20pt}
+\end{figure}
 
@@
 
-\begin{figure}[t]
-\begin{minipage}[b]{\linewidth}
-\centering
-		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{new-pics/singularity_crop.png}
-		\caption{Execution of \textit{problematic} joint-space behaviors learned through JT-DS: (left) Foot step-like motion and (right) Singularity motions.}
-		\label{fig:snapshot:c}
-\end{minipage}
-\begin{minipage}[b]{\linewidth}
-	\centering
-	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
-	\caption{End-effector trajectories for the footstep motion in Cartesian space.
-The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
-On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
-\label{fig:foot}}
-	\vspace{-10pt}
-	\end{minipage}
-	\vspace{-20pt}
-\end{figure}
-
 \subsubsection{Systematic Assessment}
@@
 One of the main advantages of the proposed DS is its ability to generate accurate paths in classical singular configurations.
-To evaluate this we generated behavior \textit{(7)}, which were demonstrated motions are entirely within a classic kinematic singularity.
+To evaluate this we generated behavior \textit{(7)}; i.e.
+joint-trajectories that transit entirely within a classic kinematic singularity.
 Fig.
@@
 The reason for this is that rather than ``determining" the velocity of the system, our $\mathcal{A}$ matrix only warps it.
-This means that adding multiple ``conflicting" synergies together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
-
-%One of the major advantages of the proposed dynamical system is that it avoids the undesirable effects of traditional Inverse Kinematic (IK) solvers.
-Fast dynamical motions require not only fast and accurate motion planning, but also precise IK mapping, which is not practically feasible.
-By utilizing the JT-DS method, even a fast system could reactively generate stable motions to its target.
-
-Since we learn a lower-dimensional embedding space $Z := \{\phi(q) ~|~q \in Q\}$, which we suggest provides a better representation for the learned joint behaviors, a careful reader might wonder: why not define a motion policy $\pi_Z$ directly in $Z$-space and then map the learned policy back out into joint space using the inverse embedding\footnote{Such an inverse is trivial to extract in the case where $\phi$ is obtained through PCA.} $\phi^{-1}(z)$?
-The answer is that we would lose any guarantee of stability with respect to a target attractor $x_t^*$.
-By construction, $\phi$ maps from a joint-position vector of size $d$ to a low-dimensional vector of size $u<d$.
-Thus the inverse $\phi^{-1}$ must be of less than full rank, so the resulting $\mathcal{A}$ matrix (derived as in \eqref{eq:ds}) will also be of less than full rank, and thus no longer positive definite.
-This means that our controller no longer provably converges to the attractor (since our convergence proof no longer holds).
-Intuitively, this is because any policy defined in a lower-dimensional space than the actuation space will forfeit certain degrees of freedom, and thus may not be able to span the configuration space.
-Instead, we use the embedding space to modulate synergies defined in joint space, and which we can thus guarantee will be positive definite and lead us to convergence to the target.
+This means that adding multiple ``conflicting" synergies amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
 
@@
 
-%A drawback of the approach is that the controller is not guaranteed to converge to the task-space target $x^*$ even if a path to a valid configuration $q^*$ exists.
-As mentioned in Sec.
-\ref{Sec:DS}, the controller will always minimize (or keep even) the warped distance to goal $(H(q) - x^*)^T(H(q) - x^*)$, and in the process may get stuck at its kinematic joint limits.
-If the only valid trajectory to reach the goal requires that we temporarily move away from the target (and increase the warped distance-to-goal), this controller will not find that trajectory.
-There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession.
-The smaller subtrajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits.
-Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.
+Since we learn a lower-dimensional embedding space $Z := \{\phi(q) ~|~q \in Q\}$, which we suggest provides a better representation for the learned joint behaviors, a careful reader might wonder: why not define a motion policy $\pi_Z$ directly in $Z$-space and then map the learned policy back out into joint space using the inverse embedding.
+$\phi^{-1}(z)$?
+The answer is that we would lose any guarantee of stability with respect to a target attractor $x_t^*$.
+By construction, $\phi$ maps from a joint-position vector of size $d$ to a low-dimensional vector of size $u<d$.
+Thus the inverse $\phi^{-1}$ must be of less than full rank, so the resulting $\mathcal{A}$ matrix (derived as in \eqref{eq:ds}) will also be of less than full rank, and thus no longer positive definite.
+This means that our controller no longer provably converges to the attractor (since our convergence proof no longer holds).
+Intuitively, this is because any policy defined in a lower-dimensional space than the actuation space will forfeit certain degrees of freedom, and thus may not be able to span the configuration space.
+Instead, we use the embedding space to modulate synergies defined in joint space, and which we can thus guarantee will be positive definite and lead us to convergence to the target.
 
@@
 In this way, we can ensure that the performance of the DS is kinematically feasible for the robot to  follow.
-
-%One application that we have discussed for this algorithm is obstacle avoidance.
-In an environment with static obstacles, the system would, through demonstration, learn joint-space behaviors that avoid those obstacles.
-One can compare this form of joint-space obstacle avoidance with other methods from the literature \cite{sciavicco1988solution,petrivc2013smooth}.
-The main difference is that these previous methods require explicit knowledge of the obstacles' position, whereas our algorithm learns the obstacle positions only implicitly by learning motions that avoid those positions.
-This lack of explicit obstacle location can be an advantage: encoding obstacles' geometry can be expensive and cumbersome.
-However, it is also a drawback: without knowing the exact location of obstacles, the algorithm cannot be guaranteed to avoid collisions, especially away from demonstrations.
 
@@
 \[ V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*) \]
-That is, $ 0 \prec V(q),~ \forall q \neq q^*$ and $ V(q^*) = 0 $. Where $q^*$ is any joint configuration such that $H(q^*) = x^*$.  The derivative of $ V $ wrt. time is as follows:
-
+That is, $ 0 \prec V(q),~ \forall q \neq q^*$ and $ V(q^*) = 0 $. Where $q^*$ is any joint configuration such that $H(q^*) = x^*$.  The derivative of $ V $ wrt. time is:
 \begin{equation}
@@
 &= -(H(q) - x^*)^TJ(q)\mathcal{A}(q)J^T(q)(H(q) - x^*)\\
-&= -(H(q) - x^*)^TJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\prec}\underbrace{A_k}_{0\prec}J^T(q)(H(q) - x^*)\leq 0
+&= -(H(q) - x^*)^TJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{> 0}\underbrace{A_k}_{\succ 0}J^T(q)(H(q) - x^*)\leq 0
 \end{aligned}
 \end{equation}
-
 Therefore, JT-DS \eqref{eq:ds} is stable and as $ q=q^* $ is the largest invariance in $ \mathbb{R}^m $, it is  globally asymptotically stable with respect to a task-space attractor; i.e.

### 2017-11-17 paper/main_RA-L_v1_final.tex
8 pages paper!

@@
 \label{Sec:Dis}
-In this paper, we have presented a dynamical system in joint space that is provably Lyapunov-stable in task space and which replicates demonstrated joint-space behaviors.
+In this paper, we have presented a dynamical system in joint space that is provably asymptotically stable in task space to a fixed target while replicating demonstrated joint-space behaviors.
 The desired motions are fast to compute, and smoothly handle singularities by avoiding the pseudo-inverse Jacobian.
-We showed the system's ability to learn different joint-space behaviors on a redundant robotic platform.
+We showed the system's ability to learn different joint-space behaviors on a robotic platform.
 
@@
  When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\frac{1}{K}\sum\limits_{k=1}^{K}A_k$ is used to move the robot, which is still guaranteed to move towards the target.
-Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably converging to the target.
+%Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably converging to the target.
 The reason for this is that rather than ``determining" the velocity of the system, our $\mathcal{A}$ matrix only warps it.
@@
 
-Finally, as the proposed DS might get stuck at its kinematic joint limits, we are currently working on improving the performance of the proposed DS, by learning its parameters wrt.
+Finally,  we are currently working on improving the performance of the proposed DS, by learning its parameters wrt.
 the kinematic constraints.

### 2017-11-20 paper/main_RA-L_v1_final.tex
Updated Video link in paper

@@
 				\end{subfigure}
-    \captionof{figure}{Demonstrated Joint-Space Behaviors with Task-Space Targets. \href{https://www.dropbox.com/sh/3vx512oldgbnkq5/AABP5iEqO0oYn7dlE-gvK1kPa?dl=0}{[videos]} \label{fig:behaviors}}
+    \captionof{figure}{Demonstrated Joint-Space Behaviors with Task-Space Targets. \label{fig:behaviors}}
   \end{minipage}
@@
 	\end{subfigure}\\	
-	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{[video link]}.}
+	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/rirQkK0SSso}{[video link]}.}
 	\label{fig:snapshot}

### 2017-11-20 paper/main_RA-L_v1_final.tex
Fixed typo in abstract, compressed final pdf, waiting for Aude's approval to submit...

@@
 \begin{abstract}
-In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while stably converging towards a its task-space attractor.
+In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while stably converging towards a task-space attractor.
 % Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.

### 2017-11-21 paper/Pic/Sources/Schematic.m
Uploading final version

@@ -13,7 +13,7 @@ x_limit=[0:0.01:4];
 y_limit=[0:0.01:4];
 X_d=[2;2];
 
-% Plot the SVM Contours
+% Plot the Lyapunov Function Contours
 level = 100; n = ceil(level/2);
 cmap1 = [linspace(1, 1, n); linspace(0, 1, n); linspace(0, 1, n)]';
 cmap2 = [linspace(1, 0, n); linspace(1, 0, n); linspace(1, 1, n)]';
@@ -32,8 +32,8 @@ subplot1 = subplot(1,2,1);
 hold(subplot1,'on');
 
 L(1) = Link([0 0 l1 0],'standard')
-L(2)= Link([0 0 l2 0],'standard')
-L(3)= Link([0 0 0 0],'standard')
+L(2) = Link([0 0 l2 0],'standard')
+L(3) = Link([0 0 0 0],'standard')
 r = SerialLink(L,'name','two link')
 r.plotopt = {'noshadow','nojaxes', 'nowrist','noname','linkcolor',0.7*[1,1,1], 'ortho','noshading','notiles','jointcolor',0.4*[1,1,1]};
 r.plot([pi/3,-pi/4,0])

### 2018-02-22 paper/old-iterations/main_v10.tex
Latest changes

--- 2017-10-29 paper/main_RA-L_v1_nadia.tex
@@
 \usepackage{authblk}
-\usepackage{tabularx}
-\usepackage{multirow}
-\usepackage[hidelinks]{hyperref}
+\usepackage{hyperref}
 % *** MATH PACKAGES ***
@@
 
-\usepackage{xcolor,colortbl}
-
-\newcommand{\mc}[2]{\multicolumn{#1}{c}{#2}}
-\definecolor{Gray}{gray}{0.85}
-\definecolor{LightCyan}{rgb}{0.88,1,1}
-
-\let\labelindent\relax
-\usepackage{enumitem}
-\usepackage{dsfont}
 \usepackage{tikz}
@@
 %\title{Learning Joint-Space Stable Dynamical Systems for Task-Space Objectives}
-\title{Learning Augmented Joint-Space Task-Oriented Dynamical Systems: \\ A
-Linear Parameter Varying and Synergetic Control Approach}
-
-%OR
-%\title{Learning Augmented Joint-Space Task-Oriented \\ Dynamical Systems in Synergy Space}
+\title{The Augmented Joint-space Task-oriented Dynamical System}
 
@@
 of Electrical Engineering and Computer Science, Massachusetts Institute of Technology, Cambridge,
-MA, 02139 USA. E-mail: yonadav.shavit@gmail.com.}% <-this % stops a space
+MA, 02139 USA. E-mail: yonadav@mit.edu.}% <-this % stops a space
 \thanks{S.
@@
 \begin{abstract}
-In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space.
-Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it Lyapunov-stably converges towards its task-space attractor.
-To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
-Specifically, we use dimensionality reduction to find low-dimensional synergy regions, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
-Our proposed approach is validated on a variety of motions for a 7-DOF KUKA LWR4+ robot arm.\end{abstract}
+In this paper, we propose a stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a reachable target in task-space with respect to the robot's kinematic constraints.
+Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, we prove that it stably moves towards its target and is guaranteed to obey the robot's kinematic joint limits.
+To encode complex joint-space behaviors while meeting these stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system, enabling us to encode complex joint-space behaviors.
+We propose a method to train such complex behaviors from kinesthetic demonstrations, by learning the parameters of the LPV system.
+We begin by applying a latent-embedding and a probabilistic approach to discover and represent a set of local behaviors from the demonstrated motions.
+Then, the corresponding local joint-space behaviors are estimated by solving a convex semi-definite optimization problem that minimizes the joint velocity error from the demonstrations.
+The effectiveness of our approach is illustrated by a variety of motions for a 7-DOF KUKA LWR4+ robot arm.\end{abstract}
 
@@
 \centering
-\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{With_Obstcle.png}
-\caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl).
-The left example avoids a known obstacle in its workspace, while the right one does not.}
+\includegraphics[scale=.23,trim={0 0 0 14.0cm},clip]{Without_Obstcle.jpg}
+\includegraphics[scale=.23,trim={0 0 0 9.cm},clip]{With_Obstcle.jpg}
+\caption{Two possible robot motions in joint space accomplishing the same behavior in task-space (pouring chips in a bowl).
+Though the \textcolor{cyan}{\textbf{task-space trajectories}} are similar, the \textcolor{orange}{\textbf{joint-space trajectories}} are different; e.g.
+the bottom example moves in joint-space to avoid an obstacle that the top one does not.}
 \label{fig:robot_example}
-\vspace{-15pt}
+\vspace{-1.0cm}
 \end{figure}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{Pic/Schematic.jpg}
+\caption{A simplified schematic comparing the standard control architecture using Cartesian dynamical systems (right) and our propose approach.
+By integrating the "inverse kinematic (IK) solver" and "motion generator" blocks into a unified block, the proposed approach not only eliminates undesirable effects of the IK solver but also is capable of generating motions that satisfy the kinematic constraints of the robot.}
+\label{fig:schematic}
+\vspace{-0.8cm}
+\end{figure*}
 
@@
 
-Although DS-based task-space learning techniques ensure target convergence and stability, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution.
-Further, for certain tasks, rather than generating arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.
+Although DS-based task-space learning techniques ensure target convergence and stability, an advantageous feat within the PbD paradigm, they still suffer from this reliance on IK methods, possibly generating erratic or discontinuous motions during execution.
+Further, for certain tasks, rather than blindly choosing the arbitrary joint-space trajectory returned by the IKs solver, it might be desirable to learn behaviors and ``styles" in joint-space directly.
 For example, throwing a ball is a motion less characterized by task-space behaviors (the ball moving backwards and then forwards) and more characterized by the joint behaviors: the elbow folds, the shoulder moves back and then swings forward, followed by the elbow straightening and the wrist snapping down.
@@
 \ref{fig:robot_example}.
-This is precisely the shortcoming associated with LfD in task-space: it fails to learn the joint-space behavior of the system.
+
+This is precisely the shortcoming associated with PbD in task-space: it fails to learn the joint-space behavior of the system.
 The current formulation of the DS-based learning techniques introduced thus far (e.g.
 SEDS, $\tau$-SEDS LMDS, etc.) do not allow for integrating combined task/joint-space behavior learning.
-
-\subsection{\textcolor{red}{Approaches that tackle the joint+task problem}}
-\textcolor{red}{*See comments in tex*}
-% TO YO:
-% ADD THESE THREE REFERENCES HERE AND SUMMARIZE THEM
-%[1] Y. Huang, D. B ̈uchler, O. Koc, B. Scholkopf and J. Peters, “Jointly
-%learning trajectory generation and hitting point prediction in robot table
-%tennis,” in Proc. IEEE International Conference on Humanoid Robots,
-%2016, pp. 650-655.
-%
-%[2] S. Calinon and A. Billard, “Statistical learning by imitation of competing constraints in joint space and task space,” Advanced Robotics,vol. 23, pp. 2059-2076, 2009.
-%
-%[3] J. Silverio, S. Calinon, L. Rozo and D. G. Caldwell, “Learning competing constraints and task priorities from demonstrations of bimanual skills,” arXiv:1707.06791, 2017.
-
 \cite{calinon2008probabilistic} notably attempted to address this problem by learning separate motion policies in task space and the null space of the Jacobian (which would not affect task-space position), and driving the robot with a weighted sum of the two.
 However, this approach no longer guarantees convergence to the desired task-space target and is still reliant on computing the pseudo-inverse Jacobian.
-
 \cite{lee2014unifying} learns from demonstration in both joint and task space within the "thin-plate spline" trajectory warping framework, but does not propose a dynamical system for generating motions.
 
-\subsection{Our proposed Approach}
-\textcolor{red}{Dynamical System, Why?\\Linear Parameter Varying, Why?\\Lower-Dimensional Manifold, Why?\\}
-In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
-\cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities.
-The resulting DS avoids singularities through generalization of the pseudo-inverse approximations.
+In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JTDS) that not only complies with the task-space constraints, but also avoids the problems generated by pseudo-inverse approximations and takes into consideration the robot's kinematic limits, see Fig.
+\ref{fig:schematic}.
+\cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities and joint kinematic limits.
+The resulting DS avoids singularities through generalization of the pseudo-inverse approximations and joint limits via online modulation.
 However, because the two DS have their own unique attractors, the non-linear interaction between them imposed by the kinematic constraints does not ensure that the combined DS has a unique attractor.
 This gives rise to spurious attractors or cycles, and thus must be carefully tuned in order to avoid them.
-In this work, we depart from any modulation or combination of DS and instead propose a novel dynamical system which:
+In this paper we depart from any modulation or combination of DS and instead propose a novel dynamical system which:
 
-\begin{enumerate}[leftmargin=*]
-\item computes a \begin{bf}complex motion in joint space\end{bf} that provably and stably converges to a \begin{bf}fixed task-space target\end{bf}.
-\item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic ``singularities".
-\item is formulated such that complex \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
+\begin{enumerate}
+\item computes a \begin{bf}motion in joint space\end{bf} that provably stably moves to a \begin{bf}fixed task-space target\end{bf} in a single DS.
+\item inherently obeys \textbf{kinematic joint limits}.
+\item \textbf{avoids} the need for computing \textbf{pseudo-inverses}, and cleanly moves through kinematic "singularities".
+\item is formulated such that it can \textbf{learn} complex \textbf{joint-space behaviors} from demonstrations.
 \end{enumerate}
@@
 It shares some of our approach's advantages: fast computation and provable task-space stability.
-However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations.
-%while obeying the robots kinematic constraints.
+However, despite some previous work integrating discontinuous joint limits \cite{sciavicco1988solution} and designing velocity adjustments by hand \cite{Shi2016}, this is the first known work to employ a JT system to learn behaviors from demonstrations while obeying the robots kinematic constraints.
 
@@
 
-%\begin{figure}[t]
-%\centering
-%\includegraphics[width=0.95\linewidth,trim={0 20 450 0},clip]{Pic/Schematic.jpg}
-%\includegraphics[width=0.9\linewidth,trim={580 20 0 0},clip]{Pic/Schematic.jpg}
-%\caption{ Standard control architecture using Cartesian dynamical systems (DS)\textbf{(top)}  and our proposed approach \textbf{(bottom)}.}
-%\label{fig:schematic}
-%\end{figure}
 \section{Problem Statement} \label{Sec:Prob}
@@
 } The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{m\times d}$.\\
-We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following two criteria:
+We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following criteria:
 \renewcommand{\labelenumi}{(\Roman{enumi})}
@@
 $V(.)$ can be thought of as a metric for the task-space distance-to-go.
+\item The dynamical system must respect the kinematic joint constraints of the robot:
+\begin{equation}
+q(0)\in Q \rightarrow q(t)\in Q~~~~\forall t>0~
+\end{equation}
 \item The dynamical system should encapsulate the desired joint-space behaviors such that the following metric is minimized
@@
 \label{sec:proposed_system}
-\textcolor{red}{Change semi-definite to definite everywhere\\Change overall $A(q)$ to $\mathcal{A}(q)$\\ Introduce 'behavior synergies' here\\}
-We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \ref{eq:Lyp} and \ref{eq:optim_error}.
+We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the three criteria presented in (I), (II), (III).
 \begin{equation}
 \label{eq:ds}
-\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
 \end{equation}
-where $P\in \mathbb{R}^{d\times d}$, and $A(q)\in \mathbb{R}^{m\times m}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
+where 
+\begin{equation}
+\label{eq:SAS}
+\mathcal{A}(q) = S(q) A(q) S^T(q). 
+\end{equation}
+$S(q)=diag(s^1(q),\dots,s^d(q))$ is a diagonal matrix and $ s^i(\cdot)\in C^0~\forall i \in \{1,\dots d\} $.
+$A(q)\in \mathbb{R}^{d\times d}$ is constructed using the Linear Parameter Varying system paradigm \cite{emedi2016fixed,7439839}, where the overall $A(q)$ is a nonlinear combination of constant linear matrices, each of which encodes a local joint-space behavior.
 \begin{equation}
@@
 \end{equation}
-where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
+where $A_k\in \mathbb{R}^{d\times d} $ and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are constant matrices and the scheduling parameters\footnote{The scheduling parameters can be a function of time $t$, the states of the
 system $q $ or external signals $d(t)$, i.e.
@@
 In this paper, we only consider it as a function of the states of the system.
-It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$.} modulating each local synergy through time as well as space.
-Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more complex multi-behavior motion.
+It is noteworthy that the presented stability and convergence proof can easily be extended for $\theta_k (t, q(t), d(t))$.} representing each local behavior, respectively.
+Based on \eqref{eq:ds}, one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.
 
@@
 \includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
-\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $A$ (emphasizing the hip, knee, and ankle respectively), A: $A(q)=diag(5,1,1) $, B: $A(q)=diag(1,5,1)$, and C: $A(q)=diag(1,1,5)$.
+\caption{Three example 3-DOF motions (A, B, C), each with a different constant behavior matrix $\mathcal{A}$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
 On the left, the task-space traces of each motion.
-On the right, the time-scaled joint positions of each joint.
+On the right, the time-scaled joint positions of each joint (with each color signifying a different trajectory).
 Each motion tends to use its ``primary" joint most and uses the other available joints to compensate for what the primary joint cannot do.}
@@
 the task target (warped according to P), and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}).
-The positive semi-definite matrix $A(q)$ warps the resulting joint-space velocity; Fig.
+The positive semi-definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
 \ref{fig:A_example} illustrates the effects of  $A(q)$ on the generated motion.
 Thus the controller can be thought of as a proportional controller in joint space.
-Lastly, we refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
+We refer to the $P$ matrix as the \begin{bf} task augmentation matrix \end{bf}(because it augments the task error, and thus the direction of motion) and $A(q)$ as the \begin{bf} joint augmentation matrix \end{bf} (because it augments the outputted joint velocities).
+Lastly, $S(q)$ intuitively enforces the joint constraints: as $q^i$ nears $q^i_{min}$, the velocity in that direction goes to $0$, and similarly when it nears $q^i_{max}$ the velocity goes to $0$.
+
 %When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle).
@@
 \label{prop:stability}
-The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if \eqref{eq:ds} meets the following constraints.
+The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) and (II), if \eqref{eq:ds} meets the following constraints.
 
@@
 & 0 \leq \theta_k(\cdot) & &\\
-%    & 0 \preceq and &\forall q \in Q\\
+%    & 0 \preceq S(q) &\forall q \in Q\\
 \end{split}
 \forall k \in \{1,\dots,K\} 
+\end{cases}
+\end{equation}    
+\begin{equation}
+\label{eq:first_criteria_joint}
+\begin{cases}
+\begin{split}
+&s^{i}(q^i_{min})=0 & &\\
+&s^{i}(q^i_{max})=0 & &
+\end{split}
+\forall i \in \{1,\dots,d\}
 \end{cases}
@@
 This happens if the target position is kinematically unreachable; i.e.
-for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or the joint has reached a kinematic limit.
-Instead, the proof in Appendix \ref{appendix:stability} shows that the system converges in the sense of Lyapunov, meaning that its distance to the attractor is always monotonically decreasing.
+for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or $S_{i,i}(q) = 0$ (the joint has reached its limit).
+Note that in practice, because $s(q)$ is continuous and designed based on \eqref{eq:first_criteria_joint}, executing JT-DS \eqref{eq:ds} will likely never cause the manipulator to reach the joint limit.
+Instead, $s(q)$ significantly shrinks the magnitude of a joint's motion when a joint nears its motion boundary.
+For a simple method of constructing $S(q)$, see Appendix \ref{appendix:limits}.
 
-Criterion (II) (i.e.
+Criterion (III) (i.e.
 encoding specific joint-space behaviors) is achieved by embedding the desired dynamics in the matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$.
@@
 
-\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} 
+\section{Learning the Parameters of JT-DS} 
 \label{Sec:Learning}
 
-The behavior of the JT-DS algorithm can be best understood through the lens of synergy control \cite{petrivc2013muscle}.
-In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the actuation-space that are sufficient to accurately recreate the desired robotic behaviors.
-In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $A(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
+JT-DS \eqref{eq:ds} is parametrized by $\theta_k(q) $ and $A_k~\forall k\in\{1,\dots,K\}$.
+One is free to determine these parameters as long as that they satisfy constraints \eqref{eq:first_criteria_stability}.
+We propose a 3-step procedure as follows:
+\begin{enumerate}
+\item We first project the demonstrated data (i.e.
+collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$ through Principal Component Analysis (PCA) \cite{jolliffe1986pca}, so as to reduce the dimensionality and extract correlations across similar joint behavior regions.
 
-A central question becomes how to modulate the synergies in different regions to yield our desired behavior.
-First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that a latent mapping $\phi(\cdot)$ and latent space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using some sub-manifold of the joint-space, called the \emph{synergy space}.
-We would like to define the robot's policy in synergy space such that in different regions of the space, we will utilize different synergies.
-We are thus left with three problems: finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(q)~\forall k\in\{1,\dots,K\}$), and finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
-Our choice of parameters must also obey the constraints laid out in \eqref{eq:first_criteria_stability}.
-We propose the following 3-step procedure.
+\item We then jointly estimate the optimal number $K$ of local behavior regions and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$.
 
-\begin{enumerate}
-\item We first construct our "synergy-space", which provides us a lower-dimensional manifold through which to control the robot, by projecting the demonstration data (i.e.
-collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$.
-In each experiment, we used either Principal Component Analysis (PCA) \cite{jolliffe1986pca} or Kernel PCA (KPCA) \cite{scholkopf1997kernel}.
-
-\item We then jointly estimate the optimal number $K$ of synergy "regions" and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$ seen in the demonstrations.
-
-\item Finally, once the local synergy regions have been found (described by each of the Gaussian distributions $\theta_k(q)$), we compute the corresponding joint synergy matrices $A_k~\forall k\in\{1,\dots,K\}$ for each region by formulating a convex optimization problem that finds the optimal set of $A_k$'s that minimize the overall velocity error with respect to the demonstrations \eqref{eq:optim_error}.
+\item Finally, once the local behavior regions have been found (described by each of the Gaussian distributions $\theta_k(q)$), we compute the corresponding joint behavior matrices $A_k~\forall k\in\{1,\dots,K\}$ for each region by formulating a convex optimization problem that finds the optimal set of $A_k$'s that minimize the overall velocity error with respect to the demonstrations \eqref{eq:optim_error}.
 \end{enumerate}
 
-\subsection{Embedding Joint Configurations in Lower-Dimensional Space}
-
-%WE ASSUME THAT, BECAUSE OUR DEMONSTRATIONS ARE SIMILAR, POINTS IN SIMILAR REGIONS WILL HAVE SIMILAR MOTION POLICIES, SO FINDING As IS EQUIVALENT TO FINDING PRIMARY SYNERGY IN REGION, AND THEN INTERPOLATING THE SYNERGIES IN THE REGIONS IN-BETWEEN
-
-The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint behaviors" (defined in Section \ref{sec:intro}), in which it would be appropriate to define synergies.
+\subsection{Embedding Joint Configurations in Latent Space}
+The search for a lower-dimensional representation of the joint positions stems from the desire to identify regions of \textit{similar} ``joint motion styles" (defined in Section \ref{sec:intro}).
 It is important to understand that distances in joint space (defined as $\| \Delta q\|$) might not numerically reflect what a human intuitively perceives as the distance between joint configurations.
@@
 
-\nocite{*}
-
-\subsection{Discovering Local Behavior Synergies}
-Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
+\subsection{Discovering Local Behavior Regions}
+Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local behaviors, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
 Moreover, we would like for our scheduling parameters $\theta'_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta'_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
@@
 \end{equation}
-which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th synergy.
-Therefore, each synergy region is associated with a Gaussian component of the GMM, cumulatively describing all the synergy regions of the dynamical system.
-We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM.
+which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th behavior.
+Therefore, each joint-space behavior region is associated with a Gaussian component of the GMM, cumulatively describing all the behavior regions of the dynamical system.
+We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM and choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.
 
-\subsection{Estimating the Synergy Matrices}
+\subsection{Estimating the Joint Behavior Matrices}
 Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $A(q)$ as a linear combination of local $A_k$ matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
@@
 
-Intuitively, each ``synergy region" in synergy space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
-Finding the appropriate synergy matrices $A_k$ for the set of local behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
+\begin{figure}
+	\begin{subfigure}[t]{0.5\textwidth}
+		\includegraphics[width=\textwidth]{Pic/cropped_Sing_1.pdf}
+		\caption{ }
+		\label{fig:Sin}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{0.5\textwidth}
+		\includegraphics[width=\textwidth]{Pic/cropped_Sing_2.pdf}
+		\caption{ }
+		\label{fig:Det}
+	\end{subfigure}\\
+	\caption{The experiment generated $K=3$ Gaussian components.
+As the first joint is the only joint which was not fixed during the demonstration, the learned augmentation matrices had only one nonzero entry $A_k(1,1)\neq0~\forall k\in \{1, 2, 3\} $.
+In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space.
+The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).
+In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
+Due to sensory noise, the determinants are not exactly zero, but are very close.
+}
+	\vspace{-20pt}
+\end{figure}
+
+Intuitively, each ``region of behavior" in joint/latent space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's behavior ($A_k$) influences the robot's current joint-space motions.
+Finding the appropriate joint augmentation matrices $A_k$ for the set of local behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
 To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:
@@
 where $f(q_{t,n})$ is calculated by combining our dynamical system formulation \eqref{eq:ds} with \eqref{eq:A}, and $x^*_{n}$ is defined as the endpoint of the $ n^{\text{th}} $ demonstrated trajectory.
+\section{Experiments} \label{Sec:Exp}
+The learning algorithm presented in the previous section, \eqref{eq:opt}, was implemented in MATLAB.
+In order to solve the semidefinite convex optimization problem, the YALMIP framework \cite{lofberg2004yalmip} was used.
+In all the experiments, the optimization is initialized multiple times, and the best resulting run is used for performance analysis.
+The source code for learning (MATLAB) and execution (C++) of the Augmented Joint-Space Task-Oriented Dynamical System together with data generated in these experiments are available on-line:\\
+\centerline{ \underline{\url{http://github.com/epfl-lasa}}}\\
 
-\begin{figure*}[!ht]
-  \begin{minipage}[b]{0.30\textwidth}
-	\centering	
-		\begin{subfigure}[t]{\linewidth}
-		\centering
-			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
-					\caption{Tennis Strokes (1-2).}
-		\end{subfigure}
-		\begin{subfigure}[t]{\linewidth}
-\centering
-	\includegraphics[width=0.45\linewidth]{new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{new-pics/pour-obst_crop.png}
-						\caption{ Pouring Motions (3-4)}
-			\end{subfigure}
-			\begin{subfigure}[t]{\linewidth}		
-\centering
-	\includegraphics[width=0.45\linewidth]{new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{new-pics/footstep.png}
-	\caption{ Pouring Motion (5) and Foot-step (6).}
-				\end{subfigure}
-    \captionof{figure}{Demonstrated Joint-Space Behaviors with Task-Space Targets. \href{https://www.dropbox.com/sh/3vx512oldgbnkq5/AABP5iEqO0oYn7dlE-gvK1kPa?dl=0}{[videos]} \label{fig:behaviors}}
-  \end{minipage}
-  \hfill
-  \begin{minipage}[b]{0.7\textwidth}
-    \scriptsize
-    \centering
-\resizebox{\linewidth}{!}{\begin{tabular}{cc|cc|cc}
-    \hline
-    \hline
-    \multicolumn{1}{c}{\multirow{4}{*}{Behavior Dataset}} & \multicolumn{1}{c}{\multirow{4}{*}{Dim.
-Red.
-Approach}} & \multicolumn{4}{c}{\multirow{2}{*}{Optimal Parameters | Joint Velocity RMSE [$rad/s$]}}
-	\\  
-    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(1) Forehand Stroke}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
-		& K-PCA ($\sigma_{opt} = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
-		\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(2) Backhand Stroke}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
-		    \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
-			& K-PCA ($\sigma_{opt} = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
-			\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(3) Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($N=9,M=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
-	& K-PCA ($\sigma_{opt} = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
-				\hline
-	\multicolumn{1}{c}{\multirow{3}{*}{(4) Pouring - Obstacle 1}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
-				& K-PCA ($\sigma_{opt} = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & \textbf{ 3.1 $\pm$ (0.316) } & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
-				\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(5) Pouring - Obstacle 2}} & None & 7 & 4.3 $\pm$ (1.418) & \textbf{ 0.122 $\pm$ (0.014) } & 3.095 $\pm$ (3.406) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($N=7,M=1406$)}} & PCA & \textbf{ 2.9 $\pm$ 0.316 } & 3.7 $\pm$ (2.213) & 0.126 $\pm$ (0.02) & \cellcolor{blue!10} 0.825 $\pm$ (0.764) \\	
-	& K-PCA ($\sigma_{opt} = 2.86$) & 3.9 $\pm$ (0.316) & \textbf{3} & 0.129 $\pm$ (0.013) & \cellcolor{blue!15} 0.738 $\pm$ (0.507)  \\
-				\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(6) Foot Step}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
-	\multicolumn{1}{c}{\multirow{3}{*}{($N=8,M=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
-	& K-PCA ($\sigma_{opt} = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
-				\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(7) Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
-	& K-PCA ($\sigma_{opt} = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
-\end{tabular}}
-      \captionof{table}{\scriptsize Performance Comparison of Learning Approach with different Dim.
-Red.
-Schemes.
-We present the \texttt{mean(std)} for the optimal $\delta$, $K$ and joint-velocity RMSE on training and testing set, found for every learning scheme over 10 runs.\label{tab:learning}}
-    \end{minipage}
-    \vspace{-20pt}
-\end{figure*}
-
-\newpage
-\section{Experimental Validation} \label{Sec:Exp}
-\subsection{JT-DS Learning Performance  Evaluation}
-To evaluate the proposed JT-DS learning algorithm we collected kinesthetic demonstrations on a 7-DOF KUKA LWR 4+ robot arm, for 7 different tasks (see Table \ref{tab:learning} and Fig.
-\ref{fig:behaviors}) which require mimicking the demonstrated joint-space behavior while reaching for a single target in task space:
-\begin{itemize}[leftmargin=*]
-\item \textit{(1-2) Tennis strokes:} For these particular behaviors, we guide the robot to perform a forehand and backhand stroke.  
-\item \textit{(3-5) Pouring motions:} We guide the robot to pour chips in a bowl with three different joint-space behaviors.
-The first is without an obstacle in the workspace.
-The second is with an obstacle in the upper hemisphere of the workspace, where the human is seen folding the elbow and lowering the shoulder of the robot arm to avoid the obstacle.
-The third instance is with an obstacle in the lower hemisphere, and now the human raises the shoulder to avoid it.
-\item \textit{(6) Footstep-like motion:} We demonstrate a footstep motion which begins with a straight leg, moving through a singularity, and finally bringing the knee up.
-\item \textit{(7) Singularity motions:} Movement was constrained to the boundary of the workspace by fixing $ q^i=0~\forall i\in\{3,\dots,7\}$, the second joint was fixed to $q^2\in\{10^\circ,20^\circ,\dots,100^\circ\}$, and only the first joint was moved by a human.
-This restricted the motion to a series of arcs of different radii along the robot's motion boundary.
-\end{itemize}
-
-\noindent For each behavior, we collected $N$ demonstrations consisting of ($14$-D) time-series $D=\{\{ q_{t,n},\dot{q}_{t,n} \}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where each demonstration has different durations $T_n$.
-We also record the final task-space configurations ${x^*}^{(n)}$, which we assume is the attractor in task-space.
-The recordings are collected at a rate of 500 Hz.
-However, our learning approach does not require such a dense representation of the trajectories, hence we down-sample them to 25 Hz.
-The total number of demonstrations $N$ and samples per \textit{dataset} $M = \sum_{n=1}^{N}T_n$ are provided in Table \ref{tab:learning}.
-By performing our evaluation on all these datasets, we seek to:
-\begin{enumerate}[leftmargin=*]
-\item Verify that the learned \textit{behavior synergies} are mimicking the demonstrated joint-space behaviors and generalize them to new initial joint configurations.
-\item Analyze the role of dimensionality reduction in our proposed approach and further find the method which yields the best trade-off between accuracy and model complexity; i.e.
-the least number of dimensions $\delta$ to represent our activation function $\theta(q)$ and the least number of local behavior synergies $K$.
-\end{enumerate}
-We thus evaluate three learning approaches with different dimensionality reduction algorithms: (1) None, (2) PCA and (3) K-PCA with RBF kernel.
-We perform 10-fold cross-validation on all datasets, for each learning approach, with a training/testing ratio of $60 \%$.
-To evaluate performance, we take the square root of the MSE defined in \eqref{eq:optim_error}, which we refer to as the joint-velocity RMSE.
-
-Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
-Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$.
-Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
-\begin{enumerate}[leftmargin=*]
-\item \textit{$\delta$ for PCA:} We choose $\delta$ such that the projection is capable of explaining $95\%$ of the variance in the data.
-\item \textit{$\delta$ and $\sigma$ for K-PCA:} The interaction of these two parameters plays a major roll in the resulting projection obtained from K-PCA.
-Hence, we do a grid search on a log-spaced range of $\sigma_{range} = [\sigma_{min}:\sigma_{max}]$ values, where
-$\sigma_{min} = \frac{1}{\kappa\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ and $\sigma_{max} = \frac{2\kappa}{\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ for $\kappa = cte.$ This yields a feasible range for $\sigma$ that is guided by the pairwise Euclidean distances between all points in the dataset.
-Moreover, K-PCA is not limited to providing a $\delta \ll d$, in fact it can generate $\delta \le M$.
-This is a nuisance as in our datasets $M\approx1000$.
-To alleviate this we truncate $\sigma_{range}$ by compute the explained variance of the eigenvectors in feature space for different values of  $\sigma \in \sigma_{range}$.
-We then remove $\sigma$ values from $\sigma_{range}$ whose number of \textit{optimal} eigenvectors $\delta > d$ and resample it.
-By doing this procedure, we can ensure that for all values in truncated $\bar{\sigma}_{range}$ we will obtain $\delta \le d$.
-We then choose $\sigma_{opt}$ by running 10-fold cross-validation for all $\sigma \in \bar{\sigma}_{range}$.
-\item \textit{$K$ for GMM:} We choose the optimal number of components $K$, by evaluating and selecting the best resulting model using the Bayesian Information Criterion (BIC) \cite{Bishop:PRM:2006}.
-Typically, one chooses the optimal $K$ manually, by visually identifying the point at which the BIC curve produced from $K_{range}$ stops changing or plateaus.
-To automate this process, we devised an approach which selects the optimal $K$ as the one which yields the highest inflection point on the second order derivative of the BIC curve.
-\end{enumerate}
-
-Once these optimal hyper-parameters are estimated for each fold, we solve for the definite convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
-We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}
-
-As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
-This suggests that using DR to represent our activation functions $\theta(q)$ yields better generalization capabilities than the contrary.
-This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learnt without DR is an order of magnitude higher than with DR.
-Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the number of $K$ needed to represent the motion.
-For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
-By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
-Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
-
-\begin{figure*}[ht]
-	\vspace{-0.05cm}
-	\begin{subfigure}[t]{0.42\linewidth}
-		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
-		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
-		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
-		\caption{Execution of Pour Obstacle 2 learned through JT-DS.}
-		\label{fig:snapshot:a}
-	\end{subfigure}
-	\begin{subfigure}[t]{0.567\linewidth}
-		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
-		\caption{Execution of Pour Obstacle 2 learned through SEDS.}
-		\label{fig:snapshot:b}
-	\end{subfigure}\\	
-	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{[video link]}.}
-	\label{fig:snapshot}
-	\vspace{-10pt}
-\end{figure*}
-
-Both PCA and K-PCA yield comparable results, with K-PCA providing a slight improvement on some datasets.
-This suggests that perhaps a linear DR method might be sufficient for such tasks.
-However, if we seek to maximize accuracy a non-linear DR method should be employed.
-One of the drawbacks of K-PCA is its computational complexity for out-of-sample evaluations, which involves computing the kernel function between the new data-point and all the samples $M$.
-This is, however, not so taxing for our method as our datasets range in $M\approx1000$ and in previous work \cite{Mirrazavi:IJRR:2017} we have experimentally found that evaluating $<3000$ RBF kernel computations (on a 3.4-GHz i7 PC with 8GB RAM)) in a control-loop with 2ms rate is feasible.
-If one requires extremely fast computation; i.e.
-a control-loop rate of $<2$ms then PCA should be used as opposed to K-PCA.
-
-\textit{Implementation Details:} The learning pipeline is implemented in MATLAB.
-We used the Matlab Toolbox for Dimensionality Reduction \cite{Maaten08dimensionalityreduction} for implementations of PCA, K-PCA and its out-of-sample extension.
-The YALMIP framework \cite{lofberg2004yalmip} was used to solve the definite convex optimization problem.
-Source code for the entire learning pipeline and cross-validation with simulation of multi-DOF robot arms can be found in:
-\footnotesize  \centerline{\underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}}
-\normalsize
-
-For execution of the learnt JT-DS models on a real 7-DOF the KUKA LWR 4+ robot arm, we provide the following code in C++:
-\footnotesize \underline{\url{https://github.com/epfl-lasa/JT-DS-lib}}\\
-\normalsize
-The robot is controlled on the joint position level (linearly interpolating from joint velocities computed from JT-DS) at a rate of $ 500 $ Hz.
+The performance of the proposed framework was evaluated on a 7-DOF robot arm, the KUKA LWR 4+, both in simulation and in reality.
+The robot is controlled on the joint position level (linearly interpolating from joint velocities) at a rate of $ 500 $ Hz.
 The resultant joint angles are filtered by a critically damped filter to avoid high torques.
 
-\subsection{JT-DS Execution Performance Evaluation}
-We now seek to elucidate the distinctive properties of our JT-DS model by comparing its performance to a DS-based Cartesian motion generator + IK solver approach for behaviors \textit{(5-7)}.
-For the Cartesian motion generator we use the SEDS approach \cite{khansari2011learning} which learns an asymptotically stable DS in Cartesian space from demonstrations.
-We then generate joint trajectories through a damped least square IK solver.
-From herein we refer to this approach as SEDS+IK.
+Our empirical validation consists of three sections, each highlighting a different advantage of the proposed dynamical system: (i) moving in singular configurations, (ii) following desired behaviors in joint and task spaces simultaneously, and (iii) fast computation and convergence time.
 
-\subsubsection{Following Desired Joint and Task-Based Behaviors}
-We compare tracking capabilities of our JT-DS method with those of SEDS+IK for behaviors \textit{(5)} and \textit{(6)}.
-In Fig.
-\ref{fig:snapshot:a} and \ref{fig:snapshot:b} we can qualitatively see the difference between these two approaches for behavior \textit{(5)}.
-The JT-DS algorithm mimicked the joint-space behavior of the demonstration (e.g.
-folding the elbow, raising the shoulder), successfully avoiding the obstacle while still converging to the desired Cartesian position.
-Meanwhile, SEDS+IK only learned the demonstrated behavior in task-space, it does not have the capabilities of constraining motion in joint-space.
-This ultimately led to one of its joints colliding with the obstacle.
-It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
-For the foot-step behavior \textit{(6)} JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig.
-\ref{fig:snapshot:c} and \ref{fig:foot}).
-This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.
-
-\begin{figure}[t]
-\begin{minipage}[b]{\linewidth}
-\centering
-		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{new-pics/singularity_crop.png}
-		\caption{Execution of \textit{problematic} joint-space behaviors learned through JT-DS: (left) Foot step-like motion and (right) Singularity motions.}
-		\label{fig:snapshot:c}
-\end{minipage}
-\begin{minipage}[b]{\linewidth}
-	\centering
-	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
-	\caption{End-effector trajectories for the footstep motion in Cartesian space.
-The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
-On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing.
-\label{fig:foot}}
-	\vspace{-10pt}
-	\end{minipage}
-	\vspace{-20pt}
-\end{figure}
-
-\subsubsection{Systematic Assessment}
-To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a SEDS+IK, we simulated $400$ simple motions on each system.
-The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
-The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
-To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
-uniform behaviors.
-The results, summarized in Table \ref{tab:sys}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
-As the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
-Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than SEDS+IK.
-
-\begin{table}[t]
-	\label{table:2}
-	\centering
-	\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
-		\hline\hline
-		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
-		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
-		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
-	\end{tabular}}
-		\caption{Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
-The convergence duration is the time required to move within $0.001$m of the target.
-The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.\label{tab:sys}}
-		\vspace{-10pt}
-\end{table}
-
-\subsubsection{Transiting through singular configurations}
-One of the main advantages of the proposed DS is its ability to generate accurate paths in classical singular configurations.
-To evaluate this we generated behavior \textit{(7)}.
-The $10$ training demonstrations were constructed as follows:
-
+\subsection{Moving in singular configurations}
+One of the main advantages of the proposed dynamical system is its ability to generate accurate paths in classical singular configurations.
+The first scenario, shown in Fig.
+\ref{fig:Sin}, was designed to assess this capability.
+The $10$ training demonstrations were constructed as follows: movement was constrained to the boundary of the workspace by fixing $ q^i=0~\forall i\in\{3,\dots,7\}$, the second joint was fixed to $q^2\in\{10^\circ,20^\circ,\dots,100^\circ\}$, and only the first joint was moved by a human demonstrator back-driving it.
+This restricted the motion to a series of arcs of different radii along the robot's motion boundary.
 In other words, the demonstrated motions were entirely within a classic kinematic singularity.
@@
 
-\section{Discussion} 
+\subsection{Systematic Assessment}
+To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a Cartesian motion generator, we simulated $400$ simple motions on each system.
+The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
+The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
+To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single Gaussian, i.e.
+uniform behaviors.
+The results, summarized in Table \ref{table:2}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
+Furthermore, as the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
+Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than the Cartesian motion generator.
+
+\begin{table}[t]
+	\tiny
+	\vspace{0.1cm}
+	\caption{  Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
+The convergence duration is the time required to move within $0.001$m of the target.
+The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.}
+	\label{table:2}
+	\centering
+	\begin{tabular}{|l|l|l|l|}
+		\hline
+		& Norm. convergence [{s}/{m}] & Comp. time {[}ms{]} & Tracking error {[}m{]} \\ \hline
+		Cart. DS  + IK solver   & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ \hline
+		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline
+	\end{tabular}
+	\hspace{0.1cm}
+	\caption{Parameters and Performance of Comparative Experiments }
+	\centering
+	\begin{tabular}{|l|c|c|c|c|}
+		\hline
+		& No. Dem. & No. K & $\phi(q)$ dim. & $e_{totol}$ [{rad}/{s}] \\ \hline
+		Obstacle avoidance (JT-DS)  &      7                &     4             &        4  &  0.0137    \\ \hline
+		Foot step  (JT-DS) &       3               &     1             &          1 & 0.01344   \\ \hline
+		Obstacle avoidance (SEDS) &      7                &     3             &        -  &   -   \\ \hline
+		Foot step (SEDS)&       3               &     1             &          - & -   \\ \hline
+	\end{tabular}
+	\label{table:1}
+	\vspace{-15pt}
+\end{table}
+
+\begin{figure}[t]
+\vspace{-0.3cm}
+	\begin{subfigure}[t]{\linewidth}
+		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
+		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
+		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
+		\caption{Desired Joint Behavior: obstacle avoidance motion. The luggage is the obstacle.}
+		\label{fig:snapshot:a}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{\linewidth}
+		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}
+		\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
+		\includegraphics[width=0.32\linewidth]{Pic/Luggage3_SEDS.png}
+		\caption{Desired Behavior: The Cartesian DS is used to generate the motion.}
+		\label{fig:snapshot:b}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{\linewidth}
+	\centering	
+		\includegraphics[width=0.45\linewidth]{Pic/1.png}~~
+		\includegraphics[width=0.45\linewidth]{Pic/3.png}\\\vspace{0.1cm}
+		\includegraphics[width=0.45\linewidth]{Pic/4.png}~~
+		\includegraphics[width=0.45\linewidth]{Pic/2.png}
+		\caption{Desired Joint Behavior: Foot step-like motion.}
+		\label{fig:snapshot:c}
+	\end{subfigure}\\	
+	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/1dgKfmN1UgE}{https://youtu.be/1dgKfmN1UgE}.}
+	\label{fig:snapshot}
+	\vspace{-0.8cm}
+\end{figure}
+
+\subsection{Following Desired Joint Behaviors}
+
+To evaluate the system's ability to track learned joint behaviors, we devised two experiments (summarized in Table \ref{table:1}) comparing the tracking capabilities of our JT-DS method with those of the Cartesian-space SEDS \cite{khansari2011learning} algorithm using damped least square IK.
+In the first experiment, the robot was guided through a complex motion task, moving an object through an environment with an obstacle.
+Human experts guided the robot through a series of demonstrations, ensuring that neither the robot's joints nor the held object intersected the obstacle.
+The results can be seen in Fig.
+\ref{fig:snapshot:a} and Fig.
+\ref{fig:snapshot:b}.
+The JT-DS algorithm mimicked the joint-space behaviors of the demonstration (e.g.
+folding the elbow, lowering the shoulder), and therefore successfully avoided the obstacle while still converging to the desired Cartesian position.
+Meanwhile, SEDS only learned the demonstrated behavior in task-space (the position of the end-effector is signified by a black ball) meaning that the robot was not constrained in joint space.
+This ultimately led to one of its joints colliding with the obstacle.
+It should be noted that the JT-DS motion did not follow the demonstrations in task-space very closely (as we would expect), but did ultimately converge to its target position.
+In the second experiment the robot was taught a footstep-like motion, beginning with a straight leg, moving through a singularity, and finally bringing the knee up (Fig.
+\ref{fig:snapshot:c}).
+JT-DS followed the demonstrations closely, while SEDS became unstable in the singularity (Fig.
+\ref{fig:foot}).
+This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities, overcoming the Achilles heel of equivalent Cartesian-based motion generators.
+
+Further inspection reveals another advantage of JT-DS over stable Cartesian motion generators: not needing to preprocess demonstrations.
+In Cartesian-space dynamical systems, in order to guarantee stability, every demonstration must have the same target position.
+As a result, all demonstrated trajectories are warped to end at a single position, introducing a non-negligible difference between the original demonstrations and those used to train the DS.
+JT-DS, on the other hand, does not require any demonstration warping to learn.
+Rather than shifting demonstrations, the algorithm learns behavior regions with respect to the base frame of the manipulator.
+This effect can be seen in the diagrams in Fig.
+\ref{fig:Sin}, which show a comparison between JT-DS and a Cartesian-space DS trained through SEDS \cite{khansari2011learning}.
+The Cartesian-space system warps the demonstrations and as a result fails to track them, while JT-DS does not modify the demonstrations and thus succeeds.
+
+\section{Discussion and Future Work} 
 \label{Sec:Dis}
 In this paper, we have presented a dynamical system in joint space that is provably Lyapunov-stable in task space and which replicates demonstrated joint-space behaviors.
-The desired motions are fast to compute, and smoothly handle singularities by avoiding the pseudo-inverse Jacobian.
+The system provably obeys the manipulator's kinematic joint limits, is fast to compute, and smoothly handles singularities by avoiding the pseudo-inverse Jacobian.
 We showed the system's ability to learn different joint-space behaviors on a redundant robotic platform.
 
+\begin{figure}
+\vspace{-10pt}
+	\centering
+	\includegraphics[width=0.8\linewidth]{Pic/Foot.pdf}
+	\caption{A plot of the end-effector trajectories for the footstep motion in Cartesian space.
+The JT-DS motion moves smoothly closely resembles the demonstrated trajectories.
+On the other hand, the SEDS-based Cartesian motion generator (whose values here are simulated because they can not physically executable) quickly becomes unstable, as evidenced by the dotted paths starting on the left side and abruptly disappearing due to becoming too large.
+\label{fig:foot}}
+	\vspace{-24pt}
+\end{figure}
+
 One of the most important points when validating a learning from demonstration method is to evaluate the system's behavior away from demonstrations.
-When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target.
+When the current joint configuration is far from any of the local behavior regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\sum\limits_{k=1}^{K}\frac{A_k}{K}$ is used to move the robot, which is still guaranteed to move towards the target.
 Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably moving to the target.
 The reason for this is that rather than ``determining" the velocity of the system, our $A$ matrix only warps it.
-This means that adding multiple ``conflicting" synergies together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
+This means that adding multiple ``conflicting" behaviors together amounts to nothing more than repeatedly warping the velocity, and still maintains the original stability property.
 
@@
 By utilizing the JT-DS method, even a fast system could reactively generate stable motions to its target.
+
+The JTDS algorithm can also be expanded to accommodate joint velocity limits (in addition to joint position limits) by scaling the controller's velocity down such that the highest-velocity component is still within the velocity limits.
+Specifically, given some set of velocity constraints for each joint $i$ s.t.
+$\dot{q}_i < \dot{q}_{i, max}$, we can design a new joint augmentation matrix $\mathcal{A}' = r \mathcal{A}$ where $r = \min \left ( 1, \max_i \left ( \frac{\dot{q}_{i, max}}{\dot{q}_i} \right ) \right )$.
+This velocity scalar would slow down the joint trajectory when the robot dynamics could not execute them, but would not change the position profile of the joint motion, preserving a major part of the joint behavior.
 
@@
 There is a simple fix, however: rather than providing a single task-space objective $x^*$, provide a series of sequential task-space objectives $\begin{bf}x^*\end{bf} = \{ x^*_0, x^*_1, \dots, x^*_n\}$ and have the controller move to each in succession.
-The smaller sub-trajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits.
+The smaller subtrajectories give the robot less of an opportunity to deviate from the learned (valid) joint behavior and thus help it avoid being caught by the robot's joint limits.
 Further, specifying a sequence of task-space targets provides greater control over the robot's task-space trajectory while still exhibiting the learned joint-space behaviors.
@@
 
-\begin{figure}[t]
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=\linewidth]{Pic/cropped_Sing_1.pdf}
-		\caption{ }
-		\label{fig:Sin}
-	\end{subfigure}\\
-	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
-		\caption{ }
-		\label{fig:Det}
-	\end{subfigure}\\
-	\caption{The experiment generated $K=3$ Gaussian components.
-As the first joint is the only joint which was not fixed during the demonstration, the learned augmentation matrices had only one nonzero entry $A_k(1,1)\neq0~\forall k\in \{1, 2, 3\} $.
-In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space.
-The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).
-In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
-Due to sensory noise, the determinants are not exactly zero, but are very close.
-}
-	\vspace{-10pt}
-\end{figure}
+Finally, we are currently working on improving the performance of this method \eqref{eq:ds} by considering a the second order dynamical system.
+This would provide numerous improvements: a guaranteed smoothness in the velocity profile, the ability to integrate critically damped filter, and the versatility to learn more complex behaviors,.
+%\footnotesize
+\appendices
 
-\footnotesize
-\appendices
 \section{Proving Stability of the Dynamical System}
 \label{appendix:stability}
-\small
 We wish to prove Proposition \ref{prop:stability}, that is, that JT-DS \eqref{eq:ds} (reproduced below)
 \begin{equation*}
-\dot{q} = f(q) = -A(q)J^T(q)P(H(q) - x^*)
+\dot{q} = f(q) = -\mathcal{A}(q)J^T(q)P(H(q) - x^*)
 \end{equation*}
-accomplishes criterion (I), where $A$ is positive semi-definite, and $P$ is positive definite.
+accomplishes criterion (I), where $\mathcal{A}$ is positive semi-definite, and $P$ is positive definite.
 \newtheorem{theorem}{Theorem}[section]
@@
 \frac{dV(q)}{dt} &= (H(q) - x^*)^TPJ(q)\dot{q}\\
-&= -(H(q) - x^*)^TPJ(q)A(q)J^T(q)P(H(q) - x^*)\\
-&= -(H(q) - x^*)^TPJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot J^T(q)P(H(q) - x^*) \leq 0
+&= -(H(q) - x^*)^TPJ(q)\mathcal{A}(q)J^T(q)P(H(q) - x^*)\\
+&= -(H(q) - x^*)^TPJ(q)S(q) \sum_{k=1}^{K}\underbrace{\theta_k(q)}_{0\leq}\underbrace{A_k}_{0\preceq}\\&~~~~~~ \cdot S^T(q)J^T(q)P(H(q) - x^*) \leq 0
 \end{aligned}
@@
 $ \| H(q)-x^*\| $ and $ \dot{q} $ are bounded.
+
+To prove the fulfillment of criterion (II), assume the contrapositive, that there exists some time $0<\tau$ such that $q^i_{max} < q^i(\tau) $.
+Because $\dot{q}^i$ given by \eqref{eq:ds} with respect to \eqref{eq:first_criteria_stability} and  \eqref{eq:first_criteria_joint}  is both bounded and continuous, we can invoke the intermediate value theorem to show that there must have been some $0<c<\tau$ s.t.
+$q^i(c) = q^i_{max}$.
+According to \eqref{eq:first_criteria_joint}, $s^i(q^i(c)) = 0$, which ensures that $\dot{q}^i(c) = f(q^i(c)) = 0$ and will be forever (by induction), which contradicts the assumption.
+The same proof holds for a $q^i(\tau) < q^i_{min} $.
+$\blacksquare$
+\section{Enforcing Joint Limits}
+\label{appendix:limits}
+We wish to find a joint-position-dependent diagonal matrix $S(q)$ which is bounded and satisfies:
+\begin{equation}
+\label{eq:last_criteria}
+\begin{cases}
+\begin{split}
+& s^{i}(q^i_{max}) = 0&&&\\
+& s^{i}(q^i_{min})= 0&&&
+\end{split}
+\forall i\in 1,\dots,m
+\end{cases}
+\end{equation}
+as previously described in Section \ref{sec:proposed_system}. We propose the using the following functions:
+\begin{equation}
+s^{i}(q) =
+1 - \left ( 2\frac{q^i - q^i_{min}}{q^i_{max} - q^i_{min}} - 1 \right ) ^{2d}  ~\forall i\in 1,\dots,m\\
+\end{equation}
+where $d$ is a positive integer. 
+This function is continuous and satisfies all of the above criteria.
+We can understand it holistically as follows: the ratio $\frac{q(v) - q_{min}}{q_{max} - q_{min}}$ maps the set of valid $q$ values to the range between $0$ and $1$.
+The rest maps these values to a parabola-like polynomial, such that at $q^i = q^i_{max}$ or $q^i_{min}$ the value is $0$, and elsewhere the value is positive (with the maximum value, $s^i = 1$, at $q^i = \frac{q_{max} - q_{min}}{2}$).
+Note as $d$ gets bigger, $S(q)$ is flatter away from the joint limits and drops off more sharply near them, which allows us to preserve a wider range of motion.
+We found $d=2$ to be an effective choice.\\
 

### 2018-02-22 paper/reply_reviews/reply_comments.tex
Latest changes

\documentclass[10pt,stdletter,dateno]{newlfm}
\usepackage{charter}

\usepackage{tikz}
\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

\makeatletter
\g@addto@macro{\ps@ltrhead}{%
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{0pt}%
}
\g@addto@macro{\ps@othhead}{%
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{0pt}%
}
\makeatother

\widowpenalty=1000
\clubpenalty=1000

\newsavebox{\Luiuc}
\sbox{\Luiuc}{%
	\parbox[b]{1.6in}{%
		\vspace{0.5in}
		\includegraphics[scale=0.25]
		{./EPFL-Logo-bitmap/JPG/EPFL-Logo-RVB-55.jpg}%
	}%
}%

\makeletterhead{Uiuc}{\Rheader{\usebox{\Luiuc}}}

\newlfmP{headermarginskip=-100pt}
\newlfmP{sigsize=0pt}
%\headermarginsize{0in}
\headermarginskip{.5in}
\unprbottom{1cm}

%\newlfmP{dateskipafter=30pt}
%\newlfmP{addrfromphone}
%\newlfmP{addrfromemail}
%\PhrPhone{Phone}
%\PhrEmail{Email}

\lthUiuc

%\namefrom{Seyed Sina Mirrazavi Salehian}
%\addrfrom{\small
%		Learning Algorithms and Systems Laboratory (LASA)\\
%       	\small \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
%       	\small Lausanne, Switzerland - 1015.
%}
%\phonefrom{\small +41 21 69 31060 }
%\emailfrom{\small nadia.figueroafernandez@epfl.ch}

%\addrto{%
%Faculty Search Committee\\
%Department of Something\\
%Some University\\
%University Address}
%\greetto{Dear Editor,}
%\closeline{With best regards,}
\begin{document}
\begin{newlfm}
\vspace{-2in}
\centerline{\textbf{Response to Reviewers: Manuscript ID 17-1142.1 entitled}} \centerline{\textbf{``Learning Augmented Joint-Space Task-Oriented Dynamical Systems:}}
\centerline{\textbf{A Linear Parameter Varying and Synergetic Control Approach" }}
First of all, we would like to thank all reviewers for the time they spent reviewing our paper. We greatly appreciate the comments, suggestions and opinions that we received. We have addressed all the necessary comments and suggestions in the attached manuscript accordingly.

\underline{Comments from Reviewer \#1:}
\begin{enumerate}
\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{Specific clarifications and typos:
\begin{itemize}
\item ..
\item ..
\end{itemize}}
\end{enumerate}

\underline{Comments from Reviewer \#2:}
\begin{enumerate}
\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{Specific clarifications and typos:
\begin{itemize}
\item ..
\item ..
\end{itemize}}

\end{enumerate}

\underline{Comments from Reviewer \#3:}
\begin{enumerate}
\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{``bla bla"}\\
\textcolor{blue}{\texttt{Response:} \small bla bla}\\

\item \textit{Specific clarifications and typos:
\begin{itemize}
\item ..
\item ..
\end{itemize}}

\end{enumerate}

\end{newlfm}

\end{document}

### 2018-02-22 paper/reply_reviews/reviews.tex
Latest changes

\documentclass[10pt,stdletter,dateno]{newlfm}
\usepackage{charter}

\makeatletter
\g@addto@macro{\ps@ltrhead}{%
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{0pt}%
}
\g@addto@macro{\ps@othhead}{%
  \renewcommand{\headrulewidth}{0pt}%
  \renewcommand{\footrulewidth}{0pt}%
}
\makeatother
\usepackage{color}
\usepackage{xcolor}
\usepackage{soul}
\widowpenalty=1000
\clubpenalty=1000

\newsavebox{\Luiuc}
\sbox{\Luiuc}{%
	\parbox[b]{1.6in}{%
		\vspace{0.5in}
		\includegraphics[scale=0.25]
		{./EPFL-Logo-bitmap/JPG/EPFL-Logo-RVB-55.jpg}%
	}%
}%
%\makeletterhead{Uiuc}{\Rheader{\usebox{\Luiuc}}}
\makeletterhead{Uiuc}{\Rheader{}}
\newlfmP{headermarginskip=-100pt}
\newlfmP{sigsize=0pt}
\newlfmP{sigskipbefore=50pt}
\headermarginskip{.5in}
%\newlfmP{dateskipafter=30pt}
%\newlfmP{addrfromphone}
%\newlfmP{addrfromemail}
%\PhrPhone{Phone}
%\PhrEmail{Email}

\lthUiuc

%\namefrom{Seyed Sina Mirrazavi Salehian}
%\addrfrom{\small
%		Learning Algorithms and Systems Laboratory (LASA)\\
%       	\small \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
%       	\small Lausanne, Switzerland - 1015.
%}
%\phonefrom{\small +41 21 69 31060 }
%\emailfrom{\small nadia.figueroafernandez@epfl.ch}

%\addrto{%
%Faculty Search Committee\\
%Department of Something\\
%Some University\\
%University Address}
\vspace{-300pt}
%\greetto{Dear Editor,}
%\closeline{With best regards,}
\begin{document}
\begin{newlfm}
\vspace{-2in}
Dear Ms. Nadia Figueroa,

Your paper

17-1142
"Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A
Linear Parameter Varying and Synergetic Control Approach"
by Yonadav Shavit, Nadia Figueroa, Seyed Sina Mirrazavi Salehian, Aude
Billard Submission for RA-L only Special Issue on Learning and Control for
Autonomous Manipulation Systems: the Role of Dimensionality Reduction submitted to the IEEE Robotics and Automation Letters (RA-L) has been reviewed by the Associate Editor and selected Reviewers. The reviews of the paper are attached.

On the basis of the reviewers' ratings and comments, we decided that
your paper cannot be published in the Letters in the present form.
However, you are encouraged to rewrite and resubmit a revised version
of your work addressing all editorial concerns.

Please notice that there is a STRICT deadline of one month (30 DAYS)
from today \hl{[i.e. March 18th, 2018]} for you to resubmit a revised version, along with a
Statement of Changes as a single pdf file indicating how comments by
the Editor and by reviewers have been addressed and a list of changes
made in the paper.

Please be aware that failure to meet the 30 days deadline implies that
the paper will be automatically moved to the Reject category. Also
notice that the next decision on this manuscript will be final, i.e.
either Accept for Publication or Reject. Unfortunately, the tight
timeline of the Letters will allow no extensions or exceptions to these
rules.

If you have any related concern, do not hesitate to contact me.

Sincerely,\\
Han Ding\\
Editor (Learning and Control for Autonomous Manipulation Systems: the
Role of Dimensionality Reduction)\\
IEEE Robotics and Automation Letters\\

\underline{Associate Editor Report:}

The paper presents a dynamical system in joint space that combines
different motor synergies to encode joint-space behaviors while
asymptotically converges to a desired position in the task space. The
paper is well written and organized. All the reviewers think the paper
is of good quality. \textcolor{red}{:) OF COURSE it is!} The reviewers’ comments are attached. Please
address the comments and improve the paper accordingly.

%\clearpage

\textbf{Reviewer(s)' Comments to Author:}\\

\underline{Reviewer: 4 (Review30069)}\\
This paper proposes a dynamical system (DS) that operates in joint
space and converges to a desired position in task space. In doing so,
the authors use a linear parameter varying system and use
dimensionality reduction to learn these parameters efficiently. I think
the paper is well written, experiments are well conducted and
explained, and the results are promising.

\textit{Detailed comments and
suggestions are as listed below:}
\begin{enumerate}
\item As mentioned by the authors, there are many reasons to plan in joint
space rather than task space. One of the important reasons to me seems
to be able to plan end-effector orientation, not just position.
Traditional movement primitives, for example DMPs, do not explicitly
encode planning in the SO(3) space of orientations. There is research
that addresses this question, for example [1], but they lose the
simplistic appeal of movement primitives. \hl{I think it would be a great
addition to this paper if the authors illustrate if their joint space
DS can be used to accomplish orientation sensitive tasks, and
generalize to end-effector orientations different that what was
demonstrated.}\textcolor{red}{[*]} \\

[1] Ude, Ale, et al. "Orientation in cartesian space dynamic movement
primitives." Robotics and Automation (ICRA), 2014 IEEE International
Conference on. IEEE, 2014.\\

\textcolor{red}{[*] We never explored this, did we? Nevertheless, I think we can address it. We have data for the orientation of the demonstrations so we can try it. If we do this and it works, we should not only mention [1] but also the work from Seungsu [2] and Sylvain [3] both of which learn motions in SE(3) but cannot prove stability.}\\
\textcolor{red}{[2] Gaussian mixture model for 3-dof orientations. S Kim, R Haschke, H Ritter. Robotics and Autonomous Systems 87, 28-37}\\
\textcolor{red}{[3] An approach for imitation learning on Riemannian manifolds
MJA Zeestraten, I Havoutis, J Silvério, S Calinon, DG Caldwell
IEEE Robotics and Automation Letters 2 (3), 1240-1247}\\

\item In the introduction, the authors stress that jacobian psuedo inverse
can be computationally expensive. Is this really true for a 7-dof robot
arm? Can the authors provide some numbers where this would really be an
issue for a robot controlled at 1000Hz? I find the argument that
jacobian inverse does not work for singularities and needs a set of
heuristics to be more compelling. Maybe the authors should rewrite the
introduction to stress more on this issue? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item In Section II, criterion I, the authors talk about a positive
symmetric matrix P. I do not see it in the equations above. Is this a
weighing matrix? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item In criterion II, the error is defined with respect to the velocity
error, rather than a position or a combination of both. This means that
the motion can be different from the demonstrated motion. I understand
that this is important to keep the later optimization convex, but is
there any disadvantage to using velocity error rather than position
error? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item I do not completely understand the significance of A(q). From Figure
3, it seems like A(q) helps resolve redundancy, while achieving the
task space goal. Do the authors have a mathematical (in constrast to
intuitive) explanation for this? Does A(q) somehow relate to the
null-space of the jacobian? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item In Section III, "One can intuitively understand the control law as
follows:$ (H(q)-x^*)$ denotes the position error w.r.t. the task target
and by multiplying that error by the transposed Jacobian $J^T(q)$, the
error is projected into joint space (similar to Jacobian transpose
control)". I find this explanation misleading as jacobian multiplied to
position error does not project the error into joint space. Jacobian
transforms velocities, not positions. Maybe the law can be seen more as
a first order approximation of the error in task space. In any case, I
would either clarify this statement more or take it out of the
manuscript all together. It seems like the DS is just a PD law that
provably converges to a fixed attractor in task space. \textcolor{red}{[Can be addressed through clarification and re-write]}

\item About figure 2 again, the position error in task and joint space are
not simply multiplied by a Jacobian. Although, task space error can be
first-order approximated using a Jacobian multiplication with joint
space error. What does the right figure in Fig 2 exactly represent? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item In Section IV, the authors say that they use EM to determine the
parameters of the GMM in Eq 6. Can they elaborate a little on this?
What is the expectation taken over and what likelihood is maximized? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item The learned dimensionality reduction and parameters seem task
dependant. I wonder if the authors have any idea on how to generalize
this across tasks. \textcolor{red}{[Can be addressed through clarification and re-write]}

\item \hl{The experiments have the same task-space goal as the
demonstrations. Can the authors include experiments that have goals
different from demonstrated, especially in orientation?} \textcolor{red}{[We already know that it won't generalize well if the target is in a completely different hemisphere of the workspace of the robot, however, we can try adding some tests and show how well it can generalize within the same region that it was demonstrated]}

\item \hl{In similar light, maybe authors could present online adaptation of
A(q) to generalize to situations different than demonstrated, for
example a slightly moved obstacle?} \textcolor{red}{[This is a really nice suggestion. But how I see it, it's basically a new paper; i.e. an extension of this one, but we can discuss about this if we have space. A quick idea that comes to mind would be to apply a modulation/activation to the LPV system in a local region. Now this is not as straight forward as in our task space approaches because our LPV system is learned in joint-space while the activation function is in the embedded space. So it still would be an open question, but an approach to tackle this problem could be along these lines.]}

\item In table II, instead of the tracking error, maybe the more useful
error to see would be the error with respect to the plan or
demonstation. One would assume that the proposed approach would be
worse at it than SEDS+IK because the minimized error is with respect to
the velocity and goal, not the position of the movement. \textcolor{red}{[This will only make sense for the tasks that are actually feasible for the IK-solver (i.e. not for foot-step or singularity). We can easily compute this and add it. However, we know already that the error will be worse than SEDS+IK, so would this make sense? ]}

\end{enumerate}

Overall, I think this is an interesting, well written paper. I think
adding experiments with different goal - position and orientation will
help the paper, as well as an analysis of tracking error with respect
to a plan or demonstration.\\

\underline{Reviewer: 5 (Review30371)}

This paper deals with the problem of controlling a manipulator towards
a desired attractor in the task space by exploiting local synergies
defined in the joint space and combined in time and space in order to
obtain desired behaviours. This method does not require the computation
of the pseudo-inverse Jacobian and hence is able to pass through
singular configurations. The most similar approach is the Jacobian
transpose control method that shares the same advantages but that does
not include the learning of behaviours from demonstrations that seems
to bring some benefits.

The paper is well written and quite easy to follow. The cited
literature is also appropriated and complete.

\textit{However, some criticisms follows:}
\begin{enumerate}
\item \hl{My first concern regards you choice of representing the demonstrated
movement as a sequence of postures and hence looking for synergies by
using PCA of K-PCA analysis. Why not using functional PCA (f-PCA)
analysis and determine the few moving synergies that make possible the
execution of the demonstrated behaviours? It would be that, when you
combine the local synergies in space and time by using functions
$\theta_k(q)$, you get what you might obtain by using directly f-PCA.
However, the last one are valid for the whole movement and can be
combined by constant coefficients.} \textcolor{red}{[Honestly, I hadn't heard of functional-PCA until now. I searched a bit and I see why the reviewer would suggest this approach (which is interesting btw). F-PCA is the application of PCA on a functional space (mainly targeted at applications where each of the datapoints is a time-series). The goal is then to find the minimal set of eigen-time-series that best describes the shape of the observed time-series.
To do this, each time-series is approximated with k basis functions and PCA is applied on these functions. The resulting eigen-functions are parametrized by a linear combination of these k basis-functions (looks very similar to k-pca/svc). First of all, I see a couple of complications: i) the basis functions are time dependent; i.e. they will only work for time-series of the same length as the datasets that they were learned from ii) the number of k basis functions is an open parameter, as well as the number of eigenfunctions to keep iii) I can understand how the reviewer would think that it is almost the same as finding the lower dimensional space via PCA/k-PCA and then fitting a GMM, as this would be equivalent to the 'k' basis functions. However, I'm not sure if the 'k' basis functions in joint space are equivalent to the 'K' clusters discovered by the GMM in the embedded space. Which is extremely important because this dictates how many A's we need. I might be wrong about these issues, I need to dig deeper into this.]}

\item I also found the paper lacking of physical intuitions, especially in
the experimental part, about the entities introduced in previous
sections. In particular, what is the lesson we learn after your
analysis of synergies and the obtained lower dimensional space for the
few movements shown in section V? Is it possible to extract some global
synergic behavior?
Moreover, in section IV.A authors introduce the function phi(q) named
the embedding that maps the joint configuration into a lower
dimensional space. Is it associated to a whole behaviour or, as
matrices $A_k$, has a local validity? I lost this point. Can you
show/report or describe what it represents in the joint space for one
or more cases in Section V? What is this lower dimensional space for
the cases in Section V? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item A similar question for the matrices $A_k$ that represent local, I
would say postural (see previous comment), synergies whose dimension is
m x m, i.e. in joint space: I would expect they were defined in the
lower dimensional space. \textcolor{red}{[We actually address this in the discussion, it is related to the stability proof; i.e. we would loose rank when projecting the A matrix back to joint space]} Moreover, are they local representation of the
distribution of postures and hence covariance matrices? What do their
eigenvectors and eigenvalues locally represent? Maybe, a comparison
with the classical and more known postural synergies for the human
hand, analyzed by PCA, could be useful. Finally, what does it mean that
they should be positive definite (apart from stability issue)? \textcolor{red}{[All of these questions can be addressed through clarification and re-write]}

\item What is the dimension of function $\theta_k(q)$ that, to some extent,
can be seen as the new control variables. Is it $K<m$? \textcolor{red}{[Can be addressed through clarification and re-write]}

\item For the test, why not comparing also with the JT control method? In
other terms, is it possible to use a SEDS + JT, especially for the
third case "Transiting through singular configurations"?

\item Minors:
- SEDS + IK: what does SEDS mean? \textcolor{red}{[We clearly say what it means on page 6: ``For the Cartesian motion generator we
use the SEDS approach [8] which learns an asymptotically
stable DS in Cartesian space from demonstrations. We then
generate joint trajectories through a damped least square IK
solver. From herein we refer to this approach as SEDS+IK."]}

\end{enumerate}

\underline{Reviewer: 7 (Review30521)}

The paper presents a kinematic controller for a robot, that drives it
towards a fixed target in task space. The controller works in joint
space and consists in a dynamical system formulated as a LPV system
which combines different motor synergies. Such synergies are learnt
from demonstrations. Theoretical proof of stability is given and
experimental results are offered for validation.

The paper is very well written and nicely organized. Results are
interesting and the technical accuracy of the paper is good. The
coverage of the state of the art is quite good.

\textit{I have the following comments:}
\begin{enumerate}
\item a considerable part of the introduction is spent in comparing joint
space trajectory generation to task space trajectory generation. Since
these are elementary concepts in robotics, the authors might consider
reducing this part; \textcolor{red}{[Can be addressed through clarification and re-write]}

\item one of the main claims of the paper is the stability proof, which in
fact is cited in the first line of the abstract. However the stability
analysis is a quite simple extension of the stability of the transpose
Jacobian scheme, where the constant gain on the task space error is
replaced by the matrix A(q). As such the emphasis on this result might
be to some extent reduced. By the way, the proof works only if the
Jacobian is full rank, which is not properly discussed in Appendix I; \textcolor{red}{[Can (should) be addressed through clarification and re-write]}

\item the meaning of variables $\dot{q}_d$ in (2) is not very clear to me, a
well as whether it takes the same meaning as $\dot{q}_{t,n}$ in (8); \textcolor{red}{[Can be addressed through clarification and re-write]}

\item Section III uses the concept of synergy without actually defining it.
An introduction to this concept is given later, in Section IV; \textcolor{red}{[Can be addressed through clarification and re-write]}

\item the expression "flow of motion" in Proposition 1 is not very clear to
me; \textcolor{red}{[Can be addressed through clarification and re-write]}

\item some details about the optimization problem in (8) might be given
(nature of the problem, complexity, feasibility). \textcolor{red}{[Can be addressed through clarification and re-write]}
\end{enumerate}

\end{newlfm}
\end{document}

### 2018-02-22 paper/submitted-v1/main_RA-L_v1_final.tex
Latest changes

--- 2017-11-20 paper/main_RA-L_v1_final.tex
@@
 \centering
-\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{With_Obstcle.png}
+\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{../Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{../With_Obstcle.png}
 \caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl).
@@
 	\centering
-	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{./Pic/Sources/Schematic_final.pdf}
+	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
 	\caption{Illustration of our proposed Lyapunov function and error projection.
@@
 	\centering
-	\includegraphics[width=\linewidth]{Pic/A_comparison_positions.pdf}
+	\includegraphics[width=\linewidth]{../Pic/A_comparison_positions.pdf}
 	\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
@@
 		\centering
-			\includegraphics[width=0.5\linewidth]{new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{new-pics/backhand_crop.png}
+			\includegraphics[width=0.5\linewidth]{../new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{../new-pics/backhand_crop.png}
 					\caption{Forward/Backward Reaching (1-2).\label{fig:forback}}
@@
 \centering
-	\includegraphics[width=0.45\linewidth]{new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{new-pics/pour-obst_crop.png}
+	\includegraphics[width=0.45\linewidth]{../new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{../new-pics/pour-obst_crop.png}
 						\caption{ Pouring Motions (3-4)}
@@
 \centering
-	\includegraphics[width=0.45\linewidth]{new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{new-pics/footstep.png}
+	\includegraphics[width=0.45\linewidth]{../new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{../new-pics/footstep.png}
 	\caption{ Pouring Motion (5) and Foot-step (6).}
@@
 	\begin{subfigure}[t]{0.42\linewidth}
-		\includegraphics[width=0.45\linewidth]{Pic/Luggage1.png}
-		\includegraphics[width=0.143\linewidth]{Pic/Luggage2.png}
-		\includegraphics[width=0.377\linewidth]{Pic/Luggage3.png}
+		\includegraphics[width=0.45\linewidth]{../Pic/Luggage1.png}
+		\includegraphics[width=0.143\linewidth]{../Pic/Luggage2.png}
+		\includegraphics[width=0.377\linewidth]{../Pic/Luggage3.png}
 		\caption{Execution of Pour Obstacle 2 learned through JT-DS.}
@@
 	\begin{subfigure}[t]{0.567\linewidth}
-		\includegraphics[width=0.33\linewidth]{Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{Pic/Luggage3_SEDS.png}
+		\includegraphics[width=0.33\linewidth]{../Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{../Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{../Pic/Luggage3_SEDS.png}
 		\caption{Execution of Pour Obstacle 2 learned through SEDS.}
@@
 	\end{subfigure}\\	
-	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/rirQkK0SSso}{[video link]}.}
+	\caption{Snapshots of the robot experiments. A corresponding video is available on-line \href{https://youtu.be/mv9u5DgIEtw}{[https://youtu.be/mv9u5DgIEtw]}.}
 	\label{fig:snapshot}
@@
 		\centering
-		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{new-pics/singularity_crop.png}
+		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{../new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{../new-pics/singularity_crop.png}
 		\caption{Execution of \textit{problematic} joint-space behaviors learned through JT-DS: (left) Foot step-like motion and (right) Singularity motions.}
@@
 		\centering
-		\includegraphics[width=0.7\linewidth]{Pic/Foot.pdf}
+		\includegraphics[width=0.7\linewidth]{../Pic/Foot.pdf}
 		\caption{End-effector trajectories for the footstep motion in Cartesian space.
@@
 %	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=\linewidth]{Pic/cropped_Sing_1.pdf}
+		\includegraphics[width=\linewidth]{../Pic/cropped_Sing_1.pdf}
 %		\caption{ }

### 2018-02-22 src/JTDS_mat_lib/crossvalidation_dr_kuka.m

@@ -18,7 +18,7 @@ switch choosen_dataset
     case 'fore'
         demos_location = strcat(data_path,'fore_hand/data.mat');
         demo_ids = [1:11];
-    case 'pour'
+    case 'pour'synergy 
         demos_location = strcat(data_path,'pour_no_obst/data.mat');
         demo_ids = [1 2 3 5 6 7 8 9 10];
     case 'pour_obst'

### 2018-02-22 src/JTDS_mat_lib/utils/preprocess_demos_jtds.m

@@ -72,7 +72,7 @@ for i=1:length(demos)
     clear tmp_full tmpq tmpq_d tmpx tmpxd
     
     % de-noising data (not necessary)
-    tmpq = demos{i}; 
+    tmdiffepq = demos{i}; 
     
     % computing the first time derivative
     if ~iscell(time)

### 2018-03-09 src/JTDS_mat_lib/Testing_orientation.m
Data-sets are added

@@ -1,0 +1,193 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Evaluation of Learning Schemes for JTDS Models (including orientation) on Different Datasets     %%
+%  Train and compare a series of JTDS models with different dimensionality                          %%   
+%  reduction schemes and GMM model fitting                                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'pour_obst_2'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour' 
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'pour_obst_2'
+        demos_location = strcat(data_path,'pour_obst_2/data.mat');
+        demo_ids = [1:7];                
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
+        demos_location = strcat(data_path,'singularity/data.mat');
+        demo_ids = [1:10];  
[…]

### 2018-03-09 src/JTDS_mat_lib/JTDS_Solver_v2.m
The orientation is added!

@@ -342,7 +342,11 @@ function [Priors, Mu, Sigma, As, latent_mapping] = JTDS_Solver_v2(Data, robotpla
     Ss = zeros(dimq, dimq, n);
     for i = 1:n
         q = Q(:, i); xt = Xt(:, i);
+        if options.orientation_flag==1
+         Qd_basis(:, i) =  robotplant.qd_basis_orientation(q, xt);
+        else
         Qd_basis(:, i) = robotplant.qd_basis(q, xt);
+        end
         if options.learn_with_bounds % if false, ignore joint limits
             Ss(:,:,i) = robotplant.compute_S(q);% the S value for each point, converted to a vector
         else

### 2018-03-09 src/JTDS_mat_lib/Testing_orientation.m

@@ -6,8 +6,8 @@
 
 clear all; close all; clc;
 do_plots  = 1;
-data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'pour_obst_2'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %
@@ -15,25 +15,25 @@ choosen_dataset = 'pour_obst_2'; % Options: 'back','fore','pour','pour_obst','fo
 
 switch choosen_dataset
     case 'back'
-        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demos_location = strcat(data_path, 'back_hand/raw_data.mat');
         demo_ids = [2:11];
     case 'fore'
-        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demos_location = strcat(data_path,'fore_hand/raw_data.mat');
         demo_ids = [1:11];
     case 'pour' 
-        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demos_location = strcat(data_path,'pour_no_obst/raw_data.mat');
         demo_ids = [1 2 3 5 6 7 8 9 10];
     case 'pour_obst'
-        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demos_location = strcat(data_path,'pour_obst/raw_data.mat');
         demo_ids = [1:10];
     case 'pour_obst_2'
-        demos_location = strcat(data_path,'pour_obst_2/data.mat');
+        demos_location = strcat(data_path,'pour_obst_2/raw_data.mat');
         demo_ids = [1:7];                
     case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
-        demos_location = strcat(data_path,'foot/data.mat');
+        demos_location = strcat(data_path,'foot/raw_data.mat');
         demo_ids = [1:8];                
     case 'singularity'   
-        demos_location = strcat(data_path,'singularity/data.mat');
+        demos_location = strcat(data_path,'singularity/raw_data.mat');
         demo_ids = [1:10];  
[…]

### 2018-03-09 src/JTDS_mat_lib/model/RobotPlant.m

@@ -20,7 +20,7 @@ classdef RobotPlant
             % custom_qd_basis takes as input the object, q, and dt,
             % and outputs the basis for our velocity policy
             obj.robot = robot;
-                
+            
             if nargin<5 % If the number of arguments suggests standard non-custom kinematics
                 if strcmp(task_type, 'end_trans')
                     obj.forward_kinematics = @obj.end_trans_forward_kinematics;
@@ -45,6 +45,17 @@ classdef RobotPlant
             basis = -1*transpose(obj.jacobian(q))*(obj.forward_kinematics(q) - xt);
         end
         
+        function basis = qd_basis_orientation(obj, q, xt)
+            % Computes the base joint velocity to move at a joint position q
+            % toward a task position xt using the jacobian transpose
+            % method (i.e. qd = J^T(H(q) - xt) ).
+            % In general, this will be multiplied by an A matrix to get the
+            % augmented joint velocity.
+            temp_end=obj.robot.fkine(q);
+            temp_end=[temp_end(1:3,1);temp_end(1:3,2);temp_end(1:3,4)];
+            basis = -1*transpose([Jacobian_X(obj,q);Jacobian_Y(obj,q);obj.jacobian(q)])*(temp_end - xt);
+        end
+        
         function x = end_trans_forward_kinematics(obj,q)
             T = obj.robot.fkine(q);
             x = T(1:3, 4);
@@ -55,7 +66,7 @@ classdef RobotPlant
         end
         
         function output = compute_S(obj, q)
-            % Compute the joint limitation matrix S as described in the 
+            % Compute the joint limitation matrix S as described in the
             % appendix of the JT-DS paper. The diagonal values are always
             % between 0 and 1, and (when correctly integrated into the
             % DS) ensure that the system will not collide with its
@@ -68,6 +79,36 @@ classdef RobotPlant
             end
             s = 1- (2*(q - Qmin)./(Qmax - Qmin) - 1).^4;
             output = diag(s);
+        end
+        
+        function J=Jacobian_X(obj,q)
+            % Compute the jacobian matrix along a specific axis
+            J=zeros(3,obj.robot.n);
+            H0F=obj.robot.fkine(q);
+            v2 = H0F(1:3,1);
+            Hoi{1}=obj.robot.A(1,q);
+            for i=2:obj.robot.n
[…]

### 2018-03-09 src/JTDS_mat_lib/utils/preprocess_demos_jtds.m

@@ -1,9 +1,9 @@
-function [Data, index] = preprocess_demos_jtds(robotplant, demos, time, tol_cutting)
+function [Data, index] = preprocess_demos_jtds(robotplant, demos, time, tol_cutting,orientation_flag)
 % This function preprocesses raw joint demonstration data and molds it into
 % a suitable format.
 % The function computes the first time derivative of demonstrations and
 % trims the data. The function can be
-% called using: 
+% called using:
 %
 %          [Data, index] = preprocess_demos(demos,time,tol_cutting)
 %
@@ -47,7 +47,7 @@ function [Data, index] = preprocess_demos_jtds(robotplant, demos, time, tol_cutt
 %              2*qdim + 1:2*qdim + xdim = target position
 %
 %              Each column of Data stands
-%              for a datapoint. All demonstrations are put next to each other 
+%              for a datapoint. All demonstrations are put next to each other
 %              along the second dimension. For example, if we have 3 demos
 %              D1, D2, and D3, then the matrix Data is:
 %                               Data = [[D1] [D2] [D3]]
@@ -57,7 +57,7 @@ function [Data, index] = preprocess_demos_jtds(robotplant, demos, time, tol_cutt
 %              that columns 1:T1-1 belongs to the first demonstration,
 %              T1:T2-1 -> 2nd demonstration, and T2:T3-1 -> 3rd
 %              demonstration.
-    
+
 %checking if a fixed time step is provided or not.
 if length(time)==1
     dt = time;
@@ -72,7 +72,7 @@ for i=1:length(demos)
     clear tmp_full tmpq tmpq_d tmpx tmpxd
     
     % de-noising data (not necessary)
-    tmdiffepq = demos{i}; 
+    tmpq = demos{i};
     
     % computing the first time derivative
     if ~iscell(time)
@@ -83,7 +83,7 @@ for i=1:length(demos)
     
     % trimming demonstrations
     ind = find(sqrt(sum(tmpq_d.*tmpq_d,1))>tol_cutting);
-    tmpq = tmpq(:,min(ind):max(ind)+1);   
+    tmpq = tmpq(:,min(ind):max(ind)+1);
     
     if iscell(time)
         tmpt = time{i}(:,min(ind):max(ind)+1);
@@ -91,8 +91,12 @@ for i=1:length(demos)
     tmpq_d = tmpq_d(:,min(ind):max(ind));
     
[…]

### 2018-03-11 Data/mat/singularity/untitled9.m
THe correct data-set is added

@@ -1,0 +1,139 @@
+clc
+clear
+close all
+
+figure1 = figure;
+subplot1 = subplot(1,8,[1 2 3 4 5],'Parent',figure1);
+% Create axes
+% axes1 = axes('Parent',subplot1);
+hold(subplot1,'on');
+xlabel(' X [m]','Interpreter','latex');
+
+% Create zlabel
+zlabel('Z [m]','Interpreter','latex');
+
+% Create ylabel
+ylabel('Y [m]','Interpreter','latex');
+
+view(subplot1,[-37.5 30]);
+grid(subplot1,'on');
+% Set the remaining axes properties
+set(subplot1,'FontSize',24,'TickLabelInterpreter','latex');
+data = [];
+for j=1:10
+    Data=importfile(sprintf('TheRobotTrajectory%d.txt',j));
+    data{j} = [Data(:,4),Data(:,5),Data(:,6),Data(:,7),Data(:,8),Data(:,9),Data(:,10)]';
+    DET=zeros(1,size(Data,1));
+    for i=1:size(Data,1)
+        Jacobian=[Data(i,15),Data(i,16),Data(i,17);
+            Data(i,18),Data(i,19),Data(i,20);
+            Data(i,21),Data(i,22),Data(i,23)];
+        DET(i)=det(Jacobian);
+    end
+    demo{j}=[Data(1:10:end,1),Data(1:10:end,2),Data(1:10:end,3)]';
+    A=sprintf('$q^{%d}=%d^{\\circ}$',j,120-10*j);
+    h3=plot3(Data(1,1),Data(1,2),Data(1,3),...
+    'MarkerFaceColor',[0.0705882385373116 0.211764708161354 0.141176477074623],...
+    'MarkerEdgeColor','none',...
+    'MarkerSize',16,...
+    'Marker','pentagram',...
+    'LineStyle','none');
+    h4=plot3(Data(end,1),Data(end,2),Data(end,3),...
+    'MarkerFaceColor',[0.749019622802734 0 0.749019622802734],...
+    'MarkerEdgeColor','none',...
+    'MarkerSize',16,...
+    'Marker','hexagram',...
+    'LineStyle','none');
+
+   h1= plot3(Data(:,1),Data(:,2),Data(:,3),'LineWidth',3,'LineStyle','--',...
+    'Color',[0 0.447058826684952 0.74117648601532]);
+   h2= plot3(Data(:,1),Data(:,2),Data(:,3),'LineWidth',1,...
[…]

### 2018-03-11 src/JTDS_mat_lib/Testing_orientation.m

@@ -15,25 +15,25 @@ choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','si
 
 switch choosen_dataset
     case 'back'
-        demos_location = strcat(data_path, 'back_hand/raw_data.mat');
+        demos_location = strcat(data_path, 'back_hand/data.mat');
         demo_ids = [2:11];
     case 'fore'
-        demos_location = strcat(data_path,'fore_hand/raw_data.mat');
+        demos_location = strcat(data_path,'fore_hand/data.mat');
         demo_ids = [1:11];
     case 'pour' 
-        demos_location = strcat(data_path,'pour_no_obst/raw_data.mat');
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
         demo_ids = [1 2 3 5 6 7 8 9 10];
     case 'pour_obst'
-        demos_location = strcat(data_path,'pour_obst/raw_data.mat');
+        demos_location = strcat(data_path,'pour_obst/data.mat');
         demo_ids = [1:10];
     case 'pour_obst_2'
-        demos_location = strcat(data_path,'pour_obst_2/raw_data.mat');
+        demos_location = strcat(data_path,'pour_obst_2/data.mat');
         demo_ids = [1:7];                
     case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
-        demos_location = strcat(data_path,'foot/raw_data.mat');
+        demos_location = strcat(data_path,'foot/data.mat');
         demo_ids = [1:8];                
     case 'singularity'   
-        demos_location = strcat(data_path,'singularity/raw_data.mat');
+        demos_location = strcat(data_path,'singularity/data.mat');
         demo_ids = [1:10];  
         fprintf('Loading demonstrations from %s \n', demos_location);
         load(demos_location)

### 2018-03-11 paper/reply_reviews/reply_comments.tex
changes in paper

@@
 
+\item \textit{``In Section II, criterion I, the authors talk about a positive symmetric matrix P. I do not see it in the
+equations above. Is this a weighing matrix?"}\\
+\textcolor{blue}{\texttt{Response:} \small Removed, that was a carry-over from a previous formulation.}\\
+
+\item \textit{``In criterion II, the error is defined with respect to the velocity error, rather than a position or a combination
+of both. This means that the motion can be different from the demonstrated motion. I understand
+that this is important to keep the later optimization convex, but is there any disadvantage to using
+velocity error rather than position error?"}\\
+\textcolor{blue}{\texttt{Response:} \small Optimizing for position distance alone would have the effect of learning a motion that converged to the average of previous motions, thus not changing the motion in new situations or fitting the joint velocity profile that is often a desired component of the motion.
+Using velocity error does mean that, for motions strongly dependent on joint position rather than joint velocity, the generated motions may generalize from the demonstrations in unexpected ways.
+However, as you mentioned, minimizing position error would, depending on the formulation, be prohibitively difficult.
+As I see it, that would mean simulating the learned motion forward in time to recompute the robot's positions, making the optimization much more complicated; this way, we can treat each point in the motion independently.}\\
+
+\item \textit{`I do not completely understand the significance of A(q). From Figure 3, it seems like A(q) helps resolve
+redundancy, while achieving the task space goal. Do the authors have a mathematical (in constrast to
+intuitive) explanation for this? Does A(q) somehow relate to the null-space of the jacobian?"}\\
+\textcolor{blue}{\texttt{Response:} \small A(q) transforms a joint velocity vector approximately aligned with the task-space error into a different joint velocity vector that remains approximately aligned with the error, but which we want to choose to more-closely approximate our demonstrations.
+In this sense, it is resolving redundancy between all the possible joint velocity vectors that would move the controller closer to the origin.
+The paragraph in Section III has been modified to improve clarity.
+A(q) doesn't relate to the null-space of the Jacobian in any clear way, as its value can and does alter the task-space trajectory as well as the joint space trajectory.
+The only property that A(q) preserves is the joint velocity's moving the task space position closer to the goal, but the direction of motion is not necessarily the same as the direction of task-space error.}\\
+
+\item \textit{``In Section III, ?One can intuitively understand the control law as follows:(H(q) ?? x) denotes the
+position error w.r.t. the task target and by multiplying that error by the transposed Jacobian JT (q),
+the error is projected into joint space (similar to Jacobian transpose control)?. I find this explanation
+misleading as jacobian multiplied to position error does not project the error into joint space. Jacobian
+transforms velocities, not positions. Maybe the law can be seen more as a first order approximation
+of the error in task space. In any case, I would either clarify this statement more or take it out of the
+manuscript all together. It seems like the DS is just a PD law that provably converges to a fixed attractor
+in task space."}\\
+\textcolor{blue}{\texttt{Response:} \small The passage has been rewritten to remove the word "project", which you appropriately noted was incorrect, and further clarifies the nature of the DS as a PD controller.
+The new text is reproduced below: \\
+One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
+the task target, which we then reinterpret as the task-space velocity of a proportional controller.
+By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control [?]), see Fig.
+[?].
+The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
+[?] illustrates the effects of  $\mathcal{A}(q)$ on the generated motion.
+Thus the controller can be thought of as a proportional controller in joint space.}\\
+
+\item \textit{``About figure 2 again, the position error in task and joint space are not simply multiplied by a Jacobian.
+Although, task space error can be first-order approximated using a Jacobian multiplication with
+joint space error. What does the right figure in Fig 2 exactly represent?"}\\
+\textcolor{blue}{\texttt{Response:} \small The figure on the left is simply the cartesian error in task space.
+The figure on the left is the Lyapunov potential function on which the jacobian transpose controller descends.
+In showing the attractive regions in joint-space corresponding to this Lyapunov function, we aim to offer insight into how the controller reformulates the task-space potential into a different yet related joint-space potential.
+The caption has been rewritten to add clarity.}\\
+
+\item \textit{``In Section IV, the authors say that they use EM to determine the parameters of the GMM in Eq 6. Can
+they elaborate a little on this? What is the expectation taken over and what likelihood is maximized?"}\\
+\textcolor{blue}{\texttt{Response:} \small The algorithm computes the expected probability of generating each of the joint positions given the current values of the GMM's means, variance matrices, and scaling values.
+The algorithm then chooses GMM parameters to maximize the posterior likelihood of the model, using previously-computed expected component membership as likelihood weights when computing the new mixture.
+Specifically, we used MATLAB's \textit{fitgmdist} implementation.
+Added a citation to clarify.}\\
+
+\item \textit{``The learned dimensionality reduction and parameters seem task dependant. I wonder if the authors
+have any idea on how to generalize this across tasks."}\\
+\textcolor{blue}{\texttt{Response:} \small The dimensionality reduction could be generalized across tasks that had the same motor synergies - that is, for example, all tasks in which the shoulder and elbow always move together.
+Similarly, one could imagine several similar motions that could be decomposed into shared submotions, thus allowing the parameters to be shared.
+However, in general, fixing some of the parameters while optimizing will simply lead to a less-flexible model, and considering that there is no intuitive scheme for a-priori identifying motions with strongly similar submotions, we did not explore this question further.
+To answer your question directly, if one wanted to, one could reuse the learned dimensionality reduction for a new task which you believed to be similar, or copy certain components of a previous motion's GMM into the new GMM and leave them fixed throughout the optimization.}\\
+
+\item \textit{``bla bla"}\\
+\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+
+\item \textit{``bla bla"}\\
+\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+
 \item \textit{``bla bla"}\\
@@
 \textcolor{blue}{\texttt{Response:} \small bla bla}\\
+
+\item \textit{``I also found the paper lacking of physical intuitions, especially in	the experimental part, about the entities introduced in previous
+	sections. In particular, what is the lesson we learn after your
+	analysis of synergies and the obtained lower dimensional space for the
+	few movements shown in section V? Is it possible to extract some global
+	synergic behavior?
+	Moreover, in section IV.A authors introduce the function phi(q) named
+	the embedding that maps the joint configuration into a lower
+	dimensional space. Is it associated to a whole behaviour or, as
+	matrices $A_k$, has a local validity? I lost this point. Can you
+	show/report or describe what it represents in the joint space for one
+	or more cases in Section V? What is this lower dimensional space for
+	the cases in Section V?"}\\
+\textcolor{blue}{\texttt{Response:} \small The result of the analysis in Section V suggests that the algorithm performs much better (lower velocity reconstruction error) and requires far fewer synergies, when the dimension of the motion hs been reduced.
+The paragraph explaining this has been clarified, and is reproduced below:\\
+"As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
+This suggests that using DR to encode our activation functions $\theta_k$ in a lower-dimensional space $\phi(q)$ yields better generalization capabilities than encoding the behaviors in using the original $q$.
+This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learned without DR is an order of magnitude higher than with DR.
+Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the higher number of $K$ needed to represent the motion without DR.
+For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
+By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
+Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations." \\
+Regarding the second half of the comment, $\phi(q)$ is a learned transformation, which is associated with a whole behavior, rather than a particular $A_k$ (as is now clarified in IV.A).
+However, the precise nature of the transformation depends on the DR technique: PCA applies a uniform matrix multiplication of joint position everywhere, whereas KPCA is a more complicated function of joint position.
+The result of the embedding is a lower-dimensional representation of the joint position.
+The following line has been added to IV.A for clarity:\\
+"For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the "shoulder" and "arm" components of $q$ into a single "shoulder-arm" component in $\phi(q)$.}"\\
+
+\item \textit{``A similar question for the matrices $A_k$ that represent local, I
+	would say postural (see previous comment), synergies whose dimension is
+	m x m, i.e. in joint space: I would expect they were defined in the
+	lower dimensional space. Moreover, are they local representation of the
+	distribution of postures and hence covariance matrices? What do their
+	eigenvectors and eigenvalues locally represent? Maybe, a comparison
+	with the classical and more known postural synergies for the human
+	hand, analyzed by PCA, could be useful. Finally, what does it mean that
+	they should be positive definite (apart from stability issue)?"}\\
+\textcolor{blue}{\texttt{Response:} \small Regarding your first comment, we address why defining $A_k$ in lower-dimensional space is problematic.
+Specifically, we would lose rank when projecting the A matrix back to joint space, invalidating our stability proof.
+The $A_k$ matrices are }\\
+
+\item \textit{``What is the dimension of function $\theta_k(q)$ that, to some extent,
+	can be seen as the new control variables. Is it $K<m$?"}\\
+\textcolor{blue}{\texttt{Response:} \small As mentioned in Sec.
+V.A., we treat K as a hyperparameter and thus choose it differently for each learned motion.
+For motions that do not vary much throughout the course of the motion, smaller K might be sufficient, but K can just as easily be greater than $m$ (though as $K$ increases, so does the learning problem's computational cost).
+If your question was about each specific $\theta_k(q)$, they are scalars, as is mentioned in Sec. III.}\\
 
@@
 \begin{enumerate}
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``a considerable part of the introduction is spent in comparing joint space trajectory generation to task
+space trajectory generation. Since these are elementary concepts in robotics, the authors might consider
+reducing this part"}\\
+\textcolor{blue}{\texttt{Response:} \small Several sentences have been removed/reduced.}\\
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``one of the main claims of the paper is the stability proof, which in fact is cited in the first line of the
+abstract. However the stability analysis is a quite simple extension of the stability of the transpose
+Jacobian scheme, where the constant gain on the task space error is replaced by the matrix A(q). As
+such the emphasis on this result might be to some extent reduced. By the way, the proof works only
+if the Jacobian is full rank, which is not properly discussed in Appendix I"}\\
+\textcolor{blue}{\texttt{Response:} \small Removed a couple mentions of the stability result, and added the dependence on the rank of the Jacobian to both the appendix and the main body of the text.}\\
+
+\item \textit{``the meaning of variables $q_d$ in (2) is not very clear to me, a well as whether it takes the same meaning
+as $q_{t;n}$ in (8)"}\\
+\textcolor{blue}{\texttt{Response:} \small $q_d$ are the velocities from the demonstrations, which we are aiming to recreate.
+Equation (8) has been updated to remove the typo (they are in fact the same).}\\
+
+\item \textit{``Section III uses the concept of synergy without actually defining it. An introduction to this concept is given later, in Section IV"}\\
+\textcolor{blue}{\texttt{Response:} \small Added the following to Section 3:\\
+Each ``synergy" is a joint-space transformation that biases the resulting motion to use particular joints; for more, see Sec. IV.}\\
+
+\item \textit{``the expression "flow of motion" in Proposition 1 is not very clear to me"}\\
+\textcolor{blue}{\texttt{Response:} \small Changed to "motion".}\\
 

### 2018-03-11 paper/reply_reviews/reviews.tex
changes in paper

@@
 \documentclass[10pt,stdletter,dateno]{newlfm}
+\usepackage{etoolbox}
 \usepackage{charter}
@@
 
-\textcolor{red}{[*] We never explored this, did we?
-Nevertheless, I think we can address it.
-We have data for the orientation of the demonstrations so we can try it.
-If we do this and it works, we should not only mention [1] but also the work from Seungsu [2] and Sylvain [3] both of which learn motions in SE(3) but cannot prove stability.}\\
+\textcolor{red}{[*] Honestly, we never explored this.
+I think we can address it, we would need to just use the Jacobian for the full end-effector pose, not just the position.
+Also, we have full joint and position+orientation of the end-effector for all the demonstrations, so we can definitely try it.
+If we do this and it works, we should not only mention [1] but also the work from Seungsu [2] and Sylvain [3] both of which learn motions in SE(3) but cannot prove stability.
+And I agree with the reviewer that it would be an even stronger contribution.}\\
 \textcolor{red}{[2] Gaussian mixture model for 3-dof orientations. S Kim, R Haschke, H Ritter. Robotics and Autonomous Systems 87, 28-37}\\

### 2018-03-11 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
changes in paper

--- 2018-02-22 paper/submitted-v1/main_RA-L_v1_final.tex
@@
 \begin{abstract}
-In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while stably converging towards a task-space attractor.
+In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a task-space attractor.
 % Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.
@@
 \cite{Pastor:ICRA:2009,Calinon:ISR:2015}.
-These approaches generally learn probabilistic model of the demonstrated motions.
+These approaches generally learn a probabilistic model of the demonstrated motions.
 %, and construct a dynamical system that chooses behaviors according to the learned model.
-Both %joint and task 
-spaces are independently useful:  task-space behaviors let us control the most task-relevant component of the robot's behavior.
-In particular, a number of approaches have pioneered stable convergence to an attractor in task space % \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015}
+%Both joint and task spaces are independently useful:  task-space behaviors let us control the most task-relevant component of the robot's behavior. 
+A number of approaches have further proposed stable convergence to an attractor in task space % \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015} 
  \cite{khansari2011learning, Neumann:RAS:2015}as a desirable capability for a dynamical system.
@@
 However, for other applications, such an approach yields significant problems \cite{buss2004introduction}.
-First of all, finding the pseudo-inverse is computationally taxing.
+First of all, finding the pseudo-inverse is computationally taxing \textbf{insert how slow it is for a 1000Hz robot}.
 Moreover, when the Jacobian matrix cannot be inverted (i.e.
@@
 In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
-To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different provably stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system.
+To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system.
 We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
@@
 \begin{enumerate}
-\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified,``stability" in this paper always refers to asymptotic stability within the workspace of the robot, and assumes no joint limits.
+\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified, ``stability" in this paper always refers to asymptotic stability within the workspace of the robot (in the regions where the Jacobian is full rank), and as such assumes no joint limits.
 We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system.} with respect to a fixed task-space target $x^*$.
@@
 is stable; i.e.
-$\dot{V}(q)\leq 0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$.
-Where  $Q=\{q|q^i_{min} < q^i < q^i_{max},~\forall i\in \{1,\dots,d\}\}$ and $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$.
-$P\in \mathbb{R}^{d\times d} $ is a symmetric and positive definite matrix.
+$\dot{V}(q) < 0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$ where $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$ and $Q=\{q|q\notin Q^*, J(q) \textit{ is full rank}\}$.
 $V(.)$ can be thought of as a metric for the task-space distance-to-go.
@@
 \end{equation}
-  where $\dot{q}_d$ is the desired ``true" velocity, and $f(\cdot)$ is the motion generation policy.
+  where $\dot{q}_d$ are the ``true" velocities from the demonstrations, and $f(\cdot)$ is the motion generation policy.
 \end{enumerate}
@@
 
-The error metric \eqref{eq:optim_error} is advantageous in a way that it mimics the magnitude and direction of the demonstrated motions.
+The error metric \eqref{eq:optim_error} is advantageous in that it mimics the magnitude and direction of the demonstrated motions.
 Hence, as suggested by \cite{gielniak2010stylized}, it is visually most similar to the human definition of ``joint motion style".
@@
 where  $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the LPV system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
+Each ``synergy" is a joint-space transformation that biases the resulting motion to use particular joints; for more, see Sec.
+\ref{Sec:Learning}.
+Explicitly,
 \begin{equation}
@@
 \end{equation}
-where $A_k\in \mathbb{R}^{m\times m} $ are the different synergies and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{%The scheduling parameters can be a function of time $t$, the states of the
+where $A_k\in \mathbb{R}^{m\times m} $ are the different synergy matrices and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{%The scheduling parameters can be a function of time $t$, the states of the
 %system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. 
@@
 One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
-the task target and by multiplying that error by the transposed Jacobian $J^T(q)$, the error is projected into joint space (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig.
+the task target, which we then reinterpret as the task-space velocity of a proportional controller.
+By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig.
 \ref{fig:error_schematic}.
@@
 \label{prop:stability}
-The flow of motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ \eqref{eq:ds} meets the following constraints. 
+The motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ \eqref{eq:ds} meets the following constraints. 
 \begin{equation}
@@
 \begin{split}
- & 0 \prec A_k, & ~~~~ 0 \leq \theta_k(\cdot) \\
+ & A_k \succ 0, & ~~~~ \theta_k(\cdot) \geq 0 \\
 \end{split}
@@
 	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
-	\caption{Illustration of our proposed Lyapunov function and error projection.
-On the left we show the task-space potential function that we are trying to descend on for a 2-DOF robot with target $x^*=[2,2]^T$.
-On the right we show the corresponding Lyapunuv function in joint space \eqref{eq:Lyp}.
-The colors on each plot correspond to each other, one can see how the error, and consequently the \textit{attractive} regions, in task-space are warped in joint-space.}
+	\caption{Illustration of our task-space potential function and proposed joint-space Lyapunov function.
+On the left we show the task-space error we are trying to minimize, represented by a potential function for a 2-DOF robot with target $x^*=[2,2]^T$.
+On the right we show the corresponding Lyapunov function in joint space defined in \eqref{eq:Lyp}.
+The colors on each plot correspond to each other.
+One can see how the error, and consequently the \textit{attractive} regions in task-space, have been warped into attractive regions in joint-space.}
 	\label{fig:error_schematic}
@@
 In this work, we assume that configurations that are nearby in joint-space should exhibit similar behaviors, and thus it is natural that we choose a low-dimensional embedding that preserves the variance in our demonstrations.
-To this end, we construct an embedding $\phi(q)$ by training dimensionality reduction techniques on the demonstrated trajectories.
+To this end, we construct a single global embedding $\phi(q)$ by training dimensionality reduction techniques on the demonstrated trajectories.
 The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^d$ into a lower-dimensional configuration $z \in \mathbb{R}^\delta$, where $\delta < d$.
+For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the ``shoulder" and ``arm" components of $q$ into a single ``shoulder-arm" component in $\phi(q)$.
 %Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}. 
@@
 Therefore, each synergy region is associated with a Gaussian component of the GMM, cumulatively describing all the synergy regions of the dynamical system.
-We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM.
+We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM \textbf{ADD CITATION FOR EM}.
 
@@
 &\underset{A_1, \dots, A_K}{\text{min}} 
-\sum_{n=1}^N\sum_{t=0}^{T_n} \left \| \dot{q}_{t,n} - f(q_{t,n}) \right \| \\
+\sum_{n=1}^N\sum_{t=0}^{T_n} \left \| \dot{q}_{d;t,n} - f(q_{t,n}) \right \| \\
 & \text{subject to}&\\
@@
 We present the \texttt{mean(std)} for the optimal $\delta$, $K$ and joint-velocity RMSE on training and testing set, found for every learning scheme over 10 runs.
-$M = \sum_{n=1}^{N}T_n$.\label{tab:learning}}
+$M = \sum_{n=1}^{N}T_n$.\label{tab:learning}}.
     \end{minipage}
@@
 Once these optimal hyper-parameters are estimated for each fold, we solve for the convex optimization problem in order to find our synergy matrices $A$'s which best minimize the objective function \eqref{eq:opt}.
-We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}
+We then select the initial joint configuration $q_0$ and target in task-space $x^*$ from each training/testing dataset, simulate the joint-space trajectories $\{\dot{q_1},\dots,\dot{q_T}\}$, with our learned $\mathcal{A}(q)$, and compute the joint-velocity RMSE between these simulations and the training/testing trajectories, as reported in Table \ref{tab:learning}.
 
 As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
-This suggests that using DR to represent our activation functions $\theta(q)$ yields better generalization capabilities than the contrary.
-This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learnt without DR is an order of magnitude higher than with DR.
-Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the number of $K$ needed to represent the motion.
+This suggests that using DR to encode our activation functions $\theta_k$ in a lower-dimensional space $\phi(q)$ yields better generalization capabilities than encoding the behaviors in using the original $q$.
+This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learned without DR is an order of magnitude higher than with DR.
+Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the higher number of $K$ needed to represent the motion without DR.
 For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
@@
 \[ V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*) \]
-That is, $ 0 \prec V(q),~ \forall q \neq q^*$ and $ V(q^*) = 0 $. Where $q^*$ is any joint configuration such that $H(q^*) = x^*$.  The derivative of $ V $ wrt. time is:
+That is, $ 0 \prec V(q),~ \forall q \neq q^*$ and $ V(q^*) = 0 $, where $q^*$ is any joint configuration such that $H(q^*) = x^*$.  The derivative of $ V $ wrt. time is:
 \begin{equation}
@@
 \end{equation}
-Therefore, JT-DS \eqref{eq:ds} is stable and as $ q=q^* $ is the largest invariance in $ \mathbb{R}^m $, it is  globally asymptotically stable with respect to a task-space attractor; i.e.
-$\lim\limits_{t\rightarrow \infty} \| H(q)-x^*\|=0 $.}
-
+When $J(q)$ is full-rank, the above inequality can be tightened to $\frac{dV(q)}{dt} > 0$.
+Therefore, as $ q=q^* $ is the largest invariance in $ \mathbb{R}^m $, JT-DS \eqref{eq:ds} is stable with respect to a task-space attractor; i.e.
+$\lim\limits_{t\rightarrow \infty} \| H(q)-x^*\|=0 $, and asymptotically stable for all regions in which the end-effector is fully manipulable ($J(q)$ is full rank).
+}
 \section*{Acknowledgment}

### 2018-03-11 src/JTDS_mat_lib/utils/preprocess_demos_jtds.m

@@ -72,7 +72,7 @@ for i=1:length(demos)
     clear tmp_full tmpq tmpq_d tmpx tmpxd
     
     % de-noising data (not necessary)
-    tmdiffepq = demos{i}; 
+    tmpq = demos{i}; 
     
     % computing the first time derivative
     if ~iscell(time)

### 2018-03-11 src/JTDS_mat_lib/utils/preprocess_demos_jtds.m
Fixed some inconsistency while merging

@@ -72,11 +72,7 @@ for i=1:length(demos)
     clear tmp_full tmpq tmpq_d tmpx tmpxd
     
     % de-noising data (not necessary)
-<<<<<<< HEAD
     tmpq = demos{i}; 
-=======
-    tmpq = demos{i};
->>>>>>> 626d08484230f67781f725e097a608835e3c8075
     
     % computing the first time derivative
     if ~iscell(time)

### 2018-03-11 paper/reply_reviews/Synergy_analysis_pcavsfpca.tex
Re-structured code, new scripts for testing with orientation in task-space, new CV results for 'None'+'PCA' and new JTDS models

--- 2018-02-22 paper/reply_reviews/reviews.tex
@@
-\documentclass[10pt,stdletter,dateno]{newlfm}
-\usepackage{charter}
+\documentclass{article}
 
-\makeatletter
-\g@addto@macro{\ps@ltrhead}{%
-  \renewcommand{\headrulewidth}{0pt}%
-  \renewcommand{\footrulewidth}{0pt}%
+\usepackage{color}
+\title{A Small \LaTeX{} Article Template\thanks{To your mother}}
+\author{Your Name  \\
+	Your Company / University  \\
+	\and 
+	The Other Dude \\
+	His Company / University \\
 	}
-\g@addto@macro{\ps@othhead}{%
-  \renewcommand{\headrulewidth}{0pt}%
-  \renewcommand{\footrulewidth}{0pt}%
-}
-\makeatother
-\usepackage{color}
-\usepackage{xcolor}
-\usepackage{soul}
-\widowpenalty=1000
-\clubpenalty=1000
 
-\newsavebox{\Luiuc}
-\sbox{\Luiuc}{%
-	\parbox[b]{1.6in}{%
-		\vspace{0.5in}
-		\includegraphics[scale=0.25]
-		{./EPFL-Logo-bitmap/JPG/EPFL-Logo-RVB-55.jpg}%
-	}%
-}%
-%\makeletterhead{Uiuc}{\Rheader{\usebox{\Luiuc}}}
-\makeletterhead{Uiuc}{\Rheader{}}
-\newlfmP{headermarginskip=-100pt}
-\newlfmP{sigsize=0pt}
-\newlfmP{sigskipbefore=50pt}
-\headermarginskip{.5in}
-%\newlfmP{dateskipafter=30pt}
-%\newlfmP{addrfromphone}
-%\newlfmP{addrfromemail}
-%\PhrPhone{Phone}
-%\PhrEmail{Email}
+\date{\today}
+% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
+% before or after the \begin{document} command 
+% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
+\begin{document}
+\vspace{-10pt}
+Following we mathematically and graphically analyze the proposed motor synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via.
+PCA or K-PCA) and grouping of local behavior regions with GMM on this lower dimensional space.
+The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$.
+\underline{Reviewer: 5 (Review30371)} is concerned with this choice and wonders why we didn't use functional-PCA to parametrize $\mathcal{A}(q)$.
+We analyze the feasibility of this.
+The main comment from the reviewer:\\
 
-\lthUiuc
-
-%\namefrom{Seyed Sina Mirrazavi Salehian}
-%\addrfrom{\small
-%		Learning Algorithms and Systems Laboratory (LASA)\\
-%       	\small \'Ecole Polytechnique F\'ed\'erale de Lausanne (EPFL)\\
-%       	\small Lausanne, Switzerland - 1015.
-%}
-%\phonefrom{\small +41 21 69 31060 }
-%\emailfrom{\small nadia.figueroafernandez@epfl.ch}
-
-%\addrto{%
-%Faculty Search Committee\\
-%Department of Something\\
-%Some University\\
-%University Address}
-\vspace{-300pt}
-%\greetto{Dear Editor,}
-%\closeline{With best regards,}
-\begin{document}
-\begin{newlfm}
-\vspace{-2in}
-Dear Ms. Nadia Figueroa,
-
-Your paper
-
-17-1142
-"Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A
-Linear Parameter Varying and Synergetic Control Approach"
-by Yonadav Shavit, Nadia Figueroa, Seyed Sina Mirrazavi Salehian, Aude
-Billard Submission for RA-L only Special Issue on Learning and Control for
-Autonomous Manipulation Systems: the Role of Dimensionality Reduction submitted to the IEEE Robotics and Automation Letters (RA-L) has been reviewed by the Associate Editor and selected Reviewers.
-The reviews of the paper are attached.
-
-On the basis of the reviewers' ratings and comments, we decided that
-your paper cannot be published in the Letters in the present form.
-However, you are encouraged to rewrite and resubmit a revised version
-of your work addressing all editorial concerns.
-
-Please notice that there is a STRICT deadline of one month (30 DAYS)
-from today \hl{[i.e. March 18th, 2018]} for you to resubmit a revised version, along with a
-Statement of Changes as a single pdf file indicating how comments by
-the Editor and by reviewers have been addressed and a list of changes
-made in the paper.
-
-Please be aware that failure to meet the 30 days deadline implies that
-the paper will be automatically moved to the Reject category. Also
-notice that the next decision on this manuscript will be final, i.e.
-either Accept for Publication or Reject. Unfortunately, the tight
-timeline of the Letters will allow no extensions or exceptions to these
-rules.
-
-If you have any related concern, do not hesitate to contact me.
-
-Sincerely,\\
-Han Ding\\
-Editor (Learning and Control for Autonomous Manipulation Systems: the
-Role of Dimensionality Reduction)\\
-IEEE Robotics and Automation Letters\\
-
-\underline{Associate Editor Report:}
-
-The paper presents a dynamical system in joint space that combines
-different motor synergies to encode joint-space behaviors while
-asymptotically converges to a desired position in the task space. The
-paper is well written and organized. All the reviewers think the paper
-is of good quality. \textcolor{red}{:) OF COURSE it is!} The reviewers’ comments are attached. Please
-address the comments and improve the paper accordingly.
-
-%\clearpage
-
-\textbf{Reviewer(s)' Comments to Author:}\\
-
-\underline{Reviewer: 4 (Review30069)}\\
-This paper proposes a dynamical system (DS) that operates in joint
-space and converges to a desired position in task space. In doing so,
-the authors use a linear parameter varying system and use
-dimensionality reduction to learn these parameters efficiently. I think
-the paper is well written, experiments are well conducted and
-explained, and the results are promising.
-
-\textit{Detailed comments and
-suggestions are as listed below:}
-\begin{enumerate}
-\item As mentioned by the authors, there are many reasons to plan in joint
-space rather than task space. One of the important reasons to me seems
-to be able to plan end-effector orientation, not just position.
-Traditional movement primitives, for example DMPs, do not explicitly
-encode planning in the SO(3) space of orientations. There is research
-that addresses this question, for example [1], but they lose the
-simplistic appeal of movement primitives. \hl{I think it would be a great
-addition to this paper if the authors illustrate if their joint space
-DS can be used to accomplish orientation sensitive tasks, and
-generalize to end-effector orientations different that what was
-demonstrated.}\textcolor{red}{[*]} \\
-
-[1] Ude, Ale, et al. "Orientation in cartesian space dynamic movement
-primitives." Robotics and Automation (ICRA), 2014 IEEE International
-Conference on. IEEE, 2014.\\
-
-\textcolor{red}{[*] We never explored this, did we?
-Nevertheless, I think we can address it.
-We have data for the orientation of the demonstrations so we can try it.
-If we do this and it works, we should not only mention [1] but also the work from Seungsu [2] and Sylvain [3] both of which learn motions in SE(3) but cannot prove stability.}\\
-\textcolor{red}{[2] Gaussian mixture model for 3-dof orientations. S Kim, R Haschke, H Ritter. Robotics and Autonomous Systems 87, 28-37}\\
-\textcolor{red}{[3] An approach for imitation learning on Riemannian manifolds
-MJA Zeestraten, I Havoutis, J Silvério, S Calinon, DG Caldwell
-IEEE Robotics and Automation Letters 2 (3), 1240-1247}\\
-
-\item In the introduction, the authors stress that jacobian psuedo inverse
-can be computationally expensive. Is this really true for a 7-dof robot
-arm? Can the authors provide some numbers where this would really be an
-issue for a robot controlled at 1000Hz? I find the argument that
-jacobian inverse does not work for singularities and needs a set of
-heuristics to be more compelling. Maybe the authors should rewrite the
-introduction to stress more on this issue? \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item In Section II, criterion I, the authors talk about a positive
-symmetric matrix P. I do not see it in the equations above. Is this a
-weighing matrix? \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item In criterion II, the error is defined with respect to the velocity
-error, rather than a position or a combination of both. This means that
-the motion can be different from the demonstrated motion. I understand
-that this is important to keep the later optimization convex, but is
-there any disadvantage to using velocity error rather than position
-error? \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item I do not completely understand the significance of A(q). From Figure
-3, it seems like A(q) helps resolve redundancy, while achieving the
-task space goal. Do the authors have a mathematical (in constrast to
-intuitive) explanation for this? Does A(q) somehow relate to the
-null-space of the jacobian? \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item In Section III, "One can intuitively understand the control law as
-follows:$ (H(q)-x^*)$ denotes the position error w.r.t. the task target
-and by multiplying that error by the transposed Jacobian $J^T(q)$, the
-error is projected into joint space (similar to Jacobian transpose
-control)". I find this explanation misleading as jacobian multiplied to
-position error does not project the error into joint space. Jacobian
-transforms velocities, not positions. Maybe the law can be seen more as
-a first order approximation of the error in task space. In any case, I
-would either clarify this statement more or take it out of the
-manuscript all together. It seems like the DS is just a PD law that
-provably converges to a fixed attractor in task space. \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item About figure 2 again, the position error in task and joint space are
-not simply multiplied by a Jacobian. Although, task space error can be
-first-order approximated using a Jacobian multiplication with joint
-space error. What does the right figure in Fig 2 exactly represent? \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item In Section IV, the authors say that they use EM to determine the
-parameters of the GMM in Eq 6. Can they elaborate a little on this?
-What is the expectation taken over and what likelihood is maximized? \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item The learned dimensionality reduction and parameters seem task
-dependant. I wonder if the authors have any idea on how to generalize
-this across tasks. \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item \hl{The experiments have the same task-space goal as the
-demonstrations. Can the authors include experiments that have goals
-different from demonstrated, especially in orientation?} \textcolor{red}{[We already know that it won't generalize well if the target is in a completely different hemisphere of the workspace of the robot, however, we can try adding some tests and show how well it can generalize within the same region that it was demonstrated]}
-
-\item \hl{In similar light, maybe authors could present online adaptation of
-A(q) to generalize to situations different than demonstrated, for
-example a slightly moved obstacle?} \textcolor{red}{[This is a really nice suggestion.
-But how I see it, it's basically a new paper; i.e.
-an extension of this one, but we can discuss about this if we have space.
-A quick idea that comes to mind would be to apply a modulation/activation to the LPV system in a local region.
-Now this is not as straight forward as in our task space approaches because our LPV system is learned in joint-space while the activation function is in the embedded space.
-So it still would be an open question, but an approach to tackle this problem could be along these lines.]}
-
-\item In table II, instead of the tracking error, maybe the more useful
-error to see would be the error with respect to the plan or
-demonstation. One would assume that the proposed approach would be
-worse at it than SEDS+IK because the minimized error is with respect to
-the velocity and goal, not the position of the movement.
-\textcolor{red}{[This will only make sense for the tasks that are actually feasible for the IK-solver (i.e.
-not for foot-step or singularity).
-We can easily compute this and add it.
-However, we know already that the error will be worse than SEDS+IK, so would this make sense?
-]}
-
-\end{enumerate}
-
-Overall, I think this is an interesting, well written paper. I think
-adding experiments with different goal - position and orientation will
-help the paper, as well as an analysis of tracking error with respect
-to a plan or demonstration.\\
-
-\underline{Reviewer: 5 (Review30371)}
-
-This paper deals with the problem of controlling a manipulator towards
-a desired attractor in the task space by exploiting local synergies
-defined in the joint space and combined in time and space in order to
-obtain desired behaviours. This method does not require the computation
-of the pseudo-inverse Jacobian and hence is able to pass through
-singular configurations. The most similar approach is the Jacobian
-transpose control method that shares the same advantages but that does
-not include the learning of behaviours from demonstrations that seems
-to bring some benefits.
-
-The paper is well written and quite easy to follow. The cited
-literature is also appropriated and complete.
-
-\textit{However, some criticisms follows:}
-\begin{enumerate}
-\item \hl{My first concern regards you choice of representing the demonstrated
+\textit{``My first concern regards you choice of representing the demonstrated
 movement as a sequence of postures and hence looking for synergies by
@@
 However, the last one are valid for the whole movement and can be
-combined by constant coefficients.} \textcolor{red}{[Honestly, I hadn't heard of functional-PCA until now.
+combined by constant coefficients."} \\
+
+Let's begin by analyzing the proposed DS-based control-law:
+\begin{equation}
+\dot{q} = -\mathcal{A}(q)J^{T}(q)(H(q)-x^*)
+\end{equation}
+The intuition behind this control-law is to
+$\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta'(\phi(q))A_k$
+
+\textcolor{red}{[Honestly, I hadn't heard of functional-PCA until now.
 I searched a bit and I see why the reviewer would suggest this approach (which is interesting btw).
@@
 
-\item I also found the paper lacking of physical intuitions, especially in
-the experimental part, about the entities introduced in previous
-sections. In particular, what is the lesson we learn after your
-analysis of synergies and the obtained lower dimensional space for the
-few movements shown in section V? Is it possible to extract some global
-synergic behavior?
-Moreover, in section IV.A authors introduce the function phi(q) named
-the embedding that maps the joint configuration into a lower
-dimensional space. Is it associated to a whole behaviour or, as
-matrices $A_k$, has a local validity? I lost this point. Can you
-show/report or describe what it represents in the joint space for one
-or more cases in Section V? What is this lower dimensional space for
-the cases in Section V? \textcolor{red}{[Can be addressed through clarification and re-write]}
+\begin{thebibliography}{9}
+\bibitem[Doe]{doe} \emph{First and last \LaTeX{} example.},
+John Doe 50 B.C. 
+\end{thebibliography}
 
-\item A similar question for the matrices $A_k$ that represent local, I
-would say postural (see previous comment), synergies whose dimension is
-m x m, i.e. in joint space: I would expect they were defined in the
-lower dimensional space.
-\textcolor{red}{[We actually address this in the discussion, it is related to the stability proof; i.e.
-we would loose rank when projecting the A matrix back to joint space]} Moreover, are they local representation of the
-distribution of postures and hence covariance matrices? What do their
-eigenvectors and eigenvalues locally represent? Maybe, a comparison
-with the classical and more known postural synergies for the human
-hand, analyzed by PCA, could be useful. Finally, what does it mean that
-they should be positive definite (apart from stability issue)? \textcolor{red}{[All of these questions can be addressed through clarification and re-write]}
-
-\item What is the dimension of function $\theta_k(q)$ that, to some extent,
-can be seen as the new control variables. Is it $K<m$? \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item For the test, why not comparing also with the JT control method? In
-other terms, is it possible to use a SEDS + JT, especially for the
-third case "Transiting through singular configurations"?
-
-\item Minors:
-- SEDS + IK: what does SEDS mean? \textcolor{red}{[We clearly say what it means on page 6: ``For the Cartesian motion generator we
-use the SEDS approach [8] which learns an asymptotically
-stable DS in Cartesian space from demonstrations. We then
-generate joint trajectories through a damped least square IK
-solver. From herein we refer to this approach as SEDS+IK."]}
-
-\end{enumerate}
-
-\underline{Reviewer: 7 (Review30521)}
-
-The paper presents a kinematic controller for a robot, that drives it
-towards a fixed target in task space. The controller works in joint
-space and consists in a dynamical system formulated as a LPV system
-which combines different motor synergies. Such synergies are learnt
-from demonstrations. Theoretical proof of stability is given and
-experimental results are offered for validation.
-
-The paper is very well written and nicely organized. Results are
-interesting and the technical accuracy of the paper is good. The
-coverage of the state of the art is quite good.
-
-\textit{I have the following comments:}
-\begin{enumerate}
-\item a considerable part of the introduction is spent in comparing joint
-space trajectory generation to task space trajectory generation. Since
-these are elementary concepts in robotics, the authors might consider
-reducing this part; \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item one of the main claims of the paper is the stability proof, which in
-fact is cited in the first line of the abstract. However the stability
-analysis is a quite simple extension of the stability of the transpose
-Jacobian scheme, where the constant gain on the task space error is
-replaced by the matrix A(q). As such the emphasis on this result might
-be to some extent reduced. By the way, the proof works only if the
-Jacobian is full rank, which is not properly discussed in Appendix I; \textcolor{red}{[Can (should) be addressed through clarification and re-write]}
-
-\item the meaning of variables $\dot{q}_d$ in (2) is not very clear to me, a
-well as whether it takes the same meaning as $\dot{q}_{t,n}$ in (8); \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item Section III uses the concept of synergy without actually defining it.
-An introduction to this concept is given later, in Section IV; \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item the expression "flow of motion" in Proposition 1 is not very clear to
-me; \textcolor{red}{[Can be addressed through clarification and re-write]}
-
-\item some details about the optimization problem in (8) might be given
-(nature of the problem, complexity, feasibility). \textcolor{red}{[Can be addressed through clarification and re-write]}
-\end{enumerate}
+\end{document}
 

### 2018-03-11 src/JTDS_mat_lib/JTDS_Solver_v2.m

@@ -343,9 +343,9 @@ function [Priors, Mu, Sigma, As, latent_mapping] = JTDS_Solver_v2(Data, robotpla
     for i = 1:n
         q = Q(:, i); xt = Xt(:, i);
         if options.orientation_flag==1
-         Qd_basis(:, i) =  robotplant.qd_basis_orientation(q, xt);
+            Qd_basis(:, i) =  robotplant.qd_basis_orientation(q, xt);
         else
-        Qd_basis(:, i) = robotplant.qd_basis(q, xt);
+            Qd_basis(:, i) = robotplant.qd_basis(q, xt);
         end
         if options.learn_with_bounds % if false, ignore joint limits
             Ss(:,:,i) = robotplant.compute_S(q);% the S value for each point, converted to a vector

### 2018-03-11 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m

@@ -1,0 +1,175 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%     Learning JTDS Models (including orientation) on Different Datasets     %%                                                       %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour' 
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'pour_obst_2'
+        demos_location = strcat(data_path,'pour_obst_2/data.mat');
+        demo_ids = [1:7];                
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
+        demos_location = strcat(data_path,'singularity/data.mat');
+        demo_ids = [1:10];  
+        fprintf('Loading demonstrations from %s \n', demos_location);
+        load(demos_location)
+end
+
+if ~strcmp(choosen_dataset,'singularity')
[…]

### 2018-03-11 src/JTDS_mat_lib/model/RobotPlant.m

@@ -47,7 +47,7 @@ classdef RobotPlant
         
         function basis = qd_basis_orientation(obj, q, xt)
             % Computes the base joint velocity to move at a joint position q
-            % toward a task position xt using the jacobian transpose
+            % toward a task position+orientation xt using the jacobian transpose
             % method (i.e. qd = J^T(H(q) - xt) ).
             % In general, this will be multiplied by an A matrix to get the
             % augmented joint velocity.

### 2018-03-11 src/JTDS_mat_lib/run_crossval_dr_kuka_wOrientation.m

@@ -7,7 +7,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %
@@ -77,7 +77,6 @@ end
 
 %%% Orientation is included
 
-
 %%% Type of DR Methods %%%
 mappings_to_compare = {'None', 'PCA'};
 
@@ -128,7 +127,7 @@ for j = 1:repetitions
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     %       Split Dataset for Training/Testing        %%
     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    tt_ratio = 0.6;
+    tt_ratio = 0.8;
     train = round(length(Qs)*tt_ratio);
     Qs_train = []; Ts_train = [];
     Qs_test = [];   Ts_test = [];

### 2018-03-11 src/JTDS_mat_lib/run_crossval_hyperparam_kpca_kuka.m

@@ -6,7 +6,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'pour_obst_2'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %

### 2018-03-11 src/JTDS_mat_lib/run_synergy_analysis.m

@@ -1,0 +1,141 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Evaluation of Learning Schemes for JTDS Models (including orientation) on Different Datasets     %%
+%  Train and compare a series of JTDS models with different dimensionality                          %%   
+%  reduction schemes and GMM model fitting                                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'pour'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour' 
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'pour_obst_2'
+        demos_location = strcat(data_path,'pour_obst_2/data.mat');
+        demo_ids = [1:7];                
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
+        demos_location = strcat(data_path,'singularity/data.mat');
+        demo_ids = [1:10];  
[…]

### 2018-03-12 paper/reply_reviews/Synergy_analysis_pcavsfpca.tex
Added learned SEDS models (pos+orient) for all datasets and latest cv results

@@
 \vspace{-10pt}
-Following we mathematically and graphically analyze the proposed motor synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via.
+In response to \underline{Reviewer: 5 (Review30371)} (Comments 1-3).
+Following we mathematically and graphically analyze the proposed local behavior synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via.
 PCA or K-PCA) and grouping of local behavior regions with GMM on this lower dimensional space.
-The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$.
-\underline{Reviewer: 5 (Review30371)} is concerned with this choice and wonders why we didn't use functional-PCA to parametrize $\mathcal{A}(q)$.
-We analyze the feasibility of this.
-The main comment from the reviewer:\\
-
-\textit{``My first concern regards you choice of representing the demonstrated
-movement as a sequence of postures and hence looking for synergies by
-using PCA of K-PCA analysis. Why not using functional PCA (f-PCA)
-analysis and determine the few moving synergies that make possible the
-execution of the demonstrated behaviours? It would be that, when you
-combine the local synergies in space and time by using functions
-$\theta_k(q)$, you get what you might obtain by using directly f-PCA.
-However, the last one are valid for the whole movement and can be
-combined by constant coefficients."} \\
-
+The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$ and analyze the feasibility of applying F-PCA to this scheme.
 Let's begin by analyzing the proposed DS-based control-law:
@@
 \end{equation}
-The intuition behind this control-law is to
-$\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta'(\phi(q))A_k$
+The intuition behind this control-law is that $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta'(\phi(q))A_k$ is a linear combination of time-invariant linear matrices $A_k$, these matrices are the ``local behavior synergy matrices" that shape the motion in joint-space.
+Meaning that, given the joint-space velocity vector representing the task-space error $J^{T}(q)(H(q)-x^*)$, the resulting motion is biased to use a particular set of joints, defined in each $A_k$.
+Each $A_k$ is thus activated depending on the current ``joint-posture" $q$ (represented in the lower-dimensional space $\phi(q)$) via the scheduling/activation function $\theta'(\phi(q))$.
+This is our proposed definition of ``local behavior synergies", which is not equivalent to the notion of ``hand posture synergies" as suggested by the reviewer.
+The fact that we use PCA or K-PCA to represent to activation function is to alleviate the complexity of extracting the local synergy regions in joint-space.
+From the classical definition of joint synergies, when transforming the joint trajectories via PCA/K-PCA to a lower dimensional-space we are indeed working in the ``joint postural synergy" space as the reviewer suggests.
+But, solely the activation function $\theta'(\phi(q))$ is represented by these joint postural synergies.
+The $A_k$ matrices themselves are the ``local behavior synergies" and these live in joint-space $A_k\in \mathbb{R}^{m \times m}$ and are time-invariant.
+Note that, if we were to bypass the DR step our state-dependent system matrix would be the following $\mathcal{A}(q)= \sum\limits_{k=1}^{K}\theta(q)A_k$, which through our definition is still a combination of ``local behavior synergy matrices".
+As shown in TABLE I of the manuscript and discussed in Section V, by using a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
+Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
+Following we demonstrate this on the pouring behavior dataset.
 
-\textcolor{red}{[Honestly, I hadn't heard of functional-PCA until now.
-I searched a bit and I see why the reviewer would suggest this approach (which is interesting btw).
-F-PCA is the application of PCA on a functional space (mainly targeted at applications where each of the datapoints is a time-series).
-The goal is then to find the minimal set of eigen-time-series that best describes the shape of the observed time-series.
-To do this, each time-series is approximated with k basis functions and PCA is applied on these functions.
-The resulting eigen-functions are parametrized by a linear combination of these k basis-functions (looks very similar to k-pca/svc).
-First of all, I see a couple of complications: i) the basis functions are time dependent; i.e.
-they will only work for time-series of the same length as the datasets that they were learned from ii) the number of k basis functions is an open parameter, as well as the number of eigenfunctions to keep iii) I can understand how the reviewer would think that it is almost the same as finding the lower dimensional space via PCA/k-PCA and then fitting a GMM, as this would be equivalent to the 'k' basis functions.
-However, I'm not sure if the 'k' basis functions in joint space are equivalent to the 'K' clusters discovered by the GMM in the embedded space.
-Which is extremely important because this dictates how many A's we need.
-I might be wrong about these issues, I need to dig deeper into this.]}
+In Fig.
 
-\begin{thebibliography}{9}
-\bibitem[Doe]{doe} \emph{First and last \LaTeX{} example.},
-John Doe 50 B.C. 
-\end{thebibliography}
+This activation is defined by  which is the posterior probability of a GMM that models these local linear motions with.
+The idea of using a DR scheme to parametrize this activation function stems from the fact that
+
+%\begin{thebibliography}{9}
+%\bibitem[Doe]{doe} \emph{First and last \LaTeX{} example.},
+%John Doe 50 B.C. 
+%\end{thebibliography}
 

### 2018-03-12 paper/reply_reviews/reply_comments.tex
Re-structured code, new scripts for testing with orientation in task-space, new CV results for 'None'+'PCA' and new JTDS models
Added learned SEDS models (pos+orient) for all datasets and latest cv results

@@
 \usepackage{tikz}
+\usepackage{amsmath}
+\usepackage{amsfonts}
 \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
@@
 \begin{enumerate}
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``As mentioned by the authors, there are many reasons to plan in joint
+space rather than task space. One of the important reasons to me seems
+to be able to plan end-effector orientation, not just position.
+Traditional movement primitives, for example DMPs, do not explicitly
+encode planning in the SO(3) space of orientations. There is research
+that addresses this question, for example [1], but they lose the
+simplistic appeal of movement primitives. I think it would be a great
+addition to this paper if the authors illustrate if their joint space
+DS can be used to accomplish orientation sensitive tasks, and
+generalize to end-effector orientations different that what was
+demonstrated.[1] Ude, Ale, et al. "Orientation in cartesian space dynamic movement
+primitives." Robotics and Automation (ICRA), 2014 IEEE International
+Conference on. IEEE, 2014.}"\\
+\textcolor{red}{\texttt{Response:} \small Yes, it works, redoing Table I for now, need to highlight this throughout the paper..}\\
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``In the introduction, the authors stress that jacobian psuedo inverse
+can be computationally expensive. Is this really true for a 7-dof robot
+arm? Can the authors provide some numbers where this would really be an
+issue for a robot controlled at 1000Hz? I find the argument that
+jacobian inverse does not work for singularities and needs a set of
+heuristics to be more compelling. Maybe the authors should rewrite the
+introduction to stress more on this issue?"}\\
+\textcolor{red}{\texttt{Response:} \small For a 7-DOF robot it is indeed negiblible.
+This only becomes an issue for higher degrees of freedom, like a multi-fingered hand or multi-arm robotic setting.
+We will remove this passage from the text as we don't have much space to discuss it.}\\
 
@@
 
-\item \textit{``In Section III, ?One can intuitively understand the control law as follows:(H(q) ?? x) denotes the
+\item \textit{``In Section III, ?One can intuitively understand the control law as follows:$(H(q) - x^*)$ denotes the
 position error w.r.t. the task target and by multiplying that error by the transposed Jacobian JT (q),
@@
 The new text is reproduced below: \\
-One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
+\textit{``One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
 the task target, which we then reinterpret as the task-space velocity of a proportional controller.
@@
 [?] illustrates the effects of  $\mathcal{A}(q)$ on the generated motion.
-Thus the controller can be thought of as a proportional controller in joint space.}\\
+Thus the controller can be thought of as a proportional controller in joint space.}}\\
 
@@
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``The experiments have the same task-space goal as the
+demonstrations. Can the authors include experiments that have goals
+different from demonstrated, especially in orientation?"}\\
+\textcolor{red}{\texttt{Response:} \small First of all, they don't..
+show some figures...
+add clarification of this in the text and explain why we cannot systematically evaluate this if we had too..}\\
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``In similar light, maybe authors could present online adaptation of
+A(q) to generalize to situations different than demonstrated, for
+example a slightly moved obstacle?"}\\
+\textcolor{red}{\texttt{Response:} \small This is a really nice suggestion.
+But how I see it, it's basically a new paper; i.e.
+an extension of this one, but we can discuss about this if we have space.
+A quick idea that comes to mind would be to apply a modulation/activation to the LPV system in a local region.
+Now this is not as straight forward as in our task space approaches because our LPV system is learned in joint-space while the activation function is in the embedded space.
+Another idea is to modify the activation function of the local behavior regions such that it is a function of an external variable, for example if force is applied/sensed, and it bias the model towards another $A_k$ matrix?
+So it still would be an open question, but an approach to tackle this problem could be along these lines....
+Add a passage in the discussion answering this question}\\
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``In table II, instead of the tracking error, maybe the more useful
+error to see would be the error with respect to the plan or
+demonstartion. One would assume that the proposed approach would be
+worse at it than SEDS+IK because the minimized error is with respect to
+the velocity and goal, not the position of the movement."}\\
+\textcolor{red}{\texttt{Response:} \small This is not necessarily the reason that it would be worse, as SEDS is also learned by minimizing the error with respect to
+the velocity and the goal.
+The reason that JT-DS is less accurate in reproducing the motion in task-space is because our optimization is focused on reproducing the motion solely in joint-space, while reaching the final target in task-space.
+Nowhere do we claim that JT-DS is capable of reproducing a motion in both joint and task space.
+Nevertheless, we add the mean swept-error-arrea (SEA) which evaluate the performance of reproduction in task-space from a reference trajectory...
+as shown in the tables....}\\
 
-\item \textit{Specific clarifications and typos:
-\begin{itemize}
-\item ..
-\item ..
-\end{itemize}}
 \end{enumerate}
@@
 \begin{enumerate}
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``My first concern regards you choice of representing the demonstrated
+movement as a sequence of postures and hence looking for synergies by
+using PCA of K-PCA analysis. Why not using functional PCA (f-PCA)
+analysis and determine the few moving synergies that make possible the
+execution of the demonstrated behaviours? It would be that, when you
+combine the local synergies in space and time by using functions
+$\theta_k(q)$, you get what you might obtain by using directly f-PCA.
+However, the last one are valid for the whole movement and can be
+combined by constant coefficients."}\\
+\textcolor{red}{\texttt{Response:} \small Working on the response to this in the other document: synergy\_analysis\_pcavsfpca.tex}\\
 
@@
 The paragraph explaining this has been clarified, and is reproduced below:\\
-"As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
+\textit{"As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
 This suggests that using DR to encode our activation functions $\theta_k$ in a lower-dimensional space $\phi(q)$ yields better generalization capabilities than encoding the behaviors in using the original $q$.
@@
 By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
-Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations." \\
+Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations."} \\
 Regarding the second half of the comment, $\phi(q)$ is a learned transformation, which is associated with a whole behavior, rather than a particular $A_k$ (as is now clarified in IV.A).
@@
 The following line has been added to IV.A for clarity:\\
-"For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the "shoulder" and "arm" components of $q$ into a single "shoulder-arm" component in $\phi(q)$.}"\\
+\textit{"For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the "shoulder" and "arm" components of $q$ into a single "shoulder-arm" component in $\phi(q)$.}}"\\
 
@@
 	they should be positive definite (apart from stability issue)?"}\\
-\textcolor{blue}{\texttt{Response:} \small Regarding your first comment, we address why defining $A_k$ in lower-dimensional space is problematic.
+\textcolor{red}{\texttt{Response:} \small Regarding your first comment, we address why defining $A_k$ in lower-dimensional space is problematic.
 Specifically, we would lose rank when projecting the A matrix back to joint space, invalidating our stability proof.
-The $A_k$ matrices are }\\
+The $A_k$ matrices are...
+this will be answered in the response to question 1..}\\
 
@@
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
+\item \textit{``For the test, why not comparing also with the JT control method? In
+other terms, is it possible to use a SEDS + JT, especially for the
+third case "Transiting through singular configurations"?"}\\
+\textcolor{blue}{\texttt{Response:} \small This could be done, but it is mathematically incorrect.
+The original JT control method is a DS that converges to a unique equilibrium point, i.e.
+the task-space target $x^*\in \mathbb{R}^{N\times 1}$, with a gain matrix $K \in \mathbb{R}^{M\times M}$ as:\begin{equation}
+\dot{q} = K J(q)^T(H(q)- x^*) 
+\label{eq:JT}
+\end{equation}
+SEDS on the other hand, is a DS-based motion generator that converges also to a unique equilibrium point, as:
+\begin{equation}
+\hfill \begin{aligned}
+&  \dot{x}  = \mathbf{f}(x)
+\rightarrow  &
+\begin{cases}
+\lim_{t \rightarrow \infty}\|x - x^*\|=0
+\end{cases}
+\end{aligned}  \hfill
+\label{eq:simple-ds}
+\end{equation}
+where $\mathbf{f}(\cdot): \mathbb{R}^M \rightarrow \mathbb{R}^M$ is a continuous differentiable vector-valued function representing a non-linear DS that converges to a single stable equilibrium point $x^*$.
+SEDS is used in the following way: Given the current state of the robot $x$ the next desired velocity $\dot{x}_d$ is computed with \eqref{eq:simple-ds} this is then fed to a velocity-based IK solver as shown in Fig.
+\ref{}:}\\
+\textcolor{red}{TO SINA: Include typical SEDS-like control-loop here}\\
+\textcolor{blue}{ \small If we were to use SEDS with the JT-control method, we would have to integrate the desired velocity to generate a desired position $x_d$ at each time-step and define it as the target in \eqref{eq:JT}, which would yield the following controller:
+\begin{equation}
+\dot{q} = K J(q)^T(H(q)- x_d) 
+\end{equation} Note that, the JT control method assumes that the target is static, hence, in order for this controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
+This is contradicting as \eqref{eq:simple-ds} is always trying to converge to $x^*$.
+Thus, merging these two approaches is theoretically ill-fitting.
+To better highlight this, we show an illustration of the supposed SEDS+JT control loop in Fig.
+\ref{}:}\\
+\textcolor{red}{TO SINA: Include SEDS+JT control-loop here}\\
+\textcolor{red}{\small Describe this control-loop and state more reasons (if any) why it's wrong...}
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
-
-\item \textit{Specific clarifications and typos:
-\begin{itemize}
-\item ..
-\item ..
-\end{itemize}}
+\item \textit{``Minors:
+- SEDS + IK: what does SEDS mean?"}\\
+\textcolor{blue}{\texttt{Response:} \small As stated in the previous comment, SEDS, short for Stable Estimator of Dynamical Systems is an approach proposed in [8] that learns a stable estimate of a DS with the form of \eqref{eq:simple-ds} from demonstrations.
+The first paragraph of Secrtion V.B was written to describe its meaning and how we use it:\\
+\textit{``We now seek to elucidate the distinctive properties of our JT-DS model by comparing its performance to a DS-based Cartesian motion generator + IK solver approach for behaviors \textit{(5-7)}.
+For the Cartesian motion generator we use the SEDS (Stable Estimator of Dynamical Systems) approach [8] which learns an asymptotically stable DS in Cartesian space from demonstrations.
+We then generate joint trajectories through a damped least square IK solver.
+From herein we refer to this approach as SEDS+IK."}}\\
 
@@
 
-\item \textit{``bla bla"}\\
-\textcolor{blue}{\texttt{Response:} \small bla bla}\\
-
-\item \textit{Specific clarifications and typos:
-\begin{itemize}
-\item ..
-\item ..
-\end{itemize}}
+\item \textit{`some details about the optimization problem in (8) might be given
+(nature of the problem, complexity, feasibility)."}\\
+\textcolor{red}{\texttt{Response:} \small TODO..}\\
 

### 2018-03-12 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
Added learned SEDS models (pos+orient) for all datasets and latest cv results

@@
 We now seek to elucidate the distinctive properties of our JT-DS model by comparing its performance to a DS-based Cartesian motion generator + IK solver approach for behaviors \textit{(5-7)}.
-For the Cartesian motion generator we use the SEDS approach \cite{khansari2011learning} which learns an asymptotically stable DS in Cartesian space from demonstrations.
+For the Cartesian motion generator we use the SEDS (Stable Estimator of Dynamical Systems) approach \cite{khansari2011learning} which learns an asymptotically stable DS in Cartesian space from demonstrations.
 We then generate joint trajectories through a damped least square IK solver.

### 2018-03-12 src/JTDS_mat_lib/crossvalidation_dr_kuka_wOrientation.m

@@ -1,0 +1,197 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Evaluation of Learning Schemes for JTDS Models (including orientation) on Different Datasets     %%
+%  Train and compare a series of JTDS models with different dimensionality                          %%   
+%  reduction schemes and GMM model fitting                                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour' 
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'pour_obst_2'
+        demos_location = strcat(data_path,'pour_obst_2/data.mat');
+        demo_ids = [1:7];                
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
+        demos_location = strcat(data_path,'singularity/data.mat');
+        demo_ids = [1:10];  
[…]

### 2018-03-12 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m

@@ -5,7 +5,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %
@@ -121,7 +121,7 @@ robotplant = RobotPlant(robot, 'end_trans');
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %       Split Dataset for Training/Testing        %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-tt_ratio = 0.8;
+tt_ratio = 0.6;
 train = round(length(Qs)*tt_ratio);
 Qs_train = []; Ts_train = [];
 Qs_test = [];   Ts_test = [];
@@ -173,3 +173,8 @@ model_dir = strcat('./learned_JTDS_models/',choosen_dataset);
 mkdir(model_dir); 
 filename = strcat(model_dir,'/JTDS_model.txt')
 out = export2JSEDS_Cpp_lib(filename,Priors,Mu,Sigma,robot, As, latent_mapping.M);
+
+model_dir = strcat('./learned_SEDS_models/',choosen_dataset);
+mkdir(model_dir); 
+cd(model_dir)
+out = exportSEDS_Cpp_lib(Priors,Mu,Sigma)

### 2018-03-12 src/JTDS_mat_lib/run_synergy_analysis.m

@@ -43,6 +43,7 @@ if ~strcmp(choosen_dataset,'singularity')
     fprintf('Loading demonstrations from %s \n', demos_location);
     [Qs_, Ts_] = ImportDemonstrations(demos_location);
 end
+
 % If the data is very dense, initializing the semidefinite program may take
 % a long time. In this case, it may help to thin down the number of
 % demonstrated points (by varying "thinning_ratio", so long as there are still sufficient points to
@@ -73,6 +74,7 @@ if do_plots
         ylabel('Angle (rad)','Interpreter', 'LaTex', 'Fontsize', 15)
     end
 end
+
 %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%    Train a JTDS model on the current dataset   %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -138,4 +140,72 @@ motion_generator = MotionGenerator(robotplant, Mu, Sigma, Priors, As, latent_map
 % Compute RMSE on training data
 rmse_train = mean(trajectory_error(motion_generator, Data_train(1:dimq, :), Data_train(dimq+1:2*dimq, :), Data_train(2*dimq+1:end, :),options.orientation_flag));
 fprintf('Using %s mapping, got prediction RMSE on training: %d \n', mapping_name, rmse_train);
+
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%    Plot Lower-Dimensional Embedding and Synergies   %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Extract Lower Dimensional Embedding of Demonstrations
+figure('Color',[1 1 1])  
+for p=1:pca_dim
+    subplot(pca_dim,1,p)    
+    for i=1:length(Qs)
+        q_ref = Qs{i};
+        phi_q = out_of_sample(q_ref', latent_mapping)';
+        plot(phi_q(p,:),'-.'); hold on;
+    end      
+    grid on;
[…]

### 2018-03-12 src/JTDS_mat_lib/utils/my_plot3dGaussian.m

@@ -1,0 +1,50 @@
+function handles = my_plot3dGaussian(Priors, Mus,Sigmas )
+% 
+
+K = size(Mus,2);
+npts = 20; 
+
+handles = zeros(K,1);
+% weights = Priors;
+weights = 1/length(Priors)*ones(length(Priors));
+
+% weights = rescale(Priors,min(Priors),max(Priors),0.01,0.8);
+
+% weights(isnan(weights)==1) = 0.8;
+K_colors = hsv(K);
+for k = 1:K    
+    [x,y,z] = sphere(npts);
+    ap = [x(:) y(:) z(:)]';
+    [v,d]=eig(Sigmas(:,:,k));
+    if any(d(:) < 0)
+        fprintf('warning: negative eigenvalues\n');
+        d = max(d,0);
+    end
+    
+    if eig(Sigmas(:,:,k)) < 1e-3
+        d = d*100;
+    end
+    
+    d =  sqrt(d); % convert variance to sdwidth*sd
+    bp = (v*d*ap) + repmat(Mus(:,k), 1, size(ap,2));
+    xp = reshape(bp(1,:), size(x));
+    yp = reshape(bp(2,:), size(y));
+    zp = reshape(bp(3,:), size(z));
+    handles(k) = surf(xp,yp,zp);
+
+    
+%     [~,D] = eig(Sigmas(:,:,k));
+%     eigV = sqrt(diag(D))
+%     [x, y, z] = ellipsoid(Mus(1,k),Mus(2,k),Mus(3,k),eigV(1,1),eigV(2,1),eigV(3,1),30);
+
+    scatter3(Mus(1,k),Mus(2,k),Mus(3,k),2,'filled','ko');
+%     sh = surfl(xp, yp, zp);
+    alpha 0.2
+%     set(handles(k),'FaceColor',K_colors(k,:),'EdgeColor','none')
+    set(handles(k),'FaceColor',K_colors(k,:))
+
+
+end
+
+end
+

### 2018-03-12 src/JTDS_mat_lib/crossvalidation_dr_kuka_wOrientation.m
Uploaded new SEDS (pos+orient) and JTDS (pos+orient) models in .txt + .mat

@@ -82,7 +82,7 @@ mappings_to_compare = {'None', 'PCA'};
 
 %%% Learning options %%%
 options = [];
-options.orientation_flag=0;
+options.orientation_flag=1;
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%

### 2018-03-12 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m

@@ -5,7 +5,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %
@@ -82,7 +82,7 @@ options = [];
 % To remove orientation from target 
 % simply set flag = 0 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-options.orientation_flag = 1; 
+options.orientation_flag = 0; 
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -169,12 +169,19 @@ fprintf('Using %s mapping, got prediction RMSE on testing: %d \n', mapping_name,
 %      If you are happy with the results, export the model       %%
 %        for execution with the rtk-DS cpp class                 %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% All parameters in 1 file
+% model_dir = strcat('./learned_JTDS_models/',choosen_dataset);
+% mkdir(model_dir); 
+% filename = strcat(model_dir,'/JTDS_model.txt')
+% out = export2JSEDS_Cpp_lib(filename,Priors,Mu,Sigma,robot, As, latent_mapping.M);
+
+% 1 file per paramater
 model_dir = strcat('./learned_JTDS_models/',choosen_dataset);
 mkdir(model_dir); 
-filename = strcat(model_dir,'/JTDS_model.txt')
-out = export2JSEDS_Cpp_lib(filename,Priors,Mu,Sigma,robot, As, latent_mapping.M);
-
-model_dir = strcat('./learned_SEDS_models/',choosen_dataset);
-mkdir(model_dir); 
 cd(model_dir)
-out = exportSEDS_Cpp_lib(Priors,Mu,Sigma)
+out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, latent_mapping.M')
+
+% save mat file of variables
+M = latent_mapping.M';
+save('model.mat','Priors','Mu','Sigma', 'As', 'M')
+

### 2018-03-12 src/JTDS_mat_lib/utils/export2JSEDS_Cpp_lib_v2.m

@@ -1,0 +1,35 @@
+function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M)
+% Export the parameters of a JT-DS model to a given file.
+% 
+% Arguments:
+% filename - the filename to write the parameters to
+% Priors - the GMM priors
+% Mu - the GMM means of the JT-DS model
+% Sigma - the GMM variances of the JT-DS model
+% robot - the Corke robot-toolkit robot which the JT-DS model was built for
+% As - the joint augmentation matrices in the JT-DS model
+% M - the PCA matrix for embedding the joint-space positions into a
+%   lower-dimensional latent space in the JT-DS model
+
+% Outputs:
+% out - whether the file was successfully fully written
+
+% The parameters are written to the file in the following order:
+% dimq, dimz, K
+% priors
+% Mus (reshaped horizontally)
+% Sigmas (one matrix per each line, reshaped horizontally)
+% As (one matrix per each line, reshaped horizontally)
+% M (the PCA matrix) reshaped horizontally
+
+dlmwrite('Priors.txt',Priors,'newline','pc','-append','Delimiter','\t','precision','%.6f');
+dlmwrite('Mu.txt', Mu,'newline','pc','-append','Delimiter','\t','precision','%.6f');
+for i=1:size(Mu,2)
+    dlmwrite('Sigma.txt',Sigma(:,:,i),'newline','pc','-append','Delimiter','\t','precision','%.6f');
+end
+for i=1:size(Mu,2)
+    dlmwrite('A_matrix.txt',As(:,:,i),'newline','pc','-append','Delimiter',' ','precision','%.6f');
+end
+dlmwrite('M_matrix.txt', M,'newline','pc','-append','Delimiter','\t','precision','%.6f');
+
+out = true;

### 2018-03-13 paper/reply_reviews/Synergy_analysis_pcavsfpca.tex
Latest model and review updates

@@
 \documentclass{article}
-
+\usepackage[left=3cm, right=3cm, top=2cm]{geometry}
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{graphicx}
 \usepackage{color}
@@
 \vspace{-10pt}
-In response to \underline{Reviewer: 5 (Review30371)} (Comments 1-3).
+In response to Comments 1-3 from \underline{Reviewer: 5 (Review30371)}.
 Following we mathematically and graphically analyze the proposed local behavior synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via.
 PCA or K-PCA) and grouping of local behavior regions with GMM on this lower dimensional space.
-The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$ and analyze the feasibility of applying F-PCA to this scheme.
+The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$.
 Let's begin by analyzing the proposed DS-based control-law:
 \begin{equation}
-\dot{q} = -\mathcal{A}(q)J^{T}(q)(H(q)-x^*)
+\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{m\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
+\label{eq:jtds}
 \end{equation}
-The intuition behind this control-law is that $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta'(\phi(q))A_k$ is a linear combination of time-invariant linear matrices $A_k$, these matrices are the ``local behavior synergy matrices" that shape the motion in joint-space.
+The intuition behind this control-law is that $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta(\phi(q))A_k$ is a linear combination of time-invariant linear matrices $A_k \in \mathbb{R}^{m \times m}$, \underline{these matrices are the ``local behavior synergy matrices"} that shape the motion in joint-space.
 Meaning that, given the joint-space velocity vector representing the task-space error $J^{T}(q)(H(q)-x^*)$, the resulting motion is biased to use a particular set of joints, defined in each $A_k$.
-Each $A_k$ is thus activated depending on the current ``joint-posture" $q$ (represented in the lower-dimensional space $\phi(q)$) via the scheduling/activation function $\theta'(\phi(q))$.
-This is our proposed definition of ``local behavior synergies", which is not equivalent to the notion of ``hand posture synergies" as suggested by the reviewer.
-The fact that we use PCA or K-PCA to represent to activation function is to alleviate the complexity of extracting the local synergy regions in joint-space.
-From the classical definition of joint synergies, when transforming the joint trajectories via PCA/K-PCA to a lower dimensional-space we are indeed working in the ``joint postural synergy" space as the reviewer suggests.
-But, solely the activation function $\theta'(\phi(q))$ is represented by these joint postural synergies.
-The $A_k$ matrices themselves are the ``local behavior synergies" and these live in joint-space $A_k\in \mathbb{R}^{m \times m}$ and are time-invariant.
-Note that, if we were to bypass the DR step our state-dependent system matrix would be the following $\mathcal{A}(q)= \sum\limits_{k=1}^{K}\theta(q)A_k$, which through our definition is still a combination of ``local behavior synergy matrices".
+Each $A_k$ is thus activated depending on the current ``joint-posture" $q$ (represented in the lower-dimensional space $\phi(q)$) via the scheduling/activation function $\theta(\phi(q))$.
+Hence, our proposed definition of ``local behavior synergies" is not analogous to the notion of ``hand posture synergies".
+Note that, if we were to bypass the DR step our state-dependent system matrix would becomes $\mathcal{A}(q)= \sum\limits_{k=1}^{K}\theta(q)A_k$, which (via our definition) is a combination of ``local behavior synergy matrices" with the activation function living in joint-space rather than postural synergy space.
+The fact that we use PCA/K-PCA to represent the activation function $\theta(\cdot)$ is to alleviate the complexity of extracting the local synergy regions in joint-space directly.
+It is also motivated by the fact that we know that multi-DOF joint postures can be accurately represented in a lower-dimensional space composed of the principal components of the postural trajectories; i.e.
+the \textit{joint postural synergies} mentioned by the reviewer.
+
+Given the raw demonstrations of a joint-space  behavior with a task-space target as shown in Fig.
+\ref{fig:raw_demos} (for the pouring task) we seek to learn a linear combination of local behavior synergies $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta(\phi(q))A_k$ such that we can use \eqref{eq:jtds} to accurately reproduce these demonstrations.
+In order to do so, we project the raw joint trajectories to a lower-dimensional space via $\phi(q)$, in the case of PCA $\phi(q)=M_p \times q$ for $M_p \in \mathbb{R}^{p \times m}$.
+By choosing the first 3 PCs we can explain 95\% of the variance in the joint posture trajectories, which can be represented in this 3-D space as shown in Fig.
+\ref{fig:raw_demos}.
+We then jointly learn the activation function $\theta(\cdot)$ and the number of $K$ local synergy regions by fitting a GMM to this representation of the joint posture trajectories.
+As can be seen in Fig.
+\ref{fig:raw_demos}, the entire joint-space motion can be represented by 3 local synergy regions in PCA-space.
+Meaning that we only need 3 synergy matrices $A_k$ to represent the complex joint-space motion accurately.
+The activation function $\theta(\cdot)$, as described in the mansucript, is the posterior probability of a joint posture in PCA-space $\phi(q)$ belonging to one of the local synergy regions $k \in \{1,\dots,K\}$.
+
+\underline{As mentioned in the reviewers 3rd comment}, since we have a representation of our joint posture trajectories and the local synergy regions in this lower-dimensional space, why not learn the $A_k$ matrices in this space; i.e.
+$\mathcal{A}(q)\in \mathbb{R}^{p\times p}$ rather than $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ ?
+This seems like the obvious method, if we were considering our synergistic approach the classical way, i.e.
+using the lower-dimensional space as the new control variables.
+However, if we were to do that, this would mean that \eqref{eq:jtds} would become:
+\begin{equation}
+\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{M_p^{-1}}_{\mathbb{R}^{m\times p}}\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{p\times p}}\underbrace{M_p}_{\mathbb{R}^{p\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
+\label{eq:jtds_low}
+\end{equation}
+This is not entirely incorrect, however, with such a control law we can no longer guarantee stability and convergence to the attractor $x^{*}$.
+Let's consider the Lyapunov function used in Appendix I to prove stability of \eqref{eq:jtds} i.e.
+$V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*)$, The derivative of $ V $ wrt.
+time with \eqref{eq:jtds_low} is:
+\begin{equation}
+\begin{aligned}
+\frac{dV(q)}{dt} &= (H(q) - x^*)^TJ(q)\dot{q}\\
+&= -(H(q) - x^*)^TJ(q)\mathcal{A}(q)J^T(q)(H(q) - x^*)\\
+&= -(H(q) - x^*)^TJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{> 0}\underbrace{A_k}_{\succ 0}J^T(q)(H(q) - x^*)\leq 0
+\end{aligned}
+\end{equation}
+
+Note that this analysis is only considering PCA as the DR approach, with K-PCA deriving such a control law is more involved as the inverse mapping $\phi(q)^{-1}$ is not straightforward to compute.
+
+\begin{figure*}[!th] 
+  \begin{minipage}{0.58\textwidth}
+     	\centering 
+     	\includegraphics[trim={1.2cm 2cm 1.7cm 2cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/raw_demos_pour.pdf}
+  \end{minipage}
+   \begin{minipage}{0.47\textwidth}
+      	\centering
+      	\includegraphics[trim={1.2cm 0.5cm 0.5cm 0.35cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/raw_demos_pca_pour.pdf}
+      	\vspace{-5pt}
+
+		\includegraphics[trim={0.8cm 0.5cm 0cm 0.35cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/local_synergy_regions_pour_annotated.pdf}
+   \end{minipage}
+   \caption{Raw demonstrations of the pouring task in original space, PCA space and the extracted local synergy regions. \label{fig:raw_demos}}
+\end{figure*}
+
 As shown in TABLE I of the manuscript and discussed in Section V, by using a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
 Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
-Following we demonstrate this on the pouring behavior dataset.
 
-In Fig.
-
-This activation is defined by  which is the posterior probability of a GMM that models these local linear motions with.
-The idea of using a DR scheme to parametrize this activation function stems from the fact that
+On the applicability of F-PCA to our learning scheme.
 

### 2018-03-13 paper/reply_reviews/Synergy_analysis_reviewer2.tex
Latest model and review updates

--- 2018-03-13 paper/reply_reviews/Synergy_analysis_pcavsfpca.tex
@@
 Hence, our proposed definition of ``local behavior synergies" is not analogous to the notion of ``hand posture synergies".
-Note that, if we were to bypass the DR step our state-dependent system matrix would becomes $\mathcal{A}(q)= \sum\limits_{k=1}^{K}\theta(q)A_k$, which (via our definition) is a combination of ``local behavior synergy matrices" with the activation function living in joint-space rather than postural synergy space.
-The fact that we use PCA/K-PCA to represent the activation function $\theta(\cdot)$ is to alleviate the complexity of extracting the local synergy regions in joint-space directly.
-It is also motivated by the fact that we know that multi-DOF joint postures can be accurately represented in a lower-dimensional space composed of the principal components of the postural trajectories; i.e.
-the \textit{joint postural synergies} mentioned by the reviewer.
 
@@
 The activation function $\theta(\cdot)$, as described in the mansucript, is the posterior probability of a joint posture in PCA-space $\phi(q)$ belonging to one of the local synergy regions $k \in \{1,\dots,K\}$.
-
-\underline{As mentioned in the reviewers 3rd comment}, since we have a representation of our joint posture trajectories and the local synergy regions in this lower-dimensional space, why not learn the $A_k$ matrices in this space; i.e.
-$\mathcal{A}(q)\in \mathbb{R}^{p\times p}$ rather than $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ ?
-This seems like the obvious method, if we were considering our synergistic approach the classical way, i.e.
-using the lower-dimensional space as the new control variables.
-However, if we were to do that, this would mean that \eqref{eq:jtds} would become:
-\begin{equation}
-\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{M_p^{-1}}_{\mathbb{R}^{m\times p}}\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{p\times p}}\underbrace{M_p}_{\mathbb{R}^{p\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
-\label{eq:jtds_low}
-\end{equation}
-This is not entirely incorrect, however, with such a control law we can no longer guarantee stability and convergence to the attractor $x^{*}$.
-Let's consider the Lyapunov function used in Appendix I to prove stability of \eqref{eq:jtds} i.e.
-$V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*)$, The derivative of $ V $ wrt.
-time with \eqref{eq:jtds_low} is:
-\begin{equation}
-\begin{aligned}
-\frac{dV(q)}{dt} &= (H(q) - x^*)^TJ(q)\dot{q}\\
-&= -(H(q) - x^*)^TJ(q)\mathcal{A}(q)J^T(q)(H(q) - x^*)\\
-&= -(H(q) - x^*)^TJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{> 0}\underbrace{A_k}_{\succ 0}J^T(q)(H(q) - x^*)\leq 0
-\end{aligned}
-\end{equation}
-
-Note that this analysis is only considering PCA as the DR approach, with K-PCA deriving such a control law is more involved as the inverse mapping $\phi(q)^{-1}$ is not straightforward to compute.
 
@@
 
+\underline{As mentioned in the reviewers 3rd comment}, since we have an accurate representation of our joint posture trajectories and the local synergy regions in this lower-dimensional space, why not learn the $A_k$ matrices in this space; i.e.
+$\mathcal{A}(q)\in \mathbb{R}^{p\times p}$ rather than $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ ?
+This seems like the obvious method, if we were considering our synergistic approach the classical way, i.e.
+using the lower-dimensional space as the new control variables.
+However, if we were to do that, this would mean that \eqref{eq:jtds} would become:
+\begin{equation}
+\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{M_p^{-1}}_{\mathbb{R}^{m\times p}}\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{p\times p}}\underbrace{M_p}_{\mathbb{R}^{p\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
+\label{eq:jtds_low}
+\end{equation}
+This is not entirely incorrect, however, with such a control law we can no longer guarantee stability and convergence to the attractor $x^{*}$.
+By considering the Lyapunov function used in Appendix I to prove stability of \eqref{eq:jtds_low} i.e.
+$V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*)$, the derivative of $ V $ wrt.
+time with is:
+\begin{equation}
+\begin{aligned}
+\frac{dV(q)}{dt} &= (H(q) - x^*)^TJ(q)\dot{q}\\
+&= -(H(q) - x^*)^TJ(q)M_p^{-1}\mathcal{A}(q)M_pJ^T(q)(H(q) - x^*)\\
+&= -(H(q) - x^*)^TJ(q)\textcolor{red}{\underbrace{M_p^{-1}\big(\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{> 0}\underbrace{A_k}_{\succ 0}\big)M_p}_{\text{This should be}~~ \succ 0}}J^T(q)(H(q) - x^*)\leq 0
+\end{aligned}
+\end{equation}
+In order for \eqref{eq:jtds_low} to be asymptotically stable wrt.
+the attractor $x^*$ the term $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p \in \mathbb{R}^{m\times m}$ should undoubtedly be $\succ 0$.
+Even if $\theta_k(q) > 0$ and $A_k \succ 0$ for all $k$, the only way for this condition to hold is if $M_p$ where a full symmetric positive definite matrix.
+However, this is not the case because $M_p\in \mathbb{R}^{p\times m}$  is a projection matrix, i.e.
+$p<m$.
+The term $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$ will no longer be full-rank, it will preserve the $p$ positive eigenvalues from the $A_k$ matrices, but the remaining eigenvalues will be 0, thus generating a semi-positive definite matrix, which invalidates our stability proof.
+Note that, as described in the Appendix, $J(q)$ is not proven to be full rank in all regions of the workspace, when this happens, this means that the end-effector pose is not manipulable.
+Hence, we can only prove asymptotic stability when the end-effector pose is manipulable.
+This is a physical artifact from the joint-space constraints of the robot arm which we cannot avoid.
+Nevertheless, since the demonstrations are acquired via kinesthetic teaching we can assume that all of the regions in task-space are manipulable.
+Hence, when $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$ has some eigenvalues $=0$ the physical meaning behind it is that we are loosing control of these DOF of the arm which will restrain us from converging to the attractor and are no longer capable of accurately following the joint-space demonstrations.
+If instead we learn the synergy matrices in joint-space following \eqref{eq:jtds} and modulate them with the activation function in PCA-space, we avoid this problem of loosing rank altogether.
+
+\noindent \textbf{Remark 1:} This analysis is only considering PCA as the DR approach.
+If we would consider K-PCA, deriving such a control law is more involved as the inverse mapping $\phi(q)^{-1}$ is not straightforward.
+
+\noindent \textbf{Remark 2:} If we were to bypass the DR step our state-dependent system matrix would become $\mathcal{A}(q)= \sum\limits_{k=1}^{K}\theta(q)A_k$, which (via our definition) is a combination of ``local behavior synergy matrices" with the activation function living in joint-space rather than the PCA space; known as the \textit{postural synergy space}.
+The fact that we use PCA/K-PCA to represent the activation function $\theta(\cdot)$ is to alleviate the complexity of extracting the local synergy regions in joint-space directly.
+It is, indeed, also motivated by the fact that we know that multi-DOF joint postures can be accurately represented in a lower-dimensional space composed of the principal components of the postural trajectories; i.e.
+the well-known \textit{joint postural synergies}.
 As shown in TABLE I of the manuscript and discussed in Section V, by using a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
 Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations.
+\\
 
-On the applicability of F-PCA to our learning scheme.
+\noindent \underline{On the applicability of F-PCA to our learning scheme:} ...
+it is not applicable as we do not want to reconstruct the full trajectory, we only want find the local synergy region that the current joint posture $q$ belongs to..
+maybe apply F-PCA to this same demonstration...
 

### 2018-03-13 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
Latest model and review updates

@@
 % Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.
- To encode joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different motor synergies, and we provide a method for learning these synergy matrices from demonstrations.
+ To encode joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different behavior synergies, and we provide a method for learning these synergy matrices from demonstrations.
 Specifically, we use dimensionality reduction to find a low-dimensional embedding space for modulating joint synergies, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
@@
 &= -(H(q) - x^*)^TJ(q)\mathcal{A}(q)J^T(q)(H(q) - x^*)\\
-&= -(H(q) - x^*)^TJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{> 0}\underbrace{A_k}_{\succ 0}J^T(q)(H(q) - x^*)\leq 0
+&= -(H(q) - x^*)^TJ(q)\sum_{k=1}^{K}\underbrace{\theta_k(\phi(q))}_{> 0}\underbrace{A_k}_{\succ 0}J^T(q)(H(q) - x^*)\leq 0
 \end{aligned}
 \end{equation}
-When $J(q)$ is full-rank, the above inequality can be tightened to $\frac{dV(q)}{dt} > 0$.
+When $J(q)$ is full-rank, the above inequality can be tightened to $\frac{dV(q)}{dt} < 0$.
 Therefore, as $ q=q^* $ is the largest invariance in $ \mathbb{R}^m $, JT-DS \eqref{eq:ds} is stable with respect to a task-space attractor; i.e.

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m

@@ -5,7 +5,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %

### 2018-03-13 paper/reply_reviews/reply_comments.tex
New changes in revision and added q_init and x_target for comparison

@@
 (nature of the problem, complexity, feasibility)."}\\
-\textcolor{red}{\texttt{Response:} \small TODO..}\\
+\textcolor{red}{\texttt{Response:} \small The following has been added near Eq. (8):\\
+``The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective..."
+``This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD), and the number of terms in the objective is $O\left(k^2d^3\left( \sum_{n=1}^N T_n\right)\right)$."\\
+To elaborate about the prior calculation: the number of terms in the sum is proportional to the number of recorded points, times the number of matrices (k) required to generate the predicted velocity from (7) squared, times the number of terms in a matrix multiplication of two $d x d$ matrices ($d^3$).}\\
 

### 2018-03-13 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
New changes in revision and added q_init and x_target for comparison

@@
 Intuitively, each ``synergy region" in synergy space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
-Finding the appropriate synergy matrices $A_k$ to accurately reproduce the observed behaviors can be reduced to a convex semidefinite optimization, with the goal of minimizing \eqref{eq:optim_error}.
-To achieve this, the following optimization is proposed which uses mean square error as a means to minimize the joint velocity error from \eqref{eq:optim_error} as follows:
+Finding the appropriate synergy matrices $A_k$ to accurately reproduce the observed behaviors can be reduced to a semidefinite program with the goal of minimizing \eqref{eq:optim_error}.
+The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective, as follows:
 \begin{equation}
@@
  \eqref{eq:ds} with \eqref{eq:A}, and $x^*_{n}$ is defined as the endpoint of the $ n^{\text{th}} $ demonstrated trajectory.
+This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD), and the number of terms in the objective is roughly $O\left(k^2d^3\left( \sum_{n=1}^N T_n\right)\right)$.
 

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m

@@ -82,7 +82,7 @@ options = [];
 % To remove orientation from target 
 % simply set flag = 0 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-options.orientation_flag = 0; 
+options.orientation_flag = 1; 
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -117,7 +117,6 @@ robot = initialize_robot(A,D,Alpha,Qmin,Qmax);
 % Create a model plant for the robot's motor controller
 robotplant = RobotPlant(robot, 'end_trans');
 
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %       Split Dataset for Training/Testing        %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@ -125,7 +124,7 @@ tt_ratio = 0.6;
 train = round(length(Qs)*tt_ratio);
 Qs_train = []; Ts_train = [];
 Qs_test = [];   Ts_test = [];
-rand_ids = randsample(length(Qs),length(Qs))';
+rand_ids = [1:length(Qs)];
 for ii=1:length(Qs)
     if ii < train
         Qs_train{ii,1} = Qs{rand_ids(ii)}; Ts_train{ii,1} = Ts{rand_ids(ii)};
@@ -179,9 +178,13 @@ fprintf('Using %s mapping, got prediction RMSE on testing: %d \n', mapping_name,
 model_dir = strcat('./learned_JTDS_models/',choosen_dataset);
 mkdir(model_dir); 
 cd(model_dir)
-out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, latent_mapping.M')
+out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, latent_mapping.M', Data_train)
 
 % save mat file of variables
 M = latent_mapping.M';
-save('model.mat','Priors','Mu','Sigma', 'As', 'M')
+save('model.mat','Priors','Mu','Sigma', 'As', 'M', 'Data_train', 'robotplant')
 
+
+%% Plot Training Data
+figure('Color',[1 1 1])
+scatter3(Data_train(end-2,:),Data_train(end-1,:), Data_train(end,:))

### 2018-03-13 src/JTDS_mat_lib/utils/export2JSEDS_Cpp_lib_v2.m

@@ -1,4 +1,4 @@
-function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M)
+function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M, Data_train)
 % Export the parameters of a JT-DS model to a given file.
 % 
 % Arguments:
@@ -32,4 +32,12 @@ for i=1:size(Mu,2)
 end
 dlmwrite('M_matrix.txt', M,'newline','pc','-append','Delimiter','\t','precision','%.6f');
 
+q_init   = Data_train(1:7,1);
+if size(Data_train,1) == 23
+    x_target = Data_train(end-8:end,1);    
+else
+    x_target = Data_train(end-2:end,1);
+end
+dlmwrite('q_init.txt',q_init,'newline','pc','-append','Delimiter','\t','precision','%.6f');
+dlmwrite('x_target.txt',x_target,'newline','pc','-append','Delimiter','\t','precision','%.6f');
 out = true;

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
Uploaded models with no DR

@@ -74,7 +74,7 @@ end
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % Choose Lower-Dimensional Mapping Technique
-mapping = {'PCA'}; % 'None', 'PCA', 'KPCA'
+mapping = {'None'}; % 'None', 'PCA', 'KPCA'
 
 %%% Learning options %%%
 options = [];
@@ -137,9 +137,9 @@ end
 [Data_train, ~] = preprocess_demos_jtds(robotplant, Qs_train, Ts_train, options.tol_cutting,options.orientation_flag);
 [Data_test, ~]  = preprocess_demos_jtds(robotplant, Qs_test, Ts_test, options.tol_cutting,options.orientation_flag);
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%       Learn JTDS variants for current fold      %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%       Learn JTDS model for the Current Dataset      %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 mapping_name = mapping{1};
 fprintf('Training JTDS generator using %s mapping...\n', mapping_name);
 options.latent_mapping_type = mapping_name;
@@ -178,6 +178,11 @@ fprintf('Using %s mapping, got prediction RMSE on testing: %d \n', mapping_name,
 model_dir = strcat('./learned_JTDS_models/',choosen_dataset);
 mkdir(model_dir); 
 cd(model_dir)
+if strcmp(mapping_name, 'None')
+    M_p = eye(7);
+else
+    M_p = latent_mapping.M';
+end
 out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, latent_mapping.M', Data_train)
 
 % save mat file of variables

### 2018-03-13 src/JTDS_mat_lib/run_synergy_analysis.m

@@ -56,7 +56,7 @@ for i = 1:length(demo_ids)
     Ts{i,1} = Ts_{demo_ids(i)}(:, 1:thinning_ratio:end);
 end
 
-% Plot Full Set of Demonstrations per DOF
+%% Plot Full Set of Demonstrations per DOF
 if do_plots
     figure('Color',[1 1 1])    
     Data_ = [];

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
Uploaded models to include x_init and x_target for SEDS

@@ -164,6 +164,7 @@ rmse_test = mean(trajectory_error(motion_generator, Data_test(1:dimq, :), Data_t
 fprintf('Using %s mapping, got prediction RMSE on testing: %d \n', mapping_name, rmse_test);
 
 
+
 %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %      If you are happy with the results, export the model       %%
 %        for execution with the rtk-DS cpp class                 %%

### 2018-03-13 src/JTDS_mat_lib/run_synergy_analysis.m

@@ -209,3 +209,26 @@ for dof=1:7
 end
 legend('Raw','Reconstructed')
 
+
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  Compare executed motion with JT-DS vs Demonstrated one %%
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+dt = 0.1; % the timestep
+q_initial = [-0.728371429443359;-1.3605418920517;2.69252680319093;0.620675325393677;0.955035626888275;0.141930669546127;-0.17979271709919];
+x_targets = [[-.5; -.5; 0.3]];
+max_trajectory_duration = 60; % How long to interpolate the trajectory
+goal_tolerance = 0.05; % How far away from the goal we need to get to accept
+
+
+[Q_traj_learned, T_traj_learned] = computeFullTrajectory(q_initial, x_targets, motion_generator_learned, goal_tolerance, max_trajectory_duration);
+[Q_traj_unlearned, T_traj_unlearned] = computeFullTrajectory(q_initial, x_targets, motion_generator_unlearned, goal_tolerance, max_trajectory_duration);
+
+% Finally, display the learned motion
+if ~ishandle(fig)
+    fig = initialize_robot_figure(robot);
+else
+    figure(fig);
+end
+plot3(x_targets(1,:),x_targets(2,:),x_targets(3,:), 'ro', 'markersize', 20);
+PlaybackTrajectory(robotplant, Q_traj_learned, T_traj_learned, fig);
+

### 2018-03-13 src/JTDS_mat_lib/JTDS_Solver_v2.m
Updated JTDS models including mean of PCA for out of sample computation

@@ -199,7 +199,9 @@ function [Priors, Mu, Sigma, As, latent_mapping] = JTDS_Solver_v2(Data, robotpla
         otherwise
             error('Incorrectly specified a latent_mapping type. Was %s, but must be one of "PCA", "KPCA", "Autoencoder", or "None"', options.latent_mapping_type);
     end
-            
+    
+    latent_mapping.TransformedTrainingData = TransformedTrainingData;
+    
     if options.verbose
         disp(sprintf('Converted demonstrations into %d-D latent space using %s.', dimz, options.latent_mapping_type));
     end

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m

@@ -74,7 +74,7 @@ end
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % Choose Lower-Dimensional Mapping Technique
-mapping = {'None'}; % 'None', 'PCA', 'KPCA'
+mapping = {'PCA'}; % 'None', 'PCA', 'KPCA'
 
 %%% Learning options %%%
 options = [];
@@ -82,7 +82,7 @@ options = [];
 % To remove orientation from target 
 % simply set flag = 0 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-options.orientation_flag = 1; 
+options.orientation_flag = 0; 
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -163,7 +163,113 @@ fprintf('Using %s mapping, got prediction RMSE on training: %d \n', mapping_name
 rmse_test = mean(trajectory_error(motion_generator, Data_test(1:dimq, :), Data_test(dimq+1:2*dimq, :), Data_test(2*dimq+1:end, :),options.orientation_flag));
 fprintf('Using %s mapping, got prediction RMSE on testing: %d \n', mapping_name, rmse_test);
 
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%    Plot Lower-Dimensional Embedding and Synergies   %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Extract Lower Dimensional Embedding of Demonstrations
+figure('Color',[1 1 1])  
+for p=1:pca_dim
+    subplot(pca_dim,1,p)    
+    for i=1:length(Qs)
+        q_ref = Qs{i};
+        phi_q = out_of_sample(q_ref', latent_mapping)';
+        plot(phi_q(p,:),'-.'); hold on;
+    end      
+    grid on;
+    title(sprintf('Raw Trajectories in PCA-space $\\phi_%d(q)$',p), 'Interpreter', 'LaTex', 'Fontsize', 15)
+    xlabel('Time (samples)','Interpreter', 'LaTex', 'Fontsize', 15)   
+end
 
+% Plot in 3D-space
+K_colors = hsv(K);
+figure('Color',[1 1 1])  
+for i=1:length(Qs)
+    q_ref = Qs{i};
[…]

### 2018-03-13 src/JTDS_mat_lib/run_synergy_analysis.m

@@ -7,7 +7,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'pour'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %

### 2018-03-13 src/JTDS_mat_lib/utils/export2JSEDS_Cpp_lib_v2.m

@@ -1,4 +1,4 @@
-function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M, Data_train)
+function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M, PCA_mean, Data_train)
 % Export the parameters of a JT-DS model to a given file.
 % 
 % Arguments:
@@ -30,8 +30,8 @@ end
 for i=1:size(Mu,2)
     dlmwrite('A_matrix.txt',As(:,:,i),'newline','pc','-append','Delimiter',' ','precision','%.6f');
 end
-dlmwrite('M_matrix.txt', M,'newline','pc','-append','Delimiter','\t','precision','%.6f');
-
+dlmwrite('M_matrix.txt', M, 'newline','pc','-append','Delimiter','\t','precision','%.6f');
+dlmwrite('PCA_mean.txt', PCA_mean,'newline','pc','-append','Delimiter','\t','precision','%.6f');
 q_init   = Data_train(1:7,1);
 if size(Data_train,1) == 23
     x_target = Data_train(end-8:end,1);    

### 2018-03-13 src/JTDS_mat_lib/utils/my_plot3dGaussian.m

@@ -21,9 +21,9 @@ for k = 1:K
         d = max(d,0);
     end
     
-    if eig(Sigmas(:,:,k)) < 1e-3
-        d = d*100;
-    end
+%     if eig(Sigmas(:,:,k)) < 1e-3
+%         d = d*100;
+%     end
     
     d =  sqrt(d); % convert variance to sdwidth*sd
     bp = (v*d*ap) + repmat(Mus(:,k), 1, size(ap,2));

### 2018-03-13 src/JTDS_mat_lib/run_synergy_analysis.m
Updated pouring JTDS models

@@ -56,7 +56,7 @@ for i = 1:length(demo_ids)
     Ts{i,1} = Ts_{demo_ids(i)}(:, 1:thinning_ratio:end);
 end
 
-%% Plot Full Set of Demonstrations per DOF
+% Plot Full Set of Demonstrations per DOF
 if do_plots
     figure('Color',[1 1 1])    
     Data_ = [];
@@ -220,15 +220,6 @@ max_trajectory_duration = 60; % How long to interpolate the trajectory
 goal_tolerance = 0.05; % How far away from the goal we need to get to accept
 
 
-[Q_traj_learned, T_traj_learned] = computeFullTrajectory(q_initial, x_targets, motion_generator_learned, goal_tolerance, max_trajectory_duration);
-[Q_traj_unlearned, T_traj_unlearned] = computeFullTrajectory(q_initial, x_targets, motion_generator_unlearned, goal_tolerance, max_trajectory_duration);
 
-% Finally, display the learned motion
-if ~ishandle(fig)
-    fig = initialize_robot_figure(robot);
-else
-    figure(fig);
-end
-plot3(x_targets(1,:),x_targets(2,:),x_targets(3,:), 'ro', 'markersize', 20);
-PlaybackTrajectory(robotplant, Q_traj_learned, T_traj_learned, fig);
+
 

### 2018-03-13 src/JTDS_mat_lib/model/RobotPlant.m
.

@@ -42,7 +42,10 @@ classdef RobotPlant
             % method (i.e. qd = J^T(H(q) - xt) ).
             % In general, this will be multiplied by an A matrix to get the
             % augmented joint velocity.
-            basis = -1*transpose(obj.jacobian(q))*(obj.forward_kinematics(q) - xt);
+            temp_end=obj.robot.fkine(q);
+            temp_end=[temp_end(1:3,1);temp_end(1:3,2);temp_end(1:3,4)];
+            %             basis = -1*transpose(obj.jacobian(q))*(obj.forward_kinematics(q) - xt);
+            basis = -1*transpose([Jacobian_X(obj,q);Jacobian_Y(obj,q);obj.jacobian(q)])*(temp_end - xt);
         end
         
         function basis = qd_basis_orientation(obj, q, xt)
@@ -60,6 +63,12 @@ classdef RobotPlant
             T = obj.robot.fkine(q);
             x = T(1:3, 4);
         end
+        
+        function temp_end = end_pos_orien(obj,q)
+            temp_end=obj.robot.fkine(q);
+            temp_end=[temp_end(1:3,1);temp_end(1:3,2);temp_end(1:3,4)];
+        end
+        
         
         function j = end_trans_jacobian(obj, q)
             j = obj.robot.jacob0(q, 'trans');
@@ -93,7 +102,7 @@ classdef RobotPlant
             for ai=1:obj.robot.n
                 J(:,ai) = cross(Hoi{ai}(1:3,3),v2);
             end
-
+            
         end
         
         function J=Jacobian_Y(obj,q)
@@ -108,7 +117,7 @@ classdef RobotPlant
             for ai=1:obj.robot.n
                 J(:,ai) = cross(Hoi{ai}(1:3,3),v2);
             end
-
+            
         end
     end
     

### 2018-03-13 src/JTDS_mat_lib/model/MotionGeneratorBounded.m
.

@@ -23,23 +23,31 @@ classdef MotionGeneratorBounded < MotionGenerator
             end
         end
         
-        function qd = get_next_motion(obj, q_old, xt, dt) % Generates the next joint velocity given the joint position,
+        function qd = get_next_motion(obj, q_old, xt, dt,orientation_flag) % Generates the next joint velocity given the joint position,
             % desired position, a GMM model, and an augmentation matrix for
             % each Gaussian. dt is the timestep for our next movement
             % We compute the velocity by integrating our joint limitation
             % matrix "S" into our classic MotionGenerator controller, all
             % the while ensuring positive-semidefiniteness.
             S = obj.plant.compute_S(q_old);
-            qd = S*obj.compute_A(q_old)*S'*obj.plant.qd_basis(q_old, xt);
+            if orientation_flag==1
+                qd = S*obj.compute_A(q_old)*S'*obj.plant.qd_basis_orientation(q_old, xt);
+            else
+                qd = S*obj.compute_A(q_old)*S'*obj.plant.qd_basis(q_old, xt);
+            end
         end
         
-        function qd_fun = ODE_fun(obj, xt, dt)  % Yields a function handle
+        function qd_fun = ODE_fun(obj, xt, dt,orientation_flag)  % Yields a function handle
             % which computes the first order differential equation
             % associated with this system, i.e. f where qd = f(t, q)
             % Used as input to an ODE solver.
             function qd = qd_fun_tmp(t, q)
                 S = obj.plant.compute_S(q);
-                qd = S*obj.compute_A(q)*S*obj.plant.qd_basis(q,xt);
+                if orientation_flag==1
+                    qd = S*obj.compute_A(q)*S'*obj.plant.qd_basis_orientation(q, xt);
+                else
+                    qd = S*obj.compute_A(q)*S'*obj.plant.qd_basis(q, xt);
+                end
             end
             qd_fun = @qd_fun_tmp;
         end

### 2018-03-13 src/JTDS_mat_lib/model/RobotPlant.m

@@ -42,10 +42,7 @@ classdef RobotPlant
             % method (i.e. qd = J^T(H(q) - xt) ).
             % In general, this will be multiplied by an A matrix to get the
             % augmented joint velocity.
-            temp_end=obj.robot.fkine(q);
-            temp_end=[temp_end(1:3,1);temp_end(1:3,2);temp_end(1:3,4)];
-            %             basis = -1*transpose(obj.jacobian(q))*(obj.forward_kinematics(q) - xt);
-            basis = -1*transpose([Jacobian_X(obj,q);Jacobian_Y(obj,q);obj.jacobian(q)])*(temp_end - xt);
+                        basis = -1*transpose(obj.jacobian(q))*(obj.forward_kinematics(q) - xt);
         end
         
         function basis = qd_basis_orientation(obj, q, xt)
@@ -54,8 +51,13 @@ classdef RobotPlant
             % method (i.e. qd = J^T(H(q) - xt) ).
             % In general, this will be multiplied by an A matrix to get the
             % augmented joint velocity.
+            %             temp_end=obj.robot.fkine(q);
+            %             temp_end=[temp_end(1:3,1);temp_end(1:3,2);temp_end(1:3,4)];
+            %             basis = -1*transpose([Jacobian_X(obj,q);Jacobian_Y(obj,q);obj.jacobian(q)])*(temp_end - xt);
+
             temp_end=obj.robot.fkine(q);
             temp_end=[temp_end(1:3,1);temp_end(1:3,2);temp_end(1:3,4)];
+            %             basis = -1*transpose(obj.jacobian(q))*(obj.forward_kinematics(q) - xt);
             basis = -1*transpose([Jacobian_X(obj,q);Jacobian_Y(obj,q);obj.jacobian(q)])*(temp_end - xt);
         end
         

### 2018-03-13 src/JTDS_mat_lib/simulate/computeFullTrajectory.m

@@ -1,4 +1,4 @@
-function [Q, T] = computeFullTrajectory(q_initial, x_targets, motion_generator, goal_tolerance, max_duration)
+function [Q, T] = computeFullTrajectory(q_initial, x_targets, motion_generator, goal_tolerance, max_duration,orientation_flag)
 %COMPUTEFULLTRAJECTORY Interpolates a JT-DS trajectory
 %   This function integrates a JT-DS problem and computes the evolution of
 %   the system over a given timespan, using an ODE solver. Each target is
@@ -29,7 +29,7 @@ function [Q, T] = computeFullTrajectory(q_initial, x_targets, motion_generator, 
     q_start = q_initial;
     for i = 1:size(x_targets, 2)
         x_target = x_targets(:, i);
-        [T_subtraj, Q_subtraj] = ode15s(motion_generator.ODE_fun(x_target, 1), ...
+        [T_subtraj, Q_subtraj] = ode15s(motion_generator.ODE_fun(x_target, 1,orientation_flag), ...
             [0, max_duration], q_start, ...
             motion_generator.ODE_options(x_target, goal_tolerance));
         if isempty(T)

### 2018-03-13 src/JTDS_mat_lib/model/MotionGenerator.m
.

@@ -68,13 +68,18 @@ classdef MotionGenerator
         end
             
         
-        function qd_fun = ODE_fun(obj, xt, dt) % Yields a function handle
+        function qd_fun = ODE_fun(obj, xt, dt,orientation_flag) % Yields a function handle
             % which computes the first order differential equation
             % associated with this system, i.e. f where qd = f(q, t)
             % (dt is a placeholder for a child class, and can be ignored in
             % this case)
             % Used when solving the system using an ODE solver.
-            qd_fun = @(t, q) obj.compute_A(q)*obj.plant.qd_basis(q, xt);
+           
+           if orientation_flag==1
+                qd_fun = @(t, q) obj.compute_A(q)*obj.plant.qd_basis_orientation(q, xt);
+            else
+                qd_fun = @(t, q) obj.compute_A(q)*obj.plant.qd_basis(q, xt);
+            end
         end
         
         function options = ODE_options(obj, xt, goal_tolerance) % provides an options struct

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
updated testing script

@@ -82,7 +82,7 @@ options = [];
 % To remove orientation from target 
 % simply set flag = 0 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-options.orientation_flag = 1; 
+options.orientation_flag = 0; 
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -208,11 +208,39 @@ zlabel('$\phi_3(q)$', 'Interpreter', 'LaTex', 'Fontsize', 15)
 %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Reproduce learned motion and compute tasks-space metrics %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%% Pre-process Data %%%
+
+% Demonstration to test
 selected_demo = 1;
+
+%%% Extract task-space demonstrations %%%
 [demos_train_task] = preprocess_demos_seds(robotplant, Qs_train,options.orientation_flag)';
 task_space_traj = demos_train_task{selected_demo}; 
-x_target = task_space_traj(end-2:end,end);
+if options.orientation_flag
+    x_target = Data_train(end-8:end,1);    
+else
+    x_target = task_space_traj(:,end);
+end
+
+% Plot reproduced trajectory vs. demonstration in joint space
+figure('Color',[1 1 1])
+q_init         = Data_train(1:dimq,1);
+goal_tolerance = 0.001;
+max_trajectory_duration = 100;
+[Q_traj_JTDS, T_traj_JTDS] = computeFullTrajectory(q_init, x_target, motion_generator, ...
+                             goal_tolerance, max_trajectory_duration,options.orientation_flag);
+q_ref = Qs{selected_demo};
+for dof = 1:7
+    subplot(7,1,dof)  
+    plot(Q_traj_JTDS(dof,:),'-.','Color',[1 0 0], 'LineWidth',2);hold on;
+    plot(q_ref(dof,:),'-','Color', [0 0 0], 'LineWidth',2); hold on;
+    xlabel('Time (samples)','Interpreter', 'LaTex', 'Fontsize', 15)
+    ylabel('Angle (rad)','Interpreter', 'LaTex', 'Fontsize', 15)
+    grid on;
+end
[…]

### 2018-03-13 src/JTDS_mat_lib/run_synergy_analysis.m

@@ -209,17 +209,8 @@ for dof=1:7
 end
 legend('Raw','Reconstructed')
 
-
 %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%  Compare executed motion with JT-DS vs Demonstrated one %%
 %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-dt = 0.1; % the timestep
-q_initial = [-0.728371429443359;-1.3605418920517;2.69252680319093;0.620675325393677;0.955035626888275;0.141930669546127;-0.17979271709919];
-x_targets = [[-.5; -.5; 0.3]];
-max_trajectory_duration = 60; % How long to interpolate the trajectory
-goal_tolerance = 0.05; % How far away from the goal we need to get to accept
-
-
-
 
 

### 2018-03-13 src/JTDS_mat_lib/model/MotionGenerator.m
.

@@ -33,9 +33,9 @@ classdef MotionGenerator
         end
         
         function A = compute_A(obj, q)
-        % This function computes the weighted A matrix based on the current
-        % robot configuration. This combines the contributions of each of
-        % the local augmentation matrices (As).
+            % This function computes the weighted A matrix based on the current
+            % robot configuration. This combines the contributions of each of
+            % the local augmentation matrices (As).
             k = length(obj.Priors);
             for i = 1:k
                 h_raw(i) = obj.Priors(i)*mvnpdf(out_of_sample(q', obj.latent_mapping)', obj.Mu(:, i), obj.Sigma(:, :, i));
@@ -66,7 +66,7 @@ classdef MotionGenerator
             qd_basis = obj.plant.qd_basis_orientation(q, xt);
             qd = obj.compute_A(q)*qd_basis;
         end
-            
+        
         
         function qd_fun = ODE_fun(obj, xt, dt,orientation_flag) % Yields a function handle
             % which computes the first order differential equation
@@ -74,21 +74,25 @@ classdef MotionGenerator
             % (dt is a placeholder for a child class, and can be ignored in
             % this case)
             % Used when solving the system using an ODE solver.
-           
-           if orientation_flag==1
+            
+            if orientation_flag==1
                 qd_fun = @(t, q) obj.compute_A(q)*obj.plant.qd_basis_orientation(q, xt);
             else
                 qd_fun = @(t, q) obj.compute_A(q)*obj.plant.qd_basis(q, xt);
             end
         end
         
-        function options = ODE_options(obj, xt, goal_tolerance) % provides an options struct
+        function options = ODE_options(obj, xt, goal_tolerance,orientation_flag) % provides an options struct
             % for the ODE solver to optionally use, to add useful
             % functionality to the ODE
             function [value, is_terminal, direction] = event_fun(t, y)
                 value = zeros(size(t));
                 for i = 1:length(t)
-                    value(i) = norm(obj.plant.end_pos_orien(y(:, i)') - xt) - goal_tolerance;
+                    if orientation_flag==1
+                        value(i) = norm(obj.plant.end_pos_orien(y(:, i)') - xt) - goal_tolerance;
+                    else
+                        value(i) = norm(obj.plant.forward_kinematics(y(:, i)') - xt) - goal_tolerance;
+                    end
                 end
                 is_terminal = ones(size(t));
                 direction = -1*ones(size(t));

### 2018-03-13 src/JTDS_mat_lib/simulate/computeFullTrajectory.m

@@ -31,7 +31,7 @@ function [Q, T] = computeFullTrajectory(q_initial, x_targets, motion_generator, 
         x_target = x_targets(:, i);
         [T_subtraj, Q_subtraj] = ode15s(motion_generator.ODE_fun(x_target, 1,orientation_flag), ...
             [0, max_duration], q_start, ...
-            motion_generator.ODE_options(x_target, goal_tolerance));
+            motion_generator.ODE_options(x_target, goal_tolerance,orientation_flag));
         if isempty(T)
             T = T_subtraj;
         else

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
Added new SEDS model for pour

@@ -5,7 +5,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'pour'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'fore'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %
@@ -82,7 +82,7 @@ options = [];
 % To remove orientation from target 
 % simply set flag = 0 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-options.orientation_flag = 0; 
+options.orientation_flag = 1; 
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -181,53 +181,55 @@ for p=1:pca_dim
 end
 
 % Plot in 3D-space
-K_colors = hsv(K);
-figure('Color',[1 1 1])  
-for i=1:length(Qs)
-    q_ref = Qs{i};
-    phi_q = out_of_sample(q_ref', latent_mapping)';
-    % Hard clustering for each local model
-    labels =  my_gmm_cluster(phi_q, Priors, Mu, Sigma, 'hard', []);
-    for k=1:K        
-        phi_q_k   = phi_q(:,labels==k);
-        scatter3(phi_q_k(1,:),phi_q_k(2,:),phi_q_k(3,:),20,K_colors(k,:),'*'); hold on;
-    end    
-    scatter3(phi_q(1,end),phi_q(2,end),phi_q(3,end),100,'o','filled','MarkerEdgeColor','k','MarkerFaceColor',[0 0 0]); hold on;
-    axis tight
+if pca_dim == 3
+    K_colors = hsv(K);
+    figure('Color',[1 1 1])
+    for i=1:length(Qs)
+        q_ref = Qs{i};
+        phi_q = out_of_sample(q_ref', latent_mapping)';
+        % Hard clustering for each local model
+        labels =  my_gmm_cluster(phi_q, Priors, Mu, Sigma, 'hard', []);
+        for k=1:K
+            phi_q_k   = phi_q(:,labels==k);
[…]

### 2018-03-13 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
Uploaded all new JTDS models with q_init, x_target and mean PCA

@@ -5,7 +5,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'fore'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %
@@ -82,7 +82,7 @@ options = [];
 % To remove orientation from target 
 % simply set flag = 0 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-options.orientation_flag = 1; 
+options.orientation_flag = 0; 
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -326,7 +326,7 @@ err_pos = norm(JTDS_position(:,end) - task_space_traj(end-2:end,end))
 %% Compare position and orientation from CPP simulation (SEDS) vs Training Data == THIS WORKS! %%
 %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 figure('Color',[1 1 1])
-TheRobotTrajectory1 = TheRobotTrajectory2;
+% TheRobotTrajectory1 = TheRobotTrajectory2;
 JTDS_orientation = zeros(6,ceil(length(TheRobotTrajectory1)/10));
 for i=1:10:length(TheRobotTrajectory1)
     axangle = [TheRobotTrajectory1(i,1:3)/norm(TheRobotTrajectory1(i,1:3))' norm(TheRobotTrajectory1(i,1:3))]';

### 2018-03-14 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
Updated all JTDS models to store initial joint and target task configurations for all demonstrations

@@ -2,14 +2,14 @@
 %%     Learning JTDS Models (including orientation) on Different Datasets     %%                                                       %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% STEP 1: Load and Process dataset %
+%%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
 choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Load and Process dataset %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 switch choosen_dataset
     case 'back'
@@ -69,9 +69,9 @@ if do_plots
     end    
 end
 
-%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%    Train a JTDS model on the current dataset   %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%  STEP 2: Prepare Data for Learning and Initial Model Parameters   %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % Choose Lower-Dimensional Mapping Technique
 mapping = {'PCA'}; % 'None', 'PCA', 'KPCA'
@@ -120,7 +120,7 @@ robotplant = RobotPlant(robot, 'end_trans');
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %       Split Dataset for Training/Testing        %%
[…]

### 2018-03-14 src/JTDS_mat_lib/utils/export2JSEDS_Cpp_lib_v2.m

@@ -1,4 +1,4 @@
-function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M, PCA_mean, Data_train)
+function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M, PCA_mean, Data_train, index_train)
 % Export the parameters of a JT-DS model to a given file.
 % 
 % Arguments:
@@ -32,12 +32,21 @@ for i=1:size(Mu,2)
 end
 dlmwrite('M_matrix.txt', M, 'newline','pc','-append','Delimiter','\t','precision','%.6f');
 dlmwrite('PCA_mean.txt', PCA_mean,'newline','pc','-append','Delimiter','\t','precision','%.6f');
-q_init   = Data_train(1:7,1);
-if size(Data_train,1) == 23
-    x_target = Data_train(end-8:end,1);    
-else
-    x_target = Data_train(end-2:end,1);
+
+q_init    = [];
+x_target  = [];
+
+for l = 1:length(index_train-1)
+    q_init = [q_init Data_train(1:7,l)]
+    
+    if size(Data_train,1) == 23
+        x_target = [x_target Data_train(end-8:end,l)];
+    else
+        x_target = [x_target Data_train(end-2:end,l)];
+    end
+    
 end
+
 dlmwrite('q_init.txt',q_init,'newline','pc','-append','Delimiter','\t','precision','%.6f');
 dlmwrite('x_target.txt',x_target,'newline','pc','-append','Delimiter','\t','precision','%.6f');
 out = true;

### 2018-03-14 src/JTDS_mat_lib/utils/preprocess_demos_jtds.m

@@ -5,7 +5,7 @@ function [Data, index] = preprocess_demos_jtds(robotplant, demos, time, tol_cutt
 % trims the data. The function can be
 % called using:
 %
-%          [Data, index] = preprocess_demos(demos,time,tol_cutting)
+%          [Data, index, data] = preprocess_demos(demos,time,tol_cutting)
 %
 % Inputs -----------------------------------------------------------------
 %
@@ -105,7 +105,7 @@ for i=1:length(demos)
     
     % saving demos next to each other
     Data = [Data tmp_full];
-    index = [index size(Data,2)+1];
+    index = [index size(Data,2)+1];    
 end
 
 end

### 2018-03-14 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
Corrected JTDS models

@@ -9,7 +9,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'foot'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 switch choosen_dataset
     case 'back'
@@ -46,7 +46,7 @@ end
 % demonstrated points (by varying "thinning_ratio", so long as there are still sufficient points to
 % satisfactorily reconstruct the shape of the trajectory.
 % In the KUKA case, we get 500 datapoints per second, so we recommend shrinking the data density considerably
-thinning_ratio = 20; % Same as demonstrations recorded at 10->50Hz, 20->25Hz
+thinning_ratio = 2; % Same as demonstrations recorded at 10->50Hz, 20->25Hz
 Qs = []; Ts= [];
 for i = 1:length(demo_ids)
     Qs{i,1} = Qs_{demo_ids(i)}(:, 1:thinning_ratio:end);
@@ -82,7 +82,7 @@ options = [];
 % To remove orientation from target 
 % simply set flag = 0 %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-options.orientation_flag = 0; 
+options.orientation_flag = 1; 
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -216,11 +216,11 @@ end
 model_dir = strcat('./learned_JTDS_models/',choosen_dataset);
 mkdir(model_dir); 
 cd(model_dir)
-if strcmp(mapping_name, 'None')
-    M_p = eye(7);
-else
-    M_p = latent_mapping.M';
-end
+% if strcmp(mapping_name, 'None')
+%     M_p = eye(7);
+% else
+%     M_p = latent_mapping.M';
+% end
 out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, latent_mapping.M', latent_mapping.mean, Data_train, index_train);
 
 % save mat file of variables

### 2018-03-14 src/JTDS_mat_lib/utils/export2JSEDS_Cpp_lib_v2.m

@@ -1,4 +1,4 @@
-function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M, PCA_mean, Data_train, index_train)
+function out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, M, PCA_mean, Data, index)
 % Export the parameters of a JT-DS model to a given file.
 % 
 % Arguments:
@@ -33,19 +33,31 @@ end
 dlmwrite('M_matrix.txt', M, 'newline','pc','-append','Delimiter','\t','precision','%.6f');
 dlmwrite('PCA_mean.txt', PCA_mean,'newline','pc','-append','Delimiter','\t','precision','%.6f');
 
-q_init    = [];
-x_target  = [];
-
-for l = 1:length(index_train-1)
-    q_init = [q_init Data_train(1:7,l)]
-    
-    if size(Data_train,1) == 23
-        x_target = [x_target Data_train(end-8:end,l)];
-    else
-        x_target = [x_target Data_train(end-2:end,l)];
-    end
-    
+if index(end) > length(Data)
+    num_demos = length(index)-1;
+else
+    num_demos = length(index);
 end
+
+q_init    = zeros(7, num_demos);
+if size(Data,1) == 23
+    x_target  = zeros(9, num_demos);
+else
+    x_target  = zeros(3, num_demos);
+end
+
+for l = 1:num_demos
+    q_init(:,l) = Data(1:7,index(l));
+    if size(Data,1) == 23
+        x_target(:,l)  = Data(end-8:end,index(l));
+    else
+        x_target(:,l)  = Data(end-2:end,index(l));
+    end
+end
+
+index
+q_init
+x_target
 
 dlmwrite('q_init.txt',q_init,'newline','pc','-append','Delimiter','\t','precision','%.6f');
 dlmwrite('x_target.txt',x_target,'newline','pc','-append','Delimiter','\t','precision','%.6f');

### 2018-03-14 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m
Updated all SEDS models with x_init, q_init, and x_targets

@@ -216,11 +216,11 @@ end
 model_dir = strcat('./learned_JTDS_models/',choosen_dataset);
 mkdir(model_dir); 
 cd(model_dir)
-% if strcmp(mapping_name, 'None')
-%     M_p = eye(7);
-% else
-%     M_p = latent_mapping.M';
-% end
+if strcmp(mapping_name, 'None')
+    M_p = eye(7);
+else
+    M_p = latent_mapping.M';
+end
 out = export2JSEDS_Cpp_lib_v2(Priors, Mu, Sigma, As, latent_mapping.M', latent_mapping.mean, Data_train, index_train);
 
 % save mat file of variables
@@ -234,7 +234,7 @@ save('model.mat','Priors','Mu','Sigma', 'As', 'M', 'Data_train', 'index_train','
 %% Reproduce learned motion and compute tasks-space metrics == Works with JTDS on position only! %%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %%% Extract task-space demonstrations %%%
-% Hacks to know which demos to use
+% Hacks to know which demos to use for comparison
 selected_demo = 1;
 num = 59;
 task_space_traj = zeros(9,num);

### 2018-03-14 src/JTDS_mat_lib/crossvalidation_dr_kuka_wOrientation.m
Corrected 'back' JTDS model

@@ -1,197 +1,0 @@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Evaluation of Learning Schemes for JTDS Models (including orientation) on Different Datasets     %%
-%  Train and compare a series of JTDS models with different dimensionality                          %%   
-%  reduction schemes and GMM model fitting                                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-clear all; close all; clc;
-do_plots  = 1;
-data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Load and Process dataset %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-switch choosen_dataset
-    case 'back'
-        demos_location = strcat(data_path, 'back_hand/data.mat');
-        demo_ids = [2:11];
-    case 'fore'
-        demos_location = strcat(data_path,'fore_hand/data.mat');
-        demo_ids = [1:11];
-    case 'pour' 
-        demos_location = strcat(data_path,'pour_no_obst/data.mat');
-        demo_ids = [1 2 3 5 6 7 8 9 10];
-    case 'pour_obst'
-        demos_location = strcat(data_path,'pour_obst/data.mat');
-        demo_ids = [1:10];
-    case 'pour_obst_2'
-        demos_location = strcat(data_path,'pour_obst_2/data.mat');
-        demo_ids = [1:7];                
-    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
-        demos_location = strcat(data_path,'foot/data.mat');
-        demo_ids = [1:8];                
-    case 'singularity'   
-        demos_location = strcat(data_path,'singularity/data.mat');
-        demo_ids = [1:10];  
[…]

### 2018-03-14 src/JTDS_mat_lib/demo_learn_JTDS_kuka.m

@@ -9,7 +9,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'foot'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 switch choosen_dataset
     case 'back'
@@ -46,7 +46,7 @@ end
 % demonstrated points (by varying "thinning_ratio", so long as there are still sufficient points to
 % satisfactorily reconstruct the shape of the trajectory.
 % In the KUKA case, we get 500 datapoints per second, so we recommend shrinking the data density considerably
-thinning_ratio = 2; % Same as demonstrations recorded at 10->50Hz, 20->25Hz
+thinning_ratio = 20; % Same as demonstrations recorded at 10->50Hz, 20->25Hz
 Qs = []; Ts= [];
 for i = 1:length(demo_ids)
     Qs{i,1} = Qs_{demo_ids(i)}(:, 1:thinning_ratio:end);

### 2018-03-14 src/JTDS_mat_lib/run_crossval_hyperparam_kpca_kuka_wOrientation.m
Added script for cross-validation of kpca DR approacg with orientation in task-space

@@ -1,0 +1,252 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Evaluation of Learning Schemes for JTDS Models on Different Datasets     %%
+%  Train and compare a series of JTDS models with different dimensionality  %%   
+%  reduction schemes and GMM model fitting                                  %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour'
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'pour_obst_2'
+        demos_location = strcat(data_path,'pour_obst_2/data.mat');
+        demo_ids = [1:7];
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
+        demos_location = strcat(data_path,'singularity/data.mat');
+        demo_ids = [1:10];  
+        fprintf('Loading demonstrations from %s \n', demos_location);
+        load(demos_location)
+end
+
+if ~strcmp(choosen_dataset,'singularity')
[…]

### 2018-03-14 Data/Comparing_SEDS_with_JT_DS/Compare_SEDS_JTDS.m
Comparison is added

@@ -1,0 +1,98 @@
+close all
+clc
+clear
+load('JT_DS_comparison.mat')
+load('SEDS_comparison.mat')
+
+%%
+
+% prepare data
+
+data=vertcat(dist_P_JT_cell,dist_P_cell);
+
+xlab={'Backward','Forward','Pouring','Obstacle 1',' Obstacle 2','Foot-Step','Singularity Motions'};
+col=[102,255,255, 100;
+0, 0, 255, 100];
+col=col/255;
+figure1 = figure;
+
+axes1 = axes('Parent',figure1);
+multiple_boxplot(data',xlab,{'JT-DS', 'SEDS+IK'},col')
+
+xlabel('Scenario','Interpreter','latex');
+title('Distance between the demonstrated end-effector position and the corresponding execution ',...
+    'Interpreter','latex');
+ylabel('$m^2$','Interpreter','latex');
+box(axes1,'on');
+grid(axes1,'on');
+% Set the remaining axes properties
+set(axes1,'FontSize',22,'TickLabelInterpreter','latex')
+
+%%
+
+data=vertcat(dist_O_JT_cell,dist_O_cell);
+
+xlab={'Backward','Forward','Pouring','Obstacle 1',' Obstacle 2','Foot-Step','Singularity Motions'};
+col=[102,255,255, 100;
+0, 0, 255, 100];
+col=col/255;
+figure1 = figure;
+
+axes1 = axes('Parent',figure1);
+multiple_boxplot(data',xlab,{'JT-DS', 'SEDS+IK'},col')
+
+xlabel('Scenario','Interpreter','latex');
+title('Distance between the demonstrated end-effector orientation and the corresponding execution ',...
+    'Interpreter','latex');
+box(axes1,'on');
+grid(axes1,'on');
+% Set the remaining axes properties
+set(axes1,'FontSize',22,'TickLabelInterpreter','latex')
+
+
+%%
+
+data=vertcat(dist_P_JT_E_JT_cell,dist_P_E_cell);
+
+xlab={'Backward','Forward','Pouring','Obstacle 1',' Obstacle 2','Foot-Step','Singularity Motions'};
+col=[102,255,255, 100;
+0, 0, 255, 100];
+col=col/255;
+figure1 = figure;
+
+axes1 = axes('Parent',figure1);
+multiple_boxplot(data',xlab,{'JT-DS', 'SEDS+IK'},col')
+
+xlabel('Scenario','Interpreter','latex');
[…]

### 2018-03-14 Data/Comparing_SEDS_with_JT_DS/Compre_the_error_JTDS.m

@@ -1,0 +1,196 @@
+clc
+close all
+clear
+List={'back','fore','pour','pour_obst','pour_obst_2','foot','singularity'};
+
+
+for ii=1:7
+    choosen_dataset=List{ii};
+    clearvars -except dist_P_JT dist_O_JT LABEL_JT Data_E_O_JT Data_E_P_JT Data_T_O_JT Data_T_P_JT choosen_dataset ii List dist_P_JT_E_JT dist_O_JT_E_JT
+    switch choosen_dataset
+        case 'back'
+            Name='Backward Reaching';
+            load('JT_DS/back/model.mat')
+            Data_E{1}=importfile_JTDS('JT_DS/back/TheRobotTrajectory1.txt');
+            Data_E{2}=importfile_JTDS('JT_DS/back/TheRobotTrajectory2.txt');
+            Data_E{3}=importfile_JTDS('JT_DS/back/TheRobotTrajectory3.txt');
+            Data_E{4}=importfile_JTDS('JT_DS/back/TheRobotTrajectory4.txt');
+            Data_E{5}=importfile_JTDS('JT_DS/back/TheRobotTrajectory5.txt');
+        case 'fore'
+            Name='Forward Reaching';
+            load('JT_DS/fore_hand/model.mat')
+            Data_E{1}=importfile_JTDS('JT_DS/fore_hand/TheRobotTrajectory1.txt');
+            Data_E{2}=importfile_JTDS('JT_DS/fore_hand/TheRobotTrajectory2.txt');
+            Data_E{3}=importfile_JTDS('JT_DS/fore_hand/TheRobotTrajectory3.txt');
+            Data_E{4}=importfile_JTDS('JT_DS/fore_hand/TheRobotTrajectory4.txt');
+            Data_E{5}=importfile_JTDS('JT_DS/fore_hand/TheRobotTrajectory5.txt');
+            Data_E{6}=importfile_JTDS('JT_DS/fore_hand/TheRobotTrajectory6.txt');
+        case 'pour'
+            Name='Pouring';
+            load('JT_DS/pour/model.mat')
+            Data_E{1}=importfile_JTDS('JT_DS/pour/TheRobotTrajectory1.txt');
+            Data_E{2}=importfile_JTDS('JT_DS/pour/TheRobotTrajectory2.txt');
[…]

### 2018-03-14 Data/Comparing_SEDS_with_JT_DS/Compre_the_error_SEDS.m

@@ -1,0 +1,197 @@
+clc
+close all
+clear
+List={'back','fore','pour','pour_obst','pour_obst_2','foot','singularity'};
+
+
+for ii=1:7
+    choosen_dataset=List{ii};
+    clearvars -except dist_P dist_O LABEL Data_E_O Data_E_P Data_T_O Data_T_P choosen_dataset ii List dist_P_E dist_O_E
+    switch choosen_dataset
+        case 'back'
+            Name='Backward Reaching';
+            load('SEDS/back/model.mat')
+            Data_E{1}=importfile_SEDS('SEDS/back/TheRobotTrajectory1.txt');
+            Data_E{2}=importfile_SEDS('SEDS/back/TheRobotTrajectory2.txt');
+            Data_E{3}=importfile_SEDS('SEDS/back/TheRobotTrajectory3.txt');
+            Data_E{4}=importfile_SEDS('SEDS/back/TheRobotTrajectory4.txt');
+            Data_E{5}=importfile_SEDS('SEDS/back/TheRobotTrajectory5.txt');
+        case 'fore'
+            Name='Forward Reaching';
+            load('SEDS/fore_hand/model.mat')
+            Data_E{1}=importfile_SEDS('SEDS/fore_hand/TheRobotTrajectory1.txt');
+            Data_E{2}=importfile_SEDS('SEDS/fore_hand/TheRobotTrajectory2.txt');
+            Data_E{3}=importfile_SEDS('SEDS/fore_hand/TheRobotTrajectory3.txt');
+            Data_E{4}=importfile_SEDS('SEDS/fore_hand/TheRobotTrajectory4.txt');
+            Data_E{5}=importfile_SEDS('SEDS/fore_hand/TheRobotTrajectory5.txt');
+            Data_E{6}=importfile_SEDS('SEDS/fore_hand/TheRobotTrajectory6.txt');
+        case 'pour'
+            Name='Pouring';
+            load('SEDS/pour/model.mat')
+            Data_E{1}=importfile_SEDS('SEDS/pour/TheRobotTrajectory1.txt');
+            Data_E{2}=importfile_SEDS('SEDS/pour/TheRobotTrajectory2.txt');
+            Data_E{3}=importfile_SEDS('SEDS/pour/TheRobotTrajectory3.txt');
[…]

### 2018-03-14 Data/Comparing_SEDS_with_JT_DS/importfile_JTDS.m

@@ -1,0 +1,79 @@
+function TheRobotTrajectory1 = importfile_JTDS(filename, startRow, endRow)
+%IMPORTFILE1 Import numeric data from a text file as a matrix.
+%   THEROBOTTRAJECTORY1 = IMPORTFILE1(FILENAME) Reads data from text file
+%   FILENAME for the default selection.
+%
+%   THEROBOTTRAJECTORY1 = IMPORTFILE1(FILENAME, STARTROW, ENDROW) Reads
+%   data from rows STARTROW through ENDROW of text file FILENAME.
+%
+% Example:
+%   TheRobotTrajectory1 = importfile1('TheRobotTrajectory1.txt', 1, 10909);
+%
+%    See also TEXTSCAN.
+
+% Auto-generated by MATLAB on 2018/03/14 21:29:58
+
+%% Initialize variables.
+delimiter = ' ';
+if nargin<=2
+    startRow = 1;
+    endRow = inf;
+end
+
+%% Format string for each line of text:
+%   column1: double (%f)
+%	column2: double (%f)
+%   column3: double (%f)
+%	column4: double (%f)
+%   column5: double (%f)
+%	column6: double (%f)
+%   column7: double (%f)
+%	column8: double (%f)
+%   column9: double (%f)
+%	column10: double (%f)
+%   column11: double (%f)
+%	column12: double (%f)
+%   column13: double (%f)
+%	column14: double (%f)
+%   column15: double (%f)
+%	column16: double (%f)
+%   column17: double (%f)
+%	column18: double (%f)
+%   column19: double (%f)
+%	column20: double (%f)
+%   column21: double (%f)
+%	column22: double (%f)
+%   column23: double (%f)
+%	column24: double (%f)
+%   column25: double (%f)
+%	column26: double (%f)
+% For more information, see the TEXTSCAN documentation.
+formatSpec = '%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%[^\n\r]';
+
+%% Open the text file.
+fileID = fopen(filename,'r');
+
+%% Read columns of data according to format string.
+% This call is based on the structure of the file used to generate this
+% code. If an error occurs for a different file, try regenerating the code
+% from the Import Tool.
[…]

### 2018-03-14 Data/Comparing_SEDS_with_JT_DS/importfile_SEDS.m

@@ -1,0 +1,73 @@
+function TheRobotTrajectory1 = importfile_SEDS(filename, startRow, endRow)
+%IMPORTFILE1 Import numeric data from a text file as a matrix.
+%   THEROBOTTRAJECTORY1 = IMPORTFILE1(FILENAME) Reads data from text file
+%   FILENAME for the default selection.
+%
+%   THEROBOTTRAJECTORY1 = IMPORTFILE1(FILENAME, STARTROW, ENDROW) Reads
+%   data from rows STARTROW through ENDROW of text file FILENAME.
+%
+% Example:
+%   TheRobotTrajectory1 = importfile1('TheRobotTrajectory1.txt', 1, 7137);
+%
+%    See also TEXTSCAN.
+
+% Auto-generated by MATLAB on 2018/03/14 16:52:40
+
+%% Initialize variables.
+delimiter = ' ';
+if nargin<=2
+    startRow = 1;
+    endRow = inf;
+end
+
+%% Format string for each line of text:
+%   column1: double (%f)
+%	column2: double (%f)
+%   column3: double (%f)
+%	column4: double (%f)
+%   column5: double (%f)
+%	column6: double (%f)
+%   column7: double (%f)
+%	column8: double (%f)
+%   column9: double (%f)
+%	column10: double (%f)
+%   column11: double (%f)
+%	column12: double (%f)
+%   column13: double (%f)
+%	column14: double (%f)
+%   column15: double (%f)
+%	column16: double (%f)
+%   column17: double (%f)
+%	column18: double (%f)
+%   column19: double (%f)
+%	column20: double (%f)
+% For more information, see the TEXTSCAN documentation.
+formatSpec = '%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%f%[^\n\r]';
+
+%% Open the text file.
+fileID = fopen(filename,'r');
+
+%% Read columns of data according to format string.
+% This call is based on the structure of the file used to generate this
+% code. If an error occurs for a different file, try regenerating the code
+% from the Import Tool.
+dataArray = textscan(fileID, formatSpec, endRow(1)-startRow(1)+1, 'Delimiter', delimiter, 'MultipleDelimsAsOne', true, 'EmptyValue' ,NaN,'HeaderLines', startRow(1)-1, 'ReturnOnError', false);
+for block=2:length(startRow)
+    frewind(fileID);
[…]

### 2018-03-14 Data/Comparing_SEDS_with_JT_DS/multiple_boxplot.m

@@ -1,0 +1,81 @@
+function multiple_boxplot(data,xlab,Mlab,colors)
+
+% data is a cell matrix of MxL where in each element there is a array of N
+% length. M is how many data for the same group, L, how many groups.
+%
+% Optional:
+% xlab is a cell array of strings of length L with the names of each
+% group
+%
+% Mlab is a cell array of strings of length M
+%
+% colors is a Mx4 matrix with normalized RGBA colors for each M.
+
+% check that data is ok.
+if ~iscell(data)
+    error('Input data is not even a cell array!');
+end
+
+% Get sizes
+M=size(data,2);
+L=size(data,1);
+if nargin>=4
+    if size(colors,2)~=M
+        error('Wrong amount of colors!');
+    end
+end
+if nargin>=2
+    if length(xlab)~=L
+        error('Wrong amount of X labels given');
+    end
+end
+
+% Calculate the positions of the boxes
+positions=1:0.25:M*L*0.25+1+0.25*L;
+positions(1:M+1:end)=[];
+
+% Extract data and label it in the group correctly
+x=[];
+group=[];
+for ii=1:L
+    for jj=1:M
+        aux=data{ii,jj};
+        x=vertcat(x,aux(:));
+        group=vertcat(group,ones(size(aux(:)))*jj+(ii-1)*M);
+    end
+end
+% Plot it
+boxplot(x,group, 'positions', positions);
+
+% Set the Xlabels
+aux=reshape(positions,M,[]);
+labelpos = sum(aux,1)./M;
+
+set(gca,'xtick',labelpos)
+if nargin>=2
+    set(gca,'xticklabel',xlab);
+else
+    idx=1:L;
+    set(gca,'xticklabel',strsplit(num2str(idx),' '));
+end
+
+
+% Get some colors
+if nargin>=4
+    cmap=colors;
+else
+    cmap = hsv(M);
+    cmap=vertcat(cmap,ones(1,M)*0.5);
+end
+color=repmat(cmap, 1, L);
+
+% Apply colors
+h = findobj(gca,'Tag','Box');
+for jj=1:length(h)
+   patch(get(h(jj),'XData'),get(h(jj),'YData'),color(1:3,jj)','FaceAlpha',color(4,jj));
+end
+
+if nargin>=3
+    legend(fliplr(Mlab));
+end
+end

### 2018-03-14 paper/reply_reviews/New_reply_to_comments.tex
Writing reply to comments and changing paper accordingly

\documentclass{article}
\usepackage[left=2cm, right=2cm, top=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\title{Response to Reviewers: Manuscript ID 17-1142.1: ``Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A Linear Parameter Varying and Synergetic Control Approach"}

%\author{Your Name  \\
%	Your Company / University  \\
%	\and 
%	The Other Dude \\
%	His Company / University \\
%	}
%
%\date{\today}
% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
% before or after the \begin{document} command 
% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
\begin{document}
\maketitle
First of all, we would like to thank all reviewers for the time they spent reviewing our paper. We greatly appreciate the comments, suggestions and opinions that we received. We have addressed all the necessary comments and suggestions in the attached manuscript accordingly.\\
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\noindent \underline{Comments from Reviewer \#1 (Reviewer: 4 (Review30069)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``As mentioned by the authors, there are many reasons to plan in joint space rather than task space. One of the important reasons to me seems to be able to plan end-effector orientation, not just position. Traditional movement primitives, for example DMPs, do not explicitly encode planning in the SO(3) space of orientations. There is research that addresses this question, for example [1], but they lose the simplistic appeal of movement primitives. I think it would be a great addition to this paper if the authors illustrate if their joint space DS can be used to accomplish orientation sensitive tasks, and generalize to end-effector orientations different that what was demonstrated.[1] Ude, Ale, et al. "Orientation in cartesian space dynamic movement primitives." Robotics and Automation (ICRA), 2014 IEEE International Conference on.}}"\\
\texttt{Response:}  We highly appreciate this comment. When we initially proposed this approach, we did not think about tackling the problem of reaching a target in orientation as well. Since our DS is learned in joint-space, indeed, we do not have to deal with the challenges that arise when trying to reach a target in both position and orientation as we do not have the constrains imposed by SO(3). There was theoretically no modification needed in our controller:
\begin{equation}
\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{m\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
\label{eq:jtds}
\end{equation}
Originally, when working solely in position $d=3$, now we represent the target with $d=9$, where the first 3 dimensions correspond to Cartesian position and the remaining 6 correspond to the first and second columns of a rotation matrix $R \in SO(3)$. This only modifies the Jacobian $J(q)$ which is now $J(q)\in \mathbb{R}^{9 \times 7}$ rather than $J(q)\in \mathbb{R}^{3 \times 7}$. The role of the $\mathcal{A}(q)$ and their meaning is the same.

\item \textcolor{blue}{\textit{``In the introduction, the authors stress that jacobian psuedo inverse
can be computationally expensive. Is this really true for a 7-dof robot
arm? Can the authors provide some numbers where this would really be an
issue for a robot controlled at 1000Hz? I find the argument that
jacobian inverse does not work for singularities and needs a set of
heuristics to be more compelling. Maybe the authors should rewrite the
introduction to stress more on this issue?"}}\\
\texttt{Response:} \small For a 7-DOF robot it is indeed negligible. Computational burden only becomes a real issue for higher degrees of freedom, like a multi-fingered hand or multi-arm robotic setting. We will remove this passage from the text as we don't have much space to discuss it....\\

\item \textcolor{blue}{\textit{``In Section II, criterion I, the authors talk about a positive symmetric matrix P. I do not see it in the
equations above. Is this a weighing matrix?"}}\\
\texttt{Response:} \small Removed, that was a carry-over from a previous formulation.\\

\item \textcolor{blue}{\textit{``In criterion II, the error is defined with respect to the velocity error, rather than a position or a combination
of both. This means that the motion can be different from the demonstrated motion. I understand
that this is important to keep the later optimization convex, but is there any disadvantage to using
velocity error rather than position error?"}}\\
\texttt{Response:} \small Optimizing for position distance alone would have the effect of learning a motion that converged to the average of previous motions, thus not changing the motion in new situations or fitting the joint velocity profile that is often a desired component of the motion. Using velocity error does mean that, for motions strongly dependent on joint position rather than joint velocity, the generated motions may generalize from the demonstrations in unexpected ways. However, as you mentioned, minimizing position error would, depending on the formulation, be prohibitively difficult. This would imply simulating the learned motion forward in time to recompute the robot's positions, making the optimization much more complicated; this way, we can treat each point in the motion independently.\\

\item \textcolor{blue}{\textit{`I do not completely understand the significance of A(q). From Figure 3, it seems like A(q) helps resolve
redundancy, while achieving the task space goal. Do the authors have a mathematical (in constrast to
intuitive) explanation for this? Does A(q) somehow relate to the null-space of the jacobian?"}}\\
\texttt{Response:} \small A(q) transforms a joint velocity vector approximately aligned with the task-space error into a different joint velocity vector that remains approximately aligned with the error, but which we want to choose to more-closely approximate our demonstrations. In this sense, it is resolving redundancy between all the possible joint velocity vectors that would move the controller closer to the origin. The paragraph in Section III has been modified to improve clarity. A(q) doesn't relate to the null-space of the Jacobian in any clear way, as its value can and does alter the task-space trajectory as well as the joint space trajectory. The only property that A(q) preserves is the joint velocity's moving the task space position closer to the goal, but the direction of motion is not necessarily the same as the direction of task-space error.\\

\item \textcolor{blue}{\textit{``In Section III, ?One can intuitively understand the control law as follows:$(H(q) - x^*)$ denotes the
position error w.r.t. the task target and by multiplying that error by the transposed Jacobian JT (q),
the error is projected into joint space (similar to Jacobian transpose control)?. I find this explanation
misleading as jacobian multiplied to position error does not project the error into joint space. Jacobian
transforms velocities, not positions. Maybe the law can be seen more as a first order approximation
of the error in task space. In any case, I would either clarify this statement more or take it out of the
manuscript all together. It seems like the DS is just a PD law that provably converges to a fixed attractor
in task space."}}\\
\texttt{Response:} \small The passage has been rewritten to remove the word "project", which you appropriately noted was incorrect, and further clarifies the nature of the DS as a PD controller. We have added the following text in the manuscript: 
\textit{``One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target, which we then reinterpret as the task-space velocity of a proportional controller. By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control [?]), see Fig. [?]. The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig. [?] illustrates the effects of  $\mathcal{A}(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space.}\\

\item \textcolor{blue}{\textit{``About figure 2 again, the position error in task and joint space are not simply multiplied by a Jacobian.
Although, task space error can be first-order approximated using a Jacobian multiplication with
joint space error. What does the right figure in Fig 2 exactly represent?"}}\\
\texttt{Response:} \small The figure on the left is simply the cartesian error in task space. The figure on the left is the Lyapunov potential function on which the jacobian transpose controller descends. In showing the attractive regions in joint-space corresponding to this Lyapunov function, we aim to offer insight into how the controller reformulates the task-space potential into a different yet related joint-space potential. The caption has been rewritten to add clarity.\\

\item \textcolor{blue}{\textit{``In Section IV, the authors say that they use EM to determine the parameters of the GMM in Eq 6. Can
they elaborate a little on this? What is the expectation taken over and what likelihood is maximized?"}}\\
\texttt{Response:} \small The algorithm computes the expected probability of generating each of the joint positions given the current values of the GMM's means, variance matrices, and scaling values. The algorithm then chooses GMM parameters to maximize the posterior likelihood of the model, using previously-computed expected component membership as likelihood weights when computing the new mixture. Specifically, we used MATLAB's \textit{fitgmdist} implementation. Added a citation to clarify.\\

\item \textcolor{blue}{\textit{``The learned dimensionality reduction and parameters seem task dependant. I wonder if the authors
have any idea on how to generalize this across tasks."}}\\
\texttt{Response:} \small The dimensionality reduction could be generalized across tasks that had the same motor synergies - that is, for example, all tasks in which the shoulder and elbow always move together. Similarly, one could imagine several similar motions that could be decomposed into shared submotions, thus allowing the parameters to be shared. However, in general, fixing some of the parameters while optimizing will simply lead to a less-flexible model, and considering that there is no intuitive scheme for a-priori identifying motions with strongly similar submotions, we did not explore this question further. To answer your question directly, if one wanted to, one could reuse the learned dimensionality reduction for a new task which you believed to be similar, or copy certain components of a previous motion's GMM into the new GMM and leave them fixed throughout the optimization.\\

\item \textcolor{red}{\textit{``The experiments have the same task-space goal as the
demonstrations. Can the authors include experiments that have goals
different from demonstrated, especially in orientation?"}}\\
\texttt{Response:} \small First of all, they don't.. show some figures... add clarification of this in the text and explain why we cannot systematically evaluate this if we had too..\\

\item \textcolor{red}{\textit{``In similar light, maybe authors could present online adaptation of
A(q) to generalize to situations different than demonstrated, for
example a slightly moved obstacle?"}}\\
\texttt{Response:} \small This is a really nice suggestion. But how I see it, it's basically a new paper; i.e. an extension of this one, but we can discuss about this if we have space. A quick idea that comes to mind would be to apply a modulation/activation to the LPV system in a local region. Now this is not as straight forward as in our task space approaches because our LPV system is learned in joint-space while the activation function is in the embedded space. Another idea is to modify the activation function of the local behavior regions such that it is a function of an external variable, for example if force is applied/sensed, and it bias the model towards another $A_k$ matrix? So it still would be an open question, but an approach to tackle this problem could be along these lines.... Add a passage in the discussion answering this question\\

\item \textcolor{red}{\textit{``In table II, instead of the tracking error, maybe the more useful
error to see would be the error with respect to the plan or
demonstartion. One would assume that the proposed approach would be
worse at it than SEDS+IK because the minimized error is with respect to
the velocity and goal, not the position of the movement."}}\\
\texttt{Response:} \small This is not necessarily the reason that it would be worse, as SEDS is also learned by minimizing the error with respect to
the velocity and the goal. The reason that JT-DS is less accurate in reproducing the motion in task-space is because our optimization is focused on reproducing the motion solely in joint-space, while reaching the final target in task-space. Nowhere do we claim that JT-DS is capable of reproducing a motion in both joint and task space. Nevertheless, we add the mean swept-error-arrea (SEA) which evaluate the performance of reproduction in task-space from a reference trajectory... as shown in the tables....\\
\end{enumerate}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\underline{Comments from Reviewer \#2 (Reviewer: 5 (Review30371)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``My first concern regards you choice of representing the demonstrated
movement as a sequence of postures and hence looking for synergies by
using PCA of K-PCA analysis. Why not using functional PCA (f-PCA)
analysis and determine the few moving synergies that make possible the
execution of the demonstrated behaviours? It would be that, when you
combine the local synergies in space and time by using functions
$\theta_k(q)$, you get what you might obtain by using directly f-PCA.
However, the last one are valid for the whole movement and can be
combined by constant coefficients."}}\\
\texttt{Response:}  See below.

\item \textcolor{blue}{\textit{``I also found the paper lacking of physical intuitions, especially in	the experimental part, about the entities introduced in previous sections. In particular, what is the lesson we learn after your analysis of synergies and the obtained lower dimensional space for the few movements shown in section V? Is it possible to extract some global	synergic behavior? Moreover, in section IV.A authors introduce the function phi(q) named the embedding that maps the joint configuration into a lower dimensional space. Is it associated to a whole behaviour or, as matrices $A_k$, has a local validity? I lost this point. Can you show/report or describe what it represents in the joint space for one or more cases in Section V? What is this lower dimensional space for the cases in Section V?"}}\\
\texttt{Response:}  See below.

\item \textcolor{blue}{\textit{``A similar question for the matrices $A_k$ that represent local, I
	would say postural (see previous comment), synergies whose dimension is
	m x m, i.e. in joint space: I would expect they were defined in the
	lower dimensional space. Moreover, are they local representation of the
	distribution of postures and hence covariance matrices? What do their
	eigenvectors and eigenvalues locally represent? Maybe, a comparison
	with the classical and more known postural synergies for the human
	hand, analyzed by PCA, could be useful. Finally, what does it mean that
	they should be positive definite (apart from stability issue)?"}}\\
\texttt{Response:}  See below.\\

\item \textcolor{blue}{\textit{``What is the dimension of function $\theta_k(q)$ that, to some extent, can be seen as the new control variables. Is it $K<m$?"}}\\

\texttt{Response:} In response to Comments 1-4, following we mathematically and graphically analyze the proposed local behavior synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via. PCA or K-PCA) and grouping of local behavior regions with GMM on this lower dimensional space. The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$. Let's begin by analyzing the proposed DS-based control-law \eqref{eq:jtds}. The intuition behind this control-law is that $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\phi(q))A_k$ is a linear combination of time-invariant linear matrices $A_k \in \mathbb{R}^{m \times m}$, \underline{these matrices are the ``local behavior synergy matrices"} that shape the motion in joint-space. Meaning that, given the joint-space velocity vector representing the task-space error $J^{T}(q)(H(q)-x^*)$, the resulting motion is biased to use a particular set of joints, defined in each $A_k$. Each $A_k$ is thus activated depending on the current ``joint-posture" $q$ (represented in the lower-dimensional space $\phi(q)$) via the scheduling/activation function $\theta_k(\phi(q))$.
$\phi(q)$ is a learned transformation, which is associated with a whole behavior, rather than a particular $A_k$ (as is now clarified in IV.A). However, the precise nature of the transformation depends on the DR technique: PCA applies a uniform matrix multiplication of joint position everywhere, whereas KPCA is a more complicated function of joint position.  The result of the embedding is a lower-dimensional representation of the joint position. The following line has been added to IV.A for clarity:
\textcolor{blue}{\small \textit{"For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the "shoulder" and "arm" components of $q$ into a single "shoulder-arm" component in $\phi(q)$.}"}

Given the raw demonstrations of a joint-space  behavior with a task-space target as shown in Fig. \ref{fig:raw_demos} (for the pouring task) we seek to learn a linear combination of local behavior synergies $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\phi(q))A_k$ such that we can use \eqref{eq:jtds} to accurately reproduce these demonstrations. In order to do so, we project the raw joint trajectories to a lower-dimensional space via $\phi(q)$, in the case of PCA $\phi(q)=M_p \times q$ for $M_p \in \mathbb{R}^{p \times m}$. By choosing the first 3 PCs we can explain 95\% of the variance in the joint posture trajectories, which can be represented in this 3-D space as shown in Fig. \ref{fig:raw_demos}. We then jointly learn the activation function $\theta_k(\cdot)$ and the number of $K$ local synergy regions by fitting a GMM to this representation of the joint posture trajectories. As can be seen in Fig. \ref{fig:raw_demos}, the entire joint-space motion can be represented by 3 local synergy regions in PCA-space. Meaning that we only need 3 synergy matrices $A_k$ to represent the complex joint-space motion accurately. The activation function $\theta_k(\cdot)$, as described in the mansucript, is the posterior probability of a joint posture in PCA-space $\phi(q)$ belonging to one of the local synergy regions $k \in \{1,\dots,K\}$. As mentioned in Sec. V.A., we treat K as a hyperparameter and thus choose it differently for each learned motion. For motions that do not vary much throughout the course of the motion, smaller K might be sufficient, but K can just as easily be greater than $m$ (though as $K$ increases, so does the learning problem's computational cost). If comment-4 was about each specific $\theta_k(q)$, they are scalars, as is mentioned in Sec. III.

\begin{figure*}[!th] 
  \begin{minipage}{0.58\textwidth}
     	\centering 
     	\includegraphics[trim={1.2cm 2cm 1.7cm 2cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/raw_demos_pour.pdf}
  \end{minipage}
   \begin{minipage}{0.47\textwidth}
      	\centering
      	\includegraphics[trim={1.2cm 0.5cm 0.5cm 0.35cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/raw_demos_pca_pour.pdf}
      	\vspace{-5pt}

		\includegraphics[trim={0.8cm 0.5cm 0cm 0.35cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/local_synergy_regions_pour_annotated.pdf}
   \end{minipage}
   \caption{Raw demonstrations of the pouring task in original space, PCA space and the extracted local synergy regions. \label{fig:raw_demos}}
\end{figure*}

\underline{As mentioned in the reviewers 3rd comment}, since we have an accurate representation of our joint posture trajectories and the local synergy regions in this lower-dimensional space, why not learn the $A_k$ matrices in this space; i.e. $\mathcal{A}(q)\in \mathbb{R}^{p\times p}$ rather than $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ ? This seems like the obvious method, if we were considering our synergistic approach the classical way, i.e. using the lower-dimensional space as the new control variables. However, if we were to do that, this would mean that \eqref{eq:jtds} would become:
\begin{equation}
\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{M_p^{-1}}_{\mathbb{R}^{m\times p}}\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{p\times p}}\underbrace{M_p}_{\mathbb{R}^{p\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
\label{eq:jtds_low}
\end{equation}
This is not entirely incorrect, however, with such a control law we can no longer guarantee stability and convergence to the attractor $x^{*}$. By considering the Lyapunov function used in Appendix I to prove stability of \eqref{eq:jtds_low} i.e. $V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*)$, the derivative of $ V $ wrt. would be:
\begin{equation}
\begin{aligned}
\frac{dV(q)}{dt} &= (H(q) - x^*)^TJ(q)\dot{q}\\
&= -(H(q) - x^*)^TJ(q)M_p^{-1}\mathcal{A}(q)M_pJ^T(q)(H(q) - x^*)\\
&= -(H(q) - x^*)^TJ(q)\textcolor{red}{\underbrace{M_p^{-1}\big(\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{> 0}\underbrace{A_k}_{\succ 0}\big)M_p}_{\text{This should be}~~ \succ 0}}J^T(q)(H(q) - x^*)\leq 0
\end{aligned}
\end{equation}
In order for \eqref{eq:jtds_low} to be asymptotically stable wrt. the attractor $x^*$ the term $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p \in \mathbb{R}^{m\times m}$ should undoubtedly be $\succ 0$. Even if $\theta_k(q) > 0$ and $A_k \succ 0$ for all $k$, the only way for this condition to hold is if $M_p$ where a full symmetric positive definite matrix. However, this is not the case because $M_p\in \mathbb{R}^{p\times m}$  is a projection matrix, i.e. $p<m$. The term $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$ will no longer be full-rank, it will preserve the $p$ positive eigenvalues from the $A_k$ matrices, but the remaining eigenvalues will be 0, thus generating a semi-positive definite matrix, which invalidates our stability proof. Note that, as described in the Appendix, $J(q)$ is not proven to be full rank in all regions of the workspace, when this happens, this means that the end-effector pose is not manipulable. Hence, we can only prove asymptotic stability when the end-effector pose is manipulable. This is a physical artifact from the joint-space constraints of the robot arm which we cannot avoid. Yet, since the demonstrations are acquired via kinesthetic teaching we can assume that they are manipulable. \underline{As asked by the reviewer, what does it physically mean to have all positive eigenvalues?}  When \\ $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$  or ($\mathcal{A}(q)$) has some eigenvalues $=0$ the physical meaning behind it is that we are forfeiting the control of these directions of motion in joint-space. Namely, the desired velocity directions in joint-space will be constrained in the corresponding DOF of the arm, this will prohibit us from converging to the attractor and will be no longer capable of accurately following the joint-space demonstrations.  If instead we learn the synergy matrices in joint-space following \eqref{eq:jtds} and modulate them with the activation function in PCA-space, we avoid all of this problems and can ensure full controllability of the motion in joint-space.

\noindent \textbf{Remark 1:} This analysis is only considering PCA as the DR approach. If we would consider K-PCA, deriving such a control law is more involved as the inverse mapping $\phi(q)^{-1}$ is not straightforward if at all possible, thus limiting the approach to linear DR techniques. By considering $\phi(q)$ solely in the activation function $\theta(\cdot)$ we do not have this problem and can use any DR technique, as long as we're able to compute out-of-sample projections.

\noindent \textbf{Remark 2:} If we were to bypass the DR step our state-dependent system matrix would become $\mathcal{A}(q)= \sum\limits_{k=1}^{K}\theta(q)A_k$, which (via our definition) is also a combination of ``local behavior synergy matrices" with the activation function living in joint-space rather than the PCA space; better-known as the \textit{joint postural synergy space}. The fact that we use PCA/K-PCA to represent the activation function $\theta(\cdot)$ is to alleviate the complexity of extracting the local synergy regions in joint-space directly. It is, indeed, also motivated by the fact that we know that multi-DOF joint postures can be accurately represented in a lower-dimensional space composed of the principal components of the postural trajectories; i.e. the well-known \textit{joint postural synergies}. As shown in TABLE I of the manuscript and discussed in Section V, by using a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations. This is explicitly highlighted in Section V:\\
\textcolor{blue}{\small \textit{"As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches. This suggests that using DR to encode our activation functions $\theta_k$ in a lower-dimensional space $\phi(q)$ yields better generalization capabilities than encoding the behaviors in using the original $q$. This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learned without DR is an order of magnitude higher than with DR. Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the higher number of $K$ needed to represent the motion without DR. For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR. By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations. Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations."}} \\
Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations and learn a \textit{time-invariant} controller that ensures reaching the task-space target while approximately following the demonstrated motion, as shown in Fig. \ref{fig:jtds_velocities}, where we plot the joint positions/velocities of 1 demonstrations of the pouring task and the joint-positions/velocities from the JT-DS controller. In these plots, we are taking the velocity generated directly by JT-DS. The joint-positions are then computed by numerical integration of these velocities. Note that our controller does not generate this entire trajectory in one-shot, it is generating the next desired velocity given the current joint posture, e.g. $\dot{q}_{t+1} = \mathcal{A}(q_{t})J^T(q_{t})(H(q_{t})-x^*)$ with $q_t$ being the current joint configuration. Moreover, our controller is state-dependent only; i.e. it has no dependence on time or the duration of the desired trajectories. Hence, it can generate motions similar to the demonstrated ones at different frequencies and time durations.

\noindent \textcolor{red}{TODO: Add clarifications in text such that no other reader gets confused with the new concept of behavior synergies that we propose as opposed to the classical joint posture synergies.}

\begin{figure*}[!th] 
  \begin{minipage}{0.5\textwidth}
     	\centering 
     	\includegraphics[trim={1.2cm 1.5cm 1.7cm 1.5cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/jtd_pos_pour1.pdf}
  \end{minipage}
    \begin{minipage}{0.5\textwidth}
       	\centering 
       	\includegraphics[trim={1.2cm 1.5cm 1.7cm 1.5cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/jtd_vel_pour1.pdf}
    \end{minipage}
   \caption{Training and executed joint-position/velocities with JT-DS for 1 pouring demonstrations. \label{fig:jtds_velocities}}
\end{figure*}
\newpage
\noindent \underline{On the applicability of F-PCA to our learning scheme:}  If we were to ignore our proposed control scheme and try to reconstruct the trajectories from the joint postural synergy components; i.e. using PCA, we would get the trajectories shown in Fig. \ref{fig:pca_velocities}. In these plots, we are taking one of the collected demonstrations projected in the PCA-space, i.e. $\phi(\{q_1,q_2,\dots,q_T\})$ and reconstructing it to generate the desired joint posture trajectories. To compute the velocities we apply numerical differentiation on the reconstructed joint positions. As can be seen, we are capable of generating similar trajectories, both in position and velocity. In this simple example we are saving the entire demonstration in PCA-space, however, if we were to use this for controlling the real robot we need to learn a model or a controller in this low-dimensional space and then map it back to joint-space. As discussed earlier, this is not feasible with our proposed control-law \eqref{eq:jtds}, due to the stability and convergence objectives we defined.

\begin{figure*}[!th] 
  \begin{minipage}{0.485\textwidth}
     	\centering 
     	\includegraphics[trim={1.2cm 1.5cm 1.7cm 1.5cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/pca_pos_pour1.pdf}
  \end{minipage}
    \begin{minipage}{0.5\textwidth}
       	\centering 
       	\includegraphics[trim={1.2cm 1.6cm 1.7cm 1.6cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/pca_vel_pour1.pdf}
    \end{minipage}
   \caption{Training and reconstructed joint-position/velocities with PCA for 1 pouring demonstrations. \label{fig:pca_velocities}}
\end{figure*}

Ignoring this limitation, one of the main draw-backs of this type of synergistic control is that the extracted representation in PCA space does not take into account the dynamic features of the motions and further assumes that the demonstrations have no time correlation. As mentioned by the reviewer and applied in \cite{Dai:IROS:2013} for modeling grasping motions and in \cite{Averta:FRAI:2017} for upper-limb motions, functional-PCA is a nice alternative to include these dynamic features and further reconstruct entire trajectories with a linear combination of a few \textit{time-dependent} functions. In this approach each demonstrated trajectory (of a single DOF) is considered as a function composed of a linear combination of $N$ basis elements (which can ve exponentials, splines, Fourier) $f_i = \sum\limits_{n=1}^{N}\theta_nb_N$. Thus, each $i$-th function (i.e. $i$-th demonstration) can be described by the vector a coefficients $\Theta = \{\theta_1,\dots,\theta_N\}$. PCA is then performed on $\Theta = \{\theta_1,\dots,\theta_N\}$ and the extracted PC are transformed to \textit{functional PCs} via the $N$ basis functions. Then $M<N$ \textit{functional PCs} are used to reconstruct the $i$-th function (i.e. $i$-th demonstration ) as $\hat{f}_i = \bar{f_i} + \sum\limits_{i=1}^{M}c_i\xi_i$, where $\bar{f}_i$ is the mean function, $c_i$ are the coefficients and $\xi_i$ are the \textit{functional PCs}. As shown in \cite{Averta:FRAI:2017} with only 1 \textit{functional PCs} 60/70\% of the demonstrations can be accurately reconstructed. This approach would be interesting if our goal were to reconstruct entire time-dependent trajectories as they do in  \cite{Dai:IROS:2013} and \cite{Averta:FRAI:2017}. This is, however, NOT our goal and applying this method to our control scheme seems not suitable and is rather an alternative approach for the following reasons: First of all, in this work we propose a DS-based controller, meaning that, we do not generate the full reference trajectory of the robot \textit{a prior} and then try to track it, we instead compute the next desired velocity on-the-fly which happens to result in the desired motion because it is encoded in $\mathcal{A}(q)$. Second, if we were to apply this method as the reviewer suggests, the only way to do it is to re-write a controller that uses 7 reconstructed functions $\hat{f}_i$ (i.e. one per joint) to track a time-indexed reference joint trajectory. This is a valid approach, however, in this work we propose a state-dependent; i.e. time-invariant approach. Finally, our goal not to accurately reproduce the demonstrations but rather to reach the target $x^*$ with a similar joint motion as the one that was demonstrated.

\item \textcolor{blue}{\textit{``For the test, why not comparing also with the JT control method? In
other terms, is it possible to use a SEDS + JT, especially for the
third case "Transiting through singular configurations"?"}}\\
\texttt{Response:} \small This could be done, but it is mathematically incorrect. The original JT control method is a DS that converges to a unique equilibrium point, i.e. the task-space target $x^*\in \mathbb{R}^{N\times 1}$, with a gain matrix $K \in \mathbb{R}^{M\times M}$ as:\begin{equation}
\dot{q} = K J(q)^T(H(q)- x^*) 
\label{eq:JT}
\end{equation}
SEDS on the other hand, is a DS-based motion generator that converges also to a unique equilibrium point, as:
\begin{equation}
\hfill \begin{aligned}
&  \dot{x}  = \mathbf{f}(x)
\rightarrow  &
\begin{cases}
\lim_{t \rightarrow \infty}\|x - x^*\|=0
\end{cases}
\end{aligned}  \hfill
\label{eq:simple-ds}
\end{equation}
where $\mathbf{f}(\cdot): \mathbb{R}^M \rightarrow \mathbb{R}^M$ is a continuous differentiable vector-valued function representing a non-linear DS that converges to a single stable equilibrium point $x^*$. SEDS is used in the following way: Given the current state of the robot $x$ the next desired velocity $\dot{x}_d$ is computed with \eqref{eq:simple-ds} this is then fed to a velocity-based IK solver as shown in Fig. \ref{}:\\
\textcolor{red}{TO SINA: Include typical SEDS-like control-loop here}\\
If we were to use SEDS with the JT-control method, we would have to integrate the desired velocity to generate a desired position $x_d$ at each time-step and define it as the target in \eqref{eq:JT}, which would yield the following controller:
\begin{equation}
\dot{q} = K J(q)^T(H(q)- x_d) 
\end{equation} Note that, the JT control method assumes that the target is static, hence, in order for this controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time. This is contradicting as \eqref{eq:simple-ds} is always trying to converge to $x^*$. Thus, merging these two approaches is theoretically ill-fitting. To better highlight this, we show an illustration of the supposed SEDS+JT control loop in Fig. \ref{}:\\
\textcolor{red}{TO SINA: Include SEDS+JT control-loop here}\\
\textcolor{red}{\small Describe this control-loop and state more reasons (if any) why it's wrong...}

\item \textcolor{blue}{\textit{``Minors:
- SEDS + IK: what does SEDS mean?"}}\\
\texttt{Response:} As stated in the previous comment, SEDS, short for Stable Estimator of Dynamical Systems is an approach proposed in \cite{khansari2011learning} that learns a stable estimate of a DS with the form of \eqref{eq:simple-ds} from demonstrations. The first paragraph of Section V.B was written to describe its meaning and how we use it:\\
\textcolor{blue}{\small \textit{``We now seek to elucidate the distinctive properties of our JT-DS model by comparing its performance to a DS-based Cartesian motion generator + IK solver approach for behaviors \textit{(5-7)}. For the Cartesian motion generator we use the SEDS (Stable Estimator of Dynamical Systems) approach [8] which learns an asymptotically stable DS in Cartesian space from demonstrations. We then generate joint trajectories through a damped least square IK solver. From herein we refer to this approach as SEDS+IK."}}\\
\end{enumerate}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\underline{Comments from Reviewer \#3 (Reviewer: 7 (Review30521)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``a considerable part of the introduction is spent in comparing joint space trajectory generation to task
space trajectory generation. Since these are elementary concepts in robotics, the authors might consider
reducing this part"}} \texttt{Response:} Several sentences in the introduction have been removed/reduced.\\

\item \textcolor{blue}{\textit{``one of the main claims of the paper is the stability proof, which in fact is cited in the first line of the
abstract. However the stability analysis is a quite simple extension of the stability of the transpose
Jacobian scheme, where the constant gain on the task space error is replaced by the matrix A(q). As
such the emphasis on this result might be to some extent reduced. By the way, the proof works only
if the Jacobian is full rank, which is not properly discussed in Appendix I"}}\\
\texttt{Response:}  Removed a couple mentions of the stability result, and added the dependence on the rank of the Jacobian to both the appendix and the main body of the text.\\

\item \textcolor{blue}{\textit{``the meaning of variables $q_d$ in (2) is not very clear to me, a well as whether it takes the same meaning
as $q_{t;n}$ in (8)"}}\\
\texttt{Response:} $q_d$ are the velocities from the demonstrations, which we are aiming to recreate. Equation (8) has been updated to remove the typo (they are in fact the same).\\

\item \textcolor{blue}{\textit{``Section III uses the concept of synergy without actually defining it. An introduction to this concept is given later, in Section IV"}}\\
\texttt{Response:} Added the following to Section 3:\\
\textcolor{blue}{\small \textit{``Each ``synergy" is a joint-space transformation that biases the resulting motion to use particular joints; for more, see Sec. IV."}}\\

\item \textcolor{blue}{\textit{``the expression "flow of motion" in Proposition 1 is not very clear to me"}}\\
\texttt{Response:}  Changed to "motion".\\

\item \textcolor{red}{\textit{`some details about the optimization problem in (8) might be given
(nature of the problem, complexity, feasibility)."}}\\
\texttt{Response:} The following has been added near Eq. (8):\\
\textcolor{blue}{\small \textit{``The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective..." ``This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD), and the number of terms in the objective is $O\left(k^2d^3\left( \sum_{n=1}^N T_n\right)\right)$."}}\\
To elaborate about the prior calculation: the number of terms in the sum is proportional to the number of recorded points, times the number of matrices (k) required to generate the predicted velocity from (7) squared, times the number of terms in a matrix multiplication of two $d x d$ matrices ($d^3$).\\

\end{enumerate}

\bibliographystyle{plain}
\bibliography{Refs.bib}

\end{document}

### 2018-03-14 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
Writing reply to comments and changing paper accordingly

@@
 \begin{abstract}
-In this paper, we propose an asymptotically stable joint-space dynamical system that captures desired behaviors in joint-space while converging towards a task-space attractor.
-% Our method is fast to compute and smoothly moves through classic kinematic singularities by avoiding the use of the pseudo-inverse Jacobian; moreover, it stably converges towards its task-space attractor.
- To encode joint-space behaviors while meeting the stability criteria, the dynamical system is constructed as a Linear Parameter Varying (LPV) system combining different behavior synergies, and we provide a method for learning these synergy matrices from demonstrations.
+In this paper, we propose an asymptotically stable joint-space dynamical system (DS) that captures desired behaviors in joint-space while converging towards a task-space attractor in both position and orientation.
+To encode joint-space behaviors while meeting the stability criteria, we propose a DS constructed as a Linear Parameter Varying (LPV) system combining different behavior synergies and provide a method for learning these synergy matrices from demonstrations.
 Specifically, we use dimensionality reduction to find a low-dimensional embedding space for modulating joint synergies, and then estimate the parameters of the corresponding synergies by solving a convex semi-definite optimization problem that minimizes the joint velocity prediction error from the demonstrations.
-Our proposed approach is empirically validated on a variety of motions.
-%for a 7-DOF KUKA LWR4+ robot arm.
+Our proposed approach is empirically validated on a variety of motions that reach a target in position and orientation, while following a desired joint-space behavior.
  \end{abstract}

### 2018-03-14 src/JTDS_mat_lib/run_crossval_hyperparam_kpca_kuka_wOrientation.m

@@ -6,7 +6,7 @@
 clear all; close all; clc;
 do_plots  = 1;
 data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+choosen_dataset = 'singularity'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Load and Process dataset %
@@ -94,11 +94,11 @@ clc;
 %%% Type of DR Methods %%%
 do_plots = 1;
 mapping_name = 'KPCA';
-rbf_range = logspace(log10(max_rbf/2),log10(max_rbf*5),10);
+rbf_range = logspace(log10(max_rbf/2),log10(max_rbf*5),10)
 
 %%% Learning options %%%
 options = [];
-options.orientation_flag=1;
+options.orientation_flag = 1;
 options.tol_cutting = 0.1;
 
 %%% Dim-Red options %%%
@@ -185,9 +185,13 @@ if do_plots
     end    
 end
 
-
 % Truncate the range to values whose exp. var < dimq
-new_rbf_range      = rbf_range(p_s <= dimq);
+new_rbf_range      = rbf_range(p_s <= dimq)
+
+if isempty(new_rbf_range)
+    warning('Try a new range!!')
+end
+
 adjusted_rbf_range = logspace(log10(new_rbf_range(1)),log10(new_rbf_range(end)),10)
 
 display('Press ENTER to continue with this range:');
@@ -197,10 +201,10 @@ pause;
 %%% Learn JTDS Model with different methods and parameters %%%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 repetitions = 10;
-K_s         = zeros(repetitions,length(rbf_range));
-rmse_train  = zeros(repetitions,length(rbf_range));
-rmse_test   = zeros(repetitions,length(rbf_range));
-kpca_dims   = zeros(repetitions,length(rbf_range));
+K_s         = zeros(repetitions,length(adjusted_rbf_range));
+rmse_train  = zeros(repetitions,length(adjusted_rbf_range));
+rmse_test   = zeros(repetitions,length(adjusted_rbf_range));
+kpca_dims   = zeros(repetitions,length(adjusted_rbf_range));
[…]

### 2018-03-14 src/JTDS_mat_lib/run_synergy_analysis.m

@@ -1,216 +1,0 @@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Evaluation of Learning Schemes for JTDS Models (including orientation) on Different Datasets     %%
-%  Train and compare a series of JTDS models with different dimensionality                          %%   
-%  reduction schemes and GMM model fitting                                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-clear all; close all; clc;
-do_plots  = 1;
-data_path = '../../Data/mat/'; % <-Insert path to datasets folder here
-choosen_dataset = 'back'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Load and Process dataset %
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-switch choosen_dataset
-    case 'back'
-        demos_location = strcat(data_path, 'back_hand/data.mat');
-        demo_ids = [2:11];
-    case 'fore'
-        demos_location = strcat(data_path,'fore_hand/data.mat');
-        demo_ids = [1:11];
-    case 'pour' 
-        demos_location = strcat(data_path,'pour_no_obst/data.mat');
-        demo_ids = [1 2 3 5 6 7 8 9 10];
-    case 'pour_obst'
-        demos_location = strcat(data_path,'pour_obst/data.mat');
-        demo_ids = [1:10];
-    case 'pour_obst_2'
-        demos_location = strcat(data_path,'pour_obst_2/data.mat');
-        demo_ids = [1:7];                
-    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
-        demos_location = strcat(data_path,'foot/data.mat');
-        demo_ids = [1:8];                
-    case 'singularity'   
[…]

### 2018-03-14 src/JTDS_mat_lib/synergy_analysis/run_synergy_analysis.m

@@ -1,0 +1,438 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Evaluation of Learning Schemes for JTDS Models (including orientation) on Different Datasets     %%
+%  Train and compare a series of JTDS models with different dimensionality                          %%   
+%  reduction schemes and GMM model fitting                                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+clear all; close all; clc;
+do_plots  = 1;
+data_path = '../../../Data/mat/'; % <-Insert path to datasets folder here
+choosen_dataset = 'pour_obst_2'; % Options: 'back','fore','pour','pour_obst','foot','singularity';
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Load and Process dataset %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+switch choosen_dataset
+    case 'back'
+        demos_location = strcat(data_path, 'back_hand/data.mat');
+        demo_ids = [2:11];
+    case 'fore'
+        demos_location = strcat(data_path,'fore_hand/data.mat');
+        demo_ids = [1:11];
+    case 'pour' 
+        demos_location = strcat(data_path,'pour_no_obst/data.mat');
+        demo_ids = [1 2 3 5 6 7 8 9 10];
+    case 'pour_obst'
+        demos_location = strcat(data_path,'pour_obst/data.mat');
+        demo_ids = [1:10];
+    case 'pour_obst_2'
+        demos_location = strcat(data_path,'pour_obst_2/data.mat');
+        demo_ids = [1:7];                
+    case 'foot'        % This dataset was recorded at 50 Hz! thinning_ratio = 1 or 2
+        demos_location = strcat(data_path,'foot/data.mat');
+        demo_ids = [1:8];                
+    case 'singularity'   
[…]

### 2018-03-14 src/JTDS_mat_lib/utils/compute_trajectory.m

@@ -1,0 +1,13 @@
+function Qd = compute_trajectory(motion_generator, Q_true, Xt, orientation_flag)
+Qd = zeros(size(Q_true));
+if orientation_flag==1
+    for i = 1:size(Q_true, 2)
+        Qd(:, i) = motion_generator.get_next_motion_orientation(Q_true(:, i), Xt);
+    end
+else
+    for i = 1:size(Q_true, 2)
+        Qd(:, i) = motion_generator.get_next_motion(Q_true(:, i), Xt);
+    end
+end
+end
+

### 2018-03-15 paper/reply_reviews/New_reply_to_comments_sina.tex
.

--- 2018-03-13 paper/reply_reviews/reply_comments.tex
@@
 \usepackage{amsfonts}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}
@@
 
-\newsavebox{\Luiuc}
-\sbox{\Luiuc}{%
-	\parbox[b]{1.6in}{%
-		\vspace{0.5in}
-		\includegraphics[scale=0.25]
-		{./EPFL-Logo-bitmap/JPG/EPFL-Logo-RVB-55.jpg}%
-	}%
-}%
-
-\makeletterhead{Uiuc}{\Rheader{\usebox{\Luiuc}}}
-
-\newlfmP{headermarginskip=-100pt}
-\newlfmP{sigsize=0pt}
-%\headermarginsize{0in}
-\headermarginskip{.5in}
-\unprbottom{1cm}
-
 %\newlfmP{dateskipafter=30pt}
@@
 %\PhrEmail{Email}
-
-\lthUiuc
 
@@
 different from demonstrated, especially in orientation?"}\\
-\textcolor{red}{\texttt{Response:} \small First of all, they don't..
-show some figures...
-add clarification of this in the text and explain why we cannot systematically evaluate this if we had too..}\\
+\textcolor{blue}{\texttt{Response:} \small Yes.
+As we have shown in Appendix I, regardless of the target location (position/orientation), the motion generated by the proposed dynamical system asymptotically converges the desired target.
+However, it is important to note that as (9) is semi-negative-definite and not negative-definite, there might be a joints configuration $ (q )$ which results in $ J(q)\sum\limits_{k=1}^{K}\theta_k(q)A_kJ^T(q)=0 $.
+In this situation, the robot is stuck in a deadlock and, consequently, cannot reach the desired target.
+}\\
 
@@
 example a slightly moved obstacle?"}\\
-\textcolor{red}{\texttt{Response:} \small This is a really nice suggestion.
+\textcolor{blue}{\texttt{Response:} \small The reviewer has pointed to a very interesting topic and authors appreciate the reviewer’s suggestion.
+In this paper, we assumed that the learned model (namely $ \pi $s, $ \mu $s and $ \Sigma $s) does not need to be updated or modified during the task execution.
+Hence, this would be out of the scope of this paper and we are planning to address it in future works.
+}\\
+%\textcolor{red}{\texttt{Response:} \small This is a really nice suggestion.
 But how I see it, it's basically a new paper; i.e.
@@
 error to see would be the error with respect to the plan or
-demonstartion. One would assume that the proposed approach would be
+demonstration. One would assume that the proposed approach would be
 worse at it than SEDS+IK because the minimized error is with respect to
 the velocity and goal, not the position of the movement."}\\
-\textcolor{red}{\texttt{Response:} \small This is not necessarily the reason that it would be worse, as SEDS is also learned by minimizing the error with respect to
-the velocity and the goal.
+\textcolor{blue}{\texttt{Response:} \small This is not in general true.
+Based on the motion of the robot and its configuration, there might be cases where the motion generated by SEDS does not even reach the target.
+To clarify this, we have systematically studied the performance of the motion generated by JT-DS and SEDS+IK in terms of (i) reaching to the desired target and (ii) following the demonstrated end-effector motions, See Figure \ref{fig:SEDSvsJTDS}.
+We used the data-sets from the seven tasks mentioned in Section V.A to construct SEDS and JT-DS.
+Both methods are tested by the same the target and the initial configurations, which are extracted from demonstrations.
+We define $ x^*=[x^*_{ori};x^*_{pos}] $ as the target configuration that includes the desired position as well as the desired  orientation of the end-effector.
+The motion generated by SEDS is converted into the 7-DOF joints state using the damped least square IK solver.\\
+In Figure \ref{fig:SEDSvsJTDSP:E} and Figure \ref{fig:SEDSvsJTDSO:E}, the error between the final location of the robot and the target is illustrated.
+As it can be seen, the motion generated by the proposed dynamical system reaches the target, whereas, the motion generated by SEDS does not, even though both models are theoretically stable.
+The main cause of failure for SEDS is due to the IK solver.
+As SEDS generates the motion of the robot without considering its joints configurations, the robot might reach its joint limits.
+Whereas, as JT-Ds directly shapes the joints' motion, by learning the desired joint-space behavior, one can avoid these situations.
+\\
+\begin{figure}[t]
+ \hspace*{-2.4cm}
+	\begin{subfigure}[t]{0.63\textwidth}
+	\centering
+		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P_End.pdf}
+		\caption{}
+		\label{fig:SEDSvsJTDSP:E}
+	\end{subfigure}%
+	\hspace{-0.2cm}
+	\begin{subfigure}[t]{0.63\textwidth}
+	\centering
+		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O_End.pdf}
+		\caption{ }
+		\label{fig:SEDSvsJTDSO:E}
+	\end{subfigure}\\		 
+	\hspace*{-2.4cm}\begin{subfigure}[t]{0.63\textwidth}
+			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P.pdf}
+		\caption{ }
+			\label{fig:SEDSvsJTDSP}
+		\end{subfigure}%
+		\begin{subfigure}[t]{0.63\textwidth}
+			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O.pdf}
+			\caption{ }	
+			\label{fig:SEDSvsJTDSO}
+		\end{subfigure}\\
+	\caption{Performance comparison between SEDS+IK and JT-DS while the initial and the target points are same.
+The error between the orientations is the difference between the first two columns of the corresponding rotation matrix representation.}
+	\label{fig:SEDSvsJTDS}
+\end{figure}
+In Figure \ref{fig:SEDSvsJTDSP} and Figure \ref{fig:SEDSvsJTDSO}, the error between the demonstrated and the excused motions are illustrated.
+As it can be seen, although JT-DS is not tailored for following the end-effector motions, its performance is mostly better than SEDS.\\
+Although including this study would highlight one of the advantages of using JT-Ds over Cartesian-motion-generators, as this not the main focus of this paper and due to the space limitations, we have decided to not include this in the paper.}
+%\textcolor{red}{\texttt{Response:} \small This is not necessarily the reason that it would be worse, as SEDS is also learned by minimizing the error with respect to
+%the velocity and the goal.
 The reason that JT-DS is less accurate in reproducing the motion in task-space is because our optimization is focused on reproducing the motion solely in joint-space, while reaching the final target in task-space.
@@
 as shown in the tables....}\\
-
 \end{enumerate}
@@
 The original JT control method is a DS that converges to a unique equilibrium point, i.e.
-the task-space target $x^*\in \mathbb{R}^{N\times 1}$, with a gain matrix $K \in \mathbb{R}^{M\times M}$ as:\begin{equation}
-\dot{q} = K J(q)^T(H(q)- x^*) 
+the task-space target $x^d\in \mathbb{R}^{N\times 1}$, with a gain matrix $K \in \mathbb{R}^{M\times M}$ as:\begin{equation}
+\dot{q} = K J(q)^T(H(q)- x^d) 
 \label{eq:JT}
 \end{equation}
-SEDS on the other hand, is a DS-based motion generator that converges also to a unique equilibrium point, as:
+SEDS, is a DS-based motion generator that converges also to a unique equilibrium point, as:
 \begin{equation}
@@
 where $\mathbf{f}(\cdot): \mathbb{R}^M \rightarrow \mathbb{R}^M$ is a continuous differentiable vector-valued function representing a non-linear DS that converges to a single stable equilibrium point $x^*$.
-SEDS is used in the following way: Given the current state of the robot $x$ the next desired velocity $\dot{x}_d$ is computed with \eqref{eq:simple-ds} this is then fed to a velocity-based IK solver as shown in Fig.
-\ref{}:}\\
-\textcolor{red}{TO SINA: Include typical SEDS-like control-loop here}\\
-\textcolor{blue}{ \small If we were to use SEDS with the JT-control method, we would have to integrate the desired velocity to generate a desired position $x_d$ at each time-step and define it as the target in \eqref{eq:JT}, which would yield the following controller:
-\begin{equation}
-\dot{q} = K J(q)^T(H(q)- x_d) 
-\end{equation} Note that, the JT control method assumes that the target is static, hence, in order for this controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
+In order to prove the stability of SEDS, one needs to assume that the dynamic of the robot is perfectly compensated by the built-in position-level controller.
+In other words, the transfer function of the dynamic of the robot+ the low-level controller and the IK solver must be one.
+\\
+Almost all the commercially available robots are equipped by a very precises joint position controller.
+Hence, the transfer function of the robot+ the low-level controller can be assumed one.
+Whereas, the transfer function of the Jacobian transpose IK solver can not be assumed one as the JT control method assumes that the target is static.
+Hence, in order for the JT controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
 This is contradicting as \eqref{eq:simple-ds} is always trying to converge to $x^*$.
-Thus, merging these two approaches is theoretically ill-fitting.
-To better highlight this, we show an illustration of the supposed SEDS+JT control loop in Fig.
-\ref{}:}\\
-\textcolor{red}{TO SINA: Include SEDS+JT control-loop here}\\
-\textcolor{red}{\small Describe this control-loop and state more reasons (if any) why it's wrong...}
+Thus, merging these two approaches is theoretically ill-fitting.\\
+ Note that,  To better highlight this, we show an illustration of the supposed SEDS+JT and JT-DS control loops in Figure \ref{fig:schematic:SEDS} and  Figure \ref{fig:schematic:jt}, respectively.}\\
+
+\begin{figure}[t]
+\centering
+	\begin{subfigure}[t]{0.8\textwidth}
+	\centering
+		\includegraphics[trim={0 0 0.25cm 1.1cm},clip,width=\textwidth]{Pics/Drawing2.png}
+		\caption{ A Cartesian motion generator (e.g., SEDS) with the Jacobian transpose IK solver.}
+		\label{fig:schematic:SEDS}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{0.8\textwidth}
+	\centering
+		\includegraphics[trim={0.25cm 0 0.0cm 1.2cm},clip,width=\textwidth]{Pics/JT_DS.png}
+		\caption{ JT-DS.}
+		\label{fig:schematic:jt}
+	\end{subfigure}\\
+	\caption{A simplified schematic comparing the standard control architecture using Cartesian dynamical systems with the Jacobian transpose IK solver (top) and our propose approach.
+By integrating the "inverse kinematic (IK) solver" and "motion generator" blocks into a unified block, the proposed approach eliminates undesirable effects of the IK solver.
+}
+\end{figure}
 

### 2018-03-15 paper/reply_reviews/New_reply_to_comments_sina.tex
.

@@
 the velocity and goal, not the position of the movement."}\\
-\textcolor{blue}{\texttt{Response:} \small This is not in general true.
-Based on the motion of the robot and its configuration, there might be cases where the motion generated by SEDS does not even reach the target.
+\textcolor{blue}{\texttt{Response:} \small This is, in general, not true.
+Based on the robot's motion and its configuration, there might be cases where the motion generated by SEDS does not even reach the target.
 To clarify this, we have systematically studied the performance of the motion generated by JT-DS and SEDS+IK in terms of (i) reaching to the desired target and (ii) following the demonstrated end-effector motions, See Figure \ref{fig:SEDSvsJTDS}.
 We used the data-sets from the seven tasks mentioned in Section V.A to construct SEDS and JT-DS.
-Both methods are tested by the same the target and the initial configurations, which are extracted from demonstrations.
+Both methods are tested by the same the target and the initial configurations, which were extracted from the demonstrations.
 We define $ x^*=[x^*_{ori};x^*_{pos}] $ as the target configuration that includes the desired position as well as the desired  orientation of the end-effector.
@@
 In Figure \ref{fig:SEDSvsJTDSP:E} and Figure \ref{fig:SEDSvsJTDSO:E}, the error between the final location of the robot and the target is illustrated.
-As it can be seen, the motion generated by the proposed dynamical system reaches the target, whereas, the motion generated by SEDS does not, even though both models are theoretically stable.
+As it can be seen, the motion generated by the proposed dynamical system reaches the target, whereas, the motion generated by SEDS does not; regardless of the fact that both models are theoretically stable.
 The main cause of failure for SEDS is due to the IK solver.
-As SEDS generates the motion of the robot without considering its joints configurations, the robot might reach its joint limits.
+As SEDS generates the motion of the robot without considering its joint configurations, the robot might reach its joint limits.
 Whereas, as JT-Ds directly shapes the joints' motion, by learning the desired joint-space behavior, one can avoid these situations.
@@
 		\end{subfigure}\\
-	\caption{Performance comparison between SEDS+IK and JT-DS while the initial and the target points are same.
+	\caption{The performance comparison between SEDS+IK and JT-DS while the initial and the target points are same.
 The error between the orientations is the difference between the first two columns of the corresponding rotation matrix representation.}
@@
 In Figure \ref{fig:SEDSvsJTDSP} and Figure \ref{fig:SEDSvsJTDSO}, the error between the demonstrated and the excused motions are illustrated.
-As it can be seen, although JT-DS is not tailored for following the end-effector motions, its performance is mostly better than SEDS.\\
-Although including this study would highlight one of the advantages of using JT-Ds over Cartesian-motion-generators, as this not the main focus of this paper and due to the space limitations, we have decided to not include this in the paper.}
+As it can be seen, even though JT-DS is not tailored for following the end-effector motions, its performance is comparably better than SEDS.\\
+Although including this study would highlight one of the advantages of using JT-DS over Cartesian-motion-generators, as this does not highlight the main focus of this paper and due to the space limitations, we have decided to not include this in the paper.}
 %\textcolor{red}{\texttt{Response:} \small This is not necessarily the reason that it would be worse, as SEDS is also learned by minimizing the error with respect to

### 2018-03-15 paper/reply_reviews/New_reply_to_comments_sina.tex
.

@@
 (nature of the problem, complexity, feasibility)."}\\
-\textcolor{red}{\texttt{Response:} \small The following has been added near Eq. (8):\\
-``The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective..."
-``This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD), and the number of terms in the objective is $O\left(k^2d^3\left( \sum_{n=1}^N T_n\right)\right)$."\\
-To elaborate about the prior calculation: the number of terms in the sum is proportional to the number of recorded points, times the number of matrices (k) required to generate the predicted velocity from (7) squared, times the number of terms in a matrix multiplication of two $d x d$ matrices ($d^3$).}\\
+\textcolor{blue}{\texttt{Response:} \small The following has been added near (8):\\
+``The resulting optimization minimizes the mean squared joint velocity error from (2), resulting in a convex semidefinite optimization with a quadratic objective..."
+``This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD)."\\
+%To elaborate about the prior calculation: the number of terms in the sum is proportional to the number of recorded points, times the number of matrices (k) required to generate the predicted velocity from (7) squared, times the number of terms in a matrix multiplication of two $d x d$ matrices ($d^3$).
+}\\
 

### 2018-03-15 paper/reply_reviews/New_reply_to_comments.tex
Reply to comments finalized, paper also almost finished.. need to update table 1

@@
 \usepackage{color}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \title{Response to Reviewers: Manuscript ID 17-1142.1: ``Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A Linear Parameter Varying and Synergetic Control Approach"}
@@
 \maketitle
+\small
 First of all, we would like to thank all reviewers for the time they spent reviewing our paper.
@@
 We have addressed all the necessary comments and suggestions in the attached manuscript accordingly.\\
+
 % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
@@
 This only modifies the Jacobian $J(q)$ which is now $J(q)\in \mathbb{R}^{9 \times 7}$ rather than $J(q)\in \mathbb{R}^{3 \times 7}$.
-The role of the $\mathcal{A}(q)$ and their meaning is the same.
+The role of $\mathcal{A}(q)$ remains  the same.
+The following paragraph has been incorporated, as well as some additions in the abstract and problem statement:
+\textcolor{blue}{\small ``Finally, by learning a behavior in joint-space we can inherently reach a task-space target, not only in the Cartesian 3-D space $\mathbb{R}^3$, but also in the space of orientations $SO(3)$.
+Planning or learning the rotational component of motion is a challenging problem.
+Representing an orientation as a vector in Euclidean space may lead to inaccurate and unstable motions, due to its directional nature and vulnerability to singularities.
+Several works have proposed tailor-made learning approaches that consider the non-Euclidean geometry of the $SO(3)$ space to generate rotational motion \cite{KIM201728} \cite{6907291} \cite{7829369}.
+These approaches, however, require an explicit coupling between position and orientation, that might cause discontinuities in the resulting end-effector motion.
+Such coupling is not necessary if the motion is encoded in joint-space."}
+
+Furthermore, we have re-learnt all demonstrated behaviors incorporating orientation in the task-space target $x^*$.
+Table I has been updated accordingly, showing no deterioration in generalization and performance.
+Finally, as will be discussed in the response to the Comments 10 and 12, we compare our approach to a Task-Space DS learnt in both position+orientation, showing a considerable improvement in accuracy while reaching the target.
 
@@
 Computational burden only becomes a real issue for higher degrees of freedom, like a multi-fingered hand or multi-arm robotic setting.
-We will remove this passage from the text as we don't have much space to discuss it....\\
+We removed this remark in order to allow space to discuss about other issues.
+\\
 
@@
 
-\item \textcolor{red}{\textit{``The experiments have the same task-space goal as the
+\item \textcolor{blue}{\textit{``The experiments have the same task-space goal as the
 demonstrations. Can the authors include experiments that have goals
 different from demonstrated, especially in orientation?"}}\\
-\texttt{Response:} \small First of all, they don't.. show some figures... add clarification of this in the text and explain why we cannot systematically evaluate this if we had too..\\
+\texttt{Response:} \small Yes.
+As we have shown in Appendix I, regardless of the target location (position/orientation), the motion generated by the proposed dynamical system asymptotically converges to the desired target.
+However, it is important to note that as (8) in Appendix I is semi-negative-definite and not negative-definite, there might be joint configuration $ (q )$ which result in $ J(q)\sum\limits_{k=1}^{K}\theta_k(q)A_kJ^T(q)=0 $.
+In this situation, the robot is stuck in a deadlock and, consequently, cannot reach the desired target.
 
-\item \textcolor{red}{\textit{``In similar light, maybe authors could present online adaptation of
+\item \textcolor{blue}{\textit{``In similar light, maybe authors could present online adaptation of
 A(q) to generalize to situations different than demonstrated, for
 example a slightly moved obstacle?"}}\\
-\texttt{Response:} \small This is a really nice suggestion.
-But how I see it, it's basically a new paper; i.e.
-an extension of this one, but we can discuss about this if we have space.
-A quick idea that comes to mind would be to apply a modulation/activation to the LPV system in a local region.
-Now this is not as straight forward as in our task space approaches because our LPV system is learned in joint-space while the activation function is in the embedded space.
-Another idea is to modify the activation function of the local behavior regions such that it is a function of an external variable, for example if force is applied/sensed, and it bias the model towards another $A_k$ matrix?
-So it still would be an open question, but an approach to tackle this problem could be along these lines....
-Add a passage in the discussion answering this question\\
+\texttt{Response:} \small The reviewer has pointed to a very interesting topic and authors appreciate the reviewer’s suggestion.
+In this paper, we assumed that the learned model (namely $ \pi $s, $ \mu $s, $ \Sigma $s and $A$s) does not need to be updated or modified during the task execution.
+Hence, this would be out of the scope of this paper and we are planning to address it in future works.
 
-\item \textcolor{red}{\textit{``In table II, instead of the tracking error, maybe the more useful
+\item \textcolor{blue}{\textit{``In table II, instead of the tracking error, maybe the more useful
 error to see would be the error with respect to the plan or
@@
 the velocity and goal, not the position of the movement."}}\\
-\texttt{Response:} \small This is not necessarily the reason that it would be worse, as SEDS is also learned by minimizing the error with respect to
-the velocity and the goal.
-The reason that JT-DS is less accurate in reproducing the motion in task-space is because our optimization is focused on reproducing the motion solely in joint-space, while reaching the final target in task-space.
-Nowhere do we claim that JT-DS is capable of reproducing a motion in both joint and task space.
-Nevertheless, we add the mean swept-error-arrea (SEA) which evaluate the performance of reproduction in task-space from a reference trajectory...
-as shown in the tables....\\
+\texttt{Response:} \small This is, in general, not true.
+Based on the robot's motion and its configuration, there might be cases where the motion generated by SEDS does not even reach the target.
+To clarify this, we have systematically studied the performance of the motion generated by JT-DS and SEDS+IK in terms of (i) reaching to the desired target and (ii) following the demonstrated end-effector motions, See Figure \ref{fig:SEDSvsJTDS}.
+We used the data-sets from the seven tasks mentioned in Section V.A to construct SEDS and JT-DS.
+Both methods are tested by the same the target and the initial configurations, which were extracted from the demonstrations.
+We define $ x^*=[x^*_{ori};x^*_{pos}] $ as the target configuration that includes the desired position as well as the desired  orientation of the end-effector.
+The motion generated by SEDS is converted into the 7-DOF joints state using the damped least square IK solver.\\
+In Figure \ref{fig:SEDSvsJTDSP:E} and Figure \ref{fig:SEDSvsJTDSO:E}, the error between the final location of the robot and the target is illustrated.
+As it can be seen, the motion generated by the proposed dynamical system reaches the target, whereas, the motion generated by SEDS does not; regardless of the fact that both models are theoretically stable.
+The main cause of failure for SEDS is due to the IK solver.
+As SEDS generates the motion of the robot without considering its joint configurations, the robot might reach its joint limits.
+Whereas, as JT-Ds directly shapes the joints' motion, by learning the desired joint-space behavior, one can avoid these situations.
+
+In Figure \ref{fig:SEDSvsJTDSP} and Figure \ref{fig:SEDSvsJTDSO}, the error between the demonstrated and the excused motions are illustrated.
+As it can be seen, even though JT-DS is not tailored for following the end-effector motions, its performance is comparably better than SEDS.
+
+Although including this study would highlight one of the advantages of using JT-DS over Cartesian-motion-generators, as this does not highlight the main focus of this paper and due to the space limitations, we have decided to not include this in the paper.
+
+\begin{figure}[!t]
+ \hspace*{-2.4cm}
+	\begin{subfigure}[t]{0.63\linewidth}
+	\centering
+		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P_End.pdf}
+		\caption{}
+		\label{fig:SEDSvsJTDSP:E}
+	\end{subfigure}%
+	\hspace{-0.2cm}
+	\begin{subfigure}[t]{0.63\linewidth}
+	\centering
+		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O_End.pdf}
+		\caption{ }
+		\label{fig:SEDSvsJTDSO:E}
+	\end{subfigure}\\		 
+	\hspace*{-2.4cm}\begin{subfigure}[t]{0.63\linewidth}
+			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P.pdf}
+		\caption{ }
+			\label{fig:SEDSvsJTDSP}
+		\end{subfigure}%
+		\begin{subfigure}[t]{0.63\linewidth}
+			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O.pdf}
+			\caption{ }	
+			\label{fig:SEDSvsJTDSO}
+		\end{subfigure}\\
+	\caption{The performance comparison between SEDS+IK and JT-DS while the initial and the target points are same.
+The error between the orientations is the difference between the first two columns of the corresponding rotation matrix representation.}
+	\label{fig:SEDSvsJTDS}
+\end{figure}
+
 \end{enumerate}
@@
 % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\newpage 
 \underline{Comments from Reviewer \#2 (Reviewer: 5 (Review30371)):}
@@
 The result of the embedding is a lower-dimensional representation of the joint position.
-The following line has been added to IV.A for clarity:
+The following line has been added to IV.A for clarity:\\
 \textcolor{blue}{\small \textit{"For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the "shoulder" and "arm" components of $q$ into a single "shoulder-arm" component in $\phi(q)$.}"}
 
-Given the raw demonstrations of a joint-space  behavior with a task-space target as shown in Fig.
+\textcolor{blue}{\small Also, Section III has been considerably re-written to clarify our proposed approach and proposed definition of ``local behavior synergies".}
+
+Continuing with the analysis, given the raw demonstrations of a joint-space  behavior with a task-space target as shown in Fig.
 \ref{fig:raw_demos} (for the pouring task) we seek to learn a linear combination of local behavior synergies $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\phi(q))A_k$ such that we can use \eqref{eq:jtds} to accurately reproduce these demonstrations.
@@
 Yet, since the demonstrations are acquired via kinesthetic teaching we can assume that they are manipulable.
-\underline{As asked by the reviewer, what does it physically mean to have all positive eigenvalues?}  When \\ $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$  or ($\mathcal{A}(q)$) has some eigenvalues $=0$ the physical meaning behind it is that we are forfeiting the control of these directions of motion in joint-space.
+\\
+
+\underline{As asked by the reviewer, what does it physically mean to have all positive eigenvalues?}  When $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$  or ($\mathcal{A}(q)$) has some eigenvalues $=0$ the physical meaning behind it is that we are forfeiting the control of these directions of motion in joint-space.
 Namely, the desired velocity directions in joint-space will be constrained in the corresponding DOF of the arm, this will prohibit us from converging to the attractor and will be no longer capable of accurately following the joint-space demonstrations.
@@
 Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations."}} \\
+
 Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations and learn a \textit{time-invariant} controller that ensures reaching the task-space target while approximately following the demonstrated motion, as shown in Fig.
@@
 Hence, it can generate motions similar to the demonstrated ones at different frequencies and time durations.
+\\
 
-\noindent \textcolor{red}{TODO: Add clarifications in text such that no other reader gets confused with the new concept of behavior synergies that we propose as opposed to the classical joint posture synergies.}
+\noindent \textcolor{blue}{We have added clarifications in the text such that no other reader gets confused with the new concept of behavior synergies that we propose as opposed to the classical joint posture synergies.}
 
@@
 where $\mathbf{f}(\cdot): \mathbb{R}^M \rightarrow \mathbb{R}^M$ is a continuous differentiable vector-valued function representing a non-linear DS that converges to a single stable equilibrium point $x^*$.
-SEDS is used in the following way: Given the current state of the robot $x$ the next desired velocity $\dot{x}_d$ is computed with \eqref{eq:simple-ds} this is then fed to a velocity-based IK solver as shown in Fig.
-\ref{}:\\
-\textcolor{red}{TO SINA: Include typical SEDS-like control-loop here}\\
-If we were to use SEDS with the JT-control method, we would have to integrate the desired velocity to generate a desired position $x_d$ at each time-step and define it as the target in \eqref{eq:JT}, which would yield the following controller:
-\begin{equation}
-\dot{q} = K J(q)^T(H(q)- x_d) 
-\end{equation} Note that, the JT control method assumes that the target is static, hence, in order for this controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
+In order to prove the stability of SEDS, one needs to assume that the dynamic of the robot is perfectly compensated by the built-in position-level controller.
+In other words, the transfer function of the dynamic of the robot+ the low-level controller and the IK solver must be one.
+\\
+Almost all the commercially available robots are equipped by a very precises joint position controller.
+Hence, the transfer function of the robot+ the low-level controller can be assumed one.
+Whereas, the transfer function of the Jacobian transpose IK solver can not be assumed one as the JT control method assumes that the target is static.
+Hence, in order for the JT controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
 This is contradicting as \eqref{eq:simple-ds} is always trying to converge to $x^*$.
 Thus, merging these two approaches is theoretically ill-fitting.
-To better highlight this, we show an illustration of the supposed SEDS+JT control loop in Fig.
-\ref{}:\\
-\textcolor{red}{TO SINA: Include SEDS+JT control-loop here}\\
-\textcolor{red}{\small Describe this control-loop and state more reasons (if any) why it's wrong...}
+Note that,  To better highlight this, we show an illustration of the supposed SEDS+JT and JT-DS control loops in Figure \ref{fig:schematic:SEDS} and  Figure \ref{fig:schematic:jt}, respectively.\\
+
+\begin{figure}[t]
+\centering
+	\begin{subfigure}[t]{0.8\textwidth}
+	\centering
+		\includegraphics[trim={0 0 0.25cm 1.1cm},clip,width=\textwidth]{Pics/Drawing2.png}
+		\caption{ A Cartesian motion generator (e.g., SEDS) with the Jacobian transpose IK solver.}
+		\label{fig:schematic:SEDS}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{0.8\textwidth}
+	\centering
+		\includegraphics[trim={0.25cm 0 0.0cm 1.2cm},clip,width=\textwidth]{Pics/JT_DS.png}
+		\caption{ JT-DS.}
+		\label{fig:schematic:jt}
+	\end{subfigure}\\
+	\caption{A simplified schematic comparing the standard control architecture using Cartesian dynamical systems with the Jacobian transpose IK solver (top) and our propose approach.
+By integrating the "inverse kinematic (IK) solver" and "motion generator" blocks into a unified block, the proposed approach eliminates undesirable effects of the IK solver.
+}
+\end{figure}
 
@@
 
-\item \textcolor{red}{\textit{`some details about the optimization problem in (8) might be given
+\item \textcolor{blue}{\textit{`some details about the optimization problem in (8) might be given
 (nature of the problem, complexity, feasibility)."}}\\
 \texttt{Response:} The following has been added near Eq. (8):\\
-\textcolor{blue}{\small \textit{``The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective..." ``This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD), and the number of terms in the objective is $O\left(k^2d^3\left( \sum_{n=1}^N T_n\right)\right)$."}}\\
-To elaborate about the prior calculation: the number of terms in the sum is proportional to the number of recorded points, times the number of matrices (k) required to generate the predicted velocity from (7) squared, times the number of terms in a matrix multiplication of two $d x d$ matrices ($d^3$).\\
+\small \textit{``The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective..." ``This optimization is always feasible, the only constraint is that the $A_k$s be independently PSD."}\\
 

### 2018-03-15 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
Reply to comments finalized, paper also almost finished.. need to update table 1

@@
 \IEEEpeerreviewmaketitle
+\vspace{-10pt}
 \section{Introduction}
@@
 Robot motion planning in joint-space has long been a major field of study \cite{kelly2006control}.
-For manipulation problems with an objective defined in the task space, %(i.e.
-target or desired trajectory)
- we can often find a myriad of joint-space trajectories to achieve the same task-space goal.
+For manipulation problems with an objective defined in task-space, we can often find a myriad of joint-space trajectories to achieve the same task-space goal.
 In many cases, however, certain joint-space trajectories are favored over others; for example, when we expect the robot to follow a desired joint-space behavior or ``style", as illustrated in Fig.
@@
 In this paper, we will explore the problem of learning a preferred joint-space behavior from previously demonstrated trajectories while still accomplishing a task-space goal, through the Learning from Demonstrations (LfD) paradigm \cite{billard2008robot,Argall:RAS:2009}.
+Most LfD approaches learn motions either in joint space
+\cite{Garrido:Neuro:2015,Schaal:AI:2002} or in task space
+\cite{Pastor:ICRA:2009,Calinon:ISR:2015} via probabilistic models.
+To ensure that the learnt models reach the desired task-space target, Dynamical System (DS) formulations have been used to ensure stability and  convergence \cite{khansari2011learning, Neumann:RAS:2015}.
 
-Many LfD approaches learn motions in either joint space %\cite{Calinon:HRI:2007, Garrido:Neuro:2015, Schaal:IROS:2003,Schaal:AI:2002}
-\cite{Garrido:Neuro:2015,Schaal:AI:2002} or task space %\cite{Pastor:ICRA:2009,Gribovskaya:IJRR:2011,Calinon:ISR:2015}
-\cite{Pastor:ICRA:2009,Calinon:ISR:2015}.
-These approaches generally learn a probabilistic model of the demonstrated motions.
-%, and construct a dynamical system that chooses behaviors according to the learned model.
-%Both joint and task spaces are independently useful:  task-space behaviors let us control the most task-relevant component of the robot's behavior. 
-A number of approaches have further proposed stable convergence to an attractor in task space % \cite{khansari2011learning, Neumann:RAS:2015, Kronander:RAS:2015} 
- \cite{khansari2011learning, Neumann:RAS:2015}as a desirable capability for a dynamical system.
+In many cases, however, one may want to specify a joint-space behavior that the physical robot should follow, while simultaneously reaching a task-space objective.
+For example, learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}.
+For a visual example of a task-space problem in which joint behavior is essential, see Fig.\ref{fig:robot_example}.
+Moreover, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations.
+Task-space motion generators, as the ones learned from demonstrations, all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}.
+When the main focus is on executing a specific task-space behavior, regardless of a joint-space constraint, this approach is sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
+However, for other applications, such an approach yields significant problems \cite{buss2004introduction}.
+Mainly, when the Jacobian matrix cannot be inverted (i.e.
+when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior.
+This encapsulates the main source of inaccuracies in tasks that require fast dynamical motions, such as catching/reaching for moving objects \cite{7439839,Salehian-RSS-16}.
+%Defining a controller in joint-space would avoid IK and all its associated drawbacks.
 
-	In many cases, one may want to specify a joint-space behavior to directly define the behavior of the physical robot while simultaneously fulfilling a task-space objective.
-For example, %throwing a ball is a motion characterized both by a task-space behavior (the ball moves backwards and then forwards) and also more obviously by its joint-space behavior (the elbow folding, the shoulder moving back and then swinging forward, and then the elbow straightening and the wrist snapping down).
-Similarly,
-learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}.
-For a visual example of a task-space problem in which joint behavior is essential, see Fig.\ref{fig:robot_example}.
-
+Finally, by learning a behavior in joint-space we can inherently reach a task-space target, not only in the Cartesian 3-D space $\mathbb{R}^3$, but also in the space of orientations $SO(3)$.
+Planning or learning the rotational component of motion is a challenging problem.
+Representing an orientation as a vector in Euclidean space may lead to inaccurate and unstable motions, due to its directional nature and vulnerability to singularities.
+Several works have proposed tailor-made learning approaches that consider the non-Euclidean geometry of the $SO(3)$ space to generate rotational motion \cite{KIM201728} \cite{6907291} \cite{7829369}.
+These approaches, however, require an explicit coupling between position and orientation, that might cause discontinuities in the resulting motion.
+Such coupling is not necessary if the motion is encoded in joint-space.
 \begin{figure}[t]
@@
 
-Furthermore, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations.
-The previously discussed task-space dynamical systems all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}.
-When the main focus is on executing a non-linear task-space behavior, regardless of a specific joint-space constraint, this approach has been deemed sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
-However, for other applications, such an approach yields significant problems \cite{buss2004introduction}.
-First of all, finding the pseudo-inverse is computationally taxing \textbf{insert how slow it is for a 1000Hz robot}.
-Moreover, when the Jacobian matrix cannot be inverted (i.e.
-when the robot is near a singularity) its behavior becomes erratic, requiring layers of additional engineering to generate smooth trajectories and ensure the desired task-space behavior.
-These problems encapsulate the main source of inaccuracies in tasks that require generating fast dynamical motions, such as catching or reaching for moving objects \cite{7439839,Salehian-RSS-16}.
-By defining a controller in joint space, one can avoid IK and all its associated drawbacks.
-
 Several approaches have tackled the problem of learning joint-space behaviors with task-space objectives.
@@
 % FACT-CHECK: is this what Calinon 2009 and Silverio et al. 2017 are doing? They don't ensure convergence, right? TP-GMM never seems to address it ->CORRECT
-\cite{lee2014unifying} learns from demonstration in both joint and task space within the ``thin-plate spline" trajectory warping framework.
+%\cite{lee2014unifying} learns from demonstration in both joint and task space within the ``thin-plate spline" trajectory warping framework.
 % but does not propose a dynamical system for generating motions.
-\cite{hersch2008reaching} proposed an approach with similar properties to this desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities.
+\cite{hersch2008reaching} proposed an approach with similar properties to our desiderata, where two concurrent DS, one in task-space and one in joint-space, are modulated by enforcing kinematic coherence constraints to avoid singularities.
 The resulting DS avoids singularities through generalization of the pseudo-inverse approximations.
@@
 This gives rise to spurious attractors or cycles, and thus must be carefully tuned in order to avoid them.
-%Most recently, \cite{huang2016jointly} proposed an approach that learns a controller for playing table tennis by combining joint-space Dynamic Movement Primitives (DMP) \cite{schaal2006dynamic} to mimic a desired task-space racket swing..
-\textcolor{red}{ must re-read this...
-I think they are doing something very similar...
-can remove if problematic}
-% Do DMPs count as learning joint space behaviors? If yes, I'm having a hard time differentiating huang2016 from us -> Yes they do
 
-%\subsection{Our Proposed Approach}
 In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
 To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system.
-We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions, and then learning a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
+We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions.
+We then learn a policy in embedding space for combining our behavior synergies to accurately reconstruct the demonstrated trajectories.
 Hence, our dynamical system:
@@
 The JT control is an IK method that yields a dynamical system in joint space
-which converges stably over time to a desired end-effector position, without the need for pseudo-inverse computations.
+which converges stably over time to a desired end-effector target, without the need for pseudo-inverse computations.
 It shares some of our approach's advantages: fast computation and provable task-space stability.
@@
 Consider a robotic system with $d$ task-space dimensions and $m$ degrees of freedom.
-We direct the system using a joint position or joint velocity controller, which can have joint position limits and a maximum joint velocity.
+The system is directed via a joint-position or joint-velocity controller.
 We are further provided with a set of $N$ demonstrated joint-space trajectories $D=\{\{q_{t,n},\dot{q}_{t,n}\}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$, where $T_n$ is the number of the sample points of the $n^{\text{th}}$ demonstration.
 We refer to the system's joint-space position as $q=\begin{bmatrix} q^1 & \dots & q^m
-\end{bmatrix}^T \in \mathbb{R}^m$, and to its task-space position as $x\in \mathbb{R}^d$.\footnote{For sake of brevity and simplicity, the time index, $t$, is dropped throughout the paper.
-} The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{m\times d}$.\\
-We wish to formulate a dynamical system $\dot{q} = f(q)$ which satisfies the following two criteria:
+\end{bmatrix}^T \in \mathbb{R}^m$, and to its task-space target as $x\in \mathbb{R}^d$.\footnote{For pure position target we consider $d=3$, for position and orientation $d=9$, where the first 3 dimensions correspond to Cartesian position and the remaining 6 correspond to the first and second columns of a rotation matrix $R \in SO(3)$.} The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{d\times m}$.
+We wish to formulate a DS $\dot{q} = f(q)$ which satisfies the following two criteria:
 \renewcommand{\labelenumi}{(\Roman{enumi})}
 \begin{enumerate}
-\item The dynamical system must be asymptotically stable\footnote{Unless otherwise specified, ``stability" in this paper always refers to asymptotic stability within the workspace of the robot (in the regions where the Jacobian is full rank), and as such assumes no joint limits.
+\item The DS must be asymptotically stable\footnote{Unless otherwise specified, ``stability" in this paper always refers to asymptotic stability within the workspace of the robot (in the regions where the Jacobian is full rank), and as such assumes no joint limits.
 We make no claim to proving global asymptotic stability, which is in fact impossible to achieve in a joint-constrained kinematic system.} with respect to a fixed task-space target $x^*$.
-This can be expressed by ensuring that the following Lyapunov function
-\begin{equation}
+This can be expressed by ensuring that the following Lyapunov function\begin{equation}
 V(q) = (H(q) - x^*)^T(H(q) - x^*) 
@@
 $V(.)$ can be thought of as a metric for the task-space distance-to-go.
-\item The dynamical system should encapsulate the desired joint-space behaviors such that the following metric is minimized
+\item The DS should encapsulate the desired joint-space behaviors such that the following metric is minimized
 \begin{equation}
@@
 
-%Lastly, we wish to capture joint-space behaviors in our dynamical system.
+\noindent The error metric \eqref{eq:optim_error} is advantageous in that it mimics the magnitude and direction of the demonstrated motions.
+This makes the reproduced motion visually most similar to the human definition of ``joint motion style" \cite{gielniak2010stylized}.
 
-%If instead of \eqref{eq:optim_error}, the error feature was the joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity.
-Alternatively, if the error feature was the joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior).
-By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
-
-The error metric \eqref{eq:optim_error} is advantageous in that it mimics the magnitude and direction of the demonstrated motions.
-Hence, as suggested by \cite{gielniak2010stylized}, it is visually most similar to the human definition of ``joint motion style".
-
-% ways. First, it implicitly suggests that behaviors are defined by expert motions of the behavior
-
-%We make two decisions by the choice of behavior error metric \eqref{eq:optim_error}.
-First, the metric implicitly suggests that behaviors are defined by expert motions of the behavior, and that fulfilling a behavior means moving similarly to the experts' motion.
-Second, the type of feature of the trajectory that is being minimized is important.
-If instead the error feature were joint position distance $\sum_i\left \| q_{d,i} - q_i \right \|$ where $q_d$ is a vector of demonstrated positions, then executing a "behavior" would imply mimicking position, but not velocity.
-Alternatively, if the error feature were joint velocity direction $\sum_i \left | \frac{\dot{q}_{d,i}}{\left \| \dot{q}_{d,i} \right \|} - \frac{\dot{q}_i}{\left \| \dot{q}_i \right \|} \right |$ mimicking a "behavior" would involve following the motion profile, but not the motion's speed (so for example a slap and a push would exhibit the same behavior).
-By choosing the combined direction and magnitude of the joint velocity as the error, we are choosing to mimic the magnitude and direction of the motion, which \cite{gielniak2010stylized} suggests is visually most similar to the human definition of "joint motion style".
 \section{Augmented \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem} \label{Sec:DS}
 \label{sec:proposed_system}
-We propose the following augmented Joint-space Task-oriented Dynamical System (JTDS) to achieve the two criteria presented in \eqref{eq:Lyp} and \eqref{eq:optim_error}.
+To achieve the two criteria presented in \eqref{eq:Lyp} and \eqref{eq:optim_error}, we propose the following DS, which we refer to as the augmented Joint-space Task-oriented DS (JT-DS): 
 \begin{equation}
@@
 \end{equation}
-where  $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the LPV system paradigm \cite{emedi2016fixed,7439839}, where the overall $\mathcal{A}(q)$ is a changing linear combination of constant  matrices, each of which encodes a local joint-space synergy.
-Each ``synergy" is a joint-space transformation that biases the resulting motion to use particular joints; for more, see Sec.
-\ref{Sec:Learning}.
-Explicitly,
-\begin{equation}
-\label{eq:A_def}
-\mathcal{A}(q) = \sum_{k=1}^{K}\theta_k(q)A_k 
-\end{equation}
-where $A_k\in \mathbb{R}^{m\times m} $ are the different synergy matrices and $\theta_k(q)\in \mathbb{R}^{1}~\forall k\in\{1,\dots,K\} $ are the scheduling parameters\footnote{%The scheduling parameters can be a function of time $t$, the states of the
-%system $q $ or external signals $d(t)$, i.e. $\theta_k (t, q(t), d(t))$. In this paper, we only consider it as a function of the states of the system. 
-It is noteworthy that the presented stability proof can easily be extended for $\theta_k (t, q(t), d(t))$, where $d(t)$ is an external signal.} modulating each local synergy through time as well as space.
-Based on \eqref{eq:ds}, one can utilize different synergies in different regions, and compose them to create a more non-linear multi-behavior motion.
+where  $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ is constructed using the LPV system paradigm \cite{emedi2016fixed,7439839}.
+$\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\cdot)A_k$ is a linear combination of time-invariant linear matrices $A_k \in \mathbb{R}^{m \times m}$, each of which encodes a ``local joint-space behavior synergy" that shapes the motion in joint-space.
+Specifically, the ``local synergy matrices" $A_k$ are joint-space transformations that bias the resulting motion to use particular joints.
+By modulating each local synergy in time as well as space through the activation function $\theta_k(\cdot)$ we can generate the desired non-linear joint-space behaviors.
+Note that, in the LPV paradigm $\theta_k(\cdot)$ can be a function of time $t$, ``joint-posture" $q$, an external signal $d(t)$ or a lower-dimensional representation of the joint posture $\phi(q)$.
+Because we seek a time-invariant controller and know that ``joint-postures" can be accurately represented in a lower-dimensional space, in this work we chose to parametrize $\theta_k(\cdot)$ with $\phi(q)$.
 
-Before proving that the proposed dynamical system satisfies all of the criteria, let us establish an intuitive understanding of the components of the system.
-While it may seem daunting at first, each of the elements has a straightforward explanation.
-One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
-the task target, which we then reinterpret as the task-space velocity of a proportional controller.
+\begin{figure}[!t]
+	\centering
+	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
+	\caption{Illustration of our task-space and proposed joint-space Lyapunov function.
+On the left we show the task-space error we are trying to minimize, represented by a potential function for a 2-DOF robot with target $x^*=[2,2]^T$.
+On the right we show the corresponding Lyapunov function in joint space defined in \eqref{eq:Lyp}.
+The colors on each plot correspond to each other.
+One can see how the error, and consequently the \textit{attractive} regions in task-space, have been warped into attractive regions in joint-space.}
+	\label{fig:error_schematic}
+	\vspace{-15pt}
+\end{figure}
+
+Before proving that \eqref{eq:ds} satisfies all of the criteria, one can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the task-space error w.r.t.
+the target, which we then reinterpret as the task-space velocity of a proportional controller.
 By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig.
@@
 $\mathcal{A}(q)$ as the \begin{bf} joint augmentation matrix \end{bf} as it augments the outputted joint velocities.
-%When using each of these matrices, the controller will augment the joint velocities to favor using one joint more than the others (in $A_1$ the hip, in $A_2$ the knee, and in $A_3$ the ankle).
-This ''augmentation" is simply a matrix multiplication ($A\dot{q}_{raw}$).
-The components of the outputted velocity are linear combinations of the ''raw" un-augmented joint velocities, where the entries in $A$ define the precise ratios.
-We can see the joint positions resulting from our experiment over time on the right side of Fig.
-\ref{fig:A_example}.
-Each motion tends to use its ``primary" joint most (the joint where the velocity is heavily scaled up) and uses the other available joints to compensate for what the primary joint cannot do.
-For example, Motion 1 starts out using mostly the hip, but once moving the hip no longer brings the system closer to the goal, the system uses the other joints to fold inwards and reach the goal.
-Ultimately, all three controllers converge on their target, but each has a different path through joint space depending on which joints we choose to emphasize in the behavior matrix $A$.
-The off-diagonal terms of $A$ have a more nuanced contribution; e.g.
-$A_{1,2}$ defines the amount that elbow velocity should be converted into hip velocity.
-
 \newtheorem{prop1}{Proposition}
@@
 \label{prop:stability}
-The motion generated by the dynamical system \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ \eqref{eq:ds} meets the following constraints. 
+The DS in \eqref{eq:ds} accomplishes criteria (I) if $ \forall k \in \{1,\dots,K\}  $ the following constraints are met. 
 \begin{equation}
@@
 \textbf{Proof}: See Appendix \ref{appendix:stability}. $\blacksquare$
-%It is worth mentioning that the constraints in \eqref{eq:first_criteria_stability}  only ensure that JT-DS \eqref{eq:ds} asymptotically converges to the desired target position $x^*$ if the robot does not have kinematic joint limits.
-If we introduce kinematic joint limits, then it's possible we may find $\exists q\in Q-Q^* $ such that both $\dot{V}(q)=0$ and $H(q) \neq x^*$.
-This happens if the target position is kinematically unreachable; i.e.
-for each task-space axis $i$,  $\big ( J^T(q)P(H(q)-x^*) \big)_i = 0 $ (meaning that moving joints increases the value of \eqref{eq:Lyp}), or the joint has reached a kinematic limit.
-In this case, the proof in Appendix \ref{appendix:stability} shows instead that the system converges in the sense of Lyapunov, meaning that its distance to the attractor is always monotonically decreasing.
 
 Criterion (II) (i.e.
-encoding specific joint-space behaviors) is achieved by embedding the desired dynamics in the matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$.
-We describe in the next section an approach to automatically learn these matrices from demonstrated data.
-
-%We achieve criterion (III), encoding joint-space behaviors in \eqref{eq:ds}, by approximating the joint augmentation matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$.
-It is important to understand how varying $A_k(q)$ will vary the system's joint space behavior.
-We know from \textit{Proposition  \ref{prop:stability}} that, regardless of choice of $A_k(q)$, the robot's motion is stable with respect to the target $x^*$.
-Yet this still leaves us with many different potential joint velocities to choose from, all of which would decrease $(H(q)-x^*)^TP(H(q)-x^*)$.
-The choice of $A_k(q)$ becomes our way to influence which of the many possible joint velocity motions is chosen.
-Thus we must clarify how the choice of $A$ affects the system's motion.
-
-%By learning a position-dependent joint augmentation matrix $\mathcal{A}(q)$, we are able to execute the joint behaviors discussed earlier by augment the relative velocities of our joints.
-\eqref{eq:A_def} constructs our overall $\mathcal{A}(q)$ as a linear combination of $A_k~\forall k\in\{1,\dots,K\}$ (each weighted by $\theta_k(q)$), where one can enforce different local joint behaviors in different regions, and compose them to create a more complex multi-behavior motion.
-In this paper the joint behavior matrix $\mathcal{A}(q)$ is determined using only the joint position $q$, though one could potentially use other information such as joint velocity $\dot{q}$, environment force or target task position $x^*$.\\
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
-	\caption{Illustration of our task-space potential function and proposed joint-space Lyapunov function.
-On the left we show the task-space error we are trying to minimize, represented by a potential function for a 2-DOF robot with target $x^*=[2,2]^T$.
-On the right we show the corresponding Lyapunov function in joint space defined in \eqref{eq:Lyp}.
-The colors on each plot correspond to each other.
-One can see how the error, and consequently the \textit{attractive} regions in task-space, have been warped into attractive regions in joint-space.}
-	\label{fig:error_schematic}
-	\vspace{-15pt}
-\end{figure}
+encoding specific joint-space behaviors) is achieved by embedding the desired joint-space behavior in the matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$.
+We describe in the next section an approach to automatically learn the number of $K$ matrices, their values and their corresponding activation functions $\theta_k(\cdot)$ from demonstrated data.
 
@@
 In robotic synergy control, a robot's movements can be decomposed into a small number of synergies: principal components of the joint-space that are sufficient to accurately recreate the desired robotic behaviors.
-In our case, the synergies are represented by $A_1, A_2, \dots, A_k$, and $\mathcal{A}(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
+In our case, the synergies are represented by the matrices $A_1, A_2, \dots, A_k$, and $\mathcal{A}(q)$ represents the resulting motion constructed from a superposition of different synergies (through \eqref{eq:ds}).
 
@@
 We would like to define the robot's policy in embedding space such that in different regions of the space, we will prioritize different synergies.
-We are thus left with three problems: (I) finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), (II) finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(q)~\forall k\in\{1,\dots,K\}$), and (III) finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
+We are thus left with three problems: (I) finding an underlying synergy-space $Z$ in which the behavior can be accurately controlled (defined by a mapping $\phi: Q \rightarrow Z$), (II) finding a policy for modulating the synergies in different regions of the synergy space (defined by $\theta_k(\phi(q))~\forall k\in\{1,\dots,K\}$), and (III) finding parameters for the synergies themselves (defined by $A_k~\forall k\in\{1,\dots,K\}$).
 Our choice of parameters must also obey the constraints laid out in \eqref{eq:first_criteria_stability}.
@@
 
-\begin{figure}[t]
+\begin{figure}[!t]
 	\centering
@@
 \item We first construct our embedding, which provides us a lower-dimensional manifold through which to control the robot, by projecting the demonstration data (i.e.
-collections of joint positions $q$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{\delta\leq d}$.
-In each experiment, we used either Principal Component Analysis (PCA) \cite{jolliffe1986pca} or Kernel PCA (KPCA) with RBF kernel \cite{scholkopf1997kernel}.
+collections of joint positions $q \in \mathbb{R}^{m}$) into a lower-dimensional embedding $\phi(q)\in \mathbb{R}^{p\leq m}$.
+In Section \ref{Sec:Exp} we evaluated Principal Component Analysis (PCA) \cite{jolliffe1986pca} or Kernel PCA (KPCA) with RBF kernel \cite{scholkopf1997kernel}.
 
-\item We then jointly estimate the optimal number $K$ of synergy ``regions" and the parameters of the scalar functions that determine the scheduling parameters $\theta_k(q)$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$ seen in the demonstrations.
+\item We then jointly estimate the optimal number $K$ of ``local synergy regions" and the parameters of the scalar functions that determine the activation parameters $\theta_k(\phi(q))$ for weighting these synergies, by fitting a Gaussian Mixture Model (GMM) on the projected joint positions $\phi(q)$ seen in the demonstrations.
 
-\item Finally, once the local synergy regions have been found (described by each of the Gaussian distributions $\theta_k(q)$), we compute the corresponding joint synergy matrices $A_k~\forall k\in\{1,\dots,K\}$ for each region by formulating a convex optimization problem that finds the optimal set of $A_k$'s that minimize the overall velocity error with respect to the demonstrations \eqref{eq:optim_error}.
+\item Once the local synergy regions have been found (described by each of the Gaussian distributions $\theta_k(\phi(q))$), we compute the corresponding joint synergy matrices $A_k~\forall k\in\{1,\dots,K\}$ for each region by formulating a convex optimization problem that finds the optimal set of $A_k$'s that minimize the overall velocity error with respect to the demonstrations \eqref{eq:optim_error}.
 \end{enumerate}
@@
 \subsection{Embedding Joint Configurations in Low-Dim. Space}
-The search for a lower-dimensional embedding of the joint space stems from the desire to identify a simplified coordinate system in which each principal component corresponded to an important source of variation in the demonstrated trajectories, and which is thus suitable for learning the demonstrated behavior.
+The search for a lower-dimensional embedding of the joint space stems from the desire to identify a simplified coordinate system in which each principal component corresponds to an important source of variation in the demonstrated trajectories, and which is thus suitable for learning the demonstrated behavior.
 Motor control studies have postulated that human arm motions like reaching or following straight/curved line trajectories, rather than utilizing the full joint space $|q|$, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
@@
 To this end, we construct a single global embedding $\phi(q)$ by training dimensionality reduction techniques on the demonstrated trajectories.
-The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^d$ into a lower-dimensional configuration $z \in \mathbb{R}^\delta$, where $\delta < d$.
-For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the ``shoulder" and ``arm" components of $q$ into a single ``shoulder-arm" component in $\phi(q)$.
-%Specifically, we experimented with Principal Component Analysis (PCA) and Kernel PCA \cite{scholkopf1997kernel}. 
-%In previous work, PCA has shown to be sufficient to encapsulate the correlations in joint-space data \cite{calinon2005recognition}, and this work further explores the usefulness of joint-space dimensionality reduction.
-
-\nocite{*}
+The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^m$ into a lower-dimensional configuration $z \in \mathbb{R}^p$, where $p < m$.
+For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(q)$ were a matrix multiplication; i.e.
+$\phi(q) = A_p\times q$ for $A_p \in \mathbb{R}^{p\times m}$, it could map the ``shoulder" and ``arm" components of $q$ into a single ``shoulder-arm" component in $\phi(q)$.
 
@@
 The next step is identifying the regions of space in which to activate different synergies.
-Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding scheduling parameter $\theta_k(q) = \theta'_k(\phi(q))$.
-Moreover, we would like for our scheduling parameters $\theta'_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta'_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
-
-Scheduling parameters for LPV systems with such properties have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}.
-Intuitively, we search for a probabilistic model that ``explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a local behavior, which the synergy will then approximate.
-In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the scheduling parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and covariance matrices that parametrize the $k$-th multivariate Gaussian distribution.
-Each distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the scheduling parameter $\theta'_k$ to the $k$th synergy of \eqref{eq:A_def}.
-We define $\theta'_k(\phi(q))$ as $p(k|\phi(q))$:
+Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding activation function $\theta_k(\phi(q))$.
+Moreover, we would like for $\theta_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
+Such activation functions, also known as scheduling parameters for LPV systems, have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}.
+Intuitively, we search for a probabilistic model that ``explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a \textit{local behavior}, which the synergy will then approximate.
+In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the activation parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and Covariance matrices that parametrize the $k$-th multivariate Gaussian distribution.
+Each distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the activation functions $\theta_k$ of the $k$th synergy matrix.
+We define $\theta_k(\phi(q))$ as $p(k|\phi(q))$:
 \begin{equation}
 \label{eq:theta}
-\theta'_k(\phi(q))= \frac{\pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}
+\theta_k(\phi(q))= \frac{\pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}{\sum_{k=1}^K \pi_k\mathcal{N} (\phi(q); \mu_k, \Sigma_k)}
 \end{equation}
-which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th synergy.
+which is the probability of the projected joint-position $\phi(q)$ belonging to the $k$-th local synergy region.
 Therefore, each synergy region is associated with a Gaussian component of the GMM, cumulatively describing all the synergy regions of the dynamical system.
-We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM \textbf{ADD CITATION FOR EM}.
+We use the standard Expectation Maximization (EM) training algorithm to estimate the parameters of the GMM \cite{Bilmes98agentle}.
+We follow a model selection approach described in Section \ref{Sec:Exp} to find the optimal $K$.
 
 \subsection{Estimating the Synergy Matrices}
-Given the parameters of $\theta_k'(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, from \eqref{eq:A_def} one can construct $\mathcal{A}(q)$ as a linear combination of local $A_k$ synergy matrices weighted by their scheduling parameters $\theta_k'(\phi(q))$ as follows:
+Given the parameters of $\theta_k(\phi(q))$ $\forall k \in \{k=1,\dots,K\}$, one can construct $\mathcal{A}(q)$ as a linear combination of local $A_k$ synergy matrices as follows: 
 \begin{equation}
@@
 \cite{nadaraya1964regress,watson1964regress}} with a Gaussian pdf as its kernel function.
-Hence \eqref{eq:A} can be considered a type of kernel estimator, with the key distinction that the weighting functions $\theta_k'(\phi(q))$ are not determined by individual points (as in the original Nadaraya-Watson kernel estimator) but by the components of a GMM, similar to the weighting functions derived in Gaussian Mixture Regression (GMR).
+Hence \eqref{eq:A} can be considered a type of kernel estimator, with the key distinction that the weighting functions $\theta_k(\phi(q))$ are not determined by individual points (as in the original Nadaraya-Watson kernel estimator) but by the components of a GMM, similar to the weighting functions derived in Gaussian Mixture Regression (GMR).
 
-Intuitively, each ``synergy region" in synergy space is defined by a Gaussian distribution, and the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
+Intuitively, each ``synergy region" is defined by a Gaussian distribution in the lower-dimensional space, the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
 Finding the appropriate synergy matrices $A_k$ to accurately reproduce the observed behaviors can be reduced to a semidefinite program with the goal of minimizing \eqref{eq:optim_error}.
@@
  \eqref{eq:ds} with \eqref{eq:A}, and $x^*_{n}$ is defined as the endpoint of the $ n^{\text{th}} $ demonstrated trajectory.
-This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD), and the number of terms in the objective is roughly $O\left(k^2d^3\left( \sum_{n=1}^N T_n\right)\right)$.
+This optimization is always feasible, the only constraint is that the $A_k$s be independently PSD.
 
@@
 	\\  
-    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $\delta$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
+    & & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $p$}} & \multicolumn{1}{c}{\multirow{2}{*}{Optimal $K$}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Train}} & \multicolumn{1}{c}{\multirow{2}{*}{RMSE Test}}	\\ \\ \hline
 
-	\multicolumn{1}{c}{\multirow{3}{*}{(1) Forward Reaching}} & None & 7 & 5.5 $\pm$ (2.99) & \textbf{ 0.271 $\pm$ (0.038) } & \cellcolor{blue!5} 0.827 $\pm$ (0.697)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(1) Forward Reaching}} & None & 7 & 5 $\pm$ (2.90) & \textbf{ 	0.234 $\pm$ (0.027) } & \cellcolor{blue!5} 0.633 $\pm$ (0.458)  \\
     \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
@@
 Schemes.
-We present the \texttt{mean(std)} for the optimal $\delta$, $K$ and joint-velocity RMSE on training and testing set, found for every learning scheme over 10 runs.
+We present the \texttt{mean(std)} for the optimal $p$, $K$ and joint-velocity RMSE on training and testing set, found for every learning scheme over 10 runs.
 $M = \sum_{n=1}^{N}T_n$.\label{tab:learning}}.
@@
 
-\section{Experimental Validation} \label{Sec:Exp}
+\section{Experimental Validation} 
+\label{Sec:Exp}
 \subsection{JT-DS Learning Performance  Evaluation}
@@
 \item Analyze the role of dimensionality reduction in our proposed approach and further find the method which yields the best trade-off between accuracy and model complexity; i.e.
-the least number of dimensions $\delta$ to represent our activation function $\theta(q)$ and the least number of local behavior synergies $K$.
+the least number of dimensions $p$ to represent our activation function $\theta(\phi(q))$ and the least number of local behavior synergies $K$.
 \end{enumerate}
@@
 
-Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $\delta$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
+Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $p$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
 Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$.
-Thus, for each fold and each learning approach we find the optimal hyper-parameters $\delta$, $K$ and $\sigma$ (when applicable), as follows:
+Thus, for each fold and each learning approach we find the optimal hyper-parameters $p$, $K$ and $\sigma$ (when applicable), as follows:
 \begin{enumerate}[leftmargin=*]
-\item \textit{$\delta$ for PCA:} We choose $\delta$ such that the projection is capable of explaining $95\%$ of the variance in the data.
-\item \textit{$\delta$ and $\sigma$ for K-PCA:} The interaction of these two parameters plays a major roll in the resulting projection obtained from K-PCA.
+\item \textit{$p$ for PCA:} We choose $p$ such that the projection is capable of explaining $95\%$ of the variance in the data.
+\item \textit{$p$ and $\sigma$ for K-PCA:} The interaction of these two parameters plays a major roll in the resulting projection obtained from K-PCA.
 Hence, we do a grid search on a log-spaced range of $\sigma_{range} = [\sigma_{min}:\sigma_{max}]$ values, where
 $\sigma_{min} = \frac{1}{\kappa\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ and $\sigma_{max} = \frac{2\kappa}{\sqrt{2}} \underset{i \in M,j\in M}{\max}\{||q_i - q_j||^2\}$ for $\kappa = cte.$ This yields a feasible range for $\sigma$ that is guided by the pairwise Euclidean distances between all points in the dataset.
-Moreover, K-PCA is not limited to providing a $\delta \ll d$, in fact it can generate $\delta \le M$.
+Moreover, K-PCA is not limited to providing a $p \ll m$, in fact it can generate $p \le M$.
 This is a nuisance as in our datasets $M\approx1000$.
 To alleviate this we truncate $\sigma_{range}$ by computing the explained variance of the eigenvectors in feature space for different values of  $\sigma \in \sigma_{range}$.
-We then remove $\sigma$ values from $\sigma_{range}$ whose number of \textit{optimal} eigenvectors $\delta > d$ and resample it.
-By doing this procedure, we can ensure that for all values in truncated $\bar{\sigma}_{range}$ we will obtain $\delta \le d$.
+We then remove $\sigma$ values from $\sigma_{range}$ whose number of \textit{optimal} eigenvectors $p > m$ and resample it.
+By doing this procedure, we can ensure that for all values in truncated $\bar{\sigma}_{range}$ we will obtain $p \le m$.
 We then choose $\sigma_{opt}$ by running 10-fold cross-validation for all $\sigma \in \bar{\sigma}_{range}$.
@@
 Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the higher number of $K$ needed to represent the motion without DR.
-For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
+For all datasets, the DR methods provided $p < m/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
 By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
@@
 This demonstrates the JT-DS algorithm's ability to move cleanly in and out of singularities.
-%, overcoming the Achilles heel of equivalent Cartesian-based motion generators.
-
-\subsubsection{Systematic Assessment}
-To systematically determine the performance properties (tracking error, computation time, and convergence time) of JT-DS \eqref{eq:ds} and compare it to a SEDS+IK, we simulated $400$ simple motions on each system.
-The systems were tasked with moving to a fixed target randomly chosen from the region $\begin{bmatrix} -0.0081\pm0.3&-0.0188\pm0.3&0.4974\pm0.21 \end{bmatrix}^T$.
-The Cartesian motion generator was SEDS-based, and mapping from Cartesian motions to joint-space motions was done using a damped least-squares IK solver.
-To save on computation time, both the JT-DS and SEDS algorithms were taught behaviors defined by a single synergy, i.e.
-uniform behaviors.
-The results, summarized in Table \ref{tab:sys}, indicate that the computation time of the proposed approach is significantly faster than the Cartesian-based DS, because the algorithm does not require calculating the Jacobian pseudo-inverse.
-As the generated joint motion by \eqref{eq:ds} is directly transmitted to the robot, the tracking error is zero.
-Nevertheless, the convergence time of  \eqref{eq:ds} is slightly higher than SEDS+IK.
-
-\begin{table}[t]
-	\label{table:2}
-	\centering
-	\resizebox{\linewidth}{!}{\begin{tabular}{l|l|l|l}
-		\hline\hline
-		& Norm. converg. [{s}/{m}] & Comp. time {[}ms{]} & Track. error {[}m{]} \\ \hline
-		SEDS  + IK    & $10.6566\pm2.1364$                          & $71.7752\pm4.9189$       & $0.0163\pm0.0063$      \\ 
-		JT-DS & $14.1926\pm5.26453$                         & $12.1736\pm0.8573$       & $0.0\pm0.0$            \\ \hline\hline
-	\end{tabular}}
-		\caption{Each simulated trajectory is initialized at $q=[0~\dots~0]^T$.
-The convergence duration is the time required to move within $0.001$m of the target.
-The normalized convergence duration is the convergence duration divided by the distance between the initial and target positions.\label{tab:sys}}
-		\vspace{-15pt}
-\end{table}
 
@@
 One of the most important points when validating a learning from demonstration method is to evaluate the system's behavior away from demonstrations.
- When the current joint configuration is far from any of the local synergy regions, computing the scheduling parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\frac{1}{K}\sum\limits_{k=1}^{K}A_k$ is used to move the robot, which is still guaranteed to move towards the target.
+ When the current joint configuration is far from any of the local synergy regions, computing the activation parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\frac{1}{K}\sum\limits_{k=1}^{K}A_k$ is used to move the robot, which is still guaranteed to move towards the target.
 %Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably converging to the target.

### 2018-05-15 paper/reply_reviews/Reply_to_reviews.tex
Final update, folder restructure, final proof of published paper

\documentclass{article}
\usepackage[left=2cm, right=2cm, top=2cm]{geometry}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{color}
\usepackage{graphicx}
\usepackage{subcaption}
\title{Response to Reviewers: Manuscript ID 17-1142.1: ``Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A Linear Parameter Varying and Synergetic Control Approach"}

%\author{Your Name  \\
%	Your Company / University  \\
%	\and 
%	The Other Dude \\
%	His Company / University \\
%	}
%
%\date{\today}
% Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
% before or after the \begin{document} command 
% BUT the \maketitle command MUST come AFTER the \begin{document} command! 
\begin{document}
\maketitle
\small
First of all, we would like to thank all reviewers for the time they spent reviewing our paper. We greatly appreciate the comments, suggestions and opinions that we received. We hope to have addressed all their comments and suggestions satisfactorily.\\

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
\noindent \underline{Comments from Reviewer \#1 (Reviewer: 4 (Review30069)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``As mentioned by the authors, there are many reasons to plan in joint space rather than task space. One of the important reasons to me seems to be able to plan end-effector orientation, not just position. Traditional movement primitives, for example DMPs, do not explicitly encode planning in the SO(3) space of orientations. There is research that addresses this question, for example [1], but they lose the simplistic appeal of movement primitives. I think it would be a great addition to this paper if the authors illustrate if their joint space DS can be used to accomplish orientation sensitive tasks, and generalize to end-effector orientations different that what was demonstrated.[1] Ude, Ale, et al. "Orientation in cartesian space dynamic movement primitives." Robotics and Automation (ICRA), 2014 IEEE International Conference on.}}"\\
\texttt{Response:}  We highly appreciate this comment. When we initially proposed this approach, we did not think about tackling the problem of reaching a target in orientation as well. Since our DS is learned in joint-space, indeed, we do not have to deal with the challenges that arise when trying to reach a target in both position and orientation as we do not have the constrains imposed by SO(3). There was theoretically no modification needed in our controller:
\begin{equation}
\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{m\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
\label{eq:jtds}
\end{equation}
Originally, when working solely in position $d=3$, now we represent the target with $d=9$, where the first 3 dimensions correspond to Cartesian position and the remaining 6 correspond to the first and second columns of a rotation matrix $R \in SO(3)$. This only modifies the Jacobian $J(q)$ which is now $J(q)\in \mathbb{R}^{9 \times 7}$ rather than $J(q)\in \mathbb{R}^{3 \times 7}$. The role of $\mathcal{A}(q)$ remains  the same. The following paragraph has been incorporated, as well as some additions in the abstract and problem statement:
\textcolor{blue}{\small ``Finally, by learning a behavior in joint-space we can inherently reach a task-space target, not only in the Cartesian 3-D space $\mathbb{R}^3$, but also in the space of orientations $SO(3)$. Planning or learning the rotational component of motion is a challenging problem. Representing an orientation as a vector in Euclidean space may lead to inaccurate and unstable motions, due to its directional nature and vulnerability to singularities. Several works have proposed tailor-made learning approaches that consider the non-Euclidean geometry of the $SO(3)$ space to generate rotational motion [18][19][20]. These approaches, however, require an explicit coupling between position and orientation, that might cause discontinuities in the resulting end-effector motion. Such coupling is not necessary if the motion is encoded in joint-space."} With the included references being:\\
$[18]$ S. Kim, R. Haschke, and H. Ritter, “Gaussian mixture model for 3-dof orientations,” \textit{Robotics and Autonomous Systems}, vol. 87, pp. 28-37, 2017.\\
$[19]$ A. Ude, B. Nemec, T. Petri, and J. Morimoto, “Orientation in cartesian space dynamic movement primitives,” in \textit{2014 IEEE International Conference on Robotics and Automation (ICRA)}, May 2014, pp. 2997-3004.\\
$[20]$ M. J. A. Zeestraten, I. Havoutis, J. Silvrio, S. Calinon, and D. G. Caldwell, “An approach for imitation learning on riemannian manifolds,” \textit{IEEE Robotics and Automation Letters}, vol. 2, no. 3, pp. 1240-1247, July 2017.

Furthermore, we have re-learnt all demonstrated behaviors incorporating orientation in the task-space target $x^*$. Table I has been updated accordingly, showing no deterioration in generalization and performance. Finally, as will be discussed in the response to the Comments 10 and 12, we compare our approach to a Task-Space DS learnt in both position+orientation, showing a considerable improvement in accuracy while reaching the target.

\item \textcolor{blue}{\textit{``In the introduction, the authors stress that jacobian psuedo inverse
can be computationally expensive. Is this really true for a 7-dof robot
arm? Can the authors provide some numbers where this would really be an
issue for a robot controlled at 1000Hz? I find the argument that
jacobian inverse does not work for singularities and needs a set of
heuristics to be more compelling. Maybe the authors should rewrite the
introduction to stress more on this issue?"}}\\
\texttt{Response:} \small For a 7-DOF robot it is indeed negligible. Computational burden only becomes a real issue for higher degrees of freedom, like a multi-fingered hand or multi-arm robotic setting. We removed this remark in order to allow space to discuss about other issues. \\

\item \textcolor{blue}{\textit{``In Section II, criterion I, the authors talk about a positive symmetric matrix P. I do not see it in the
equations above. Is this a weighing matrix?"}}\\
\texttt{Response:} \small Removed, that was a carry-over from a previous formulation.\\

\item \textcolor{blue}{\textit{``In criterion II, the error is defined with respect to the velocity error, rather than a position or a combination
of both. This means that the motion can be different from the demonstrated motion. I understand
that this is important to keep the later optimization convex, but is there any disadvantage to using
velocity error rather than position error?"}}\\
\texttt{Response:} \small Optimizing for position distance alone would have the effect of learning a motion that converged to the average of previous motions, thus not changing the motion in new situations or fitting the joint velocity profile that is often a desired component of the motion. Using velocity error does mean that, for motions strongly dependent on joint position rather than joint velocity, the generated motions may generalize from the demonstrations in unexpected ways. However, as you mentioned, minimizing position error would, depending on the formulation, be prohibitively difficult. This would imply simulating the learned motion forward in time to recompute the robot's positions, making the optimization much more complicated; this way, we can treat each point in the motion independently.\\

\item \textcolor{blue}{\textit{`I do not completely understand the significance of A(q). From Figure 3, it seems like A(q) helps resolve
redundancy, while achieving the task space goal. Do the authors have a mathematical (in constrast to
intuitive) explanation for this? Does A(q) somehow relate to the null-space of the jacobian?"}}\\
\texttt{Response:} \small A(q) transforms a joint velocity vector approximately aligned with the task-space error into a different joint velocity vector that remains approximately aligned with the error, but which we want to choose to more-closely approximate our demonstrations. In this sense, it is resolving redundancy between all the possible joint velocity vectors that would move the controller closer to the origin. The paragraph in Section III has been modified to improve clarity. A(q) doesn't relate to the null-space of the Jacobian in any clear way, as its value can and does alter the task-space trajectory as well as the joint space trajectory. The only property that A(q) preserves is the joint velocity's moving the task space position closer to the goal, but the direction of motion is not necessarily the same as the direction of task-space error.\\

\item \textcolor{blue}{\textit{``In Section III, ?One can intuitively understand the control law as follows:$(H(q) - x^*)$ denotes the
position error w.r.t. the task target and by multiplying that error by the transposed Jacobian JT (q),
the error is projected into joint space (similar to Jacobian transpose control)?. I find this explanation
misleading as jacobian multiplied to position error does not project the error into joint space. Jacobian
transforms velocities, not positions. Maybe the law can be seen more as a first order approximation
of the error in task space. In any case, I would either clarify this statement more or take it out of the
manuscript all together. It seems like the DS is just a PD law that provably converges to a fixed attractor
in task space."}}\\
\texttt{Response:} \small The passage has been rewritten to remove the word "project", which you appropriately noted was incorrect, and further clarifies the nature of the DS as a PD controller. We have added the following text in the manuscript: 
\textcolor{blue}{\textit{``One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t. the task target, which we then reinterpret as the task-space velocity of a proportional controller. By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control [23][27]), see Fig. 2. The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig. 3 illustrates the effects of  $\mathcal{A}(q)$ on the generated motion. Thus the controller can be thought of as a proportional controller in joint space.}}\\

\item \textcolor{blue}{\textit{``About figure 2 again, the position error in task and joint space are not simply multiplied by a Jacobian.
Although, task space error can be first-order approximated using a Jacobian multiplication with
joint space error. What does the right figure in Fig 2 exactly represent?"}}\\
\texttt{Response:} \small The figure on the left is simply the cartesian error in task space. The figure on the left is the Lyapunov potential function on which the jacobian transpose controller descends. In showing the attractive regions in joint-space corresponding to this Lyapunov function, we aim to offer insight into how the controller reformulates the task-space potential into a different yet related joint-space potential. The caption has been rewritten to add clarity.\\

\item \textcolor{blue}{\textit{``In Section IV, the authors say that they use EM to determine the parameters of the GMM in Eq 6. Can
they elaborate a little on this? What is the expectation taken over and what likelihood is maximized?"}}\\
\texttt{Response:} \small The algorithm computes the expected probability of generating each of the joint positions given the current values of the GMM's means, variance matrices, and scaling values. The algorithm then chooses GMM parameters to maximize the posterior likelihood of the model, using previously-computed expected component membership as likelihood weights when computing the new mixture. Specifically, we used MATLAB's \textit{fitgmdist} implementation. Added a citation to clarify.\\

\item \textcolor{blue}{\textit{``The learned dimensionality reduction and parameters seem task dependant. I wonder if the authors
have any idea on how to generalize this across tasks."}}\\
\texttt{Response:} \small The dimensionality reduction could be generalized across tasks that had the same joint postural synergies (i.e. same lower dimensional space) - that is, for example, all tasks in which the shoulder and elbow always move together. Similarly, one could imagine several similar motions that could be decomposed into shared sub-motions, thus allowing the parameters to be shared. However, in general, fixing some of the parameters while optimizing will simply lead to a less-flexible model, and considering that there is no intuitive scheme for \textit{a-priori} identifying motions with strongly similar sub-motions, we did not explore this question further. Given that the local behavior regions are defined by the GMM components, one could use the likelihood (or KL divergence) to determine the similarity of the motions when face with a new task. To answer your question directly, if one wanted to, one could reuse the learned dimensionality reduction for a new task which you believed to be similar, or copy certain components of a previous motion's GMM into the new GMM and leave them fixed throughout the optimization.

\item \textcolor{blue}{\textit{``The experiments have the same task-space goal as the
demonstrations. Can the authors include experiments that have goals
different from demonstrated, especially in orientation?"}}\\
\texttt{Response:} \small Yes. As we have shown in Appendix I, regardless of the target location (position/orientation), the motion generated by the proposed dynamical system asymptotically converges to the desired target. However, it is important to note that as (8) in Appendix I is semi-negative-definite and not negative-definite, there might be joint configuration $ (q )$ which result in $ J(q)\sum\limits_{k=1}^{K}\theta_k(q)A_kJ^T(q)=0 $. In this situation, the robot is stuck in a deadlock and, consequently, cannot reach the desired target.

\item \textcolor{blue}{\textit{``In similar light, maybe authors could present online adaptation of
A(q) to generalize to situations different than demonstrated, for
example a slightly moved obstacle?"}}\\
\texttt{Response:} \small The reviewer has pointed to a very interesting topic and the authors appreciate the reviewer’s suggestion. In this paper, we assumed that the learned model (namely $ \pi $'s, $ \mu $'s, $ \Sigma $'s and $A$'s) does not need to be updated or modified during the task execution. Hence, this would be out of the scope of this paper and we are planning to address it in future works.

\item \textcolor{blue}{\textit{``In table II, instead of the tracking error, maybe the more useful
error to see would be the error with respect to the plan or
demonstartion. One would assume that the proposed approach would be
worse at it than SEDS+IK because the minimized error is with respect to
the velocity and goal, not the position of the movement."}}\\
\texttt{Response:} \small This is, in general, not true. Based on the robot's motion and its configuration, there might be cases where the motion generated by SEDS does not even reach the target. To clarify this, we have systematically studied the performance of the motion generated by JT-DS and SEDS+IK in terms of (i) reaching to the desired target and (ii) following the demonstrated end-effector motions, See Figure \ref{fig:SEDSvsJTDS}. We used the data-sets from the seven tasks mentioned in Section V.A to construct SEDS and JT-DS. Both methods are tested by the same the target and the initial configurations, which were extracted from the demonstrations. We define $ x^*=[x^*_{ori};x^*_{pos}] $ as the target configuration that includes the desired position as well as the desired  orientation of the end-effector. The motion generated by SEDS is converted into the 7-DOF joints state using the damped least square IK solver.\\ 
In Figure \ref{fig:SEDSvsJTDSP:E} and Figure \ref{fig:SEDSvsJTDSO:E}, the error between the final location of the robot and the target is illustrated. As it can be seen, the motion generated by the proposed dynamical system reaches the target, whereas, the motion generated by SEDS does not; regardless of the fact that both models are theoretically stable. The main cause of failure for SEDS is due to the IK solver. As SEDS generates the motion of the robot without considering its joint configurations, the robot might reach its joint limits.  Whereas, as JT-Ds directly shapes the joints' motion, by learning the desired joint-space behavior, one can avoid these situations.

In Figure \ref{fig:SEDSvsJTDSP} and Figure \ref{fig:SEDSvsJTDSO}, the error between the demonstrated and the excused motions are illustrated. As it can be seen, even though JT-DS is not tailored for following the end-effector motions, its performance is comparably better than SEDS.

Although including this study would highlight one of the advantages of using JT-DS over Cartesian-motion-generators, as this does not highlight the main focus of this paper and due to the space limitations, we have decided to not include this in the paper.

\begin{figure}[!t]
 \hspace*{-2.4cm}
	\begin{subfigure}[t]{0.63\linewidth}
	\centering
		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P_End.pdf}
		\caption{}
		\label{fig:SEDSvsJTDSP:E}
	\end{subfigure}%
	\hspace{-0.2cm}
	\begin{subfigure}[t]{0.63\linewidth}
	\centering
		\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O_End.pdf}
		\caption{ }
		\label{fig:SEDSvsJTDSO:E}
	\end{subfigure}\\		 
	\hspace*{-2.4cm}\begin{subfigure}[t]{0.63\linewidth}
			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_P.pdf}
		\caption{ }
			\label{fig:SEDSvsJTDSP}
		\end{subfigure}%
		\begin{subfigure}[t]{0.63\linewidth}
			\includegraphics[trim={2cm 0.0cm 3.25cm 0cm},clip,width=\textwidth]{Pics/SEDS_JT_O.pdf}
			\caption{ }	
			\label{fig:SEDSvsJTDSO}
		\end{subfigure}\\
	\caption{The performance comparison between SEDS+IK and JT-DS while the initial and the target points are same. The error between the orientations is the difference between the first two columns of the corresponding rotation matrix representation.}
	\label{fig:SEDSvsJTDS}
\end{figure}

\end{enumerate}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\newpage 
\underline{Comments from Reviewer \#2 (Reviewer: 5 (Review30371)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``My first concern regards you choice of representing the demonstrated
movement as a sequence of postures and hence looking for synergies by
using PCA of K-PCA analysis. Why not using functional PCA (f-PCA)
analysis and determine the few moving synergies that make possible the
execution of the demonstrated behaviours? It would be that, when you
combine the local synergies in space and time by using functions
$\theta_k(q)$, you get what you might obtain by using directly f-PCA.
However, the last one are valid for the whole movement and can be
combined by constant coefficients."}}\\
\texttt{Response:}  See below.

\item \textcolor{blue}{\textit{``I also found the paper lacking of physical intuitions, especially in	the experimental part, about the entities introduced in previous sections. In particular, what is the lesson we learn after your analysis of synergies and the obtained lower dimensional space for the few movements shown in section V? Is it possible to extract some global	synergic behavior? Moreover, in section IV.A authors introduce the function phi(q) named the embedding that maps the joint configuration into a lower dimensional space. Is it associated to a whole behaviour or, as matrices $A_k$, has a local validity? I lost this point. Can you show/report or describe what it represents in the joint space for one or more cases in Section V? What is this lower dimensional space for the cases in Section V?"}}\\
\texttt{Response:}  See below.

\item \textcolor{blue}{\textit{``A similar question for the matrices $A_k$ that represent local, I
	would say postural (see previous comment), synergies whose dimension is
	m x m, i.e. in joint space: I would expect they were defined in the
	lower dimensional space. Moreover, are they local representation of the
	distribution of postures and hence covariance matrices? What do their
	eigenvectors and eigenvalues locally represent? Maybe, a comparison
	with the classical and more known postural synergies for the human
	hand, analyzed by PCA, could be useful. Finally, what does it mean that
	they should be positive definite (apart from stability issue)?"}}\\
\texttt{Response:}  See below.\\

\item \textcolor{blue}{\textit{``What is the dimension of function $\theta_k(q)$ that, to some extent, can be seen as the new control variables. Is it $K<m$?"}}\\

\texttt{Response:} In response to Comments 1-4, following we mathematically and graphically analyze the proposed local behavior synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via. PCA or K-PCA) and grouping of local behavior regions with GMM on this lower dimensional space. The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$. Let's begin by analyzing the proposed DS-based control-law \eqref{eq:jtds}. The intuition behind this control-law is that $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\phi(q))A_k$ is a linear combination of time-invariant linear matrices $A_k \in \mathbb{R}^{m \times m}$, \underline{these matrices are the ``local behavior synergy matrices"} that shape the motion in joint-space. Meaning that, given the joint-space velocity vector representing the task-space error $J^{T}(q)(H(q)-x^*)$, the resulting motion is biased to use a particular set of joints, defined in each $A_k$. Each $A_k$ is thus activated depending on the current ``joint-posture" $q$ (represented in the lower-dimensional space $\phi(q)$) via the scheduling/activation function $\theta_k(\phi(q))$.
$\phi(q)$ is a learned transformation, which is associated with a whole behavior, rather than a particular $A_k$ (as is now clarified in IV.A). However, the precise nature of the transformation depends on the DR technique: PCA applies a uniform matrix multiplication of joint position everywhere, whereas KPCA is a more complicated function of joint position.  The result of the embedding is a lower-dimensional representation of the joint position. The following line has been added to IV.A for clarity:\\
\textcolor{blue}{\small \textit{"For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the "shoulder" and "arm" components of $q$ into a single "shoulder-arm" component in $\phi(q)$.}"}

\textcolor{blue}{\small Also, Section III has been considerably re-written to clarify our proposed approach and proposed definition of ``local behavior synergies".}

Continuing with the analysis, given the raw demonstrations of a joint-space  behavior with a task-space target as shown in Fig. \ref{fig:raw_demos} (for the pouring task) we seek to learn a linear combination of local behavior synergies $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\phi(q))A_k$ such that we can use \eqref{eq:jtds} to accurately reproduce these demonstrations. In order to do so, we project the raw joint trajectories to a lower-dimensional space via $\phi(q)$, in the case of PCA $\phi(q)=M_p \times q$ for $M_p \in \mathbb{R}^{p \times m}$. By choosing the first 3 PCs we can explain 95\% of the variance in the joint posture trajectories, which can be represented in this 3-D space as shown in Fig. \ref{fig:raw_demos}. We then jointly learn the activation function $\theta_k(\cdot)$ and the number of $K$ local synergy regions by fitting a GMM to this representation of the joint posture trajectories. As can be seen in Fig. \ref{fig:raw_demos}, the entire joint-space motion can be represented by 3 local synergy regions in PCA-space. Meaning that we only need 3 synergy matrices $A_k$ to represent the complex joint-space motion accurately. The activation function $\theta_k(\cdot)$, as described in the mansucript, is the posterior probability of a joint posture in PCA-space $\phi(q)$ belonging to one of the local synergy regions $k \in \{1,\dots,K\}$. As mentioned in Sec. V.A., we treat K as a hyperparameter and thus choose it differently for each learned motion. For motions that do not vary much throughout the course of the motion, smaller K might be sufficient, but K can just as easily be greater than $m$ (though as $K$ increases, so does the learning problem's computational cost). If comment-4 was about each specific $\theta_k(q)$, they are scalars, as is mentioned in Sec. III.

\begin{figure*}[!th] 
  \begin{minipage}{0.58\textwidth}
     	\centering 
     	\includegraphics[trim={1.2cm 2cm 1.7cm 2cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/raw_demos_pour.pdf}
  \end{minipage}
   \begin{minipage}{0.47\textwidth}
      	\centering
      	\includegraphics[trim={1.2cm 0.5cm 0.5cm 0.35cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/raw_demos_pca_pour.pdf}
      	\vspace{-5pt}

		\includegraphics[trim={0.8cm 0.5cm 0cm 0.35cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/local_synergy_regions_pour_annotated.pdf}
   \end{minipage}
   \caption{Raw demonstrations of the pouring task in original space, PCA space and the extracted local synergy regions. \label{fig:raw_demos}}
\end{figure*}

\underline{As mentioned in the reviewer's 3rd comment}, since we have an accurate representation of our joint posture trajectories and the local synergy regions in this lower-dimensional space, why not learn the $A_k$ matrices in this space; i.e. $\mathcal{A}(q)\in \mathbb{R}^{p\times p}$ rather than $\mathcal{A}(q)\in \mathbb{R}^{m\times m}$ ? This seems like the obvious method, if we were considering our synergistic approach the classical way, i.e. using the lower-dimensional space as the new control variables. However, if we were to do that, this would mean that \eqref{eq:jtds} would become:
\begin{equation}
\underbrace{\dot{q}}_{\mathbb{R}^{m\times 1}} = -\underbrace{M_p^{-1}}_{\mathbb{R}^{m\times p}}\underbrace{\mathcal{A}(q)}_{\mathbb{R}^{p\times p}}\underbrace{M_p}_{\mathbb{R}^{p\times m}}\underbrace{J^{T}(q)}_{\mathbb{R}^{m\times d}}\underbrace{(H(q)-x^*)}_{\mathbb{R}^{d\times 1}}
\label{eq:jtds_low}
\end{equation}
This is not entirely incorrect, however, with such a control law we can no longer guarantee stability and convergence to the attractor $x^{*}$. By considering the Lyapunov function used in Appendix I to prove stability of \eqref{eq:jtds_low} i.e. $V(q) = \frac{1}{2}(H(q) - x^*)^T(H(q) - x^*)$, the derivative of $ V $ wrt. would be:
\begin{equation}
\begin{aligned}
\frac{dV(q)}{dt} &= (H(q) - x^*)^TJ(q)\dot{q}\\
&= -(H(q) - x^*)^TJ(q)M_p^{-1}\mathcal{A}(q)M_pJ^T(q)(H(q) - x^*)\\
&= -(H(q) - x^*)^TJ(q)\textcolor{red}{\underbrace{M_p^{-1}\big(\sum_{k=1}^{K}\underbrace{\theta_k(q)}_{> 0}\underbrace{A_k}_{\succ 0}\big)M_p}_{\text{This should be}~~ \succ 0}}J^T(q)(H(q) - x^*)\leq 0
\end{aligned}
\end{equation}
In order for \eqref{eq:jtds_low} to be asymptotically stable wrt. the attractor $x^*$ the term $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p \in \mathbb{R}^{m\times m}$ should undoubtedly be $\succ 0$. Even if $\theta_k(q) > 0$ and $A_k \succ 0$ for all $k$, the only way for this condition to hold is if $M_p$ where a full symmetric positive definite matrix. However, this is not the case because $M_p\in \mathbb{R}^{p\times m}$  is a projection matrix, i.e. $p<m$. The term $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$ will no longer be full-rank, it will preserve the $p$ positive eigenvalues from the $A_k$ matrices, but the remaining eigenvalues will be 0, thus generating a semi-positive definite matrix, which invalidates our stability proof. Note that, as described in the Appendix, $J(q)$ is not proven to be full rank in all regions of the workspace, when this happens, this means that the end-effector pose is not manipulable. Hence, we can only prove asymptotic stability when the end-effector pose is manipulable. This is a physical artifact from the joint-space constraints of the robot arm which we cannot avoid. Yet, since the demonstrations are acquired via kinesthetic teaching we can assume that they are manipulable. \\

\underline{As asked by the reviewer, what does it physically mean to have all positive eigenvalues?}  When $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$  or ($\mathcal{A}(q)$) has some eigenvalues $=0$ the physical meaning behind it is that we are forfeiting the control of these directions of motion in joint-space. Namely, the desired velocity directions in joint-space will be constrained in the corresponding DOF of the arm, this will prohibit us from converging to the attractor and will be no longer capable of accurately following the joint-space demonstrations.  If instead we learn the synergy matrices in joint-space following \eqref{eq:jtds} and modulate them with the activation function in PCA-space, we avoid all of this problems and can ensure full controllability of the motion in joint-space.

\noindent \textbf{Remark 1:} This analysis is only considering PCA as the DR approach. If we would consider K-PCA, deriving such a control law is more involved as the inverse mapping $\phi(q)^{-1}$ is not straightforward if at all possible, thus limiting the approach to linear DR techniques. By considering $\phi(q)$ solely in the activation function $\theta(\cdot)$ we do not have this problem and can use any DR technique, as long as we're able to compute out-of-sample projections.

\noindent \textbf{Remark 2:} If we were to bypass the DR step our state-dependent system matrix would become $\mathcal{A}(q)= \sum\limits_{k=1}^{K}\theta(q)A_k$, which (via our definition) is also a combination of ``local behavior synergy matrices" with the activation function living in joint-space rather than the PCA space; better-known as the \textit{joint postural synergy space}. The fact that we use PCA/K-PCA to represent the activation function $\theta(\cdot)$ is to alleviate the complexity of extracting the local synergy regions in joint-space directly. It is, indeed, also motivated by the fact that we know that multi-DOF joint postures can be accurately represented in a lower-dimensional space composed of the principal components of the postural trajectories; i.e. the well-known \textit{joint postural synergies}. As shown in TABLE I of the manuscript and discussed in Section V, by using a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations. This is explicitly highlighted in Section V:\\
\textcolor{blue}{\small \textit{"As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches. This suggests that using DR to encode our activation functions $\theta_k$ in a lower-dimensional space $\phi(q)$ yields better generalization capabilities than encoding the behaviors in using the original $q$. This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learned without DR is an order of magnitude higher than with DR. Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the higher number of $K$ needed to represent the motion without DR. For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR. By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations. Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations."}} \\

Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations and learn a \textit{time-invariant} controller that ensures reaching the task-space target while approximately following the demonstrated motion, as shown in Fig. \ref{fig:jtds_velocities}, where we plot the joint positions/velocities of 1 demonstrations of the pouring task and the joint-positions/velocities from the JT-DS controller. In these plots, we are taking the velocity generated directly by JT-DS. The joint-positions are then computed by numerical integration of these velocities. Note that our controller does not generate this entire trajectory in one-shot, it is generating the next desired velocity given the current joint posture, e.g. $\dot{q}_{t+1} = \mathcal{A}(q_{t})J^T(q_{t})(H(q_{t})-x^*)$ with $q_t$ being the current joint configuration. Moreover, our controller is state-dependent only; i.e. it has no dependence on time or the duration of the desired trajectories. Hence, it can generate motions similar to the demonstrated ones at different frequencies and time durations. \\

\noindent \textcolor{blue}{We have added clarifications in the text such that no other reader gets confused with the new concept of behavior synergies that we propose as opposed to the classical joint posture synergies.}

\begin{figure*}[!th] 
  \begin{minipage}{0.5\textwidth}
     	\centering 
     	\includegraphics[trim={1.2cm 1.5cm 1.7cm 1.5cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/jtd_pos_pour1.pdf}
  \end{minipage}
    \begin{minipage}{0.5\textwidth}
       	\centering 
       	\includegraphics[trim={1.2cm 1.5cm 1.7cm 1.5cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/jtd_vel_pour1.pdf}
    \end{minipage}
   \caption{Training and executed joint-position/velocities with JT-DS for 1 pouring demonstrations. \label{fig:jtds_velocities}}
\end{figure*}

\noindent \underline{On the applicability of F-PCA to our learning scheme:}  If we were to ignore our proposed control scheme and try to reconstruct the trajectories from the joint postural synergy components; i.e. using PCA, we would get the trajectories shown in Fig. \ref{fig:pca_velocities}. In these plots, we are taking one of the collected demonstrations projected in the PCA-space, i.e. $\phi(\{q_1,q_2,\dots,q_T\})$ and reconstructing it to generate the desired joint posture trajectories. To compute the velocities we apply numerical differentiation on the reconstructed joint positions. As can be seen, we are capable of generating similar trajectories, both in position and velocity. In this simple example we are saving the entire demonstration in PCA-space, however, if we were to use this for controlling the real robot we need to learn a model or a controller in this low-dimensional space and then map it back to joint-space. As discussed earlier, this is not feasible with our proposed control-law \eqref{eq:jtds}, due to the stability and convergence objectives we defined.

\begin{figure*}[!th] 
  \begin{minipage}{0.485\textwidth}
     	\centering 
     	\includegraphics[trim={1.2cm 1.5cm 1.7cm 1.5cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/pca_pos_pour1.pdf}
  \end{minipage}
    \begin{minipage}{0.5\textwidth}
       	\centering 
       	\includegraphics[trim={1.2cm 1.6cm 1.7cm 1.6cm},clip,width=\linewidth]{../../src/JTDS_mat_lib/figures/pca_vel_pour1.pdf}
    \end{minipage}
   \caption{Training and reconstructed joint-position/velocities with PCA for 1 pouring demonstrations. \label{fig:pca_velocities}}
\end{figure*}

Ignoring this limitation, one of the main draw-backs of this type of synergistic control is that the extracted representation in PCA space does not take into account the dynamic features of the motions and further assumes that the demonstrations have no time correlation. As mentioned by the reviewer and applied in \cite{Dai:IROS:2013} for modeling grasping motions and in \cite{Averta:FRAI:2017} for upper-limb motions, functional-PCA is a nice alternative to include these dynamic features and further reconstruct entire trajectories with a linear combination of a few \textit{time-dependent} functions. In this approach each demonstrated trajectory (of a single DOF) is considered as a function composed of a linear combination of $N$ basis elements (which can ve exponentials, splines, Fourier) $f_i = \sum\limits_{n=1}^{N}\theta_nb_N$. Thus, each $i$-th function (i.e. $i$-th demonstration) can be described by the vector a coefficients $\Theta = \{\theta_1,\dots,\theta_N\}$. PCA is then performed on $\Theta = \{\theta_1,\dots,\theta_N\}$ and the extracted PC are transformed to \textit{functional PCs} via the $N$ basis functions. Then $M<N$ \textit{functional PCs} are used to reconstruct the $i$-th function (i.e. $i$-th demonstration ) as $\hat{f}_i = \bar{f_i} + \sum\limits_{i=1}^{M}c_i\xi_i$, where $\bar{f}_i$ is the mean function, $c_i$ are the coefficients and $\xi_i$ are the \textit{functional PCs}. As shown in \cite{Averta:FRAI:2017} with only 1 \textit{functional PCs} 60/70\% of the demonstrations can be accurately reconstructed. This approach would be interesting if our goal were to reconstruct entire time-dependent trajectories as they do in  \cite{Dai:IROS:2013} and \cite{Averta:FRAI:2017}. This is, however, NOT our goal and applying this method to our control scheme seems not suitable and is rather an alternative approach for the following reasons: First of all, in this work we propose a DS-based controller, meaning that, we do not generate the full reference trajectory of the robot \textit{a prior} and then try to track it, we instead compute the next desired velocity on-the-fly which happens to result in the desired motion because it is encoded in $\mathcal{A}(q)$. Second, if we were to apply this method as the reviewer suggests, the only way to do it is to re-write a controller that uses 7 reconstructed functions $\hat{f}_i$ (i.e. one per joint) to track a time-indexed reference joint trajectory. This is a valid approach, however, in this work we propose a state-dependent; i.e. time-invariant approach. A time-invariant approach is advantage over a time-dependent approach because it is inherently robust to  perturbations,  i.e.  it can ensure that a  motion may be quickly adapted to perturbation and changes in a dynamic environment. Perturbations can be caused by the physical inability of a robot to perform a pre-planned trajectory or due to the inherent uncertainties of the environment. Thus, such robustness is particularly important when it comes to couple the different joints in a synergy, for examples and more detailed discussion of this aspect refer to \cite{EPFL-ARTICLE-148817}. Finally, our goal is not to accurately reproduce the demonstrations but rather to reach the target $x^*$ with a similar joint motion as the one that was demonstrated.

\item \textcolor{blue}{\textit{``For the test, why not comparing also with the JT control method? In
other terms, is it possible to use a SEDS + JT, especially for the
third case "Transiting through singular configurations"?"}}\\
\texttt{Response:} \small This could be done, but it is mathematically incorrect. The original JT control method is a DS that converges to a unique equilibrium point, i.e. the task-space target $x^*\in \mathbb{R}^{N\times 1}$, with a gain matrix $K \in \mathbb{R}^{M\times M}$ as:\begin{equation}
\dot{q} = - KJ(q)^T(H(q)- x^*) 
\label{eq:JT}
\end{equation}
SEDS, is a DS-based motion generator that converges also to a unique equilibrium point, as:
\begin{equation}
\hfill \begin{aligned}
&  \dot{x}  = \mathbf{f}(x)
\rightarrow  &
\begin{cases}
\lim_{t \rightarrow \infty}\|x - x^*\|=0
\end{cases}
\end{aligned}  \hfill
\label{eq:simple-ds}
\end{equation}
where $\mathbf{f}(\cdot): \mathbb{R}^M \rightarrow \mathbb{R}^M$ is a continuous differentiable vector-valued function representing a non-linear DS that converges to a single stable equilibrium point $x^*$. In order to prove the stability of SEDS, one needs to assume that the dynamic of the robot is perfectly compensated by the built-in position-level controller. In other words, the transfer function of the dynamic of the robot+ the low-level controller and the IK solver must be one.

Almost all the commercially available robots are equipped with a very precise joint position controller. Hence, the transfer function of the robot+ the low-level controller can be assumed to be one. Whereas, the transfer function of the Jacobian transpose IK solver can not be assumed one as the JT control method assumes that the target is static. Hence, in order for the JT controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time. This is contradicting as \eqref{eq:simple-ds} is always trying to converge to $x^*$. Thus, merging these two approaches is theoretically ill-fitting. Note that,  To better highlight this, we show an illustration of the supposed SEDS+JT and JT-DS control loops in Figure \ref{fig:schematic:SEDS} and  Figure \ref{fig:schematic:jt}, respectively.\\

\begin{figure}[t]
\centering
	\begin{subfigure}[t]{0.8\textwidth}
	\centering
		\includegraphics[trim={0 0 0.25cm 1.1cm},clip,width=\textwidth]{Pics/Drawing2.png}
		\caption{ A Cartesian motion generator (e.g., SEDS) with the Jacobian transpose IK solver.}
		\label{fig:schematic:SEDS}
	\end{subfigure}\\
	\begin{subfigure}[t]{0.8\textwidth}
	\centering
		\includegraphics[trim={0.25cm 0 0.0cm 1.2cm},clip,width=\textwidth]{Pics/JT_DS.png}
		\caption{ JT-DS.}
		\label{fig:schematic:jt}
	\end{subfigure}\\
	\caption{A simplified schematic comparing the standard control architecture using Cartesian dynamical systems with the Jacobian transpose IK solver (top) and our propose approach. By integrating the "inverse kinematic (IK) solver" and "motion generator" blocks into a unified block, the proposed approach eliminates undesirable effects of the IK solver. }
\end{figure}

\item \textcolor{blue}{\textit{``Minors:
- SEDS + IK: what does SEDS mean?"}}\\
\texttt{Response:} As stated in the previous comment, SEDS, short for Stable Estimator of Dynamical Systems is an approach proposed in \cite{khansari2011learning} that learns a stable estimate of a DS with the form of \eqref{eq:simple-ds} from demonstrations. The first paragraph of Section V.B was written to describe its meaning and how we use it:\\
\textcolor{blue}{\small \textit{``We now seek to elucidate the distinctive properties of our JT-DS model by comparing its performance to a DS-based Cartesian motion generator + IK solver approach for behaviors \textit{(5-7)}. For the Cartesian motion generator we use the SEDS (Stable Estimator of Dynamical Systems) approach [8] which learns an asymptotically stable DS in Cartesian space from demonstrations. We then generate joint trajectories through a damped least square IK solver. From herein we refer to this approach as SEDS+IK."}}\\
\end{enumerate}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %

\underline{Comments from Reviewer \#3 (Reviewer: 7 (Review30521)):}
\begin{enumerate}
\item \textcolor{blue}{\textit{``a considerable part of the introduction is spent in comparing joint space trajectory generation to task
space trajectory generation. Since these are elementary concepts in robotics, the authors might consider
reducing this part"}} \texttt{Response:} Several sentences in the introduction have been removed/reduced.\\

\item \textcolor{blue}{\textit{``one of the main claims of the paper is the stability proof, which in fact is cited in the first line of the
abstract. However the stability analysis is a quite simple extension of the stability of the transpose
Jacobian scheme, where the constant gain on the task space error is replaced by the matrix A(q). As
such the emphasis on this result might be to some extent reduced. By the way, the proof works only
if the Jacobian is full rank, which is not properly discussed in Appendix I"}}\\
\texttt{Response:}  Removed a couple mentions of the stability result, and added the dependence on the rank of the Jacobian to both the appendix and the main body of the text.\\

\item \textcolor{blue}{\textit{``the meaning of variables $q_d$ in (2) is not very clear to me, a well as whether it takes the same meaning
as $q_{t;n}$ in (8)"}}\\
\texttt{Response:} $q_d$ are the velocities from the demonstrations, which we are aiming to recreate. Equation (8) has been updated to remove the typo (they are in fact the same).\\

\item \textcolor{blue}{\textit{``Section III uses the concept of synergy without actually defining it. An introduction to this concept is given later, in Section IV"}}\\
\texttt{Response:} Added the following to Section 3:\\
\textcolor{blue}{\small \textit{``Each ``synergy" is a joint-space transformation that biases the resulting motion to use particular joints; for more, see Sec. IV."}}\\

\item \textcolor{blue}{\textit{``the expression "flow of motion" in Proposition 1 is not very clear to me"}}\\
\texttt{Response:}  Changed to "motion".\\

\item \textcolor{blue}{\textit{`some details about the optimization problem in (8) might be given
(nature of the problem, complexity, feasibility)."}}\\
\texttt{Response:} The following has been added near Eq. (8):\\
\textcolor{blue}{\small \textit{``The resulting optimization minimizes the mean squared joint velocity error from (2), resulting in a convex semidefinite optimization with a quadratic objective..." ``This optimization is always feasible, the only constraint is that the $A_k$s be independently PSD."}}\\

\end{enumerate}

\underline{List of Changes in the Re-Submitted Manuscript:}
\begin{enumerate}
\item Clarified text, variables and inconsistencies throughout the text as pointed out by the reviewers. Exact location of re-writes are mentioned in the responses above.
\item Re-wrote abstract, introduction and parts of problem statement to highlight the fact that our controller can reach targets in both \underline{position and orientation}.
\item Updated Table I, by redoing the experiments and reporting performance of the learning approach in both \underline{position} \underline{and orientation}.
\end{enumerate}

\bibliographystyle{plain}
\bibliography{Refs.bib}

\end{document}

### 2018-05-15 paper/final-source-for-zip/main_RA-L_final.tex
Final update, folder restructure, final proof of published paper

--- 2018-03-15 paper/submitted-v2-revised/main_RA-L_v2_revised.tex
@@
 %\documentclass[9pt,technote,fleqn]{IEEEtran}
-\documentclass[letterpaper, 10 pt, conference,fleqn]{ieeeconf}
+\documentclass[letterpaper, 10 pt, journal, twoside, fleqn]{IEEEtran}
 
@@
 \begin{document}
+
+% Paper headers 
+\markboth{IEEE Robotics and Automation Letters. Preprint Version. Accepted April, 2018}
+{Shavit \MakeLowercase{\textit{et al.}}: Learning Augmented Joint-Space Task-Oriented Dynamical Systems}
 %
@@
 %\title{Learning Joint-Space Stable Dynamical Systems for Task-Space Objectives}
-\title{Learning Augmented Joint-Space Task-Oriented Dynamical Systems: \\ A
+\title{Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A
 Linear Parameter Varying and Synergetic Control Approach}
@@
 %
-\author{Yonadav~Shavit*, Nadia~Figueroa*, Seyed~Sina~Mirrazavi~Salehian*, Aude~Billard
+\author{Yonadav~Shavit$^{*1}$, Nadia~Figueroa$^{*2}$, Seyed~Sina~Mirrazavi~Salehian$^{*2}$, Aude~Billard$^{2}$
 % <-this % stops a space
-\thanks{*These authors contributed equally to this work.}
-\thanks{Y. Shavit is with the Department
-of Electrical Engineering and Computer Science, Massachusetts Institute of Technology, Cambridge,
-MA, 02139 USA. E-mail: yonadav.shavit@gmail.com.}% <-this % stops a space
-\thanks{N.
-Figueroa, S.
-S.
-M.
-Salehian, and A.
-Billard are with the Swiss Federal Institute of Technology (EPFL), 1015 Lausanne, Switzerland.
-\\E-mail: \{nadia.figueroafernandez,sina.mirrazavi,aude.billard\}@epfl.ch.}
-}% <-this %
+\thanks{Manuscript received: November 21, 2017; Revised March 16, 2018; Accepted April 21, 2018.}%Use only for final RAL version
+\thanks{This paper was recommended for publication by Editor Han Ding upon evaluation of the Associate Editor and Reviewers' comments. 
+This work was supported by EU project Cogimon H2020\textendash ICT\textendash 23\textendash2014.}
+\thanks{$^*$These authors contributed equally to this work.}
+\thanks{$^{1}$Y. Shavit is with the Department of Electrical Engineering and Computer Science, Massachusetts Institute of Technology, Cambridge,
+MA, 02139 USA {\tt\footnotesize yonadav.shavit@gmail.com}}% <-this % stops a space
+\thanks{$^{2}$ N. Figueroa, S. S. M. Salehian, and A. Billard are with the Swiss Federal Institute of Technology (EPFL), 1015 Lausanne, Switzerland
+{\tt\footnotesize\{nadia.figueroafernandez,sina.mirrazavi,aude.billard\}\\ \tt\footnotesize @epfl.ch}}%
+\thanks{Digital Object Identifier (DOI): see top of this page.}%
+}
+% <-this %
 
 \maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
+%\thispagestyle{empty}
+%\pagestyle{empty}
 
@@
 
-%\begin{IEEEkeywords}
-%Dynamical Systems, Imitation learning, Robot Kinematics, Gaussian Mixture Models
-%\end{IEEEkeywords}
+\begin{IEEEkeywords}
+Motion Control, Learning from Demonstration, Kinematics, Gaussian Mixture Models, Joint-Space Control
+\end{IEEEkeywords}
 
@@
 \label{sec:intro}
-Robot motion planning in joint-space has long been a major field of study \cite{kelly2006control}.
+\IEEEPARstart{R}{obot} motion planning in joint-space has long been a major field of study \cite{kelly2006control}.
 For manipulation problems with an objective defined in task-space, we can often find a myriad of joint-space trajectories to achieve the same task-space goal.
@@
 
-In many cases, however, one may want to specify a joint-space behavior that the physical robot should follow, while simultaneously reaching a task-space objective.
+In many cases, however, one may want to specify a joint-space behavior that the physical robot should follow while simultaneously reaching a task-space objective.
 For example, learned joint-space behaviors are helpful in playing ping-pong \cite{huang2016jointly}, grasping \cite{calinon2010learning}, and avoiding self-collisions of bi-manual manipulators \cite{silverio2017learning}.
@@
 Moreover, learning a motion in joint space allows us to avoid inverse kinematic (IK) approximations.
-Task-space motion generators, as the ones learned from demonstrations, all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}.
+Task-space motion generators learned from demonstrations all rely on projecting the desired task-space velocity into joint-space via Jacobian Pseudo-Inverse IK approximations and variants thereof \cite{kelly2006control}.
 When the main focus is on executing a specific task-space behavior, regardless of a joint-space constraint, this approach is sufficient \cite{figueroa2016HRIrolling,ureche2015taskconst}.
@@
 \centering
-\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{../Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{../With_Obstcle.png}
+\includegraphics[scale=.155,trim={0 0 0 0cm},clip]{./figures/Without_Obstcle_white_new.png}\includegraphics[scale=.125,trim={0 0 0 0cm},clip]{./figures/With_Obstcle.png}
 \caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl).
@@
 However, because the two DS have their own unique attractors, the non-linear interaction between them imposed by the kinematic constraints does not ensure that the combined DS has a unique attractor.
-This gives rise to spurious attractors or cycles, and thus must be carefully tuned in order to avoid them.
+This gives rise to spurious attractors or cycles, and thus requires careful tuning to avoid them.
 
 In this work, we seek to devise an augmented \textbf{J}oint-space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem (JT-DS) that not only incorporates task-space attractors, but also avoids the problems generated by pseudo-inverse approximations.
-To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different stable behavior in joint space, and modulates the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system.
+To approximate this DS, we further propose an algorithm to learn a set of \emph{behavior synergies}, each of which corresponds to a different stable behavior in joint space, and modulate the use of these synergies throughout joint-space using a learned Linear Parameter Varying (LPV) system.
 We determine the scheduling parameters for the LPV by finding a \emph{lower-dimensional embedding} of the joint space, which accounts for the variation in the demonstrated motions.
@@
 \item computes a \begin{bf}motion in joint-space\end{bf} that provably and asymptotically converges to a \begin{bf}task-space \end{bf}target.
-\item is formulated such that \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations in \textbf{synergy space}.
+\item is formulated such that \textbf{joint-space behaviors} can be \textbf{learned} from demonstrations as \textbf{synergies}.
 \item can transit through kinematic \textbf{singularities}.
@@
 However, despite some previous work designing velocity adjustments by hand \cite{Shi2016}, this is, to the best of our knowledge the first work to employ a JT system to learn behaviors from demonstrations.
-Furthermore, by formulating our LPV system on a latent embedding (via dimensionality reduction schemes), we are capable of discovering meaningful local behavior synergy regions, while being robust to outliers, noise and redundancies that might arise from raw demonstrations.
+Furthermore, by formulating our LPV system on a latent embedding (via dimensionality reduction schemes), we are able to discover meaningful local behavior synergy regions, while being robust to outliers, noise and redundancies that might arise from raw demonstrations.
 This leads to a compelling improvement in generalization of the demonstrated behavior, as opposed to learning the LPV system solely in joint space.
@@
 We refer to the system's joint-space position as $q=\begin{bmatrix} q^1 & \dots & q^m
-\end{bmatrix}^T \in \mathbb{R}^m$, and to its task-space target as $x\in \mathbb{R}^d$.\footnote{For pure position target we consider $d=3$, for position and orientation $d=9$, where the first 3 dimensions correspond to Cartesian position and the remaining 6 correspond to the first and second columns of a rotation matrix $R \in SO(3)$.} The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics is indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{d\times m}$.
+\end{bmatrix}^T \in \mathbb{R}^m$, and to its task-space target as $x\in \mathbb{R}^d$.\footnote{For pure position targets we consider $d=3$, for position and orientation $d=9$, where the first 3 dimensions correspond to Cartesian position and the remaining 6 correspond to the first and second columns of a rotation matrix $R \in SO(3)$.} The kinematics of the robot are assumed to be known, hence, the robot's forward kinematics are indicated by $x = H(q)$ and its Jacobian is $J(q) = \frac{dx}{dq}\in \mathbb{R}^{d\times m}$.
 We wish to formulate a DS $\dot{q} = f(q)$ which satisfies the following two criteria:
@@
 $\dot{V}(q) < 0 ~\forall q\in Q$ and $V(q)=0~ \forall q\in Q^*$ where $Q^*=\{q|H(q)=x^*\wedge q\in Q\}$ and $Q=\{q|q\notin Q^*, J(q) \textit{ is full rank}\}$.
-$V(.)$ can be thought of as a metric for the task-space distance-to-go.
+$V(\cdot)$ can be thought of as a metric for the task-space distance-to-go.
 \item The DS should encapsulate the desired joint-space behaviors such that the following metric is minimized
@@
 Specifically, the ``local synergy matrices" $A_k$ are joint-space transformations that bias the resulting motion to use particular joints.
-By modulating each local synergy in time as well as space through the activation function $\theta_k(\cdot)$ we can generate the desired non-linear joint-space behaviors.
-Note that, in the LPV paradigm $\theta_k(\cdot)$ can be a function of time $t$, ``joint-posture" $q$, an external signal $d(t)$ or a lower-dimensional representation of the joint posture $\phi(q)$.
+By modulating each local synergy in time as well as space through the activation functions $\theta_k(\cdot)$ we can generate the desired non-linear joint-space behaviors.
+Note that in the LPV paradigm $\theta_k(\cdot)$ can be a function of time $t$, ``joint-posture" $q$, an external signal $d(t)$ or a lower-dimensional representation of the joint posture $\phi(q)$.
 Because we seek a time-invariant controller and know that ``joint-postures" can be accurately represented in a lower-dimensional space, in this work we chose to parametrize $\theta_k(\cdot)$ with $\phi(q)$.
@@
 	\centering
-	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
+	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{./figures/Schematic_final.pdf}
 	\caption{Illustration of our task-space and proposed joint-space Lyapunov function.
@@
 
-Before proving that \eqref{eq:ds} satisfies all of the criteria, one can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the task-space error w.r.t.
-the target, which we then reinterpret as the task-space velocity of a proportional controller.
-By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig.
+Before proving that \eqref{eq:ds} satisfies all of the criteria, one can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the task-space error wrt.
+the target\footnote{If the target defines an orientation, this is the Euclidean distance between the two $9$-dimensional position-and-orientation vectors.}.
+We derive the task-space velocity of a proportional controller minimizing that error by multiplying by $-1$.
+Then, by multiplying this task-space velocity by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control \cite{wolovich1984computational,sciavicco1988solution}), see Fig.
 \ref{fig:error_schematic}.
@@
 encoding specific joint-space behaviors) is achieved by embedding the desired joint-space behavior in the matrices $A_k(q)~ \forall k \in\{1,\dots,K\}$.
-We describe in the next section an approach to automatically learn the number of $K$ matrices, their values and their corresponding activation functions $\theta_k(\cdot)$ from demonstrated data.
+We describe in the next section an approach to automatically learn the number of matrices $K$, their values and their corresponding activation functions $\theta_k(\cdot)$ from demonstrated data.
 
-\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystem in Synergy Space} 
+\section{Learning \textbf{J}oint-Space \textbf{T}ask-oriented \textbf{D}ynamical \textbf{S}ystems in Synergy Space} 
 \label{Sec:Learning}
@@
 
-A central question becomes, how to modulate the synergies in different regions to yield our desired behavior?
-First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that an embedding $\phi(\cdot)$ and embedding space $B: \{b = \phi(a) | a \in A\}$ "efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using a sub-manifold of the joint-space, called the \emph{embedding space}.
+A central question arises: how do we modulate the synergies in different regions to yield our desired behavior?
+First, we assume that our desired behavior can be efficiently described\footnote{Given some original space $A$ and some behavior policy $\pi_A(a)$ in $A$, we say that an embedding $\phi(\cdot)$ and embedding space $B: \{b = \phi(a) | a \in A\}$ ``efficiently describe" $\pi_A$ if there exists some policy $\pi_B(b)$ in $B$ such that we can deterministically reconstruct $\pi_A(a)$ given $\pi_B(\phi(a))$} using a sub-manifold of the joint-space, called the \emph{embedding space}.
 We would like to define the robot's policy in embedding space such that in different regions of the space, we will prioritize different synergies.
@@
 	\centering
-	\includegraphics[width=\linewidth]{../Pic/A_comparison_positions.pdf}
+	\includegraphics[width=\linewidth]{./figures/A_comparison_positions.pdf}
 	\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
@@
 \subsection{Embedding Joint Configurations in Low-Dim. Space}
-The search for a lower-dimensional embedding of the joint space stems from the desire to identify a simplified coordinate system in which each principal component corresponds to an important source of variation in the demonstrated trajectories, and which is thus suitable for learning the demonstrated behavior.
+The search for a lower-dimensional embedding of the joint space stems from the desire to identify a simplified coordinate system in which each principal component corresponds to an important source of variation in the demonstrated trajectories, and which is thus suitable for parameterizing the demonstrated behavior.
 Motor control studies have postulated that human arm motions like reaching or following straight/curved line trajectories, rather than utilizing the full joint space $|q|$, are the result of compromising between planning a straight line in the task space and a straight line in the joint space \cite{Cruse1987humanarm,Okadome1999arm}.
@@
 The learned embedding $\phi(\cdot)$ maps a joint configuration $q \in \mathbb{R}^m$ into a lower-dimensional configuration $z \in \mathbb{R}^p$, where $p < m$.
-For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(q)$ were a matrix multiplication; i.e.
-$\phi(q) = A_p\times q$ for $A_p \in \mathbb{R}^{p\times m}$, it could map the ``shoulder" and ``arm" components of $q$ into a single ``shoulder-arm" component in $\phi(q)$.
+For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(q)$ were a matrix multiplication (i.e.
+$\phi(q) = A_p\times q$ for $A_p \in \mathbb{R}^{p\times m}$), it could map the ``shoulder" and ``arm" components of $q$ into a single ``shoulder-arm" component in $\phi(q)$.
 
@@
 The next step is identifying the regions of space in which to activate different synergies.
-Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding activation function $\theta_k(\phi(q))$.
-Moreover, we would like for $\theta_k(\phi(q))$ to have the following properties: (i) $0 \prec \theta_k(\phi(q))$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
+Given the set of projected joint position trajectories $D=\{\{\phi(q_{t,n})\}_{t=1,\dots, T_n}\}_{n=1,\dots,N}$ where $\phi(q_{t,n})$ is the lower-dimensional embedding of $q_{t,n}$, $t$ is the time-step and $N$ is the number of demonstrations, we seek to learn a set of regions of distinct local synergies, each defined by their corresponding activation function $\theta_k(\phi(q))$.
+Moreover, we would like for $\theta_k(\phi(q))$ to have the following properties: (i) $\theta_k(\phi(\cdot))>0$ and (ii) $\sum_{k=1}^{K}\theta'_k(\phi(q)) = 1$.
 Such activation functions, also known as scheduling parameters for LPV systems, have been modeled in previous work as probability distributions \cite{7439839, Salehian-RSS-16}.
 Intuitively, we search for a probabilistic model that ``explains" the variance in the demonstrated trajectories, and treat each cluster as expressing a \textit{local behavior}, which the synergy will then approximate.
-In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the activation parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and Covariance matrices that parametrize the $k$-th multivariate Gaussian distribution.
+In this work, we adopt this approach and use a GMM to estimate the joint distribution over the projected joint positions\footnote{It must be noted that, although we present GMM as the approach to estimate the activation parameters, alternative algorithms can be used.}, $p(\phi(q)) = \sum_{k=1}^K\pi_k\mathcal{N}(\phi(q);\mu_k,\Sigma_k)$, where $\pi_k$ are the prior probabilities and $\{\mu_k,\Sigma_k\}$ are the mean and covariance matrices that parametrize the $k$-th multivariate Gaussian distribution.
 Each distribution represents a local region of projected joint positions $\phi(q)$, and will be used to construct the activation functions $\theta_k$ of the $k$th synergy matrix.
@@
 
-Intuitively, each ``synergy region" is defined by a Gaussian distribution in the lower-dimensional space, the closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
-Finding the appropriate synergy matrices $A_k$ to accurately reproduce the observed behaviors can be reduced to a semidefinite program with the goal of minimizing \eqref{eq:optim_error}.
+Intuitively, each ``synergy region" is defined by a Gaussian distribution in the lower-dimensional space.
+The closer the robot is to a region, the more that region's synergy ($A_k$) influences the robot's current joint-space motions.
+Finding the appropriate synergy matrices $A_k$ to accurately reproduce the observed demonstrations can be reduced to a semidefinite program with the goal of minimizing \eqref{eq:optim_error}.
 The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective, as follows:
@@
  \eqref{eq:ds} with \eqref{eq:A}, and $x^*_{n}$ is defined as the endpoint of the $ n^{\text{th}} $ demonstrated trajectory.
-This optimization is always feasible, the only constraint is that the $A_k$s be independently PSD.
+This optimization is always feasible; the only constraint is that the $A_k$s be independently PSD.
 
@@
 		\centering
-			\includegraphics[width=0.5\linewidth]{../new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{../new-pics/backhand_crop.png}
+			\includegraphics[width=0.5\linewidth]{./figures/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{./figures/backhand_crop.png}
 					\caption{Forward/Backward Reaching (1-2).\label{fig:forback}}
@@
 \centering
-	\includegraphics[width=0.45\linewidth]{../new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{../new-pics/pour-obst_crop.png}
+	\includegraphics[width=0.45\linewidth]{./figures/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{./figures/pour-obst_crop.png}
 						\caption{ Pouring Motions (3-4)}
@@
 \centering
-	\includegraphics[width=0.45\linewidth]{../new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{../new-pics/footstep.png}
+	\includegraphics[width=0.45\linewidth]{./figures/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{./figures/footstep.png}
 	\caption{ Pouring Motion (5) and Foot-step (6).}
@@
 	\multicolumn{1}{c}{\multirow{3}{*}{(1) Forward Reaching}} & None & 7 & 5 $\pm$ (2.90) & \textbf{ 	0.234 $\pm$ (0.027) } & \cellcolor{blue!5} 0.633 $\pm$ (0.458)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1424$)}} & PCA & 4.1 $\pm$ (0.316) & 4.8 $\pm$ (2.347) & 0.276 $\pm$ (0.033)  & \cellcolor{blue!10} 0.589 $\pm$ (0.367) \\	
-		& K-PCA ($\sigma_{opt} = 5.488$) & \textbf{3.9 $\pm$ (0.316)} & \textbf{3.2 $\pm$ (0.421)} & 0.305 $\pm$ (0.021) & \cellcolor{blue!15} 0.433 $\pm$ (0.104)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1424$)}} & PCA & 3.1 $\pm$ (0.316) & 3.8 $\pm$ (1.549) & 0.247 $\pm$ (0.017)  & \cellcolor{blue!10} 0.484 $\pm$ (0.263) \\	
+		& K-PCA ($\sigma_{opt} = 5.111$) & \textbf{3 $\pm$ (0)} & \textbf{3.5 $\pm$ (1.204)} & 0.2532 $\pm$ (0.019) & \cellcolor{blue!15} 0.385 $\pm$ (0.098)  \\
 		\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(2) Backward Reaching}} & None & 7 & 5.1 $\pm$ (2.643) &\textbf{ 0.192 $\pm$ (0.033)} &  \cellcolor{blue!5} 0.728 $\pm$ (0.709)  \\
-		    \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1223$)}} & PCA & 3.1 $\pm$ (0.316) & 3.5 $\pm$ (0.971) & 0.216 $\pm$ (0.021) & \cellcolor{blue!10} 0.469 $\pm$ (0.164) \\	
-			& K-PCA ($\sigma_{opt} = 6.238$) & \textbf{3.1 $\pm$ (0.316)} & \textbf{3.1 $\pm$ (0.316)} & 0.219 $\pm$ (0.013) & \cellcolor{blue!15} 0.355 $\pm$ (0.068)   \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(2) Backward Reaching}} & None & 7 & 5.8 $\pm$ (2.097) &\textbf{ 0.281 $\pm$ (0.015)} &  \cellcolor{blue!5} 0.647 $\pm$ (0.354)  \\
+		    \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1223$)}} & PCA & 4.4 $\pm$ (0.516) & 6 $\pm$ (3.091) & 0.282 $\pm$ (0.044) & \cellcolor{blue!10} 0.466 $\pm$ (0.102) \\	
+			& K-PCA ($\sigma_{opt} = 6.238$) & \textbf{3.5 $\pm$ (0.5)} & \textbf{3.1 $\pm$ (0.3)} & 0.319 $\pm$ (0.025) & \cellcolor{blue!15} 0.482 $\pm$ (0.114)   \\
 			\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(3) Pouring - Free}} & None & 7 & \textbf{ 4.1 $\pm$ (1.286)} & 0.163 $\pm$ (0.026) & 7.609 $\pm$ (18.99) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($N=9,M=1032$)}} & PCA & \textbf{ 2.3 $\pm$ (0.483)} & 4.8 $\pm$ (2.044) & \textbf{ 0.161 $\pm$ (0.034) } & \cellcolor{blue!10} 0.623 $\pm$ (0.301) \\	
-	& K-PCA ($\sigma_{opt} = 3.92$) & 3 & 4.3 $\pm$ (1.767) & 0.171 $\pm$ (0.106) & \cellcolor{blue!15} 0.511 $\pm$ (0.305)   \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(3) Pouring - Free}} & None & 7 &  4.6 $\pm$ (1.897) & \textbf{0.181 $\pm$ (0.019)} & 1.099 $\pm$ (1.140) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($N=9,M=1032$)}} & PCA & 2.6 $\pm$ (0.5164) & 4 $\pm$ (1.054) &  0.186 $\pm$ (0.026)  & \cellcolor{blue!10} 0.419 $\pm$ (0.119) \\	
+	& K-PCA ($\sigma_{opt} = 3.92$) & \textbf{2} & \textbf{3.5 $\pm$ (0.5)} & 0.183 $\pm$ (0.0204) & \cellcolor{blue!15} 0.397 $\pm$ (0.105)   \\
 				\hline
-	\multicolumn{1}{c}{\multirow{3}{*}{(4) Pouring - Obstacle 1}} & None & 7 & 6.2 $\pm$ (2.573) & \textbf{0.247 $\pm$ (0.039)} & 3.049 $\pm$ (3.113) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1232$)}} & PCA & 3.8 $\pm$ (0.421) & 3.8 $\pm$ (1.549) & 0.292 $\pm$ (0.018) & \cellcolor{blue!10} 0.486 $\pm$ (0.125) \\	
-				& K-PCA ($\sigma_{opt} = 7.695$) & \textbf{ 3.1 $\pm$ (0.421) } & \textbf{ 3.1 $\pm$ (0.316) } & 0.288 $\pm$ (0.020) & \cellcolor{blue!15} 0.482 $\pm$ (0.078)   \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(4) Pouring - Obstacle 1}} & None & 7 & 7 $\pm$ (2.357) & \textbf{0.296 $\pm$ (0.029)} & 0.984 $\pm$ (0.55) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($N=11,M=1232$)}} & PCA & 3.9 $\pm$ (0.316) & 3.6 $\pm$ (0.699) & 0.321 $\pm$ (0.028) & \cellcolor{blue!10} 0.402 $\pm$ (0.0587) \\	
+				& K-PCA ($\sigma_{opt} = 7.695$) & \textbf{ 3 } & \textbf{ 3.2 $\pm$ (0.4) } & 0.311 $\pm$ (0.011) & \cellcolor{blue!15} 0.388 $\pm$ (0.040)   \\
 				\hline	
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(5) Pouring - Obstacle 2}} & None & 7 & 4.3 $\pm$ (1.418) & \textbf{ 0.122 $\pm$ (0.014) } & 3.095 $\pm$ (3.406) \\
-   \multicolumn{1}{c}{\multirow{3}{*}{($N=7,M=1406$)}} & PCA & \textbf{ 2.9 $\pm$ 0.316 } & 3.7 $\pm$ (2.213) & 0.126 $\pm$ (0.02) & \cellcolor{blue!10} 0.825 $\pm$ (0.764) \\	
-	& K-PCA ($\sigma_{opt} = 2.86$) & 3.9 $\pm$ (0.316) & \textbf{3} & 0.129 $\pm$ (0.013) & \cellcolor{blue!15} 0.738 $\pm$ (0.507)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(5) Pouring - Obstacle 2}} & None & 7 & 4.1 $\pm$ (1.7288) & \textbf{ 0.1242 $\pm$ (0.0169) } & 1.0907 $\pm$ (1.0686) \\
+   \multicolumn{1}{c}{\multirow{3}{*}{($N=7,M=1406$)}} & PCA & \textbf{ 2.8 $\pm$ 0.4216 } & 3.6 $\pm$ (1.8974) & 0.1345 $\pm$ (0.0218) & \cellcolor{blue!10} 0.693 $\pm$ (1.0541) \\	
+	& K-PCA ($\sigma_{opt} = 2.86$) & 3.4 $\pm$ (0.4899) & \textbf{3.1 $\pm$ (0.3)} & 0.1345 $\pm$ (0.0182) & \cellcolor{blue!15} 0.4028 $\pm$ (0.1377)  \\
 				\hline
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(6) Foot Step}} & None & 7 & 4.4 $\pm$ (2.366) & \textbf{ 0.080 $\pm$ (0.008) } & 0.368 $\pm$ (0.265)  \\
-	\multicolumn{1}{c}{\multirow{3}{*}{($N=8,M=1058$)}} & PCA & \textbf{1} & 3.2 $\pm$ (0.421) & 0.084 $\pm$ (0.004) & \cellcolor{blue!10}  0.123 $\pm$ (0.011) \\	
-	& K-PCA ($\sigma_{opt} = 1.513$) & 3.2 $\pm$ (0.421) & \textbf{3} &  0.091 $\pm$ (0.007) & \cellcolor{blue!15} 0.115 $\pm$ (0.015)   \\
+	\multicolumn{1}{c}{\multirow{3}{*}{(6) Foot Step}} & None & 7 & 4.2 $\pm$ (1.4757) & \textbf{ 0.1396 $\pm$ (0.0252) } & 1.0697 $\pm$ (0.6574)  \\
+	\multicolumn{1}{c}{\multirow{3}{*}{($N=8,M=1058$)}} & PCA & \textbf{1} & 3.1 $\pm$ (0.3162) & 0.1494 $\pm$ (0.0143) & \cellcolor{blue!10}  0.2578 $\pm$ (0.0795) \\	
+	& K-PCA ($\sigma_{opt} = 1.513$) & 2 & \textbf{3} &  0.1557 $\pm$ (0.0116) & \cellcolor{blue!15} 0.2271 $\pm$ (0.0466)   \\
 				\hline		
-
-	\multicolumn{1}{c}{\multirow{3}{*}{(7) Singularity Motions}} & None & 7 & 6.2 $\pm$ (2.898) & 0.054 $\pm$ (0.009) & 0.137 $\pm$ (0.098)  \\
-    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1467$)}} & PCA & \textbf{2} & 6 $\pm$ (2.357) & \textbf{ 0.053 $\pm$ (0.010) } & \cellcolor{blue!10} 0.105 $\pm$ (0.046) \\	
-	& K-PCA ($\sigma_{opt} = 1.769$) & 3.6 $\pm$ (0.516) & \textbf{5.3 $\pm$ (2.452)} & 0.055 $\pm$ (0.018) & \cellcolor{blue!15} 0.071 $\pm$ (0.010)   \\ \hline\hline
+	\multicolumn{1}{c}{\multirow{3}{*}{(7) Singularity Motions}} & None & 7 & 6.4 $\pm$ (1.5055) & \textbf{0.048 $\pm$ (0.0158)} & 0.2365 $\pm$ (0.0768)  \\
+    \multicolumn{1}{c}{\multirow{3}{*}{($N=10,M=1467$)}} & PCA & \textbf{1.9 $\pm$ (0.3162)} & 5.7 $\pm$ (1.6364) &  0.0503 $\pm$ (0.0153) & \cellcolor{blue!10} 0.1802 $\pm$ (0.0748) \\	
+	& K-PCA ($\sigma_{opt} = 1.769$) & 3.9 $\pm$ (0.3) & \textbf{5.2 $\pm$ (2.1817)} & 0.0593 $\pm$ (0.0122) & \cellcolor{blue!15} 0.1276 $\pm$ (0.0591)   \\ \hline\hline
 \end{tabular}}
@@
 \begin{itemize}[leftmargin=*]
-\item \textit{(1-2) Forward/Backward reaching:} We guide the robot to  forward and backward reaching motions towards a specific target in task-space with joint-space trajectories mimicking forehand/backhand strokes as shown in Fig.
+\item \textit{(1-2) Forward/Backward reaching:} We guide the robot in forward and backward reaching motions towards a specific target in task-space with joint-space trajectories mimicking forehand/backhand strokes as shown in Fig.
 \ref{fig:forback}.
@@
 	\begin{subfigure}[t]{0.42\linewidth}
-		\includegraphics[width=0.45\linewidth]{../Pic/Luggage1.png}
-		\includegraphics[width=0.143\linewidth]{../Pic/Luggage2.png}
-		\includegraphics[width=0.377\linewidth]{../Pic/Luggage3.png}
+		\includegraphics[width=0.45\linewidth]{./figures/Luggage1.png}
+		\includegraphics[width=0.143\linewidth]{./figures/Luggage2.png}
+		\includegraphics[width=0.377\linewidth]{./figures/Luggage3.png}
 		\caption{Execution of Pour Obstacle 2 learned through JT-DS.}
@@
 	\begin{subfigure}[t]{0.567\linewidth}
-		\includegraphics[width=0.33\linewidth]{../Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{../Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{../Pic/Luggage3_SEDS.png}
+		\includegraphics[width=0.33\linewidth]{./figures/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{./figures/Luggage2_SEDS.png}\includegraphics[width=0.33\linewidth]{./figures/Luggage3_SEDS.png}
 		\caption{Execution of Pour Obstacle 2 learned through SEDS.}
@@
 
-Performing such cross-validation in our case, is however not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $p$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
+Performing such cross-validation in our case is not trivial, as our $\mathcal{A}(q)$ formulation has several hyper-parameters: the dimensionality $p$ of our lower-dimensional embedding $\phi(q)$ and the number of local behavior synergies $K$.
 Moreover, when using K-PCA (with RBF kernel $k(q,q') = \exp(-\frac{||q-q'||^2}{2\sigma^2})$) we must also find the optimal width $\sigma$.
@@
 		\centering
-		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{../new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{../new-pics/singularity_crop.png}
+		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{./figures/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{../new-pics/singularity_crop.png}
 		\caption{Execution of \textit{problematic} joint-space behaviors learned through JT-DS: (left) Foot step-like motion and (right) Singularity motions.}
@@
 		\centering
-		\includegraphics[width=0.7\linewidth]{../Pic/Foot.pdf}
+		\includegraphics[width=0.7\linewidth]{./figures/Foot.pdf}
 		\caption{End-effector trajectories for the footstep motion in Cartesian space.
@@
 The YALMIP framework \cite{lofberg2004yalmip} was used to solve the definite convex optimization problem.
-Source code for the entire learning pipeline and cross-validation with simulation of multi-DOF robot arms can be found in:
+Source code for the entire learning pipeline and simulated behavior on multi-DOF robot arms can be found in:
 \footnotesize  \centerline{\underline{\url{https://github.com/epfl-lasa/JT-DS-Learning}}}
 \normalsize
-
 For execution of the learnt JT-DS models on a real 7-DOF the KUKA LWR 4+ robot arm, we provide the following code in C++:
@@
 The resultant joint angles are filtered by a critically damped filter to avoid high torques.
-
 \subsection{JT-DS Execution Performance Evaluation}
@@
 We then generate joint trajectories through a damped least square IK solver.
-From herein we refer to this approach as SEDS+IK.
+From hereon we refer to this approach as SEDS+IK.
 
@@
 folding the elbow, raising the shoulder), successfully avoiding the obstacle while still converging to the desired Cartesian position.
-Meanwhile, SEDS+IK only learned the demonstrated behavior in task-space, it does not have the capabilities of constraining motion in joint-space.
+Meanwhile, SEDS+IK only learned the demonstrated behavior in task-space; it is incapable of constraining motion in joint-space.
 This ultimately led to one of its joints colliding with the obstacle.
@@
 One of the most important points when validating a learning from demonstration method is to evaluate the system's behavior away from demonstrations.
- When the current joint configuration is far from any of the local synergy regions, computing the activation parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so $\frac{1}{K}\sum\limits_{k=1}^{K}A_k$ is used to move the robot, which is still guaranteed to move towards the target.
+ When the current joint configuration is far from any of the local synergy regions, computing the activation parameters $\theta_k(\cdot)$ becomes numerically infeasible (all the Gaussians in \eqref{eq:A} $\rightarrow 0$), and so we default to $\mathcal{A} = \frac{1}{K}\sum\limits_{k=1}^{K}A_k$, which still guarantees that the robot moves towards the target.
 %Moreover, in overlapping local regions where multiple $A_k$'s might be in conflict, the presented system compromises between them while still stably converging to the target.
@@
 \begin{figure}[t]
-%	\begin{subfigure}[t]{\linewidth}
-		\includegraphics[width=\linewidth]{../Pic/cropped_Sing_1.pdf}
-%		\caption{ }
-%	\end{subfigure}\\
-%	\begin{subfigure}[t]{\linewidth} %
-%		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
-%		\caption{ }
-%		\label{fig:Det}
-%	\end{subfigure}\\
+		\includegraphics[width=\linewidth]{./figures/cropped_Sing_1.pdf}
 	\caption{The experiment generated $K=3$ Gaussian components.
@@
 In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space.
-The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).} %In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
-Due to sensory noise, the determinants are not exactly zero, but are very close.
-}
+The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).}
 			\label{fig:Sin}
@@
 
-Since we learn a lower-dimensional embedding space $Z := \{\phi(q) ~|~q \in Q\}$, which we suggest provides a better representation for the learned joint behaviors, a careful reader might wonder: why not define a motion policy $\pi_Z$ directly in $Z$-space and then map the learned policy back out into joint space using the inverse embedding.
-$\phi^{-1}(z)$?
+Since we learn a lower-dimensional embedding space $Z := \{\phi(q) ~|~q \in Q\}$, which we suggest provides a better representation for the learned joint behaviors, a careful reader might wonder: why not define a motion policy $\pi_Z$ directly in $Z$-space and then map the learned policy back out into joint space using the inverse embedding $\phi^{-1}(z)$?
 The answer is that we would lose any guarantee of stability with respect to a target attractor $x_t^*$.
@@
 Intuitively, this is because any policy defined in a lower-dimensional space than the actuation space will forfeit certain degrees of freedom, and thus may not be able to span the configuration space.
-Instead, we use the embedding space to modulate synergies defined in joint space, and which we can thus guarantee will be positive definite and lead us to convergence to the target.
+Instead, we use the embedding space to modulate synergies defined in joint space, which we can guarantee will be positive definite and lead us to convergence to the target.
 
@@
 \footnotesize
-This work was supported by EU project Cogimon H2020\textendash ICT \textendash 23\textendash2014.
-The authors would like to thank A.
-Karimi for his insightful comments about formulating the convex optimization problem.
+The authors would like to thank A. Karimi for his insightful comments about formulating the convex optimization problem.
 

### 2018-05-15 paper/reply_reviews/old_iterations/Synergy_analysis_reviewer2.tex
Final update, folder restructure, final proof of published paper

--- 2018-03-14 paper/reply_reviews/New_reply_to_comments.tex
@@
 \documentclass{article}
-\usepackage[left=2cm, right=2cm, top=2cm]{geometry}
+\usepackage[left=3cm, right=3cm, top=2cm]{geometry}
 \usepackage{amsmath}
@@
 \usepackage{color}
-\title{Response to Reviewers: Manuscript ID 17-1142.1: ``Learning Augmented Joint-Space Task-Oriented Dynamical Systems: A Linear Parameter Varying and Synergetic Control Approach"}
+\title{A Small \LaTeX{} Article Template\thanks{To your mother}}
+\author{Your Name  \\
+	Your Company / University  \\
+	\and 
+	The Other Dude \\
+	His Company / University \\
+	}
 
-%\author{Your Name  \\
-%	Your Company / University  \\
-%	\and 
-%	The Other Dude \\
-%	His Company / University \\
-%	}
-%
-%\date{\today}
+\date{\today}
 % Hint: \title{what ever}, \author{who care} and \date{when ever} could stand 
@@
 \begin{document}
-\maketitle
-First of all, we would like to thank all reviewers for the time they spent reviewing our paper.
-We greatly appreciate the comments, suggestions and opinions that we received.
-We have addressed all the necessary comments and suggestions in the attached manuscript accordingly.\\
-% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
-\noindent \underline{Comments from Reviewer \#1 (Reviewer: 4 (Review30069)):}
-\begin{enumerate}
-\item \textcolor{blue}{\textit{``As mentioned by the authors, there are many reasons to plan in joint space rather than task space.
-One of the important reasons to me seems to be able to plan end-effector orientation, not just position.
-Traditional movement primitives, for example DMPs, do not explicitly encode planning in the SO(3) space of orientations.
-There is research that addresses this question, for example [1], but they lose the simplistic appeal of movement primitives.
-I think it would be a great addition to this paper if the authors illustrate if their joint space DS can be used to accomplish orientation sensitive tasks, and generalize to end-effector orientations different that what was demonstrated.[1] Ude, Ale, et al.
-"Orientation in cartesian space dynamic movement primitives." Robotics and Automation (ICRA), 2014 IEEE International Conference on.}}"\\
-\texttt{Response:}  We highly appreciate this comment.
-When we initially proposed this approach, we did not think about tackling the problem of reaching a target in orientation as well.
-Since our DS is learned in joint-space, indeed, we do not have to deal with the challenges that arise when trying to reach a target in both position and orientation as we do not have the constrains imposed by SO(3).
-There was theoretically no modification needed in our controller:
+\vspace{-10pt}
+In response to Comments 1-3 from \underline{Reviewer: 5 (Review30371)} following we mathematically and graphically analyze the proposed local behavior synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via.
+PCA or K-PCA) and grouping of local behavior regions with GMM on this lower dimensional space.
+The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$.
+Let's begin by analyzing the proposed DS-based control-law:
 \begin{equation}
@@
 \end{equation}
-Originally, when working solely in position $d=3$, now we represent the target with $d=9$, where the first 3 dimensions correspond to Cartesian position and the remaining 6 correspond to the first and second columns of a rotation matrix $R \in SO(3)$.
-This only modifies the Jacobian $J(q)$ which is now $J(q)\in \mathbb{R}^{9 \times 7}$ rather than $J(q)\in \mathbb{R}^{3 \times 7}$.
-The role of the $\mathcal{A}(q)$ and their meaning is the same.
-
-\item \textcolor{blue}{\textit{``In the introduction, the authors stress that jacobian psuedo inverse
-can be computationally expensive. Is this really true for a 7-dof robot
-arm? Can the authors provide some numbers where this would really be an
-issue for a robot controlled at 1000Hz? I find the argument that
-jacobian inverse does not work for singularities and needs a set of
-heuristics to be more compelling. Maybe the authors should rewrite the
-introduction to stress more on this issue?"}}\\
-\texttt{Response:} \small For a 7-DOF robot it is indeed negligible.
-Computational burden only becomes a real issue for higher degrees of freedom, like a multi-fingered hand or multi-arm robotic setting.
-We will remove this passage from the text as we don't have much space to discuss it....\\
-
-\item \textcolor{blue}{\textit{``In Section II, criterion I, the authors talk about a positive symmetric matrix P. I do not see it in the
-equations above. Is this a weighing matrix?"}}\\
-\texttt{Response:} \small Removed, that was a carry-over from a previous formulation.\\
-
-\item \textcolor{blue}{\textit{``In criterion II, the error is defined with respect to the velocity error, rather than a position or a combination
-of both. This means that the motion can be different from the demonstrated motion. I understand
-that this is important to keep the later optimization convex, but is there any disadvantage to using
-velocity error rather than position error?"}}\\
-\texttt{Response:} \small Optimizing for position distance alone would have the effect of learning a motion that converged to the average of previous motions, thus not changing the motion in new situations or fitting the joint velocity profile that is often a desired component of the motion.
-Using velocity error does mean that, for motions strongly dependent on joint position rather than joint velocity, the generated motions may generalize from the demonstrations in unexpected ways.
-However, as you mentioned, minimizing position error would, depending on the formulation, be prohibitively difficult.
-This would imply simulating the learned motion forward in time to recompute the robot's positions, making the optimization much more complicated; this way, we can treat each point in the motion independently.\\
-
-\item \textcolor{blue}{\textit{`I do not completely understand the significance of A(q). From Figure 3, it seems like A(q) helps resolve
-redundancy, while achieving the task space goal. Do the authors have a mathematical (in constrast to
-intuitive) explanation for this? Does A(q) somehow relate to the null-space of the jacobian?"}}\\
-\texttt{Response:} \small A(q) transforms a joint velocity vector approximately aligned with the task-space error into a different joint velocity vector that remains approximately aligned with the error, but which we want to choose to more-closely approximate our demonstrations.
-In this sense, it is resolving redundancy between all the possible joint velocity vectors that would move the controller closer to the origin.
-The paragraph in Section III has been modified to improve clarity.
-A(q) doesn't relate to the null-space of the Jacobian in any clear way, as its value can and does alter the task-space trajectory as well as the joint space trajectory.
-The only property that A(q) preserves is the joint velocity's moving the task space position closer to the goal, but the direction of motion is not necessarily the same as the direction of task-space error.\\
-
-\item \textcolor{blue}{\textit{``In Section III, ?One can intuitively understand the control law as follows:$(H(q) - x^*)$ denotes the
-position error w.r.t. the task target and by multiplying that error by the transposed Jacobian JT (q),
-the error is projected into joint space (similar to Jacobian transpose control)?. I find this explanation
-misleading as jacobian multiplied to position error does not project the error into joint space. Jacobian
-transforms velocities, not positions. Maybe the law can be seen more as a first order approximation
-of the error in task space. In any case, I would either clarify this statement more or take it out of the
-manuscript all together. It seems like the DS is just a PD law that provably converges to a fixed attractor
-in task space."}}\\
-\texttt{Response:} \small The passage has been rewritten to remove the word "project", which you appropriately noted was incorrect, and further clarifies the nature of the DS as a PD controller.
-We have added the following text in the manuscript:
-\textit{``One can intuitively understand the control law as follows: $(H(q) - x^*)$ denotes the position error w.r.t.
-the task target, which we then reinterpret as the task-space velocity of a proportional controller.
-By multiplying that error by the transposed Jacobian $J^T(q)$, it is transformed into a joint-space velocity vector correlated with the error (similar to Jacobian transpose control [?]), see Fig.
-[?].
-The positive definite matrix $\mathcal{A}(q)$ warps the resulting joint-space velocity; Fig.
-[?] illustrates the effects of  $\mathcal{A}(q)$ on the generated motion.
-Thus the controller can be thought of as a proportional controller in joint space.}\\
-
-\item \textcolor{blue}{\textit{``About figure 2 again, the position error in task and joint space are not simply multiplied by a Jacobian.
-Although, task space error can be first-order approximated using a Jacobian multiplication with
-joint space error. What does the right figure in Fig 2 exactly represent?"}}\\
-\texttt{Response:} \small The figure on the left is simply the cartesian error in task space.
-The figure on the left is the Lyapunov potential function on which the jacobian transpose controller descends.
-In showing the attractive regions in joint-space corresponding to this Lyapunov function, we aim to offer insight into how the controller reformulates the task-space potential into a different yet related joint-space potential.
-The caption has been rewritten to add clarity.\\
-
-\item \textcolor{blue}{\textit{``In Section IV, the authors say that they use EM to determine the parameters of the GMM in Eq 6. Can
-they elaborate a little on this? What is the expectation taken over and what likelihood is maximized?"}}\\
-\texttt{Response:} \small The algorithm computes the expected probability of generating each of the joint positions given the current values of the GMM's means, variance matrices, and scaling values.
-The algorithm then chooses GMM parameters to maximize the posterior likelihood of the model, using previously-computed expected component membership as likelihood weights when computing the new mixture.
-Specifically, we used MATLAB's \textit{fitgmdist} implementation.
-Added a citation to clarify.\\
-
-\item \textcolor{blue}{\textit{``The learned dimensionality reduction and parameters seem task dependant. I wonder if the authors
-have any idea on how to generalize this across tasks."}}\\
-\texttt{Response:} \small The dimensionality reduction could be generalized across tasks that had the same motor synergies - that is, for example, all tasks in which the shoulder and elbow always move together.
-Similarly, one could imagine several similar motions that could be decomposed into shared submotions, thus allowing the parameters to be shared.
-However, in general, fixing some of the parameters while optimizing will simply lead to a less-flexible model, and considering that there is no intuitive scheme for a-priori identifying motions with strongly similar submotions, we did not explore this question further.
-To answer your question directly, if one wanted to, one could reuse the learned dimensionality reduction for a new task which you believed to be similar, or copy certain components of a previous motion's GMM into the new GMM and leave them fixed throughout the optimization.\\
-
-\item \textcolor{red}{\textit{``The experiments have the same task-space goal as the
-demonstrations. Can the authors include experiments that have goals
-different from demonstrated, especially in orientation?"}}\\
-\texttt{Response:} \small First of all, they don't.. show some figures... add clarification of this in the text and explain why we cannot systematically evaluate this if we had too..\\
-
-\item \textcolor{red}{\textit{``In similar light, maybe authors could present online adaptation of
-A(q) to generalize to situations different than demonstrated, for
-example a slightly moved obstacle?"}}\\
-\texttt{Response:} \small This is a really nice suggestion.
-But how I see it, it's basically a new paper; i.e.
-an extension of this one, but we can discuss about this if we have space.
-A quick idea that comes to mind would be to apply a modulation/activation to the LPV system in a local region.
-Now this is not as straight forward as in our task space approaches because our LPV system is learned in joint-space while the activation function is in the embedded space.
-Another idea is to modify the activation function of the local behavior regions such that it is a function of an external variable, for example if force is applied/sensed, and it bias the model towards another $A_k$ matrix?
-So it still would be an open question, but an approach to tackle this problem could be along these lines....
-Add a passage in the discussion answering this question\\
-
-\item \textcolor{red}{\textit{``In table II, instead of the tracking error, maybe the more useful
-error to see would be the error with respect to the plan or
-demonstartion. One would assume that the proposed approach would be
-worse at it than SEDS+IK because the minimized error is with respect to
-the velocity and goal, not the position of the movement."}}\\
-\texttt{Response:} \small This is not necessarily the reason that it would be worse, as SEDS is also learned by minimizing the error with respect to
-the velocity and the goal.
-The reason that JT-DS is less accurate in reproducing the motion in task-space is because our optimization is focused on reproducing the motion solely in joint-space, while reaching the final target in task-space.
-Nowhere do we claim that JT-DS is capable of reproducing a motion in both joint and task space.
-Nevertheless, we add the mean swept-error-arrea (SEA) which evaluate the performance of reproduction in task-space from a reference trajectory...
-as shown in the tables....\\
-\end{enumerate}
-
-% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % 
-\underline{Comments from Reviewer \#2 (Reviewer: 5 (Review30371)):}
-\begin{enumerate}
-\item \textcolor{blue}{\textit{``My first concern regards you choice of representing the demonstrated
-movement as a sequence of postures and hence looking for synergies by
-using PCA of K-PCA analysis. Why not using functional PCA (f-PCA)
-analysis and determine the few moving synergies that make possible the
-execution of the demonstrated behaviours? It would be that, when you
-combine the local synergies in space and time by using functions
-$\theta_k(q)$, you get what you might obtain by using directly f-PCA.
-However, the last one are valid for the whole movement and can be
-combined by constant coefficients."}}\\
-\texttt{Response:}  See below.
-
-\item \textcolor{blue}{\textit{``I also found the paper lacking of physical intuitions, especially in	the experimental part, about the entities introduced in previous sections.
-In particular, what is the lesson we learn after your analysis of synergies and the obtained lower dimensional space for the few movements shown in section V?
-Is it possible to extract some global	synergic behavior?
-Moreover, in section IV.A authors introduce the function phi(q) named the embedding that maps the joint configuration into a lower dimensional space.
-Is it associated to a whole behaviour or, as matrices $A_k$, has a local validity?
-I lost this point.
-Can you show/report or describe what it represents in the joint space for one or more cases in Section V?
-What is this lower dimensional space for the cases in Section V?"}}\\
-\texttt{Response:}  See below.
-
-\item \textcolor{blue}{\textit{``A similar question for the matrices $A_k$ that represent local, I
-	would say postural (see previous comment), synergies whose dimension is
-	m x m, i.e. in joint space: I would expect they were defined in the
-	lower dimensional space. Moreover, are they local representation of the
-	distribution of postures and hence covariance matrices? What do their
-	eigenvectors and eigenvalues locally represent? Maybe, a comparison
-	with the classical and more known postural synergies for the human
-	hand, analyzed by PCA, could be useful. Finally, what does it mean that
-	they should be positive definite (apart from stability issue)?"}}\\
-\texttt{Response:}  See below.\\
-
-\item \textcolor{blue}{\textit{``What is the dimension of function $\theta_k(q)$ that, to some extent, can be seen as the new control variables. Is it $K<m$?"}}\\
-
-\texttt{Response:} In response to Comments 1-4, following we mathematically and graphically analyze the proposed local behavior synergy extraction scheme which is composed of dimensionality reduction of the joint trajectories (via.
-PCA or K-PCA) and grouping of local behavior regions with GMM on this lower dimensional space.
-The $K$ clusters found in this embedded space correspond to the number of $K$ synergy matrices that will parametrize $\mathcal{A}(q)$.
-Let's begin by analyzing the proposed DS-based control-law \eqref{eq:jtds}.
-The intuition behind this control-law is that $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\phi(q))A_k$ is a linear combination of time-invariant linear matrices $A_k \in \mathbb{R}^{m \times m}$, \underline{these matrices are the ``local behavior synergy matrices"} that shape the motion in joint-space.
+The intuition behind this control-law is that $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta(\phi(q))A_k$ is a linear combination of time-invariant linear matrices $A_k \in \mathbb{R}^{m \times m}$, \underline{these matrices are the ``local behavior synergy matrices"} that shape the motion in joint-space.
 Meaning that, given the joint-space velocity vector representing the task-space error $J^{T}(q)(H(q)-x^*)$, the resulting motion is biased to use a particular set of joints, defined in each $A_k$.
-Each $A_k$ is thus activated depending on the current ``joint-posture" $q$ (represented in the lower-dimensional space $\phi(q)$) via the scheduling/activation function $\theta_k(\phi(q))$.
-$\phi(q)$ is a learned transformation, which is associated with a whole behavior, rather than a particular $A_k$ (as is now clarified in IV.A).
-However, the precise nature of the transformation depends on the DR technique: PCA applies a uniform matrix multiplication of joint position everywhere, whereas KPCA is a more complicated function of joint position.
-The result of the embedding is a lower-dimensional representation of the joint position.
-The following line has been added to IV.A for clarity:
-\textcolor{blue}{\small \textit{"For example, if the shoulder and arm joints are coupled throughout the motion, and $\phi(\cdot)$ were a matrix multiplication, it could map the "shoulder" and "arm" components of $q$ into a single "shoulder-arm" component in $\phi(q)$.}"}
+Each $A_k$ is thus activated depending on the current ``joint-posture" $q$ (represented in the lower-dimensional space $\phi(q)$) via the scheduling/activation function $\theta(\phi(q))$.
+Hence, our proposed definition of ``local behavior synergies" is \underline{not analogous} to the notion of ``hand posture synergies".
 
 Given the raw demonstrations of a joint-space  behavior with a task-space target as shown in Fig.
-\ref{fig:raw_demos} (for the pouring task) we seek to learn a linear combination of local behavior synergies $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta_k(\phi(q))A_k$ such that we can use \eqref{eq:jtds} to accurately reproduce these demonstrations.
+\ref{fig:raw_demos} (for the pouring task) we seek to learn a linear combination of local behavior synergies $\mathcal{A}(q) = \sum\limits_{k=1}^{K}\theta(\phi(q))A_k$ such that we can use \eqref{eq:jtds} to accurately reproduce these demonstrations.
 In order to do so, we project the raw joint trajectories to a lower-dimensional space via $\phi(q)$, in the case of PCA $\phi(q)=M_p \times q$ for $M_p \in \mathbb{R}^{p \times m}$.
@@
 \ref{fig:raw_demos}.
-We then jointly learn the activation function $\theta_k(\cdot)$ and the number of $K$ local synergy regions by fitting a GMM to this representation of the joint posture trajectories.
+We then jointly learn the activation function $\theta(\cdot)$ and the number of $K$ local synergy regions by fitting a GMM to this representation of the joint posture trajectories.
 As can be seen in Fig.
@@
 Meaning that we only need 3 synergy matrices $A_k$ to represent the complex joint-space motion accurately.
-The activation function $\theta_k(\cdot)$, as described in the mansucript, is the posterior probability of a joint posture in PCA-space $\phi(q)$ belonging to one of the local synergy regions $k \in \{1,\dots,K\}$.
-As mentioned in Sec.
-V.A., we treat K as a hyperparameter and thus choose it differently for each learned motion.
-For motions that do not vary much throughout the course of the motion, smaller K might be sufficient, but K can just as easily be greater than $m$ (though as $K$ increases, so does the learning problem's computational cost).
-If comment-4 was about each specific $\theta_k(q)$, they are scalars, as is mentioned in Sec.
-III.
+The activation function $\theta(\cdot)$, as described in the mansucript, is the posterior probability of a joint posture in PCA-space $\phi(q)$ belonging to one of the local synergy regions $k \in \{1,\dots,K\}$.
 
@@
 Yet, since the demonstrations are acquired via kinesthetic teaching we can assume that they are manipulable.
-\underline{As asked by the reviewer, what does it physically mean to have all positive eigenvalues?}  When \\ $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$  or ($\mathcal{A}(q)$) has some eigenvalues $=0$ the physical meaning behind it is that we are forfeiting the control of these directions of motion in joint-space.
-Namely, the desired velocity directions in joint-space will be constrained in the corresponding DOF of the arm, this will prohibit us from converging to the attractor and will be no longer capable of accurately following the joint-space demonstrations.
+
+\noindent \underline{As asked by the reviewer, what does it physically mean to have all positive eigenvalues?}  When \\ $M_p^{-1}\big(\sum_{k=1}^{K}\theta_k(q)A_k\big)M_p$  or ($\mathcal{A}(q)$) has some eigenvalues $=0$ the physical meaning behind it is that we are forfeiting the control of these directions of motion in joint-space.
+Namely, the desired velocity directions in joint-space will be constrained in the corresponding DOF of the arm, this will prohibit us from converging to the attractor and no longer capable of accurately following the joint-space demonstrations.
 If instead we learn the synergy matrices in joint-space following \eqref{eq:jtds} and modulate them with the activation function in PCA-space, we avoid all of this problems and can ensure full controllability of the motion in joint-space.
@@
 As shown in TABLE I of the manuscript and discussed in Section V, by using a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
-This is explicitly highlighted in Section V:\\
-\textcolor{blue}{\small \textit{"As can be seen, for all datasets there is a significant increase in performance on the testing sets when using either dimensionality reduction (DR) approaches.
-This suggests that using DR to encode our activation functions $\theta_k$ in a lower-dimensional space $\phi(q)$ yields better generalization capabilities than encoding the behaviors in using the original $q$.
-This is most notable for the three pouring motions, where the joint-velocity RMSE testing error for a JT-DS model learned without DR is an order of magnitude higher than with DR.
-Such an error indicates that the demonstrated joint-behavior was over-fitted on the training set, which is also exhibited in the higher number of $K$ needed to represent the motion without DR.
-For all datasets, the DR methods provided $\delta < d/2$, either comparable or less number of local behaviors synergies $K$ and better RMSE errors on testing sets as opposed to no DR.
-By finding a lower-dimensional manifold to represent the joint trajectories, we are getting rid of outliers, noise and redundancies that might arise from the raw joint demonstrations.
-Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations."}} \\
 Hence, through DR we are capable of robustly extracting the local behavior synergies from raw demonstrations and learn a \textit{time-invariant} controller that ensures reaching the task-space target while approximately following the demonstrated motion, as shown in Fig.
@@
 The joint-positions are then computed by numerical integration of these velocities.
-Note that our controller does not generate this entire trajectory in one-shot, it is generating the next desired velocity given the current joint posture, e.g.
-$\dot{q}_{t+1} = \mathcal{A}(q_{t})J^T(q_{t})(H(q_{t})-x^*)$ with $q_t$ being the current joint configuration.
-Moreover, our controller is state-dependent only; i.e.
-it has no dependence on time or the duration of the desired trajectories.
-Hence, it can generate motions similar to the demonstrated ones at different frequencies and time durations.
-
-\noindent \textcolor{red}{TODO: Add clarifications in text such that no other reader gets confused with the new concept of behavior synergies that we propose as opposed to the classical joint posture synergies.}
 
@@
 \end{figure*}
-\newpage
+
+Note that our controller is not generating the entire trajectory in one-shot, it is generating the next desired velocity given the current joint posture, e.g.
+$\dot{q}_{t+1} = \mathcal{A}(q_{t})J^T(q_{t})(H(q_{t})-x^*)$ with $q_t$ being the current joint configuration.
+Moreover, our controller is state-dependent only; i.e.
+it has no dependence on time or the duration of the desired trajectories.
+Hence, it can generate motions similar to the demonstrated ones at different frequencies and time durations.
+\\
+
 \noindent \underline{On the applicability of F-PCA to our learning scheme:}  If we were to ignore our proposed control scheme and try to reconstruct the trajectories from the joint postural synergy components; i.e.
@@
 
-Ignoring this limitation, one of the main draw-backs of this type of synergistic control is that the extracted representation in PCA space does not take into account the dynamic features of the motions and further assumes that the demonstrations have no time correlation.
+Ignoring our limitation, one of the main draw-backs of this type of synergistic control is that the extracted representation in PCA space does not take into account the dynamic features of the motions and further assumes that the demonstrations have no time correlation.
 As mentioned by the reviewer and applied in \cite{Dai:IROS:2013} for modeling grasping motions and in \cite{Averta:FRAI:2017} for upper-limb motions, functional-PCA is a nice alternative to include these dynamic features and further reconstruct entire trajectories with a linear combination of a few \textit{time-dependent} functions.
@@
 
-\item \textcolor{blue}{\textit{``For the test, why not comparing also with the JT control method? In
-other terms, is it possible to use a SEDS + JT, especially for the
-third case "Transiting through singular configurations"?"}}\\
-\texttt{Response:} \small This could be done, but it is mathematically incorrect.
-The original JT control method is a DS that converges to a unique equilibrium point, i.e.
-the task-space target $x^*\in \mathbb{R}^{N\times 1}$, with a gain matrix $K \in \mathbb{R}^{M\times M}$ as:\begin{equation}
-\dot{q} = K J(q)^T(H(q)- x^*) 
-\label{eq:JT}
-\end{equation}
-SEDS on the other hand, is a DS-based motion generator that converges also to a unique equilibrium point, as:
-\begin{equation}
-\hfill \begin{aligned}
-&  \dot{x}  = \mathbf{f}(x)
-\rightarrow  &
-\begin{cases}
-\lim_{t \rightarrow \infty}\|x - x^*\|=0
-\end{cases}
-\end{aligned}  \hfill
-\label{eq:simple-ds}
-\end{equation}
-where $\mathbf{f}(\cdot): \mathbb{R}^M \rightarrow \mathbb{R}^M$ is a continuous differentiable vector-valued function representing a non-linear DS that converges to a single stable equilibrium point $x^*$.
-SEDS is used in the following way: Given the current state of the robot $x$ the next desired velocity $\dot{x}_d$ is computed with \eqref{eq:simple-ds} this is then fed to a velocity-based IK solver as shown in Fig.
-\ref{}:\\
-\textcolor{red}{TO SINA: Include typical SEDS-like control-loop here}\\
-If we were to use SEDS with the JT-control method, we would have to integrate the desired velocity to generate a desired position $x_d$ at each time-step and define it as the target in \eqref{eq:JT}, which would yield the following controller:
-\begin{equation}
-\dot{q} = K J(q)^T(H(q)- x_d) 
-\end{equation} Note that, the JT control method assumes that the target is static, hence, in order for this controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
-This is contradicting as \eqref{eq:simple-ds} is always trying to converge to $x^*$.
-Thus, merging these two approaches is theoretically ill-fitting.
-To better highlight this, we show an illustration of the supposed SEDS+JT control loop in Fig.
-\ref{}:\\
-\textcolor{red}{TO SINA: Include SEDS+JT control-loop here}\\
-\textcolor{red}{\small Describe this control-loop and state more reasons (if any) why it's wrong...}
-
-\item \textcolor{blue}{\textit{``Minors:
-- SEDS + IK: what does SEDS mean?"}}\\
-\texttt{Response:} As stated in the previous comment, SEDS, short for Stable Estimator of Dynamical Systems is an approach proposed in \cite{khansari2011learning} that learns a stable estimate of a DS with the form of \eqref{eq:simple-ds} from demonstrations.
-The first paragraph of Section V.B was written to describe its meaning and how we use it:\\
-\textcolor{blue}{\small \textit{``We now seek to elucidate the distinctive properties of our JT-DS model by comparing its performance to a DS-based Cartesian motion generator + IK solver approach for behaviors \textit{(5-7)}.
-For the Cartesian motion generator we use the SEDS (Stable Estimator of Dynamical Systems) approach [8] which learns an asymptotically stable DS in Cartesian space from demonstrations.
-We then generate joint trajectories through a damped least square IK solver.
-From herein we refer to this approach as SEDS+IK."}}\\
-\end{enumerate}
-% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-
-\underline{Comments from Reviewer \#3 (Reviewer: 7 (Review30521)):}
-\begin{enumerate}
-\item \textcolor{blue}{\textit{``a considerable part of the introduction is spent in comparing joint space trajectory generation to task
-space trajectory generation. Since these are elementary concepts in robotics, the authors might consider
-reducing this part"}} \texttt{Response:} Several sentences in the introduction have been removed/reduced.\\
-
-\item \textcolor{blue}{\textit{``one of the main claims of the paper is the stability proof, which in fact is cited in the first line of the
-abstract. However the stability analysis is a quite simple extension of the stability of the transpose
-Jacobian scheme, where the constant gain on the task space error is replaced by the matrix A(q). As
-such the emphasis on this result might be to some extent reduced. By the way, the proof works only
-if the Jacobian is full rank, which is not properly discussed in Appendix I"}}\\
-\texttt{Response:}  Removed a couple mentions of the stability result, and added the dependence on the rank of the Jacobian to both the appendix and the main body of the text.\\
-
-\item \textcolor{blue}{\textit{``the meaning of variables $q_d$ in (2) is not very clear to me, a well as whether it takes the same meaning
-as $q_{t;n}$ in (8)"}}\\
-\texttt{Response:} $q_d$ are the velocities from the demonstrations, which we are aiming to recreate. Equation (8) has been updated to remove the typo (they are in fact the same).\\
-
-\item \textcolor{blue}{\textit{``Section III uses the concept of synergy without actually defining it. An introduction to this concept is given later, in Section IV"}}\\
-\texttt{Response:} Added the following to Section 3:\\
-\textcolor{blue}{\small \textit{``Each ``synergy" is a joint-space transformation that biases the resulting motion to use particular joints; for more, see Sec. IV."}}\\
-
-\item \textcolor{blue}{\textit{``the expression "flow of motion" in Proposition 1 is not very clear to me"}}\\
-\texttt{Response:}  Changed to "motion".\\
-
-\item \textcolor{red}{\textit{`some details about the optimization problem in (8) might be given
-(nature of the problem, complexity, feasibility)."}}\\
-\texttt{Response:} The following has been added near Eq. (8):\\
-\textcolor{blue}{\small \textit{``The resulting optimization minimizes the mean squared joint velocity error from \eqref{eq:optim_error}, resulting in a convex semidefinite optimization with a quadratic objective..." ``This optimization is always feasible (the only constraint is that the $A_k$s be independently PSD), and the number of terms in the objective is $O\left(k^2d^3\left( \sum_{n=1}^N T_n\right)\right)$."}}\\
-To elaborate about the prior calculation: the number of terms in the sum is proportional to the number of recorded points, times the number of matrices (k) required to generate the predicted velocity from (7) squared, times the number of terms in a matrix multiplication of two $d x d$ matrices ($d^3$).\\
-
-\end{enumerate}
-
 \bibliographystyle{plain}

### 2018-05-15 paper/reply_reviews/old_iterations/reply_comments.tex
Final update, folder restructure, final proof of published paper

--- 2018-03-13 paper/reply_reviews/reply_comments.tex
@@
 where $\mathbf{f}(\cdot): \mathbb{R}^M \rightarrow \mathbb{R}^M$ is a continuous differentiable vector-valued function representing a non-linear DS that converges to a single stable equilibrium point $x^*$.
-SEDS is used in the following way: Given the current state of the robot $x$ the next desired velocity $\dot{x}_d$ is computed with \eqref{eq:simple-ds} this is then fed to a velocity-based IK solver as shown in Fig.
-\ref{}:}\\
-\textcolor{red}{TO SINA: Include typical SEDS-like control-loop here}\\
-\textcolor{blue}{ \small If we were to use SEDS with the JT-control method, we would have to integrate the desired velocity to generate a desired position $x_d$ at each time-step and define it as the target in \eqref{eq:JT}, which would yield the following controller:
-\begin{equation}
-\dot{q} = K J(q)^T(H(q)- x_d) 
-\end{equation} Note that, the JT control method assumes that the target is static, hence, in order for this controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
+In order to prove the stability of SEDS, one needs to assume that the dynamic of the robot is perfectly compensated by the built-in position-level controller.
+In other words, the transfer function of the dynamic of the robot+ the low-level controller and the IK solver must be one.
+\\
+Almost all the commercially available robots are equipped by a very precises joint position controller.
+Hence, the transfer function of the robot+ the low-level controller can be assumed one.
+Whereas, the transfer function of the Jacobian transpose IK solver can not be assumed one as the JT control method assumes that the target is static.
+Hence, in order for the JT controller to actually converge, the output of \eqref{eq:simple-ds} must seem constant for a period of time.
 This is contradicting as \eqref{eq:simple-ds} is always trying to converge to $x^*$.
-Thus, merging these two approaches is theoretically ill-fitting.
-To better highlight this, we show an illustration of the supposed SEDS+JT control loop in Fig.
-\ref{}:}\\
-\textcolor{red}{TO SINA: Include SEDS+JT control-loop here}\\
-\textcolor{red}{\small Describe this control-loop and state more reasons (if any) why it's wrong...}
+Thus, merging these two approaches is theoretically ill-fitting.\\
+ Note that,  To better highlight this, we show an illustration of the supposed SEDS+JT and JT-DS control loops in Figure \ref{fig:schematic:SEDS} and  Figure \ref{fig:schematic:jt}, respectively.}\\
+
+\begin{figure}[t]
+\centering
+	\begin{subfigure}[t]{0.8\textwidth}
+	\centering
+		\includegraphics[trim={0 0 0.25cm 0.0cm},clip,width=\textwidth]{Pics/Drawing2.png}
+		\caption{ A Cartesian motion generator (e.g., SEDS) with the Jacobian transpose IK solver.}
+		\label{fig:schematic:SEDS}
+	\end{subfigure}\\
+	\begin{subfigure}[t]{0.8\textwidth}
+	\centering
+		\includegraphics[trim={0.25cm 0 0.0cm 0.0cm},clip,width=\textwidth]{Pics/JT_DS.png}
+		\caption{ JT-DS.}
+		\label{fig:schematic:jt}
+	\end{subfigure}\\
+	\caption{A simplified schematic comparing the standard control architecture using Cartesian dynamical systems with the Jacobian transpose IK solver (top) and our propose approach.
+By integrating the "inverse kinematic (IK) solver" and "motion generator" blocks into a unified block, the proposed approach eliminates undesirable effects of the IK solver.
+}
+\end{figure}
 

### 2018-05-15 paper/submitted-v3/main_RA-L_v3.tex
Final update, folder restructure, final proof of published paper

--- 2018-05-15 paper/final-source-for-zip/main_RA-L_final.tex
@@
 \centering
-\includegraphics[scale=.155,trim={0 0 0 0cm},clip]{./figures/Without_Obstcle_white_new.png}\includegraphics[scale=.125,trim={0 0 0 0cm},clip]{./figures/With_Obstcle.png}
+\includegraphics[scale=.16,trim={0 0 0 0cm},clip]{../Without_Obstcle_white_new.png}\includegraphics[scale=.135,trim={0 0 0 0cm},clip]{../With_Obstcle.png}
 \caption{Two robot motions in \textcolor{orange}{\textbf{joint-space}} accomplishing \textit{similar} behavior in \textcolor{cyan}{\textbf{task-space}} (pouring chips in a bowl).
@@
 	\centering
-	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{./figures/Schematic_final.pdf}
+	\includegraphics[width=\linewidth,trim={2.5cm 1.75cm 2.5cm 2cm},clip]{../Pic/Sources/Schematic_final.pdf}
 	\caption{Illustration of our task-space and proposed joint-space Lyapunov function.
@@
 	\centering
-	\includegraphics[width=\linewidth]{./figures/A_comparison_positions.pdf}
+	\includegraphics[width=\linewidth]{../Pic/A_comparison_positions.pdf}
 	\caption{Three example 3-DOF motions (A, B, C), each with a different constant joint augmentation matrix $\mathcal{A}(q)$ (emphasizing the hip, knee, and ankle respectively), A: $\mathcal{A}(q)=diag(5,1,1) $, B: $\mathcal{A}(q)=diag(1,5,1)$, and C: $\mathcal{A}(q)=diag(1,1,5)$.
@@
 		\centering
-			\includegraphics[width=0.5\linewidth]{./figures/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{./figures/backhand_crop.png}
+			\includegraphics[width=0.5\linewidth]{../new-pics/forehand_crop.png}\hspace{5pt}\includegraphics[width=0.43\linewidth]{../new-pics/backhand_crop.png}
 					\caption{Forward/Backward Reaching (1-2).\label{fig:forback}}
@@
 \centering
-	\includegraphics[width=0.45\linewidth]{./figures/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{./figures/pour-obst_crop.png}
+	\includegraphics[width=0.45\linewidth]{../new-pics/pour-free_crop.png}\hspace{5pt}\includegraphics[width=0.477\linewidth]{../new-pics/pour-obst_crop.png}
 						\caption{ Pouring Motions (3-4)}
@@
 \centering
-	\includegraphics[width=0.45\linewidth]{./figures/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{./figures/footstep.png}
+	\includegraphics[width=0.45\linewidth]{../new-pics/pour-obst-2_crop.png}\hspace{5pt}\includegraphics[width=0.5\linewidth]{../new-pics/footstep.png}
 	\caption{ Pouring Motion (5) and Foot-step (6).}
@@
 	\begin{subfigure}[t]{0.42\linewidth}
-		\includegraphics[width=0.45\linewidth]{./figures/Luggage1.png}
-		\includegraphics[width=0.143\linewidth]{./figures/Luggage2.png}
-		\includegraphics[width=0.377\linewidth]{./figures/Luggage3.png}
+		\includegraphics[width=0.45\linewidth]{../Pic/Luggage1.png}
+		\includegraphics[width=0.143\linewidth]{../Pic/Luggage2.png}
+		\includegraphics[width=0.377\linewidth]{../Pic/Luggage3.png}
 		\caption{Execution of Pour Obstacle 2 learned through JT-DS.}
@@
 	\begin{subfigure}[t]{0.567\linewidth}
-		\includegraphics[width=0.33\linewidth]{./figures/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{./figures/Luggage2_SEDS.png}\includegraphics[width=0.33\linewidth]{./figures/Luggage3_SEDS.png}
+		\includegraphics[width=0.33\linewidth]{../Pic/Luggage4_SEDS.png}\includegraphics[width=0.33\linewidth]{../Pic/Luggage3_SEDS.png}\includegraphics[width=0.33\linewidth]{../Pic/Luggage3_SEDS.png}
 		\caption{Execution of Pour Obstacle 2 learned through SEDS.}
@@
 		\centering
-		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{./figures/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{../new-pics/singularity_crop.png}
+		\includegraphics[angle=180,origin=c,width=0.5\linewidth]{../new-pics/footstep_execution_crop.png}\hspace{5pt}\includegraphics[width=0.4\linewidth]{../new-pics/singularity_crop.png}
 		\caption{Execution of \textit{problematic} joint-space behaviors learned through JT-DS: (left) Foot step-like motion and (right) Singularity motions.}
@@
 		\centering
-		\includegraphics[width=0.7\linewidth]{./figures/Foot.pdf}
+		\includegraphics[width=0.7\linewidth]{../Pic/Foot.pdf}
 		\caption{End-effector trajectories for the footstep motion in Cartesian space.
@@
 \begin{figure}[t]
-		\includegraphics[width=\linewidth]{./figures/cropped_Sing_1.pdf}
+%	\begin{subfigure}[t]{\linewidth}
+		\includegraphics[width=\linewidth]{../Pic/cropped_Sing_1.pdf}
+%		\caption{ }
+%	\end{subfigure}\\
+%	\begin{subfigure}[t]{\linewidth} %
+%		\includegraphics[width=\linewidth]{Pic/cropped_Sing_2.pdf}
+%		\caption{ }
+%		\label{fig:Det}
+%	\end{subfigure}\\
 	\caption{The experiment generated $K=3$ Gaussian components.
@@
 In (a), the end-effector positions for the demonstrations and executed motions are plotted in Cartesian space.
-The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).}
+The JT-DS trajectory was generated closed-loop, while the SEDS trajectory was generated open-loop (otherwise it would be unstable).} %In (b), the determinants of the Jacobian $(|J(q)J^T(q)|)$ for all $10$ JT-DS motions are superimposed.
+Due to sensory noise, the determinants are not exactly zero, but are very close.
+}
 			\label{fig:Sin}
